### 2022-11-24 paper-1/manuscript.tex
Added draft main text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\RequirePackage{tikz}

\documentclass[pdflatex,iicol,sn-basic]{sn-jnl}

%%%% Standard Packages
\usepackage{xcolor,hyperref}
\usepackage[autostyle=false, style=english]{csquotes}
%%%%

% Avoid having to ``quote'' verbosely (csquotes package)
\MakeOuterQuote{"}

% Make Orcid icon
\definecolor{lime}{HTML}{A6CE39}
\DeclareRobustCommand{\orcidicon}{%
	\begin{tikzpicture}
	\draw[lime, fill=lime] (0,0) 
	circle [radius=0.16] 
	node[white] {{\fontfamily{qag}\selectfont \tiny ID}};
	\draw[white, fill=white] (-0.0625,0.095) 
	circle [radius=0.007];
	\end{tikzpicture}
	\hspace{-2mm}
}

\foreach \x in {A, ..., Z}{%
	\expandafter\xdef\csname orcid\x\endcsname{\noexpand\href{https://orcid.org/\csname orcidauthor\x\endcsname}{\noexpand\orcidicon}}
}

% Define the ORCID iD command for each author separately.
\newcommand{\orcidauthorA}{0000-0002-3420-4576}
\newcommand{\orcidauthorB}{0000-0001-9402-0851}

%%%%%=============================================================================%%%%
%%%%  Remarks: This template is provided to aid authors with the preparation
%%%%  of original research articles intended for submission to journals published 
%%%%  by Springer Nature. The guidance has been prepared in partnership with 
%%%%  production teams to conform to Springer Nature technical requirements. 
%%%%  Editorial and presentation requirements differ among journal portfolios and 
%%%%  research disciplines. You may find sections in this template are irrelevant 
%%%%  to your work and are empowered to omit any such section if allowed by the 
%%%%  journal you intend to submit to. The submission guidelines and policies 
%%%%  of the journal take precedence. A detailed User Manual is available in the 
%%%%  template package for technical guidance.
%%%%%=============================================================================%%%%

\jyear{2021}%

%% as per the requirement new theorem styles can be included as shown below
\theoremstyle{thmstyleone}%
\newtheorem{theorem}{Theorem}%  meant for continuous numbers
%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
\newtheorem{proposition}[theorem]{Proposition}% 
%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.

\theoremstyle{thmstyletwo}%
\newtheorem{example}{Example}%
\newtheorem{remark}{Remark}%

\theoremstyle{thmstylethree}%
\newtheorem{definition}{Definition}%

\raggedbottom
%%\unnumbered% uncomment this for unnumbered level heads

\begin{document}

\title{Short-term synaptic plasticity and threshold adaptation act in synergy to give rise to deviance detection in a spiking model of neuronal culture}

%%=============================================================%%
%% Prefix	-> \pfx{Dr}
%% GivenName	-> \fnm{Joergen W.}
%% Particle	-> \spfx{van der} -> surname prefix
%% FamilyName	-> \sur{Ploeg}
%% Suffix	-> \sfx{IV}
%% NatureName	-> \tanm{Poet Laureate} -> Title after name
%% Degrees	-> \dgr{MSc, PhD}
%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
%%=============================================================%%

\author*[1]{\fnm{Felix Benjamin} \sur{Kern} \orcidA{}}\email{kernfel@gmail.com}

\author[1]{\fnm{Zenas C.} \sur{Chao} \orcidB{}}\email{zenas.c.chao@gmail.com}

\affil[1]{\orgdiv{International Research Center for Neurointelligence (WPI-IRCN)}, \orgname{The University of Tokyo}, \orgaddress{\city{Tokyo}, \country{Japan}}}

%%==================================%%
%% sample for unstructured abstract %%
%%==================================%%

\abstract{TODO}

\keywords{keyword1, Keyword2, Keyword3, Keyword4}

\maketitle

\section{Introduction}\label{sec-intro}

\section{Methods}\label{sec-methods}

\subsection{Model}\label{sec-model}

Our research is based on loose approximation of a matured culture of dissociated neurons. We modeled neurons as leaky integrate-and-fire units whose membrane potential $v$ followed
\begin{equation}
    \tau_m \frac{dv}{dt} = (v_{rest}-v) + I_{syn}
\end{equation}
where
\begin{align}
    I_{syn} &= g_e(E_e-v) + g_i(E_i-v) \nonumber \\
    \tau_e \frac{dg_e}{dt} &= -g_e + \sum_{j \in \boldsymbol E} U x_j w \delta(t - \hat{t_j}) \nonumber \\
    \tau_i \frac{dg_i}{dt} &= -g_i + \sum_{j \in \boldsymbol I} w \delta(t - \hat{t_j}) \label{eqn-syn}
\end{align}
with $\boldsymbol E$ and $\boldsymbol I$ the sets of presynaptic excitatory and inhibitory neurons, respectively, $\delta(\cdot)$ the Dirac delta function, and $\hat{t}$ the time of spikes.

Excitatory, but not inhibitory, synapses were subject to short-term depression (STD), simplified from \cite{Tsodyks1997-qt}; since synaptic transmission was not stochastic, we modeled the depression variable $x_j$ as a property of the presynaptic neuron $j$,
\begin{equation}
    \tau_x \frac{dx_j}{dt} = (1-x_j) - U x_j \delta(t - \hat{t})
\end{equation}
In ablated experiments without STD, $x$ remained constant at $1$, equivalent to $\tau_x = 0$, but $U$ in Equation \ref{eqn-syn} was left unchanged for balance purposes.

When a neuron's membrane potential reached $v \geq v_\theta$, a spike was emitted, and the potential was clamped to $v = v_{reset}$ for a fixed refractory period of 3~ms (excitatory neurons) or 2~ms (inhibitory neurons). For inhibitory neurons, the firing threshold remained constant at $v_\theta = \theta_0$, while for excitatory neurons, the threshold included an adaptive component, following
\begin{align}
    v_\theta &= \theta_0 + \theta \nonumber \\
    \tau_{TA} \frac{dV_{TA}}{dt} &= -V_{TA} + \hat{\theta} \delta(t - \hat{t})
\end{align}
In ablated experiments without TA, $\hat{\theta}$ was set to $0$ to abolish adaptation.

Figure \ref{fig1}C shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration. The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \cite{Stimberg2020-go}, and simulated with an integration time step of 1~ms. Synaptic delay was not modeled explicitly, but spikes were delivered, and voltages reset, in the time step following spike emission. We simulated neurons without any stochasticity in either their inputs or their parameters, other than network structure (see below), in order to ensure that our results were driven purely by the experimental paradigm, rather than any model-internal sources of noise.

800 excitatory and 200 inhibitory neurons were placed randomly in a two-dimensional circular space with a radius of 4~mm, mimicking a petri dish, and formed synapses of weight $w=1$ with 50 randomly selected postsynaptic partners within a range of 2~mm (excitatory) or 1~mm (inhibitory). Connectivity and weights remained fixed throughout simulation. Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each. See Figure \ref{fig1}AB for a representative example of the resulting network structure. Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response, discarding networks where less than half of all neurons in the fully recovered network responded to stimulation at any of the five sites.

\subsection{Paradigm}\label{sec-paradigm}

To investigate deviance detection, we used a classical oddball paradigm \citep{TODO} as illustrated in Figure \ref{fig2}A, presenting a total of 500 stimuli at regular intervals of 500~ms. Target stimuli (labeled A hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (std), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (dev) with 100 A and 400 B, and the many-standards control sequence (msc) with 100 presentations of each of the stimuli A through E. In this way, A is presented as the predominant and therefore expected stimulus (std), as an infrequent violation of an expectation (of B, dev), and, to control for pure adaptation effects, as an equally infrequent stimulus with no strong expectation of any particular input (msc).

In each network, we simulated oddball sequences with two A/B stimulus pairs (sites 1 and 2 in one pair, and sites 3 and 5 in the other pair), where each stimulus site in a pair was used as both target (A) and non-target (B). In addition, we simulated one msc sequence, yielding 4 complete data sets per network, or a total of 120 data sets across all networks. Networks were reset to a fully recovered state between sequence presentations to guarantee their independence.

\section{Results}\label{sec-results}

\subsection{Deviance detection}\label{sec-dd}

We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs. Following related research \citep{TODO}, we were careful to exclude effects due to stimulus identity (e.g. comparing responses to A and B in the dev sequence) and to adaptation to frequent stimulus presentation (e.g. comparing A in the dev sequence to A in the std sequence). In other words, we used the msc sequence as a neutral baseline where the target stimulus was presented no more often than in the dev sequence, but no expectation of a competing stimulus B could be formed. To quantify deviance detection, we compared the mean number of spikes fired across the network in response to target stimulus A in the dev and msc sequences, defining a deviance detection index (DDI) as
\begin{equation}
    DDI = \frac{R^A_{dev} - R^A_{msc}}{R^A_{dev} + R^A_{msc}}
\end{equation}
We define adaptation (msc - std) and mismatch (dev - std) indices in similar fashion, capturing respectively the effect of adaptation, and of both adaptation and deviance detection together.

Our first aim was to replicate deviance detection properties reported in dissociated neuronal cell culture by \cite{Kubota2021-dx}, reproduced in Figure \ref{fig2}B. With minimal tuning of global parameters, we found that most networks (and 92 of 120 target stimuli) reliably produced not only mismatch responses based on adaptation alone, but also showed robust deviance detection, as summarized in Figure \ref{fig2}D.

The model responses developed on a compressed time scale compared to neuronal culture, which we tentatively attribute to unfaithful aspects of the model, including a lack of conduction delay or spontaneous activity, a much smaller number of participating neurons, and structural differences. However, we could readily identify target stimuli whose response profiles looked otherwise qualitatively similar to the cell culture results.

Having confirmed the presence of deviance detection in our model, we then turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model -- synaptic short-term depression (STD) and threshold adaptation (TA) -- enable the networks to learn relevant regularities. As shown in Figure \ref{fig3}, networks with neither plasticity mechanism did not show any difference in response, indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance. Networks with only STD likewise did not show any deviance detection, though they did exhibit robust adaptation effects (data not shown), leading to a smaller target response in std than in either dev or msc sequences. Conversely, with only TA, networks produced both adaptation and deviance detection effects.

Finally, with both STD and TA applied, deviance detection was also reliable, as shown before; however, we were surprised to find that the effect was significantly stronger than that of either plasticity mechanism alone (STD: t = 6.7, p = 4.53e-10; TA: t = 3.2, p = 0.00103), and greater too than the sum of both STD only and TA only DDI values (t = 3.6, p = 0.000199). This suggests that TA and STD act not as independent mechanisms in this paradigm, but rather interact constructively to enhance the network's ability to encode temporal regularities.

In the following, we will use a sample network and target stimulus to first build an understanding of how TA leads to deviance detection when acting alone. Then, we will build on this understanding to elucidate how the addition of STD, which is ineffective on its own, can lead to an increased deviant response. The particular network and target stimulus site used in the following individual analyses was selected for the response size to target and non-target stimuli (A, B), both of which were average, as well as its deviance detection performance, which likewise reflected the tendencies observed across all networks.

\subsection{The role of TA in deviance detection}\label{sec-ta}

To understand why the same stimulus A evoked a higher response in dev than in msc, we first compared the average responses in the two conditions. As shown in Figure \ref{fig4}A, we found that the response developed in a qualitatively similar fashion, with a tendency of dev spikes to occur earlier, particularly shortly after stimulation, and more frequently, particularly towards the later part of the response. Since the stimulus was identical in both conditions, this difference had to be due to either changed firing thresholds, or changed synaptic currents as a consequence of spike differences earlier in the response. We quantified the former with the strength of adaptation $V_{TA}$, which showed a clear and widespread lowering of thresholds at the start of dev trials relative to msc, followed by an increase beyond the msc baseline in many neurons as they spiked (Figure \ref{fig4}B).

To quantify the synaptic contributions to a neuron's membrane potential $v$, we calculated a quantity $V_{syn}$, which captures all synaptic inputs, but excludes voltage deflections caused by direct stimulation and by the self-inhibition induced by resetting to $v_{reset} < v_{rest}$ after a spike (see Figure S\ref{SI-fig1}). Being induced by spiking activity, this measure closely followed the activity pattern, making the earlier activation of the dev response clearly visible across the network (Figure \ref{fig4}C). In addition, since $V_{syn}$ decays with the membrane timescale $\tau_m$, but resets to $0$ on postsynaptic spikes, many of the neurons that spiked more frequently showed lower $V_{syn}$ towards the end of network activity.

Based on these measures, we then asked why there were more spikes in dev trials. To answer this, we weighted the contrasts in $V_{TA}$ and $V_{syn}$ by the spike probability contrast, discarding any instances of equal or lower firing in dev, and summed over neurons, directly linking network-wide increased activity to its immediate underlying causes, de-adaptation and depolarization. We found that the increased depolarization induced by synaptic input greatly outweighed the impact of reduced adaptation throughout most of the response (Figure \ref{fig4}D). Only in the very early stages of the response did adaptation dominate the increase in firing (Figure \ref{fig4}E), whereas the late response was enlarged primarily as a result of increased presynaptic excitation. This trend was reflected in most networks investigated, as shown in the summary plot in Figure \ref{fig4}F. To summarize, while most added spikes arose from increased prior activity within a trial, the initial impetus -- the increase in the early response -- was largely caused by lower TA.

Having established that the increased response to A in dev trials was driven by lower thresholds, we next turned to the cause of this reduction. Firstly, plotting $V_{TA}$ at the start of A trials in Figure \ref{fig5}A, we found that most of the network was indeed less adapted in dev, with only a few neurons in the vicinity of B showing higher thresholds. We hypothesized that this could be due to the different non-target stimulation pattern, since B was stimulated more frequently in dev than msc. Indeed, as shown in Figure \ref{fig5}B, the non-target response activity in the dev sequence -- in other words, the response to B -- was concentrated around the site where B was delivered, while in the msc sequence, activity in response to the various non-target stimuli was dispersed across most of the network. Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (but 300 C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, and most of these were very close to the stimulation site of B.

We quantified the correlation between the differences in non-target response magnitude and strength of adaptation before target trials across excitatory neurons, and found an almost perfect relationship (Pearson's r = 0.99, n = 800, see Figure \ref{fig5}C). This effect was robust across networks: Greater reduction in the non-target response was closely correlated with a greater reduction in $V_{TA}$ at the start of A trials. In addition, we found a robust relationship between the reduction in $V_{TA}$ in a given neuron and its increased response in target trials in dev (r = -0.42), corroborating the notion that lower thresholds yield higher responses.

Finally, we examined the cause of the reduced response in non-target trials. By moving from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation, and secondly, the more frequent presentation of B changes adaptation, reducing the average response. We contrasted the response to B in the msc sequence with the response to other non-target stimuli in the same sequence, and with the response to B in dev, to estimate the effect of stimulus replacement and adaptation, respectively. We found that, in this particular instance, adaptation weakly, but consistently, decreased the response, whereas stimulus replacement reduced many responses more dramatically while permitting larger responses in many neurons in the general vicinity of the stimulus sites of A and B. Across networks, however, the effect of adaptation was more pronounced and more consistent, removing an average of 71 spikes per trial, whereas stimulus replacement had no effect on average.

To conclude, TA plays two complementary roles in deviance detection, affecting the standard and deviant response separately: On the one hand, it reduces the response to the frequently presented standard stimulus, leading to a larger deviant response merely by contrast with the corresponding standard. On the other hand, due to the overlap in receptive fields, TA acts as blanket suppression in the many-standards control condition, against which the deviant response stands out, relieved from this suppression and encountering a less fatigued network.

\subsection{The role of STD in deviance detection}\label{sec-std}

As noted above, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA. To understand the basis of this apparent synergy, we contrasted the TA-only condition examined in the previous section against the same network with both TA and STD. As shown in Figure \ref{fig6}, we found that the non-target response magnitude was reduced by the introduction of STD, as we would expect. Indeed, across networks, the response to all stimuli was reduced, regardless of context. Concomitantly, and in keeping with the above analysis, the levels of $V_{TA}$ before A trials also dropped.

However, we noticed a systematic difference in the reduction of both response magnitude and $V_{TA}$ between the dev and msc conditions: The non-target response in the dev sequence was reduced more than that in the msc sequence, leading to stronger de-adaptation and, as a consequence, a larger A trial response in dev compared to TA-only in large parts of the network. Across networks, only few cases of an increased dev response were seen; however, the response to target stimulus was reduced less in dev than msc, and less in msc than std. This differential reduction was directly responsible for the larger DDI reported in Figure \ref{fig3}, but how did it arise?

We hypothesized that, due to its relatively short time scale, the direct effect of STD across trials is limited to repeated trials of the same stimulus, or in other words, to the standard stimulus in oddball sequences. However, by reducing the standard response -- or, shifting the perspective, the non-target dev response -- in the full model, STD also affects adaptation levels, causing de-adaptation and thereby indirectly increasing the target dev response.

In direct support of the notion that STD does carry over between (repeated) trials, we noted first that the response sizes in the STD-only condition was less in std than both msc and dev, see Figure \ref{fig6}D, indicating robust SSA. To show that the across-trial effect of STD in the full model is limited to the dev non-target trials, however, we needed a more sophisticated tool, showing the level of depression experienced by each neuron. Since the effect of STD is to reduce excitatory postsynaptic potentials (EPSP), we estimated this reduction per synapse in a given trial type $S$ (e.g. A dev) as the product of the expected number of spikes transmitted $R^S$, the peak EPSP of the synapse ($k = 1.4 mV$, recall that all weights are 1), and its average depression coefficient $(1-x)$. Summing these EPSP estimates onto each postsynaptic neuron, we arrived at a rough estimate

\begin{equation}
    D^S = \sum_{i \in pre} R^S_i k (1-x_i^{t=0})
\end{equation}

of the total membrane potential lost due to STD.

In Figure \ref{fig7}, we show that, in A dev and A msc trials, depression has relatively evenly distributed effects across the network. Neither sequence seemed to have a stronger overall effect, despite a slight trend towards stronger depression near the stimulus site in the msc sequence. This trend was also seen in B msc trials. In sharp contrast, in B dev trials, depression was very strong and very clearly localized to the stimulation site. This location dependence was closely mirrored in the B dev network response, which was likewise largely confined to the area near the stimulation site (data not shown). To quantify this result while avoiding spurious events, we averaged the depression value $D$ over neurons with no less than 1 spike in 5 trials. Across networks, we found that depression in the dev sequence was systematically stronger for B than for A (t = 7.5, p = 7.07e-12). More importantly, the level of depression in A trials changed considerably less between msc and dev sequences when compared to B trials (t = -6.7, p = 3.92e-10). Taken together, this shows that, indeed, the dominant direct effect of STD was to weaken the response to the frequently presented standard stimulus.

Finally, to complete the picture and confirm our hypothesis in full, we compared the small, but significant relief from depression in A dev trials relative to msc (t = -3.2, p = 0.001) to the difference in $V_{TA}$, applying the same neuron exclusion criteria. We found that the relief from suppression that was shown above to be a key ingredient to deviance detection was driven almost exclusively by lower $V_{TA}$ (t = 11, p = 4.06e-19). More importantly, adding STD to the model induced a much larger additional difference in $V_{TA}$, compared to the TA-only condition, than in directly STD-related $D$ (t = 2.1, p = 0.0195), indicating that, indeed, STD affected the deviant response primarily through indirect effects.

\section{Discussion}\label{sec-discussion}

\backmatter

\bmhead{Supplementary information}
TODO

\bmhead{Author contributions}
All authors contributed to the study conception and design. Material preparation, data collection and analysis were performed by Felix Benjamin Kern. The first draft of the manuscript was written by Felix Benjamin Kern and all authors commented on previous versions of the manuscript. All authors read and approved the final manuscript.

\bmhead{Code availability}
TODO

\section*{Declarations}

\bmhead{Funding}
This study was funded by a World Premier International Research Center Initiative research startup grant.

\bmhead{Competing interests}
The authors have no relevant financial or non-financial interests to disclose.

\bmhead{Ethics approval}
Not applicable.

\bmhead{Consent to participate}
Not applicable.

\bmhead{Consent for publication}
Not applicable.

\begin{appendices}

\section{Section title of first appendix}\label{secA1}

An appendix contains supplementary information that is not an essential part of the text itself but which may be helpful in providing a more comprehensive understanding of the research problem or it is information that is too cumbersome to be included in the body of the paper.

%%=============================================%%
%% For submissions to Nature Portfolio Journals %%
%% please use the heading ``Extended Data''.   %%
%%=============================================%%

%%=============================================================%%
%% Sample for another appendix section			       %%
%%=============================================================%%

%% \section{Example of another appendix section}\label{secA2}%
%% Appendices may be used for helpful, supporting or essential material that would otherwise 
%% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
%% tables and equations etc.

\end{appendices}

%%===========================================================================================%%
%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
%% system, please include the references within the manuscript file itself. You may do this  %%
%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
%% file, and delete the associated \verb+\bibliography+ commands.                            %%
%%===========================================================================================%%

\bibliography{sn-bibliography}% common bib file
%% if required, the content of .bbl file can be included here once bbl is generated
%%\input sn-article.bbl

%% Default %%
%%\input sn-sample-bib.tex%

\end{document}

### 2022-12-01 dish.py
Lotsa figure updates

@@ -1,0 +1,87 @@
+import numpy as np
+from brian2.only import ms
+dish = {
+    'dt': 10*ms,
+    'dev': np.asarray([
+        3.587385019710907,
+        5.153745072273325,
+        8.885676741130093,
+        9.579500657030225,
+        9.442838370565047,
+        9.127463863337715,
+        8.643889618922472,
+        7.971090670170829,
+        6.909329829172142,
+        5.816031537450724,
+        4.743758212877793,
+        3.608409986859395,
+        2.5676741130091987,
+        1.6215505913271997,
+        0.9382391590013146,
+        0.5072273324572922,
+        0.27595269382391585,
+        0.16031537450722766,
+        0.07621550591327164,
+        0.03416557161629363,
+        0.023653088042049575,
+        0.0026281208935596823,
+        0,0,0,0,0,
+    ]),
+
+    'msc': np.asarray([
+        3.0722733245729312,
+        1.8948751642575559,
+        1.8738501971090678,
+        2.052562417871222,
+        2.4730617608409986,
+        2.83048620236531,
+        2.977660972404731,
+        2.9671484888304853,
+        2.756898817345598,
+        2.4730617608409986,
+        2.157687253613666,
+        1.7792378449408677,
+        1.3902759526938233,
+        0.9592641261498027,
+        0.6859395532194483,
+        0.423127463863338,
+        0.27595269382391585,
+        0.18134034165571578,
+        0.13929040735873777,
+        0.0867279894875157,
+        0.055190538764783525,
+        0.04467805519053769,
+        0.013140604467805517,
+        0,0,0
+    ]),
+
+    'std': np.asarray([
+        2.935611038107753,
+        1.4848883048620234,
+        1.2956636005256232,
+        1.264126149802891,
+        1.3377135348226012,
+        1.3902759526938233,
+        1.4323258869908013,
+        1.4638633377135353,
+        1.4743758212877793,
+        1.4428383705650472,
+        1.3797634691195793,
+        1.2431011826544012,
+        1.0854139290407367,
+        0.9382391590013146,
+        0.7805519053876466,
+        0.6438896189224703,
+        0.5282522996057821,
+        0.41261498028909216,
+        0.32851511169513614,
+        0.2654402102496718,
+        0.1918528252299616,
+        0.1498028909329836,
+        0.10775295663600382,
+        0.07621550591327164,
+        0.04993429697766061,
+        0.03416557161629363,
+        0.023653088042049575,
+    ])
+}

### 2022-12-01 plot_utils.py

@@ -1,11 +1,13 @@
 import matplotlib.pyplot as plt
 import numpy as np
 from matplotlib.colors import LinearSegmentedColormap
+from matplotlib.colors import colorConverter
 from brian2.units import msecond
+import styling
 
 
 colors = {
-    'std': 'k',
+    'std': styling.offblack,
     'dev': 'r',
     'msc': 'b',
     'Depression': '#9b19f5',
@@ -93,3 +95,16 @@ def plot_pulse_hist(histograms, index_N, index_t, dt, figsize=(10,15), grid=Fals
         return fig, axs, cb, orders
     else:
         return fig, axs, cb
+
+def alpha_to_color(c, alpha, bg='white'):
+    c = np.asarray(colorConverter.to_rgb(c))
+    bg = np.asarray(colorConverter.to_rgb(bg))
+    rgb = (1-alpha)*bg + alpha*c
+    return rgb
+
+def fill_ratios(*ratios, to=100):
+    ratios = np.asarray(ratios)
+    total = ratios[ratios>0].sum()
+    remainder = to-total
+    ratios[ratios<0] = remainder / (ratios<0).sum()
+    return ratios

### 2022-12-01 styling.py

@@ -19,6 +19,7 @@ sns.set_theme(style='ticks', font_scale=0.75, rc={
     'axes.titlepad': 4,
     'grid.linewidth': 0.4,
     'lines.linewidth': 0.5,
+    'patch.linewidth': 0.5,
     'legend.fontsize': 8,
     'legend.title_fontsize': 9,
     'xtick.labelsize': 6,
@@ -56,7 +57,8 @@ sns.set_theme(style='ticks', font_scale=0.75, rc={
     'axes.edgecolor': offblack,
     'axes.labelcolor': offblack,
     'xtick.color': offblack,
-    'ytick.color': offblack
+    'ytick.color': offblack,
 })
 
 label_kwargs = {'fontsize': 12, 'fontweight': 'bold'}
+annotation_kwargs = {'fontsize': 8}

### 2022-12-01 paper-1/manuscript.tex
+ Main text draft

@@
 
-\documentclass[pdflatex,iicol,sn-basic]{sn-jnl}
+\documentclass[pdflatex,referee,iicol,sn-basic]{sn-jnl}
 
@@
 
-\title{Short-term synaptic plasticity and threshold adaptation act in synergy to give rise to deviance detection in a spiking model of neuronal culture}
+\title[Synaptic and somatic short-term plasticity collaborate to yield deviance detection]{Short-term synaptic plasticity and threshold adaptation act in synergy to give rise to deviance detection in a spiking model of neuronal culture}
 
@@
 
-Figure \ref{fig1}C shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
-The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \cite{Stimberg2020-go}, and simulated with an integration time step of 1~ms.
+\begin{figure*}%
+    \centering
+    \includegraphics{fig1.eps}
+    \caption{\textbf{a} Sample network layout, showing excitatory and inhibitory neurons and all outgoing synaptic connections from a small subset.
+\textbf{b} Distance from stimulation site 0 in terms of minimum number of excitatory synapses.
+Stimulation is provided to the 10 neurons nearest to each of 5 stimulation sites, which are regularly spaced on a circle with a radius of 2.5 mm.
+\textbf{c} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes.
+The excitatory connection has weight $w = 3$ for demonstration purposes only; a single connection with $w = 1$ is normally unable to evoke postsynaptic firing.
+Top: Membrane voltage (black, $\tau$ = 30 ms), firing threshold baseline (gray) and adaptive threshold (green, $\tau$ = 1 s) demonstrating TA.
+Middle: Excitatory ($g_e$, $\tau$ = 2 ms) and inhibitory ($g_i$, $\tau$ = 4 ms) synaptic conductances, demonstrating excitatory STD.
+Bottom: Inputs to $g_e$ and $g_i$ (vertical bars), and the depression variable $x$ of the excitatory presynaptic neuron (dashed line)}
+    \label{fig1}
+\end{figure*}
+
+Figure \ref{fig1}c shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
+The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \citep{Stimberg2020-go}, and simulated with an integration time step of 1~ms.
 Synaptic delay was not modeled explicitly, but spikes were delivered, and voltages reset, in the time step following spike emission.
@@
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-See Figure \ref{fig1}AB for a representative example of the resulting network structure.
+See Figure \ref{fig1}a,b for a representative example of the resulting network structure.
 Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response, discarding networks where less than half of all neurons in the fully recovered network responded to stimulation at any of the five sites.
@@
 
-To investigate deviance detection, we used a classical oddball paradigm \citep{TODO} as illustrated in Figure \ref{fig2}A, presenting a total of 500 stimuli at regular intervals of 500~ms.
+To investigate deviance detection, we used a classical oddball paradigm as illustrated in Figure \ref{fig2}a, presenting a total of 500 stimuli at regular intervals of 500~ms.
 Target stimuli (labeled A hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (std), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (dev) with 100 A and 400 B, and the many-standards control sequence (msc) with 100 presentations of each of the stimuli A through E.
@@
 
-In each network, we simulated oddball sequences with two A/B stimulus pairs (sites 1 and 2 in one pair, and sites 3 and 5 in the other pair), where each stimulus site in a pair was used as both target (A) and non-target (B).
-In addition, we simulated one msc sequence, yielding 4 complete data sets per network, or a total of 120 data sets across all networks.
+In each network, we simulated oddball sequences with two A/B stimulus pairs (sites 1 and 2 in one pair, and sites 3 and 5 in the other pair), where each stimulus site in a pair was used once as target (A) and once as non-target (B), and a single msc sequence involving all sites.
 Networks were reset to a fully recovered state between sequence presentations to guarantee their independence.
+This yielded 4 complete data sets per network, or a total of 120 data sets across all networks.
 
@@
 
+\begin{figure*}%
+    \centering
+    \includegraphics{fig2.pdf}
+    \caption{\textbf{a} Illustration of the oddball paradigm used.
+Each square represents a stimulus, with the target (A) in dark blue, and other stimuli in lighter colors.
+The temporal extent of stimulation is exaggerated; in reality, stimulation was delivered in a single time step, and the response allowed to develop freely for 500 ms.
+\textbf{b} Response of a neuronal cell culture in vitro, adapted from \cite{Kubota2021-dx}.
+\textbf{c} Model response, showing average network-wide spike counts after stimulation with target (A) in each sequence.
+\textbf{d} Mismatch, SSA and deviance detection indices across networks and stimulus pairs (n = 120).
+In this and later figures, notches indicate the confidence interval for the median, calculated by bootstrap with 10000 iterations.
+All indices were positive (mismatch: t = 11, p = 6.7e-21, SSA: t = 7.8, p = 1.05e-12, deviance detection: t = 6.5, p = 1.05e-09)}
+    \label{fig2}
+\end{figure*}
+
 We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs.
-Following related research \citep{TODO}, we were careful to exclude effects due to stimulus identity (e.g.
-comparing responses to A and B in the dev sequence) and to adaptation to frequent stimulus presentation (e.g.
-comparing A in the dev sequence to A in the std sequence).
+Following related research \citep{TODO}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the dev sequence to A in the std sequence).
 In other words, we used the msc sequence as a neutral baseline where the target stimulus was presented no more often than in the dev sequence, but no expectation of a competing stimulus B could be formed.
@@
 \end{equation}
-We define adaptation (msc - std) and mismatch (dev - std) indices in similar fashion, capturing respectively the effect of adaptation, and of both adaptation and deviance detection together.
+We define stimulus-specific adaptation (SSA, msc - std) and mismatch (dev - std) indices analogously, capturing respectively the effect of adaptation, and of both adaptation and deviance detection together.
 
-Our first aim was to replicate deviance detection properties reported in dissociated neuronal cell culture by \cite{Kubota2021-dx}, reproduced in Figure \ref{fig2}B.
-With minimal tuning of global parameters, we found that most networks (and 92 of 120 target stimuli) reliably produced not only mismatch responses based on adaptation alone, but also showed robust deviance detection, as summarized in Figure \ref{fig2}D.
+Our first aim was to replicate deviance detection properties reported in dissociated neuronal cell culture by \cite{Kubota2021-dx}, reproduced in Figure \ref{fig2}b.
+With minimal tuning of global parameters, we found that most networks (92 out of 120 target stimuli) reliably produced not only mismatch responses based on adaptation alone, but also showed robust deviance detection, as summarized in Figure \ref{fig2}d.
 
@@
 
-Having confirmed the presence of deviance detection in our model, we then turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model -- synaptic short-term depression (STD) and threshold adaptation (TA) -- enable the networks to learn relevant regularities.
+Having confirmed the presence of deviance detection in our model, we then turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model -- synaptic short-term depression (STD) and threshold adaptation (TA) -- enabled the networks to learn relevant regularities.
 As shown in Figure \ref{fig3}, networks with neither plasticity mechanism did not show any difference in response, indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
-Networks with only STD likewise did not show any deviance detection, though they did exhibit robust adaptation effects (data not shown), leading to a smaller target response in std than in either dev or msc sequences.
+Networks with only STD likewise did not show any deviance detection, though they did exhibit robust stimulus-specific adaptation effects, leading to a smaller target response in std than in either dev or msc sequences.
 Conversely, with only TA, networks produced both adaptation and deviance detection effects.
 
-Finally, with both STD and TA applied, deviance detection was also reliable, as shown before; however, we were surprised to find that the effect was significantly stronger than that of either plasticity mechanism alone (STD: t = 6.7, p = 4.53e-10; TA: t = 3.2, p = 0.00103), and greater too than the sum of both STD only and TA only DDI values (t = 3.6, p = 0.000199).
-This suggests that TA and STD act not as independent mechanisms in this paradigm, but rather interact constructively to enhance the network's ability to encode temporal regularities.
+\begin{figure*}%
+    \centering
+    \includegraphics{fig3.pdf}
+    \caption{Deviance detection and SSA indices under model ablation.
+\textbf{a} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
+\textbf{b, c} Deviance detection and SSA indices across networks and stimulus pairs.
+Significance was assessed by 1-sample t-tests.
+TODO p-values: text or table?}
+    \label{fig3}
+\end{figure*}
 
-In the following, we will use a sample network and target stimulus to first build an understanding of how TA leads to deviance detection when acting alone.
+Finally, with both STD and TA applied, deviance detection was also reliable, as shown before; however, we were surprised to find that the effect was significantly stronger than that of either plasticity mechanism alone (STD: t = 6.7, p = 4.53e-10; TA: t = 3.2, p = 0.00103), and greater too than the sum of the DDI values across both STD only and TA only models (t = 3.6, p = 0.000199).
+This suggests that TA and STD act not as independent mechanisms in this paradigm, but rather interact constructively to enhance the network's ability to encode input regularities.
+
+In the following, we will use the sample network and target stimulus shown in Figure \ref{fig3}a to first build an understanding of how TA leads to deviance detection when acting alone.
 Then, we will build on this understanding to elucidate how the addition of STD, which is ineffective on its own, can lead to an increased deviant response.
@@
 To understand why the same stimulus A evoked a higher response in dev than in msc, we first compared the average responses in the two conditions.
-As shown in Figure \ref{fig4}A, we found that the response developed in a qualitatively similar fashion, with a tendency of dev spikes to occur earlier, particularly shortly after stimulation, and more frequently, particularly towards the later part of the response.
-Since the stimulus was identical in both conditions, this difference had to be due to either changed firing thresholds, or changed synaptic currents as a consequence of spike differences earlier in the response.
-We quantified the former with the strength of adaptation $V_{TA}$, which showed a clear and widespread lowering of thresholds at the start of dev trials relative to msc, followed by an increase beyond the msc baseline in many neurons as they spiked (Figure \ref{fig4}B).
+As shown in Figure \ref{fig4}a, we found that the response developed in a qualitatively similar fashion, with a tendency of dev spikes to occur earlier, particularly shortly after stimulation, and more frequently, particularly towards the later part of the response.
+Since the stimulus was identical in both conditions, this difference had to be due to either changed firing thresholds, or changed synaptic currents as a consequence of activity differences in earlier parts of the response.
+We quantified the former with the strength of adaptation $V_{TA}$ (Figure \ref{fig4}b), which showed a clear and widespread lowering of thresholds at the start of dev trials relative to msc, followed by an increase beyond the msc baseline in many neurons as they spiked.
+Correlating the difference in $V_{TA}$ at the beginning of trials with the difference in trial spike count, we found that, across neurons, lower thresholds are strongly predictive of higher activity (r = -0.424, p = 3.1e-36, see Figure \ref{fig4}d).
 
-To quantify the synaptic contributions to a neuron's membrane potential $v$, we calculated a quantity $V_{syn}$, which captures all synaptic inputs, but excludes voltage deflections caused by direct stimulation and by the self-inhibition induced by resetting to $v_{reset} < v_{rest}$ after a spike (see Figure S\ref{SI-fig1}).
-Being induced by spiking activity, this measure closely followed the activity pattern, making the earlier activation of the dev response clearly visible across the network (Figure \ref{fig4}C).
+To quantify the synaptic contributions to a neuron's membrane potential $v$, we calculated a quantity $V_{syn}$, shown in Figure \ref{fig4}c, which captures all synaptic inputs, but excludes voltage deflections caused by direct stimulation and by the self-inhibition induced by resetting to $v_{reset} < v_{rest}$ after a spike.
+Being induced by spiking activity, this measure closely followed the activity pattern, making the earlier activation of the dev response clearly visible across the network.
 In addition, since $V_{syn}$ decays with the membrane timescale $\tau_m$, but resets to $0$ on postsynaptic spikes, many of the neurons that spiked more frequently showed lower $V_{syn}$ towards the end of network activity.
 
+\begin{figure*}%
+    \centering
+    \includegraphics{fig4.pdf}
+    \caption{Responses and threshold adaptation in a TA-only ablation.
+\textbf{a, b, c} Post-stimulus histograms across target trials, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
+Left and middle column, target trials in dev and msc sequence, respectively; right column, contrast between these two.
+\textbf{d} Per-neuron relationship between the contrast (dev - msc) in $V_{TA}$ at the start of trials, and the contrast in average number of spikes fired, excluding inhibitory neurons, and associated linear regression (r = -0.424, p = 3.1e-36).
+\textbf{e} Contribution of $V_{TA}$ and $V_{syn}$ to bins with increased firing in dev (i.e., red bins in the spike probability contrast, \textbf{a}), weighted by that same contrast.
+\textbf{f} Contribution of $V_{TA}$ relative to the sum of the weighted contributions in \textbf{e}.
+\textbf{g} Relative contribution of $V_{TA}$ across networks, showing median (solid line) and inter-quartile range (shaded area).}
+    \label{fig4}
+\end{figure*}
+
 Based on these measures, we then asked why there were more spikes in dev trials.
-To answer this, we weighted the contrasts in $V_{TA}$ and $V_{syn}$ by the spike probability contrast, discarding any instances of equal or lower firing in dev, and summed over neurons, directly linking network-wide increased activity to its immediate underlying causes, de-adaptation and depolarization.
-We found that the increased depolarization induced by synaptic input greatly outweighed the impact of reduced adaptation throughout most of the response (Figure \ref{fig4}D).
-Only in the very early stages of the response did adaptation dominate the increase in firing (Figure \ref{fig4}E), whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
-This trend was reflected in most networks investigated, as shown in the summary plot in Figure \ref{fig4}F.
+To answer this, we weighted the contrasts in $V_{TA}$ and $V_{syn}$ by the spike probability contrast, discarding any instances of equal or lower firing in dev, and summed over neurons.
+The resulting time courses (Figure \ref{fig4}e) reveal the immediate underlying causes of network-wide activity increases, namely de-adaptation and depolarization.
+We found that increased depolarization induced by synaptic input greatly outweighed the impact of reduced adaptation throughout most of the response.
+Only in the very early stages of the response did adaptation dominate the increase in firing (Figure \ref{fig4}f), whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
+This trend was reflected in most networks investigated, as shown in the summary plot in Figure \ref{fig4}g.
 To summarize, while most added spikes arose from increased prior activity within a trial, the initial impetus -- the increase in the early response -- was largely caused by lower TA.
@@
 Having established that the increased response to A in dev trials was driven by lower thresholds, we next turned to the cause of this reduction.
-Firstly, plotting $V_{TA}$ at the start of A trials in Figure \ref{fig5}A, we found that most of the network was indeed less adapted in dev, with only a few neurons in the vicinity of B showing higher thresholds.
+Firstly, plotting $V_{TA}$ at the start of A trials in Figure \ref{fig5}a, we found that most of the network was indeed less adapted in dev, with only a few neurons in the vicinity of B showing higher thresholds.
 We hypothesized that this could be due to the different non-target stimulation pattern, since B was stimulated more frequently in dev than msc.
-Indeed, as shown in Figure \ref{fig5}B, the non-target response activity in the dev sequence -- in other words, the response to B -- was concentrated around the site where B was delivered, while in the msc sequence, activity in response to the various non-target stimuli was dispersed across most of the network.
-Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (but 300 C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, and most of these were very close to the stimulation site of B.
+Indeed, as shown in Figure \ref{fig5}b, the non-target response activity in the dev sequence -- in other words, the response to B -- was concentrated around the site where B was delivered, while in the msc sequence, activity in response to the various non-target stimuli was dispersed across most of the network.
+Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (alongside 300 C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, most of these very close to the stimulation site of B.
 
-We quantified the correlation between the differences in non-target response magnitude and strength of adaptation before target trials across excitatory neurons, and found an almost perfect relationship (Pearson's r = 0.99, n = 800, see Figure \ref{fig5}C).
+\begin{figure*}%
+    \centering
+    \includegraphics{fig5.pdf}
+    \caption{}
+    \label{fig5}
+\end{figure*}
+
+We quantified the correlation between the differences in non-target response magnitude and strength of adaptation before target trials across excitatory neurons, and found an almost perfect relationship (Pearson's r = 0.99, n = 800, see Figure \ref{fig5}d).
 This effect was robust across networks: Greater reduction in the non-target response was closely correlated with a greater reduction in $V_{TA}$ at the start of A trials.
-In addition, we found a robust relationship between the reduction in $V_{TA}$ in a given neuron and its increased response in target trials in dev (r = -0.42), corroborating the notion that lower thresholds yield higher responses.
 
 Finally, we examined the cause of the reduced response in non-target trials.
-By moving from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation, and secondly, the more frequent presentation of B changes adaptation, reducing the average response.
-We contrasted the response to B in the msc sequence with the response to other non-target stimuli in the same sequence, and with the response to B in dev, to estimate the effect of stimulus replacement and adaptation, respectively.
-We found that, in this particular instance, adaptation weakly, but consistently, decreased the response, whereas stimulus replacement reduced many responses more dramatically while permitting larger responses in many neurons in the general vicinity of the stimulus sites of A and B.
-Across networks, however, the effect of adaptation was more pronounced and more consistent, removing an average of 71 spikes per trial, whereas stimulus replacement had no effect on average.
+Going from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation, and secondly, the more frequent presentation of B changes adaptation, reducing the average response.
+To estimate the effect of stimulus replacement and adaptation, respectively, we contrasted the response to B in the msc sequence with the response to other non-target stimuli in the same sequence (replacement), and with the response to B in dev (adaptation).
+We found that, in this particular network and stimulus pair, adaptation weakly but consistently decreased the response, whereas stimulus replacement reduced many responses more dramatically while permitting larger responses in many neurons in the general vicinity of the stimulus sites of A and B.
+Across networks, however, the effect of adaptation was more pronounced and more consistent, removing an average of 71 spikes per trial, whereas stimulus replacement had no effect on average (t = -0.69, p = 0.246).
 
-To conclude, TA plays two complementary roles in deviance detection, affecting the standard and deviant response separately: On the one hand, it reduces the response to the frequently presented standard stimulus, leading to a larger deviant response merely by contrast with the corresponding standard.
-On the other hand, due to the overlap in receptive fields, TA acts as blanket suppression in the many-standards control condition, against which the deviant response stands out, relieved from this suppression and encountering a less fatigued network.
+To conclude, TA plays two complementary roles in deviance detection, affecting the standard and deviant response separately: On the one hand, it reduces the response to the frequently presented standard stimulus, leading to a larger deviant response merely by contrast with the corresponding standard (SSA).
+On the other hand, due to the overlap in receptive fields, TA acts as blanket suppression in the many-standards control condition, against which the deviant response, relieved from this suppression and encountering a less fatigued network, stands out.
 
@@
 
-However, we noticed a systematic difference in the reduction of both response magnitude and $V_{TA}$ between the dev and msc conditions: The non-target response in the dev sequence was reduced more than that in the msc sequence, leading to stronger de-adaptation and, as a consequence, a larger A trial response in dev compared to TA-only in large parts of the network.
-Across networks, only few cases of an increased dev response were seen; however, the response to target stimulus was reduced less in dev than msc, and less in msc than std.
+\begin{figure*}%
+    \centering
+    \includegraphics{fig6.pdf}
+    \caption{}
+    \label{fig6}
+\end{figure*}
+
+However, we noticed a systematic difference in the reduction of both response magnitude and $V_{TA}$ between the dev and msc conditions: The non-target response in the dev sequence was reduced more than that in the msc sequence, leading to a larger difference between the responses in these two contexts.
+$V_{TA}$ before A trials followed suit, leading to stronger de-adaptation and, as a consequence, a larger A trial response in dev in large parts of the network in the full model compared to TA-only.
+Across networks, only few cases of an increased dev response were seen; however, the response to the target stimulus (Figure \ref{fig6}e) was reduced less in dev than msc, and less in msc than std.
 This differential reduction was directly responsible for the larger DDI reported in Figure \ref{fig3}, but how did it arise?
 
-We hypothesized that, due to its relatively short time scale, the direct effect of STD across trials is limited to repeated trials of the same stimulus, or in other words, to the standard stimulus in oddball sequences.
-However, by reducing the standard response -- or, shifting the perspective, the non-target dev response -- in the full model, STD also affects adaptation levels, causing de-adaptation and thereby indirectly increasing the target dev response.
+Due to its relatively short time scale, the direct effect of STD across trials was likely limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.
+However, by reducing the standard (i.e., the dev non-target) response in the full model, we hypothesized that STD also affects adaptation levels, causing de-adaptation and thereby indirectly increasing the target dev response.
 
-In direct support of the notion that STD does carry over between (repeated) trials, we noted first that the response sizes in the STD-only condition was less in std than both msc and dev, see Figure \ref{fig6}D, indicating robust SSA.
-To show that the across-trial effect of STD in the full model is limited to the dev non-target trials, however, we needed a more sophisticated tool, showing the level of depression experienced by each neuron.
+In direct support of the notion that STD indeed carried over between (repeated) trials, we noted first that the response size in the STD-only condition was slightly less in std than in msc or dev (cf.
+Figure \ref{fig3}c), indicating some amount of SSA.
+To show that the across-trial effect of STD in the full model was limited to the dev non-target trials, however, we needed a more sophisticated tool, showing the level of depression experienced by each neuron.
 Since the effect of STD is to reduce excitatory postsynaptic potentials (EPSP), we estimated this reduction per synapse in a given trial type $S$ (e.g.
-A dev) as the product of the expected number of spikes transmitted $R^S$, the peak EPSP of the synapse ($k = 1.4 mV$, recall that all weights are 1), and its average depression coefficient $(1-x)$.
-Summing these EPSP estimates onto each postsynaptic neuron, we arrived at a rough estimate
+A dev) as the product of the expected number of transmitted spikes $R^S$, the peak EPSP of the synapse ($k = 1.4 mV$, recall that all weights are 1), and its average depression coefficient $(1-x)$ at the start of the trial.
+Summing these estimates onto each postsynaptic neuron, we arrived at a rough estimate
 
@@
 
-of the total membrane potential lost due to STD.
+of the average depolarization lost due to STD.
 
-In Figure \ref{fig7}, we show that, in A dev and A msc trials, depression has relatively evenly distributed effects across the network.
-Neither sequence seemed to have a stronger overall effect, despite a slight trend towards stronger depression near the stimulus site in the msc sequence.
-This trend was also seen in B msc trials.
+\begin{figure*}%
+    \centering
+    \includegraphics{fig7.pdf}
+    \caption{}
+    \label{fig7}
+\end{figure*}
+
+In Figure \ref{fig7}, we show that, in A dev and A msc trials, depression had relatively evenly distributed effects across the network.
+Neither sequence seemed to have a stronger overall effect, despite a slight trend towards weaker depression near the stimulus site in the dev sequence.
 In sharp contrast, in B dev trials, depression was very strong and very clearly localized to the stimulation site.
-This location dependence was closely mirrored in the B dev network response, which was likewise largely confined to the area near the stimulation site (data not shown).
-To quantify this result while avoiding spurious events, we averaged the depression value $D$ over neurons with no less than 1 spike in 5 trials.
+This location dependence was closely mirrored in the B dev network response, which was largely confined to the area near the stimulation site (data not shown).
+To quantify this result while avoiding spurious events, we averaged the depression value $D$ over neurons with at least 1 spike per 5 trials in the given context.
 Across networks, we found that depression in the dev sequence was systematically stronger for B than for A (t = 7.5, p = 7.07e-12).
@@
 Finally, to complete the picture and confirm our hypothesis in full, we compared the small, but significant relief from depression in A dev trials relative to msc (t = -3.2, p = 0.001) to the difference in $V_{TA}$, applying the same neuron exclusion criteria.
-We found that the relief from suppression that was shown above to be a key ingredient to deviance detection was driven almost exclusively by lower $V_{TA}$ (t = 11, p = 4.06e-19).
-More importantly, adding STD to the model induced a much larger additional difference in $V_{TA}$, compared to the TA-only condition, than in directly STD-related $D$ (t = 2.1, p = 0.0195), indicating that, indeed, STD affected the deviant response primarily through indirect effects.
+We found that the relief from suppression, which was shown in the TA-only analysis above to be a key ingredient to deviance detection, was driven almost exclusively by lower $V_{TA}$ (t = 11, p = 4.06e-19).
+More importantly, adding STD to the model induced a much larger additional difference in $V_{TA}$ compared to the TA-only condition, than in directly STD-related $D$ (t = 2.1, p = 0.0195), indicating that, indeed, STD affected the deviant response primarily through indirect effects.
 
@@
 
-\begin{appendices}
+% \begin{appendices}
 
-\section{Section title of first appendix}\label{secA1}
+% \section{Section title of first appendix}\label{secA1}
 
-An appendix contains supplementary information that is not an essential part of the text itself but which may be helpful in providing a more comprehensive understanding of the research problem or it is information that is too cumbersome to be included in the body of the paper.
+% %%=============================================%%
+% %% For submissions to Nature Portfolio Journals %%
+% %% please use the heading ``Extended Data''.   %%
+% %%=============================================%%
 
-%%=============================================%%
-%% For submissions to Nature Portfolio Journals %%
-%% please use the heading ``Extended Data''.   %%
-%%=============================================%%
+% %%=============================================================%%
+% %% Sample for another appendix section			       %%
+% %%=============================================================%%
 
-%%=============================================================%%
-%% Sample for another appendix section			       %%
-%%=============================================================%%
+% %% \section{Example of another appendix section}\label{secA2}%
+% %% Appendices may be used for helpful, supporting or essential material that would otherwise 
+% %% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
+% %% tables and equations etc.
 
-%% \section{Example of another appendix section}\label{secA2}%
-%% Appendices may be used for helpful, supporting or essential material that would otherwise 
-%% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
-%% tables and equations etc.
-
-\end{appendices}
+% \end{appendices}
 
@@
 
-\bibliography{sn-bibliography}% common bib file
+\bibliography{bibliography}% common bib file
 %% if required, the content of .bbl file can be included here once bbl is generated

### 2022-12-02 paper-1/manuscript.tex
Added remaining figure legends

@@
 \end{align}
-with $\boldsymbol E$ and $\boldsymbol I$ the sets of presynaptic excitatory and inhibitory neurons, respectively, $\delta(\cdot)$ the Dirac delta function, and $\hat{t}$ the time of spikes.
+with $\boldsymbol E$ and $\boldsymbol I$ the sets of presynaptic excitatory and inhibitory neurons, respectively, $w$ the synaptic weight, $\delta(\cdot)$ the Dirac delta function, and $\hat{t}$ the time of spikes.
 
@@
 \textbf{f} Contribution of $V_{TA}$ relative to the sum of the weighted contributions in \textbf{e}.
-\textbf{g} Relative contribution of $V_{TA}$ across networks, showing median (solid line) and inter-quartile range (shaded area).}
+\textbf{g} Relative contribution of $V_{TA}$ across networks, showing median (solid line) and inter-quartile range (shaded area)}
     \label{fig4}
@@
     \includegraphics{fig5.pdf}
-    \caption{}
+    \caption{\textbf{a} $V_{TA}$ at the start of A trials (mean over trials), with each neuron shown in its spatial location.
+Left and middle column, raw averages for dev and msc sequences, respectively; right column, contrast between dev and msc averages.
+A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the msc sequence.
+\textbf{b} Average response (in spikes per trial) to non-target trials (B in dev, B/C/D/E in msc).
+\textbf{c} Left and middle: Replacement and adaptation components of the non-target response contrast (\textbf{b}, right column), where replacement is the contrast (B msc - C/D/E msc), and adaptation is the contrast (B dev - B msc).
+Right: Network-wide mean replacement and adaptation values across networks and target stimuli.
+The adaptation effect is less than zero (t = -9, p = 2.02e-15), and less than the replacement effect (t = 2.3, p = 0.0119), which is not different from zero (t = -0.69, p = 0.246).
+\textbf{d} Per-neuron relationship between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation, and the contrast in $V_{TA}$ at the start of target trials, excluding inhibitory neurons, and associated linear regression (r = 0.99, p = 0).
+\textbf{e} The same relationship as \textbf{d}, but using the medians across excitatory neurons, and plotting one point per network and target stimulus.
+The values are highly correlated (r = 0.956, p = 5.98e-65)}
     \label{fig5}
@@
 We found that, in this particular network and stimulus pair, adaptation weakly but consistently decreased the response, whereas stimulus replacement reduced many responses more dramatically while permitting larger responses in many neurons in the general vicinity of the stimulus sites of A and B.
-Across networks, however, the effect of adaptation was more pronounced and more consistent, removing an average of 71 spikes per trial, whereas stimulus replacement had no effect on average (t = -0.69, p = 0.246).
+Across networks, however, the effect of adaptation was more pronounced and more consistent, removing an average of 71 spikes per trial, whereas stimulus replacement had no effect on average.
 
@@
     \includegraphics{fig6.pdf}
-    \caption{}
+    \caption{Contrast between the ablated model with only TA (shown in Figure \ref{fig5}) and the full model with both STD and TA.
+\textbf{a}, \textbf{b} Data shown in Figure \ref{fig5}ab, dev and msc columns, subtracted from the corresponding data for the full model.
+\textbf{c} Equivalent contrast (full - ablated model) in the response to A, in columns as above.
+\textbf{d} Network-wide mean $V_{TA}$ before A trials across networks.
+Left: Ablated and full model averages.
+Right: Difference between these model conditions.
+With the introduction of STD, $V_{TA}$ decreased more in dev than in msc (t = -3.49, p = 0.000342).
+\textbf{e} Mean number of spikes per target trial across networks.
+Left: Ablated and full model averages.
+Right: Difference between these model conditions.
+With the introduction of STD, the response decreased more in the std sequence than msc (t = -1.9, p = 0.0297), and more in msc than in dev (t = -2.99, p = 0.00168)}
     \label{fig6}
@@
 $V_{TA}$ before A trials followed suit, leading to stronger de-adaptation and, as a consequence, a larger A trial response in dev in large parts of the network in the full model compared to TA-only.
-Across networks, only few cases of an increased dev response were seen; however, the response to the target stimulus (Figure \ref{fig6}e) was reduced less in dev than msc, and less in msc than std.
+Across networks, only few cases of an increased dev response were seen; however, the response to the target stimulus (Figure \ref{fig6}e) was reduced least in dev, and most in std.
 This differential reduction was directly responsible for the larger DDI reported in Figure \ref{fig3}, but how did it arise?
@@
     \includegraphics{fig7.pdf}
-    \caption{}
+    \caption{\textbf{a} Estimated average STD-derived postsynaptic depression $D$ in target (A) trials (top row) and B trials (bottom row).
+Note that the difference between A and B in msc (middle column) is due only to the network configuration.
+\textbf{b} Statistics across networks of the contrast (dev - msc) in $D$ (cf.
+\textbf{a}, right column), and in $V_{TA}$ at the start of A trials (cf.
+Figure \ref{fig6}d).
+In each network, averages were taken over neurons with at least 0.2 spikes per dev trial (B trials for $\Delta D^B$, A trials for $\Delta D^A$ and $\Delta V_{TA}$).
+Whereas the third box ($\Delta V_{TA}$ total) indicates the averages in the full model, the final box ($\Delta V_{TA}$ added) indicates the difference between the full and the ablated (TA-only) model.
+This additional amount of $\Delta V_{TA}$, caused indirectly by the addition of STD, was less (i.e., more negative) than the directly STD-related $\Delta D^A$ itself (t = -2.09, p = 0.0195)}
     \label{fig7}
@@
 This location dependence was closely mirrored in the B dev network response, which was largely confined to the area near the stimulation site (data not shown).
-To quantify this result while avoiding spurious events, we averaged the depression value $D$ over neurons with at least 1 spike per 5 trials in the given context.
+To quantify this result while avoiding spurious events, we averaged the depression value $D$ over neurons with at least 1 spike per 5 trials of the corresponding type in the dev sequence.
 Across networks, we found that depression in the dev sequence was systematically stronger for B than for A (t = 7.5, p = 7.07e-12).
@@
 We found that the relief from suppression, which was shown in the TA-only analysis above to be a key ingredient to deviance detection, was driven almost exclusively by lower $V_{TA}$ (t = 11, p = 4.06e-19).
-More importantly, adding STD to the model induced a much larger additional difference in $V_{TA}$ compared to the TA-only condition, than in directly STD-related $D$ (t = 2.1, p = 0.0195), indicating that, indeed, STD affected the deviant response primarily through indirect effects.
+More importantly, adding STD to the model induced a much larger additional difference in $V_{TA}$ compared to the TA-only condition, than in directly STD-related $D$ (t = -2.09, p = 0.0195), indicating that, indeed, STD affected the deviant response primarily through indirect effects.
 

### 2022-12-05 paper-1/manuscript.tex
Adjusted tex for pandoc word export

@@
     \centering
-    \includegraphics{fig1.eps}
+    \includegraphics{fig1.pdf}
     \caption{\textbf{a} Sample network layout, showing excitatory and inhibitory neurons and all outgoing synaptic connections from a small subset.
@@
 \end{figure*}
+
+\textbf{Fig.
+1 a} Sample network layout, showing excitatory and inhibitory neurons and all outgoing synaptic connections from a small subset.
+\textbf{b} Distance from stimulation site 0 in terms of minimum number of excitatory synapses.
+Stimulation is provided to the 10 neurons nearest to each of 5 stimulation sites, which are regularly spaced on a circle with a radius of 2.5 mm.
+\textbf{c} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes.
+The excitatory connection has weight $w = 3$ for demonstration purposes only; a single connection with $w = 1$ is normally unable to evoke postsynaptic firing.
+Top: Membrane voltage (black, $\tau$ = 30 ms), firing threshold baseline (gray) and adaptive threshold (green, $\tau$ = 1 s) demonstrating TA.
+Middle: Excitatory ($g_e$, $\tau$ = 2 ms) and inhibitory ($g_i$, $\tau$ = 4 ms) synaptic conductances, demonstrating excitatory STD.
+Bottom: Inputs to $g_e$ and $g_i$ (vertical bars), and the depression variable $x$ of the excitatory presynaptic neuron (dashed line)
 
@@
 
+\textbf{Fig.
+2 a} Illustration of the oddball paradigm used.
+Each square represents a stimulus, with the target (A) in dark blue, and other stimuli in lighter colors.
+The temporal extent of stimulation is exaggerated; in reality, stimulation was delivered in a single time step, and the response allowed to develop freely for 500 ms.
+\textbf{b} Response of a neuronal cell culture in vitro, adapted from \cite{Kubota2021-dx}.
+\textbf{c} Model response, showing average network-wide spike counts after stimulation with target (A) in each sequence.
+\textbf{d} Mismatch, SSA and deviance detection indices across networks and stimulus pairs (n = 120).
+In this and later figures, notches indicate the confidence interval for the median, calculated by bootstrap with 10000 iterations.
+All indices were positive (mismatch: t = 11, p = 6.7e-21, SSA: t = 7.8, p = 1.05e-12, deviance detection: t = 6.5, p = 1.05e-09)
+
 We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs.
@@
 \end{figure*}
+
+\textbf{Fig.
+3} Deviance detection and SSA indices under model ablation.
+\textbf{a} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
+\textbf{b, c} Deviance detection and SSA indices across networks and stimulus pairs.
+Significance was assessed by 1-sample t-tests.
+TODO p-values: text or table?
 
@@
 
+\textbf{Fig.
+4} Responses and threshold adaptation in a TA-only ablation.
+\textbf{a, b, c} Post-stimulus histograms across target trials, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
+Left and middle column, target trials in dev and msc sequence, respectively; right column, contrast between these two.
+\textbf{d} Per-neuron relationship between the contrast (dev - msc) in $V_{TA}$ at the start of trials, and the contrast in average number of spikes fired, excluding inhibitory neurons, and associated linear regression (r = -0.424, p = 3.1e-36).
+\textbf{e} Contribution of $V_{TA}$ and $V_{syn}$ to bins with increased firing in dev (i.e., red bins in the spike probability contrast, \textbf{a}), weighted by that same contrast.
+\textbf{f} Contribution of $V_{TA}$ relative to the sum of the weighted contributions in \textbf{e}.
+\textbf{g} Relative contribution of $V_{TA}$ across networks, showing median (solid line) and inter-quartile range (shaded area)
+
 Based on these measures, we then asked why there were more spikes in dev trials.
@@
 \end{figure*}
+
+\textbf{Fig.
+5 a} $V_{TA}$ at the start of A trials (mean over trials), with each neuron shown in its spatial location.
+Left and middle column, raw averages for dev and msc sequences, respectively; right column, contrast between dev and msc averages.
+A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the msc sequence.
+\textbf{b} Average response (in spikes per trial) to non-target trials (B in dev, B/C/D/E in msc).
+\textbf{c} Left and middle: Replacement and adaptation components of the non-target response contrast (\textbf{b}, right column), where replacement is the contrast (B msc - C/D/E msc), and adaptation is the contrast (B dev - B msc).
+Right: Network-wide mean replacement and adaptation values across networks and target stimuli.
+The adaptation effect is less than zero (t = -9, p = 2.02e-15), and less than the replacement effect (t = 2.3, p = 0.0119), which is not different from zero (t = -0.69, p = 0.246).
+\textbf{d} Per-neuron relationship between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation, and the contrast in $V_{TA}$ at the start of target trials, excluding inhibitory neurons, and associated linear regression (r = 0.99, p = 0).
+\textbf{e} The same relationship as \textbf{d}, but using the medians across excitatory neurons, and plotting one point per network and target stimulus.
+The values are highly correlated (r = 0.956, p = 5.98e-65)
 
@@
 
+\textbf{Fig.
+6} Contrast between the ablated model with only TA (shown in Figure \ref{fig5}) and the full model with both STD and TA.
+\textbf{a}, \textbf{b} Data shown in Figure \ref{fig5}ab, dev and msc columns, subtracted from the corresponding data for the full model.
+\textbf{c} Equivalent contrast (full - ablated model) in the response to A, in columns as above.
+\textbf{d} Network-wide mean $V_{TA}$ before A trials across networks.
+Left: Ablated and full model averages.
+Right: Difference between these model conditions.
+With the introduction of STD, $V_{TA}$ decreased more in dev than in msc (t = -3.49, p = 0.000342).
+\textbf{e} Mean number of spikes per target trial across networks.
+Left: Ablated and full model averages.
+Right: Difference between these model conditions.
+With the introduction of STD, the response decreased more in the std sequence than msc (t = -1.9, p = 0.0297), and more in msc than in dev (t = -2.99, p = 0.00168)
+
 However, we noticed a systematic difference in the reduction of both response magnitude and $V_{TA}$ between the dev and msc conditions: The non-target response in the dev sequence was reduced more than that in the msc sequence, leading to a larger difference between the responses in these two contexts.
@@
 \end{figure*}
+
+\textbf{Fig.
+7 a} Estimated average STD-derived postsynaptic depression $D$ in target (A) trials (top row) and B trials (bottom row).
+Note that the difference between A and B in msc (middle column) is due only to the network configuration.
+\textbf{b} Statistics across networks of the contrast (dev - msc) in $D$ (cf.
+\textbf{a}, right column), and in $V_{TA}$ at the start of A trials (cf.
+Figure \ref{fig6}d).
+In each network, averages were taken over neurons with at least 0.2 spikes per dev trial (B trials for $\Delta D^B$, A trials for $\Delta D^A$ and $\Delta V_{TA}$).
+Whereas the third box ($\Delta V_{TA}$ total) indicates the averages in the full model, the final box ($\Delta V_{TA}$ added) indicates the difference between the full and the ablated (TA-only) model.
+This additional amount of $\Delta V_{TA}$, caused indirectly by the addition of STD, was less (i.e., more negative) than the directly STD-related $\Delta D^A$ itself (t = -2.09, p = 0.0195)
 
@@
 
-\bibliography{bibliography}% common bib file
+\bibliography{refs}% common bib file
 %% if required, the content of .bbl file can be included here once bbl is generated

### 2022-12-07 paper-1/manuscript.tex
Changed corresponding author

@@
 
-\author*[1]{\fnm{Felix Benjamin} \sur{Kern} \orcidA{}}\email{kernfel@gmail.com}
+\author[1]{\fnm{Felix Benjamin} \sur{Kern} \orcidA{}}\email{kernfel@gmail.com}
 
-\author[1]{\fnm{Zenas C.} \sur{Chao} \orcidB{}}\email{zenas.c.chao@gmail.com}
+\author*[1]{\fnm{Zenas C.} \sur{Chao} \orcidB{}}\email{zenas.c.chao@gmail.com}
 

### 2022-12-07 paper-1/manuscript.tex
Rev sec Methods.Model

@@
 
-Our research is based on loose approximation of a matured culture of dissociated neurons. We modeled neurons as leaky integrate-and-fire units whose membrane potential $v$ followed
+We modeled neurons as leaky integrate-and-fire units whose membrane potential $v$ at time $t$ followed
 \begin{equation}
-    \tau_m \frac{dv}{dt} = (v_{rest}-v) + I_{syn}
+    \tau_m \frac{dv}{dt} = (v_{rest}-v) + I_{syn}(t)
 \end{equation}
-where
+with $tau_m = 30$~ms the membrane time constant and $v_{rest} = -60$~mV the resting membrane potential. Synaptic currents were modeled in a conductance-based manner, following
+\begin{equation}
+    I_{syn} &= g_e(E_e-v) + g_i(E_i-v)
+\end{equation}
+with $E_e = 0$~mV and $E_i = -100$~mV the excitatory and inhibitory reversal potentials, respectively. Synaptic conductances evolved according to
 \begin{align} 
-    I_{syn} &= g_e(E_e-v) + g_i(E_i-v) \nonumber \\
     \tau_e \frac{dg_e}{dt} &= -g_e + \sum_{j \in \boldsymbol E} U x_j w \delta(t - \hat{t_j}) \nonumber \\
-    \tau_i \frac{dg_i}{dt} &= -g_i + \sum_{j \in \boldsymbol I} w \delta(t - \hat{t_j}) \label{eqn-syn}
+    \tau_i \frac{dg_i}{dt} &= -g_i + \sum_{j \in \boldsymbol I} w \delta(t - \hat{t_j}) \label{eqn-gsyn}
 \end{align}
-with $\boldsymbol E$ and $\boldsymbol I$ the sets of presynaptic excitatory and inhibitory neurons, respectively, $w$ the synaptic weight, $\delta(\cdot)$ the Dirac delta function, and $\hat{t}$ the time of spikes.
+with $\tau_e = 2$~ms the excitatory time constant, echoing AMPA receptor dynamics \citep{Hausser1997-cn}, $\tau_i = 4$~ms the inhibitory time constant, echoing GABA-A receptor dynamics \citep{Destexhe1994-oc}, $\boldsymbol E$ and $\boldsymbol I$ the sets of presynaptic excitatory and inhibitory neurons, respectively, $w$ the synaptic weight, $\delta(\cdot)$ the Dirac delta function, and $\hat{t}$ the spike times.
 
@@
 \begin{equation}
-    \tau_x \frac{dx_j}{dt} = (1-x_j) - U x_j \delta(t - \hat{t})
+    \tau_x \frac{dx_j}{dt} = (1-x_j) - U x_j \delta(t - \hat{t}) \label{eqn-xsyn}
 \end{equation}
-In ablated experiments without STD, $x$ remained constant at $1$, equivalent to $\tau_x = 0$, but $U$ in Equation \ref{eqn-syn} was left unchanged for balance purposes.
+with recovery time constant $\tau_x = 150$~ms and release fraction $U = 0.4$.
 
-When a neuron's membrane potential reached $v \geq v_\theta$, a spike was emitted, and the potential was clamped to $v = v_{reset}$ for a fixed refractory period of 3~ms (excitatory neurons) or 2~ms (inhibitory neurons).
-For inhibitory neurons, the firing threshold remained constant at $v_\theta = \theta_0$, while for excitatory neurons, the threshold included an adaptive component, following
+When a neuron's membrane potential reached the firing threshold $v_\theta$, a spike was emitted, and the potential was clamped to $v = v_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons).
+Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $v_\theta = \theta_0 = -54$~mV \refp{Mensi2012-au}.
+In contrast, excitatory neurons were modeled with threshold adaptation (TA) as in \cite{Teeter2018-iz}, following
 \begin{align}
-    v_\theta &= \theta_0 + \theta \nonumber \\
-    \tau_{TA} \frac{dV_{TA}}{dt} &= -V_{TA} + \hat{\theta} \delta(t - \hat{t})
+    v_\theta &= \theta_0 + V_{TA}(t) \nonumber \\
+    \tau_{\theta} \frac{dV_{TA}}{dt} &= -V_{TA} + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
 \end{align}
-In ablated experiments without TA, $\hat{\theta}$ was set to $0$ to abolish adaptation.
+with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \refp{Pozzorini2015-ei}.
 
@@
     \includegraphics{fig1.pdf}
-    \caption{\textbf{a} Sample network layout, showing excitatory and inhibitory neurons and all outgoing synaptic connections from a small subset.
-\textbf{b} Distance from stimulation site 0 in terms of minimum number of excitatory synapses.
-Stimulation is provided to the 10 neurons nearest to each of 5 stimulation sites, which are regularly spaced on a circle with a radius of 2.5 mm.
-\textbf{c} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes.
-The excitatory connection has weight $w = 3$ for demonstration purposes only; a single connection with $w = 1$ is normally unable to evoke postsynaptic firing.
-Top: Membrane voltage (black, $\tau$ = 30 ms), firing threshold baseline (gray) and adaptive threshold (green, $\tau$ = 1 s) demonstrating TA.
-Middle: Excitatory ($g_e$, $\tau$ = 2 ms) and inhibitory ($g_i$, $\tau$ = 4 ms) synaptic conductances, demonstrating excitatory STD.
-Bottom: Inputs to $g_e$ and $g_i$ (vertical bars), and the depression variable $x$ of the excitatory presynaptic neuron (dashed line)}
+    \caption{}
     \label{fig1}
@@
 \textbf{Fig.
-1 a} Sample network layout, showing excitatory and inhibitory neurons and all outgoing synaptic connections from a small subset.
-\textbf{b} Distance from stimulation site 0 in terms of minimum number of excitatory synapses.
-Stimulation is provided to the 10 neurons nearest to each of 5 stimulation sites, which are regularly spaced on a circle with a radius of 2.5 mm.
-\textbf{c} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes.
+1}Model and paradigm.
+\textbf{a} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron, illustrated above the plots, firing pre-determined spikes.
 The excitatory connection has weight $w = 3$ for demonstration purposes only; a single connection with $w = 1$ is normally unable to evoke postsynaptic firing.
-Top: Membrane voltage (black, $\tau$ = 30 ms), firing threshold baseline (gray) and adaptive threshold (green, $\tau$ = 1 s) demonstrating TA.
-Middle: Excitatory ($g_e$, $\tau$ = 2 ms) and inhibitory ($g_i$, $\tau$ = 4 ms) synaptic conductances, demonstrating excitatory STD.
-Bottom: Inputs to $g_e$ and $g_i$ (vertical bars), and the depression variable $x$ of the excitatory presynaptic neuron (dashed line)
+Top: Membrane voltage (black), firing threshold baseline (gray) and adaptive threshold (green) demonstrating TA.
+Middle: Excitatory ($g_e$, blue) and inhibitory ($g_i$, red) synaptic conductances, demonstrating excitatory STD.
+Bottom: Inputs to $g_e$ and $g_i$ (vertical bars in the appropriate colors), and the depression variable $x$ of the excitatory presynaptic neuron (dashed line).
+\textbf{b} Sample network layout, showing excitatory and inhibitory neurons and all outgoing synaptic connections from two neurons of each type.
+\textbf{c} Distance from stimulation site 0 in terms of minimum number of excitatory synapses.
+Stimulation is provided to the 10 neurons nearest to the center of each of 5 stimulation sites (round markers), which are regularly spaced on a circle with a radius of 2.5 mm.
 
-Figure \ref{fig1}c shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
+Figure \ref{fig1}a shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
 The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \citep{Stimberg2020-go}, and simulated with an integration time step of 1~ms.
 Synaptic delay was not modeled explicitly, but spikes were delivered, and voltages reset, in the time step following spike emission.
-We simulated neurons without any stochasticity in either their inputs or their parameters, other than network structure (see below), in order to ensure that our results were driven purely by the experimental paradigm, rather than any model-internal sources of noise.
+We simulated neurons without any stochasticity in either their inputs or their parameters, other than network structure (described below), in order to ensure that our results were driven purely by the experimental paradigm, rather than any model-internal sources of noise.
 
-800 excitatory and 200 inhibitory neurons were placed randomly in a two-dimensional circular space with a radius of 4~mm, mimicking a petri dish, and formed synapses of weight $w=1$ with 50 randomly selected postsynaptic partners within a range of 2~mm (excitatory) or 1~mm (inhibitory).
-Connectivity and weights remained fixed throughout simulation.
+800 excitatory and 200 inhibitory neurons were placed randomly in a two-dimensional circular space with a radius of 4~mm, and formed synapses of weight $w=1$ with 50 randomly selected postsynaptic partners within a range of 2~mm (excitatory) or 1~mm (inhibitory), reflecting the notion that excitatory neurons project over longer distances, while inhibitory neurons mainly connect to their local neighborhood.
+Connectivity and weights remained fixed throughout.
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-See Figure \ref{fig1}a,b for a representative example of the resulting network structure.
-Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response, discarding networks where less than half of all neurons in the fully recovered network responded to stimulation at any of the five sites.
+See Figure \ref{fig1}b and c for a representative example of the resulting network structure.
+Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response: Networks where fewer than 500 neurons responded to stimulation at any of the five sites after full recovery were discarded.
+In total, 30 networks were used for the experiments described below.
 

### 2022-12-08 paper-1/manuscript.tex
Rev sec Methods.Paradigm

@@
 \end{figure*}
-
 \textbf{Fig.
@@
 
-To investigate deviance detection, we used a classical oddball paradigm as illustrated in Figure \ref{fig2}a, presenting a total of 500 stimuli at regular intervals of 500~ms.
+To investigate deviance detection, we used a classical oddball paradigm as illustrated in Figure \ref{fig1}d, presenting a total of 500 stimuli at regular intervals of 500~ms.
 Target stimuli (labeled A hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (std), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (dev) with 100 A and 400 B, and the many-standards control sequence (msc) with 100 presentations of each of the stimuli A through E.
@@
 This yielded 4 complete data sets per network, or a total of 120 data sets across all networks.
+
+In order to disentangle the effects of STD and TA, we ran all simulations in four conditions: Without any short-term plasticity, with either TA only or STD only, and with both STD and TA (labeled STD+TA in the following, and corresponding to the full model as described above).
+To turn off STD, we replaced Equation \ref{eqn-xsyn} with a constant $x = 1$, but retained the scaling of inputs to $g_e$ with $U$ in Equation \ref{eqn-gsyn}.
+This corresponds to immediate recovery ($\tau_x = 0$) and was done to maintain the magnitude of the excitatory postsynaptic potential (EPSP) in the recovered state regardless of whether STD was turned on or off.
+To turn off TA, we replaced Equation \ref{eqn-TA} with a constant $V_{TA} = 0$, making the firing threshold completely unadaptive.
 

### 2022-12-08 paper-1/manuscript.tex
Rev sec Results.DD

@@
     \includegraphics{fig2.pdf}
-    \caption{\textbf{a} Illustration of the oddball paradigm used.
-Each square represents a stimulus, with the target (A) in dark blue, and other stimuli in lighter colors.
-The temporal extent of stimulation is exaggerated; in reality, stimulation was delivered in a single time step, and the response allowed to develop freely for 500 ms.
-\textbf{b} Response of a neuronal cell culture in vitro, adapted from \cite{Kubota2021-dx}.
-\textbf{c} Model response, showing average network-wide spike counts after stimulation with target (A) in each sequence.
-\textbf{d} Mismatch, SSA and deviance detection indices across networks and stimulus pairs (n = 120).
-In this and later figures, notches indicate the confidence interval for the median, calculated by bootstrap with 10000 iterations.
-All indices were positive (mismatch: t = 11, p = 6.7e-21, SSA: t = 7.8, p = 1.05e-12, deviance detection: t = 6.5, p = 1.05e-09)}
+    \caption{}
     \label{fig2}
 \end{figure*}
-
 \textbf{Fig.
-2 a} Illustration of the oddball paradigm used.
-Each square represents a stimulus, with the target (A) in dark blue, and other stimuli in lighter colors.
-The temporal extent of stimulation is exaggerated; in reality, stimulation was delivered in a single time step, and the response allowed to develop freely for 500 ms.
-\textbf{b} Response of a neuronal cell culture in vitro, adapted from \cite{Kubota2021-dx}.
-\textbf{c} Model response, showing average network-wide spike counts after stimulation with target (A) in each sequence.
-\textbf{d} Mismatch, SSA and deviance detection indices across networks and stimulus pairs (n = 120).
-In this and later figures, notches indicate the confidence interval for the median, calculated by bootstrap with 10000 iterations.
-All indices were positive (mismatch: t = 11, p = 6.7e-21, SSA: t = 7.8, p = 1.05e-12, deviance detection: t = 6.5, p = 1.05e-09)
+2} Deviance detection in vitro and in silico.
+\textbf{a} Response of a 2 \times 4~mm neuronal cell culture on a multi-electrode array, adapted from \cite{Kubota2021-dx}.
+Curves represent the average number of spikes per electrode in each 10~ms bin in response to target stimulation in the indicated sequence.
+Unlike in our study, oddball sequences here were split 10\%/90\%, and 10 equally frequent stimuli were used for the msc sequence.
+\textbf{b} Sample model response, showing average network-wide spike counts after stimulation with target (A) in the indicated sequence (same color code as panel \textbf{a}).
+Both in vitro and in the model, the dev response is very large, followed by an intermediate msc response and a low std response.
+\textbf{c} Mismatch, SSA and deviance detection indices (see Equation \ref{eqn-ddi}) across networks and stimuli (n = 120).
+In this and later boxplots, the median is indicated with an orange line, notches indicate the confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points
 
 We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs.
-Following related research \citep{TODO}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the dev sequence to A in the std sequence).
+Following related research \citep{Kubota2021-dx,Harms2014-ah,Jacobsen2001-sc}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the dev sequence to A in the std sequence).
 In other words, we used the msc sequence as a neutral baseline where the target stimulus was presented no more often than in the dev sequence, but no expectation of a competing stimulus B could be formed.
-To quantify deviance detection, we compared the mean number of spikes fired across the network in response to target stimulus A in the dev and msc sequences, defining a deviance detection index (DDI) as
+To quantify deviance detection, we compared the mean number of spikes $R^A$ fired across the network in response to target stimulus A in the dev and msc sequences, defining a deviance detection index (DDI) as
 \begin{equation}
-    DDI = \frac{R^A_{dev} - R^A_{msc}}{R^A_{dev} + R^A_{msc}}
+    DDI = \frac{R^A_{dev} - R^A_{msc}}{R^A_{dev} + R^A_{msc}} \label{eqn-ddi}
 \end{equation}
-We define stimulus-specific adaptation (SSA, msc - std) and mismatch (dev - std) indices analogously, capturing respectively the effect of adaptation, and of both adaptation and deviance detection together.
+We define stimulus-specific adaptation (SSA, $R^A_{msc} - R^A_{std}$) and mismatch ($R^A_{dev} - R^A_{std}$) indices analogously, capturing respectively the effect of adaptation, and of both adaptation and deviance detection together.
 
-Our first aim was to replicate deviance detection properties reported in dissociated neuronal cell culture by \cite{Kubota2021-dx}, reproduced in Figure \ref{fig2}b.
-With minimal tuning of global parameters, we found that most networks (92 out of 120 target stimuli) reliably produced not only mismatch responses based on adaptation alone, but also showed robust deviance detection, as summarized in Figure \ref{fig2}d.
+Our first aim was to replicate deviance detection properties reported in dissociated neuronal cell culture by \cite{Kubota2021-dx}, reproduced in Figure \ref{fig2}a.
+In their case, cell cultures -- like our model, randomly connected networks with two-dimensional extent -- were stimulated at 10 different sites in msc sequences, and at two sites with 10\%/90\% probability in dev and std sequences.
+Deviance detection was assessed analogously to our definition of DDI and reached an average level of around 0.3 across two cultures.
 
+In our model, we could readily identify networks and stimuli whose responses developed in a qualitatively similar fashion, with high activity in response to target in the dev sequence, low activity in the std sequence, and intermediate activity in the msc sequence, as illustrated with an example in Figure \ref{fig2}b.
 The model responses developed on a compressed time scale compared to neuronal culture, which we tentatively attribute to unfaithful aspects of the model, including a lack of conduction delay or spontaneous activity, a much smaller number of participating neurons, and structural differences.
-However, we could readily identify target stimuli whose response profiles looked otherwise qualitatively similar to the cell culture results.
 
-Having confirmed the presence of deviance detection in our model, we then turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model -- synaptic short-term depression (STD) and threshold adaptation (TA) -- enabled the networks to learn relevant regularities.
-As shown in Figure \ref{fig3}, networks with neither plasticity mechanism did not show any difference in response, indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
-Networks with only STD likewise did not show any deviance detection, though they did exhibit robust stimulus-specific adaptation effects, leading to a smaller target response in std than in either dev or msc sequences.
-Conversely, with only TA, networks produced both adaptation and deviance detection effects.
+Across networks and stimuli (Figure \ref{fig2}c), we found that our model reliably produced mismatch responses (t = 11.3, p = 6.7e-21; one-sided t-test for mismatch index > 0, n = 120) while both SSA index (t = 7.84, p = 1.05e-12) and DDI (t = 6.49, p = 1.05e-09) were also positive.
+The average DDI value reached around 0.1, which was less than the values reported by \cite{Kubota2021-dx}, but note that our target stimuli were presented more frequently (20\%) than theirs (10\%) in the dev and msc sequences.
 
@@
     \includegraphics{fig3.pdf}
-    \caption{Deviance detection and SSA indices under model ablation.
-\textbf{a} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
-\textbf{b, c} Deviance detection and SSA indices across networks and stimulus pairs.
-Significance was assessed by 1-sample t-tests.
-TODO p-values: text or table?}
+    \caption{}
     \label{fig3}
 \end{figure*}
+\textbf{Fig.
+3} Deviance detection under model ablation.
+\textbf{a} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
+The color code is the same as in Figure \ref{fig2}.
+\textbf{b} SSA indices across all networks and stimuli.
+Asterisks indicate data greater than 0 at a significance level of p<0.05, see main text for detailed statistics.
+\textbf{c} Deviance detection indices across all networks and stimuli.
 
-\textbf{Fig.
-3} Deviance detection and SSA indices under model ablation.
-\textbf{a} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
-\textbf{b, c} Deviance detection and SSA indices across networks and stimulus pairs.
-Significance was assessed by 1-sample t-tests.
-TODO p-values: text or table?
+Having confirmed the presence of deviance detection in our model, we then turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
+Figure \ref{fig3}a shows the responses to target stimulation of a sample network under ablation.
+With no plasticity, the responses are indistinguishable between sequences, yielding a DDI of 0, and indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
+With STD only, the response in the dev sequence developed earlier and very slightly larger, yielding a DDI of 0.03.
+With TA only, the response became noticeably smaller, but the sequences became readily distinguishable, with high dev, intermediate msc, and low std responses, yielding a DDI of 0.19.
+Finally, in the full model, the dev response stood out more clearly, leaving much smaller msc and std responses and yielding a DDI of 0.4.
 
-Finally, with both STD and TA applied, deviance detection was also reliable, as shown before; however, we were surprised to find that the effect was significantly stronger than that of either plasticity mechanism alone (STD: t = 6.7, p = 4.53e-10; TA: t = 3.2, p = 0.00103), and greater too than the sum of the DDI values across both STD only and TA only models (t = 3.6, p = 0.000199).
+The SSA index (Figure \ref{fig3}b) was exactly 0 as expected in the ablation with no plasticity, and positive with any plasticity applied (STD only: t = 3.17, p = 0.000983, median 0.0039; TA only: t = 6.74, p = 3e-10, median 0.097; STD+TA: t = 7.84, p = 1.05e-12, median 0.135; one-sided t-tests, n = 120).
+Since the SSA index measures the degree of adaptation to frequent stimulus presentation, this indicates that both plasticity mechanisms were capable of imbuing the networks with some level of history dependence.
+
+Finally, across networks and stimuli, the DDI (Figure \ref{fig3}c) was exactly 0 with no plasticity, not significantly different from 0 with STD only (t = -1.89, p = 0.0609, median -2.1e-04; two-sided t-test, n = 120), but clearly positive for TA only (t = 5.52, p = 1.01e-07, median 0.077; one-sided t-test, n = 120) and the full model (t = 6.49, p = 1.05e-09, median 0.114).
+To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: t = 6.7, p = 4.53e-10; TA: t = 3.2, p = 0.00103), and greater too than the sum of the DDI values across both STD only and TA only models (t = 3.6, p = 0.000199).
 This suggests that TA and STD act not as independent mechanisms in this paradigm, but rather interact constructively to enhance the network's ability to encode input regularities.
 
-In the following, we will use the sample network and target stimulus shown in Figure \ref{fig3}a to first build an understanding of how TA leads to deviance detection when acting alone.
-Then, we will build on this understanding to elucidate how the addition of STD, which is ineffective on its own, can lead to an increased deviant response.
-The particular network and target stimulus site used in the following individual analyses was selected for the response size to target and non-target stimuli (A, B), both of which were average, as well as its deviance detection performance, which likewise reflected the tendencies observed across all networks.
+In the following, we will use a sample network and target stimulus to first build an understanding of how TA leads to deviance detection when acting alone, i.e.
+in the ablated TA only model.
+Then, we will build on this understanding to elucidate how the addition of STD, which was shown above to be ineffective for deviance detection on its own, can lead to an increased deviant response in the full model.
 

### 2022-12-09 paper-1/manuscript.tex
Rev title

@@
 
-\title[Synaptic and somatic short-term plasticity collaborate to yield deviance detection]{Short-term synaptic plasticity and threshold adaptation act in synergy to give rise to deviance detection in a spiking model of neuronal culture}
+\title{Short-term synaptic and neuronal plasticity act in synergy for deviance detection in spiking networks}
 

### 2022-12-09 paper-1/manuscript.tex
Rev sec Results.DD (sample selection)

@@
 Having confirmed the presence of deviance detection in our model, we then turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-Figure \ref{fig3}a shows the responses to target stimulation of a sample network under ablation.
+In the remainder of the paper, we will proceed to detail the processes leading to deviance detection on a sample network, then confirm the highlighted trends statistically with reference to the full set of 120 networks and stimuli.
+The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the msc sequence and calculated z-scores with respect to the population.
+We then focused on networks whose absolute z-score was less than 1 for both A and B responses, and hand-picked a sample that exhibited both an increased DDI in STD+TA over TA only, and positive SSA indices in both STD+TA and TA only conditions.
+
+Figure \ref{fig3}a shows the responses to target stimulation of the sample network under ablation.
 With no plasticity, the responses are indistinguishable between sequences, yielding a DDI of 0, and indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.

### 2022-12-09 digest.py
Added raw Vm to readout:get_voltages

@@ -23,11 +23,13 @@ def get_voltages(params, dynamics, overflow=None):
     threshold = ensure_unit(dynamics['th_adapt'], volt)
     synapses = ensure_unit(dynamics['vsyn'], volt) + depression  # vsyn + depr = vsyn + u-v = usyn
     reset = ensure_unit(dynamics['v'] - dynamics['vsyn'], volt) - params['v_rest']
+    vm = ensure_unit(dynamics['v'], volt)
     return {
         'Depression': depression,
         'Threshold': threshold,
         'Synapses': synapses,
-        'Reset': reset}
+        'Reset': reset,
+        'Vm': vm}
 
 
 def get_voltage_histograms(params, rundata, overflow=None):

### 2022-12-12 plot_utils.py
Replaced Vsyn with Vm for Fig4

@@ -13,7 +13,8 @@ colors = {
     'Depression': '#9b19f5',
     'Threshold': 'C2',
     'Synapses': 'orange',
-    'Reset': 'slategrey'
+    'Reset': 'slategrey',
+    'Vm': 'orange'
 }
 
 labels = {
@@ -22,6 +23,7 @@ labels = {
     'Threshold': '$V_{TA}$ (mV)',
     'Synapses': '$V_{syn}$ (mV)',
     'Reset': '$V_{reset}$ (mV)',
+    'Vm': '$V_m$ (mV)'
 }
 
 

### 2022-12-12 paper-1/manuscript.tex
Update v -> V_m

@@
 
-We modeled neurons as leaky integrate-and-fire units whose membrane potential $v$ at time $t$ followed
+We modeled neurons as leaky integrate-and-fire units whose membrane potential $V_m$ at time $t$ followed
 \begin{equation}
-    \tau_m \frac{dv}{dt} = (v_{rest}-v) + I_{syn}(t)
+    \tau_m \frac{dV_m}{dt} = (v_{rest}-V_m) + I_{syn}(t)
 \end{equation}
@@
 \begin{equation}
-    I_{syn} &= g_e(E_e-v) + g_i(E_i-v)
+    I_{syn} = g_e(E_e-V_m) + g_i(E_i-V_m)
 \end{equation}
@@
 
-When a neuron's membrane potential reached the firing threshold $v_\theta$, a spike was emitted, and the potential was clamped to $v = v_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons).
+When a neuron's membrane potential reached the firing threshold $v_\theta$, a spike was emitted, and the potential was clamped to $V_m = v_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons).
 Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $v_\theta = \theta_0 = -54$~mV \refp{Mensi2012-au}.

### 2022-12-12 paper-1/manuscript.tex
Include CI level in boxplot caption

@@
 \textbf{c} Mismatch, SSA and deviance detection indices (see Equation \ref{eqn-ddi}) across networks and stimuli (n = 120).
-In this and later boxplots, the median is indicated with an orange line, notches indicate the confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points
+In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points
 

### 2022-12-12 paper-1/manuscript.tex
Rev sec Results.TA (F4)

@@
 
-To understand why the same stimulus A evoked a higher response in dev than in msc, we first compared the average responses in the two conditions.
-As shown in Figure \ref{fig4}a, we found that the response developed in a qualitatively similar fashion, with a tendency of dev spikes to occur earlier, particularly shortly after stimulation, and more frequently, particularly towards the later part of the response.
-Since the stimulus was identical in both conditions, this difference had to be due to either changed firing thresholds, or changed synaptic currents as a consequence of activity differences in earlier parts of the response.
-We quantified the former with the strength of adaptation $V_{TA}$ (Figure \ref{fig4}b), which showed a clear and widespread lowering of thresholds at the start of dev trials relative to msc, followed by an increase beyond the msc baseline in many neurons as they spiked.
-Correlating the difference in $V_{TA}$ at the beginning of trials with the difference in trial spike count, we found that, across neurons, lower thresholds are strongly predictive of higher activity (r = -0.424, p = 3.1e-36, see Figure \ref{fig4}d).
+To understand why the same stimulus A evoked a higher response in dev than in msc with TA only, we first compared the average responses in the two conditions, sorting neurons by their response onset to permit a high-resolution view of the population activity.
+As shown in Figure \ref{fig4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both conditions.
+The contrast (right) revealed a tendency of dev spikes to occur earlier, particularly shortly after stimulation, as shown by the succession of orange (higher dev activity) by blue (higher msc activity) in the lower parts of the panel.
+In later parts of the response (top half of the panel), the dev response was larger, but not obviously shifted in time.
+Since the stimulus was identical in both conditions, and TA was the only plsaticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.
 
-To quantify the synaptic contributions to a neuron's membrane potential $v$, we calculated a quantity $V_{syn}$, shown in Figure \ref{fig4}c, which captures all synaptic inputs, but excludes voltage deflections caused by direct stimulation and by the self-inhibition induced by resetting to $v_{reset} < v_{rest}$ after a spike.
-Being induced by spiking activity, this measure closely followed the activity pattern, making the earlier activation of the dev response clearly visible across the network.
-In addition, since $V_{syn}$ decays with the membrane timescale $\tau_m$, but resets to $0$ on postsynaptic spikes, many of the neurons that spiked more frequently showed lower $V_{syn}$ towards the end of network activity.
+We quantified the former with the strength of adaptation $V_{TA}$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
+As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of dev trials relative to msc, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+In contrast, once the bulk of activity had occurred, many neurons showed higher thresholds in dev (right panel, later orange portions), since they spiked more often.
+Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $V_{TA}$.
+
+To assess the difference in synaptic currents in a way that allowed direct comparison to $V_{TA}$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs (resets after spikes notwithstanding).
+As the direct cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
+The contrast (right panel) shows not only a notable relative increase in membrane voltages in dev around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
 
@@
     \includegraphics{fig4.pdf}
-    \caption{Responses and threshold adaptation in a TA-only ablation.
-\textbf{a, b, c} Post-stimulus histograms across target trials, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
-Left and middle column, target trials in dev and msc sequence, respectively; right column, contrast between these two.
-\textbf{d} Per-neuron relationship between the contrast (dev - msc) in $V_{TA}$ at the start of trials, and the contrast in average number of spikes fired, excluding inhibitory neurons, and associated linear regression (r = -0.424, p = 3.1e-36).
-\textbf{e} Contribution of $V_{TA}$ and $V_{syn}$ to bins with increased firing in dev (i.e., red bins in the spike probability contrast, \textbf{a}), weighted by that same contrast.
-\textbf{f} Contribution of $V_{TA}$ relative to the sum of the weighted contributions in \textbf{e}.
-\textbf{g} Relative contribution of $V_{TA}$ across networks, showing median (solid line) and inter-quartile range (shaded area)}
+    \caption{}
     \label{fig4}
 \end{figure*}
-
-\textbf{Fig.
-4} Responses and threshold adaptation in a TA-only ablation.
-\textbf{a, b, c} Post-stimulus histograms across target trials, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
-Left and middle column, target trials in dev and msc sequence, respectively; right column, contrast between these two.
-\textbf{d} Per-neuron relationship between the contrast (dev - msc) in $V_{TA}$ at the start of trials, and the contrast in average number of spikes fired, excluding inhibitory neurons, and associated linear regression (r = -0.424, p = 3.1e-36).
-\textbf{e} Contribution of $V_{TA}$ and $V_{syn}$ to bins with increased firing in dev (i.e., red bins in the spike probability contrast, \textbf{a}), weighted by that same contrast.
-\textbf{f} Contribution of $V_{TA}$ relative to the sum of the weighted contributions in \textbf{e}.
-\textbf{g} Relative contribution of $V_{TA}$ across networks, showing median (solid line) and inter-quartile range (shaded area)
+\textbf{Fig. 4} Responses and threshold adaptation with TA only.
+\textbf{a} Target trial average post-stimulus spike histogram, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
+Left and middle column, target trials in dev sequence and msc sequence, respectively; right column, contrast between these two.
+\textbf{b} Target trial average of the threshold adaptation voltage $V_{TA}$, with axes, neuron order and columns as in panel a.
+\textbf{c} Target trial average of the membrane potential $V_m$, with axes, neuron order and columns as in panel a.
+\textbf{d} Relationship across excitatory neurons between the contrast (dev - msc) in $V_{TA}$ at the start of target trials, and the contrast in average number of spikes fired in target trials, and associated linear regression.
+\textbf{e} Median contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\\Delta V_{TA}$ vs response size contrast across excitatory neurons).
+Each marker corresponds to one network and target stimulus.
+Red markers indicate a significant correlation (p < 0.05).
+Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
+\textbf{f} In the sample network, contribution of the contrast in $V_{TA}$ and $V_m$ to bins with increased firing in dev (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
+\textbf{g} Contribution of $V_{TA}$ relative to the sum of the absolute values of the weighted contributions in panel f.
+\textbf{h} Relative contribution of $V_{TA}$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
 
 Based on these measures, we then asked why there were more spikes in dev trials.
-To answer this, we weighted the contrasts in $V_{TA}$ and $V_{syn}$ by the spike probability contrast, discarding any instances of equal or lower firing in dev, and summed over neurons.
-The resulting time courses (Figure \ref{fig4}e) reveal the immediate underlying causes of network-wide activity increases, namely de-adaptation and depolarization.
-We found that increased depolarization induced by synaptic input greatly outweighed the impact of reduced adaptation throughout most of the response.
-Only in the very early stages of the response did adaptation dominate the increase in firing (Figure \ref{fig4}f), whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
-This trend was reflected in most networks investigated, as shown in the summary plot in Figure \ref{fig4}g.
-To summarize, while most added spikes arose from increased prior activity within a trial, the initial impetus -- the increase in the early response -- was largely caused by lower TA.
+We first focused on the contribution of $V_{TA}$ alone in an effort to confirm that thresholds are lower in dev, and that this reduction in thresholds causes increased firing.
+We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response of each excitatory neuron in terms of spikes fired.
+In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (t = -17.3, p = 2.57e-57, median -0.166 mV; one-sided t-test, n=800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson correlation, $\rho$ = -0.424, p = 3.1e-36).
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Median $V_{TA}$ was lower in dev than msc (t = -11.2, p = 1.08e-20, median = -0.094 mV; one-sided t-test, n=120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (t = -23.1, p = 3.72e-46, median = -0.357).
+This indicates that networks were less adapted during oddball sequences, and that this lack of adaptation played an important role in the increased response to deviant inputs.
+
+Yet, as we saw in Figure \ref{fig4}c, the membrane voltages also differed between the two conditions.
+In order to assess the relative contribution of $V_{TA}$ to the increased dev response, we chose to focus exclusively on time points with higher firing probability in the dev sequence, discarding data points with equal or lower firing probability relative to the msc sequence.
+We then weighted the remaining contrasts in $V_{TA}$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
+The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased dev response.
+While thresholds were slighly lowered throughout the response, the membrane potential was dramatically increased.
+In relative terms (Figure \ref{fig4}g), $V_{TA}$ constituted a major driver of increased firing only right at the onset of the response, with later contributions amounting to less than 20\% of the total voltage difference.
+This pattern was consistent across networks and stimuli, as shown in Figure \ref{fig4}h.
+This indicates that adaptation initiated and dominated the increase in firing in the beginning, whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
 

### 2022-12-13 paper-1/manuscript.tex
Tweak Fig4 (mean, not med, dV_TA)

@@
 We quantified the former with the strength of adaptation $V_{TA}$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
-As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of dev trials relative to msc, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target dev trials relative to msc, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 In contrast, once the bulk of activity had occurred, many neurons showed higher thresholds in dev (right panel, later orange portions), since they spiked more often.
@@
 \textbf{c} Target trial average of the membrane potential $V_m$, with axes, neuron order and columns as in panel a.
+The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the color mask.
 \textbf{d} Relationship across excitatory neurons between the contrast (dev - msc) in $V_{TA}$ at the start of target trials, and the contrast in average number of spikes fired in target trials, and associated linear regression.
-\textbf{e} Median contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\\Delta V_{TA}$ vs response size contrast across excitatory neurons).
+\textbf{e} Mean contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\\Delta V_{TA}$ vs response size contrast across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response of each excitatory neuron in terms of spikes fired.
-In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (t = -17.3, p = 2.57e-57, median -0.166 mV; one-sided t-test, n=800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson correlation, $\rho$ = -0.424, p = 3.1e-36).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Median $V_{TA}$ was lower in dev than msc (t = -11.2, p = 1.08e-20, median = -0.094 mV; one-sided t-test, n=120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (t = -23.1, p = 3.72e-46, median = -0.357).
+In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (t = -17.3, p = 2.57e-57, median -0.166 mV; one-sided t-test, n=800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 3.1e-36).
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Average $V_{TA}$ was lower in dev than msc (t = -7.32, p = 1.58e-11, median = -0.074 mV; one-sided t-test, n=120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (t = -23.1, p = 3.72e-46, median = -0.357).
 This indicates that networks were less adapted during oddball sequences, and that this lack of adaptation played an important role in the increased response to deviant inputs.

### 2022-12-13 spatial.py
Refactored get_distance_to_stim

@@ -87,3 +87,19 @@ def get_stimulated(X, Y, Xstim, Ystim, params):
         sorted = np.argsort(dist)
         idx[i] = sorted[:params['neurons_per_stim']]
     return idx
+
+def get_distance_to_stim(stimid, W, X, Y, params):
+    Wb = W>0
+    d = np.zeros_like(X, int)
+    Xstim, Ystim = create_stimulus_locations(params)
+    d[get_stimulated(X, Y, Xstim, Ystim, params)[stimid]] = 1
+    for i in range(2, 10):
+        pre = (d == i-1) & (np.arange(len(d)) < params['N_exc'])
+        post = np.unique(np.nonzero(Wb[pre])[1])
+        for p in post:
+            if d[p] == 0:
+                d[p] = i
+        if np.sum(d==0) == 0:
+            break
+    d -= 1
+    return d

### 2022-12-13 spatial.py
Bugfix

@@ -90,7 +90,7 @@ def get_stimulated(X, Y, Xstim, Ystim, params):
 
 def get_distance_to_stim(stimid, W, X, Y, params):
     Wb = W>0
-    d = np.zeros_like(X, int)
+    d = np.zeros(X.shape, int)
     Xstim, Ystim = create_stimulus_locations(params)
     d[get_stimulated(X, Y, Xstim, Ystim, params)[stimid]] = 1
     for i in range(2, 10):

### 2022-12-14 paper-1/manuscript.tex
Rev sec Results.TA (F5~6)

@@
 The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the color mask.
-\textbf{d} Relationship across excitatory neurons between the contrast (dev - msc) in $V_{TA}$ at the start of target trials, and the contrast in average number of spikes fired in target trials, and associated linear regression.
+\textbf{d} Relationship across excitatory neurons between the contrast (dev - msc) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
 \textbf{e} Mean contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\\Delta V_{TA}$ vs response size contrast across excitatory neurons).
@@
 Having established that the increased response to A in dev trials was driven by lower thresholds, we next turned to the cause of this reduction.
-Firstly, plotting $V_{TA}$ at the start of A trials in Figure \ref{fig5}a, we found that most of the network was indeed less adapted in dev, with only a few neurons in the vicinity of B showing higher thresholds.
-We hypothesized that this could be due to the different non-target stimulation pattern, since B was stimulated more frequently in dev than msc.
-Indeed, as shown in Figure \ref{fig5}b, the non-target response activity in the dev sequence -- in other words, the response to B -- was concentrated around the site where B was delivered, while in the msc sequence, activity in response to the various non-target stimuli was dispersed across most of the network.
+A priori, since threshold adaptation is a direct reflection of activity, and since the response to target trials was larger in dev than in msc, the lower thresholds in dev must have been caused by lower activity in non-target trials, i.e.
+lower responses to B in dev than to B through E in msc.
+Yet, with activity modulated by thresholds as shown above, we might expect to observe higher dev responses regardless of stimulus.
+To resolve this apparent paradox, we will turn to an analysis of the spatial arrangement of activity and adaptation.
+
+First, we mapped the average value of $V_{TA}$ at the start of target trials to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
+In the msc sequence, $V_{TA}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
+Conversely, in the dev sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
+Next, we turned to the likely cause of this difference, the non-target responses (Figure \ref{fig5}b), mapped into space as the average number of spikes fired in each neuron.
+As expected, the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds.
 Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (alongside 300 C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, most of these very close to the stimulation site of B.
+
+Correlating the non-target response and thresholds excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev (t = -7.61, p = 3.5e-12, median = -0.044; one-sided t-test, n=800) almost perfectly predicted the resulting reduction in threshold adaptation levels (Pearson's $\rho$ = 0.99, p = 0).
+This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.994 and a consistent reduction in the average non-target response between msc and dev sequences (t = -11.6, p = 1.46e-21, median = -0.058; one-sided t-test, n=120), clearly validating our notion that the non-target trial response is the primary driver of threshold adaptation.
 
@@
     \includegraphics{fig5.pdf}
-    \caption{\textbf{a} $V_{TA}$ at the start of A trials (mean over trials), with each neuron shown in its spatial location.
+    \caption{}
+    \label{fig5}
+\end{figure*}
+\textbf{Fig. 5} Non-target response magnitude determines thresholds.
+\textbf{a} Sample network $V_{TA}$ at the start of A trials (mean over trials), with each neuron shown in its spatial location.
 Left and middle column, raw averages for dev and msc sequences, respectively; right column, contrast between dev and msc averages.
 A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the msc sequence.
-\textbf{b} Average response (in spikes per trial) to non-target trials (B in dev, B/C/D/E in msc).
-\textbf{c} Left and middle: Replacement and adaptation components of the non-target response contrast (\textbf{b}, right column), where replacement is the contrast (B msc - C/D/E msc), and adaptation is the contrast (B dev - B msc).
-Right: Network-wide mean replacement and adaptation values across networks and target stimuli.
-The adaptation effect is less than zero (t = -9, p = 2.02e-15), and less than the replacement effect (t = 2.3, p = 0.0119), which is not different from zero (t = -0.69, p = 0.246).
-\textbf{d} Per-neuron relationship between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation, and the contrast in $V_{TA}$ at the start of target trials, excluding inhibitory neurons, and associated linear regression (r = 0.99, p = 0).
-\textbf{e} The same relationship as \textbf{d}, but using the medians across excitatory neurons, and plotting one point per network and target stimulus.
-The values are highly correlated (r = 0.956, p = 5.98e-65)}
-    \label{fig5}
+\textbf{b} Average response (in spikes per trial) to non-target trials (B in dev, B, C, D, and E in msc). Columns are arranged as in panel a.
+\textbf{c} Relationship across excitatory neurons between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation and the contrast in $V_{TA}$ at the start of target trials, and associated linear regression.
+\textbf{d} Mean contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\\Delta V_{TA}$ across excitatory neurons).
+Each marker corresponds to one network and target stimulus.
+Red markers indicate a significant correlation (p < 0.05).
+Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations
+
+\begin{figure*}%
+    \centering
+    \includegraphics{fig6.pdf}
+    \caption{}
+    \label{fig6}
 \end{figure*}
-
-\textbf{Fig.
-5 a} $V_{TA}$ at the start of A trials (mean over trials), with each neuron shown in its spatial location.
-Left and middle column, raw averages for dev and msc sequences, respectively; right column, contrast between dev and msc averages.
-A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the msc sequence.
-\textbf{b} Average response (in spikes per trial) to non-target trials (B in dev, B/C/D/E in msc).
-\textbf{c} Left and middle: Replacement and adaptation components of the non-target response contrast (\textbf{b}, right column), where replacement is the contrast (B msc - C/D/E msc), and adaptation is the contrast (B dev - B msc).
-Right: Network-wide mean replacement and adaptation values across networks and target stimuli.
-The adaptation effect is less than zero (t = -9, p = 2.02e-15), and less than the replacement effect (t = 2.3, p = 0.0119), which is not different from zero (t = -0.69, p = 0.246).
-\textbf{d} Per-neuron relationship between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation, and the contrast in $V_{TA}$ at the start of target trials, excluding inhibitory neurons, and associated linear regression (r = 0.99, p = 0).
-\textbf{e} The same relationship as \textbf{d}, but using the medians across excitatory neurons, and plotting one point per network and target stimulus.
-The values are highly correlated (r = 0.956, p = 5.98e-65)
-
-We quantified the correlation between the differences in non-target response magnitude and strength of adaptation before target trials across excitatory neurons, and found an almost perfect relationship (Pearson's r = 0.99, n = 800, see Figure \ref{fig5}d).
-This effect was robust across networks: Greater reduction in the non-target response was closely correlated with a greater reduction in $V_{TA}$ at the start of A trials.
+\textbf{Fig. 6} Stimulus replacement and adaptation both lead to reduced dev non-target response.
+\textbf{a} Non-target response contrasts: The replacement contrast (top right) compares the response to B in msc (top left) to all non-target responses in msc (top center).
+The adaptation contrast (bottom right) compares the response to B in dev (bottom left) to the response to B in msc (bottom center).
+Both contrasts together add up to the non-target contrast (dev - msc) shown in Figure \ref{fig5}b.
+Note that the contrast column (right) shares a color scale to allow a direct comparison.
+\textbf{b} Mean replacement (left) and adaptation (right) contrasts across networks and stimuli, as well as the mean replacement contrast filtered for neurons near B, i.e., neurons within one excitatory synapse of the neurons under direct stimulation of B (center).
+\textbf{c} Mean $V_{TA}$ contrast (dev - msc) at the start of A trials, filtered for neurons near B.
+Each marker corresponds to one network; the sample network is highlighted in red.
+The histogram in the background shows the distribution of the data, with the orange portion indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations
 
 Finally, we examined the cause of the reduced response in non-target trials.
-Going from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation, and secondly, the more frequent presentation of B changes adaptation, reducing the average response.
-To estimate the effect of stimulus replacement and adaptation, respectively, we contrasted the response to B in the msc sequence with the response to other non-target stimuli in the same sequence (replacement), and with the response to B in dev (adaptation).
-We found that, in this particular network and stimulus pair, adaptation weakly but consistently decreased the response, whereas stimulus replacement reduced many responses more dramatically while permitting larger responses in many neurons in the general vicinity of the stimulus sites of A and B.
-Across networks, however, the effect of adaptation was more pronounced and more consistent, removing an average of 71 spikes per trial, whereas stimulus replacement had no effect on average.
+Going from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation.
+Secondly, the more frequent presentation of B changes adaptation, likely reducing the average response to B.
+In order to separate these two, we split the non-target contrast shown in Figure \ref{fig5}b into two components: a replacement effect, contrasting the response to B in the msc sequence with the response to all non-target stimuli in the same sequence; and an adaptation effect, contrasting the response to B in dev with the response to B in msc.
+For the sample network, these contrasts are shown in Figure \ref{fig6}a.
+Unsurprisingly, replacement yielded a decreased response near the stimulation sites for C, D and E, and an increased response in areas near the stimulation site for B.
+By contrast, adaptation caused a decreased response in virtually the entire network, particularly in the vicinity of the stimulation site for B.
 
-To conclude, TA plays two complementary roles in deviance detection, affecting the standard and deviant response separately: On the one hand, it reduces the response to the frequently presented standard stimulus, leading to a larger deviant response merely by contrast with the corresponding standard (SSA).
-On the other hand, due to the overlap in receptive fields, TA acts as blanket suppression in the many-standards control condition, against which the deviant response, relieved from this suppression and encountering a less fatigued network, stands out.
+Across networks and stimuli (Figure \ref{fig6}b), we saw a wide range of replacement effects, including both increased and decreased firing, likely as a result of differing sensitivities of the networks to the various non-target stimuli.
+On average, stimulus replacement did not have a significant effect (t = -0.255, p = 0.799; two-sided t-test).
+In contrast, adaptation yielded a reduced response on average (t = -9.01, p = 2.02e-15; one-sided t-test), indicating that the more frequent presentation of B indeed lowered the network's response to this stimulus.
+
+Finally, given the clear spatial pattern exhibited by the replacement effect in the sample network, we surmised that adaptation is driven mainly by neurons close to the stimulation site B responding more frequently due to the more frequent presentation of the stimulus.
+To show this, we first identified all neurons located within a single excitatory synapse of the stimulated neurons (cf.
+Figure \ref{1}c), and calculated the average replacement effect for these alone.
+In this subpopulation, replacement yielded a clear response increase, shown in the central box of Figure \ref{fig6}b (t = 8.21, p = 1.5e-13; one-sided t-test).
+As shown in Figure \ref{fig6}c, this response increase directly translated into elevated thresholds in dev trials compared to msc (t = 5.01, p = 9.59e-07).
+
+Earlier, we showed that the early response is strongly influenced by TA (Figure \ref{fig4}h).
+Here, we showed that $V_{TA}$ near the stimulation site was increased in dev as a result of the more frequent presentation of B.
+We posit that this increase in TA affects the early response, reducing the response further downstream and freeing up the rest of the network to respond more strongly to the infrequent target stimulus.
+
+To conclude, TA plays two complementary roles in deviance detection, affecting the standard and deviant response separately: In its first role, as we have shown with the adaptation contrast above, TA reduces the response to frequently presented stimuli regardless of context, leading to a stronger response to infrequently presented stimuli (SSA).
+Since this happens without the need for large-scale adaptation -- only a small fraction of the network near the stimulation site adapts strongly to frequently presented inputs -- the network as a whole is allowed to relax in the oddball condition.
+In its second role, TA causes network-wide adaptation in the control condition, where many different stimuli are presented, all of which are infrequently presented and therefore able to evoke large responses.
+Deviant detection then arises as a relief from this blanket suppression, or as a result of adaptation shifting from the entire network to a small region around the frequently presented standard.
 

### 2022-12-14 paper-1/manuscript.tex
Rev sec Results.STD (F7)

@@
 
-As noted above, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA.
-To understand the basis of this apparent synergy, we contrasted the TA-only condition examined in the previous section against the same network with both TA and STD.
-As shown in Figure \ref{fig6}, we found that the non-target response magnitude was reduced by the introduction of STD, as we would expect.
-Indeed, across networks, the response to all stimuli was reduced, regardless of context.
-Concomitantly, and in keeping with the above analysis, the levels of $V_{TA}$ before A trials also dropped.
+% adding STD: What happens?
+As noted previously, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA.
+To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks with both TA and STD.
+On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (data not shown).
+However, this reduction was not uniform: Target responses were reduced most in the std sequence, and least in the dev sequence, this difference leading directly to the greater DD and SSA indices reported in Figure \ref{fig3}.
+
+% Why does dXA increase?:
+To understand the reason for this difference, we first focused on the target response and asked how and why the target response contrast (dev - msc) changed as STD was added to the model.
+Firstly, as shown in Figure \ref{fig7}a, the target response was greater in dev than msc in both the full and the ablated model (indicating a positive DDI).
+In most neurons, the contrast was greater in the full model, such that the difference (full - ablated) was positive, as indicated by the largely orange plot in the right column.
+Following the logic established in the previous section, we may expect a greater target response in dev to be caused by lower $V_{TA}$ relative to msc.
+Indeed, comparing the contrast (dev - msc) in $V_{TA}$ before target trials across models (Figure \ref{fig7}b), we see that, while the contrast is negative (i.e., lower in dev) in both models, the magnitude of the contrast is greater in the full model.
+For clarity, the difference is therefore negative, as indicated by the largely blue plot in the right column.
+Lastly, we trace this one step further back in Figure \ref{fig7}c: Again as established in the previous section, a greater contrast in $V_{TA}$ may be explained by a similarly enhanced contrast (dev - msc) in the non-target response, which is exactly what we found: The non-target response was smaller in dev than msc by a greater margin in the full model, as indicated by the largely blue plot in the right column.
+
+This pattern is perhaps simpler to grasp in the aggregate across networks and stimuli.
+Firstly, the contrast (dev - msc) in the target response (Figure \ref{fig7}d, left) was positive in both models, but the full model contrast was greater than the ablated model contrast (t = 2.99, p = 0.00168; one-sided t-test, n = 120).
+Referring to the previous section, we locate the cause for this in an enhanced contrast in $V_{TA}$, as shown in Figure \ref{fig7}e; the contrast (dev - msc) in the full model was more negative than that in the ablated model (t = -3.49, p = 0.000342).
+Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921).
+In other words, the reason for the larger target dev response, relative to msc, is the smaller non-target dev response, relative to msc.
 
@@
     \centering
-    \includegraphics{fig6.pdf}
-    \caption{Contrast between the ablated model with only TA (shown in Figure \ref{fig5}) and the full model with both STD and TA.
-\textbf{a}, \textbf{b} Data shown in Figure \ref{fig5}ab, dev and msc columns, subtracted from the corresponding data for the full model.
-\textbf{c} Equivalent contrast (full - ablated model) in the response to A, in columns as above.
-\textbf{d} Network-wide mean $V_{TA}$ before A trials across networks.
-Left: Ablated and full model averages.
-Right: Difference between these model conditions.
-With the introduction of STD, $V_{TA}$ decreased more in dev than in msc (t = -3.49, p = 0.000342).
-\textbf{e} Mean number of spikes per target trial across networks.
-Left: Ablated and full model averages.
-Right: Difference between these model conditions.
-With the introduction of STD, the response decreased more in the std sequence than msc (t = -1.9, p = 0.0297), and more in msc than in dev (t = -2.99, p = 0.00168)}
-    \label{fig6}
+    \includegraphics{fig7.pdf}
+    \caption{}
+    \label{fig7}
 \end{figure*}
+\textbf{Fig. 7} Adding STD to the model enhances the effect of TA.
+\textbf{a} Contrast (dev - msc) of the response to A, in terms of mean number of spikes per trial, in the sample network, with each neuron shown in its spatial location.
+Left and middle column, full model with STD and TA, and ablated model with only TA, respectively.
+Right column, subtraction of these two.
+Stimulus locations are as shown in Figure \ref{fig5}a.
+\textbf{b} Contrast of mean $V_{TA}$ at the start of target trials, in columns as in panel a.
+\textbf{c} Contrast of the response to non-target stimuli, in columns as in panel a.
+\textbf{d} Network-wide mean contrast of the response to target (left) and non-target stimuli (right) across networks and stimuli, arranged equivalently to the columns in panels a-c.
+Blue and green boxes represent the full and ablated model, respectively, and white boxes the difference between these two.
+\textbf{e} Network-wide mean contrast of $V_{TA}$ before A trials across networks and stimuli, arranged as in panel d
 

### 2022-12-14 paper-1/manuscript.tex
Editing

@@
 To investigate deviance detection, we used a classical oddball paradigm as illustrated in Figure \ref{fig1}d, presenting a total of 500 stimuli at regular intervals of 500~ms.
-Target stimuli (labeled A hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (std), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (dev) with 100 A and 400 B, and the many-standards control sequence (msc) with 100 presentations of each of the stimuli A through E.
+Target stimuli (labeled "A" or "target" hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (std), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (dev) with 100 A and 400 B, and the many-standards control sequence (msc) with 100 presentations of each of the stimuli A through E.
 In this way, A is presented as the predominant and therefore expected stimulus (std), as an infrequent violation of an expectation (of B, dev), and, to control for pure adaptation effects, as an equally infrequent stimulus with no strong expectation of any particular input (msc).
@@
 \textbf{Fig. 4} Responses and threshold adaptation with TA only.
-\textbf{a} Target trial average post-stimulus spike histogram, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
+\textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
 Left and middle column, target trials in dev sequence and msc sequence, respectively; right column, contrast between these two.
@@
 The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the color mask.
-\textbf{d} Relationship across excitatory neurons between the contrast (dev - msc) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
-\textbf{e} Mean contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\\Delta V_{TA}$ vs response size contrast across excitatory neurons).
+\textbf{d} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
+\textbf{e} Network-wide mean contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\\Delta V_{TA}$ vs response size contrast across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 \textbf{Fig. 5} Non-target response magnitude determines thresholds.
-\textbf{a} Sample network $V_{TA}$ at the start of A trials (mean over trials), with each neuron shown in its spatial location.
+\textbf{a} $V_{TA}$ at the start of A trials (mean over trials) in the sample network, with each neuron shown in its spatial location.
 Left and middle column, raw averages for dev and msc sequences, respectively; right column, contrast between dev and msc averages.
@@
 \textbf{b} Average response (in spikes per trial) to non-target trials (B in dev, B, C, D, and E in msc). Columns are arranged as in panel a.
-\textbf{c} Relationship across excitatory neurons between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation and the contrast in $V_{TA}$ at the start of target trials, and associated linear regression.
-\textbf{d} Mean contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\\Delta V_{TA}$ across excitatory neurons).
+\textbf{c} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation and the contrast in $V_{TA}$ at the start of target trials, and associated linear regression.
+\textbf{d} Network-wide mean contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\\Delta V_{TA}$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 \end{figure*}
-\textbf{Fig. 6} Stimulus replacement and adaptation both lead to reduced dev non-target response.
-\textbf{a} Non-target response contrasts: The replacement contrast (top right) compares the response to B in msc (top left) to all non-target responses in msc (top center).
+\textbf{Fig. 6} Stimulus replacement and adaptation both lead to a reduced dev non-target response.
+\textbf{a} Non-target response contrasts in the sample network: The replacement contrast (top right) compares the response to B in msc (top left) to all non-target responses in msc (top center).
 The adaptation contrast (bottom right) compares the response to B in dev (bottom left) to the response to B in msc (bottom center).
@@
 Note that the contrast column (right) shares a color scale to allow a direct comparison.
-\textbf{b} Mean replacement (left) and adaptation (right) contrasts across networks and stimuli, as well as the mean replacement contrast filtered for neurons near B, i.e., neurons within one excitatory synapse of the neurons under direct stimulation of B (center).
+\textbf{b} Network-wide mean replacement (left) and adaptation (right) contrasts across networks and stimuli, as well as the mean replacement contrast filtered for neurons near B, i.e., neurons within one excitatory synapse of the neurons under direct stimulation of B (center).
 \textbf{c} Mean $V_{TA}$ contrast (dev - msc) at the start of A trials, filtered for neurons near B.

### 2022-12-15 paper-1/manuscript.tex
Rev Results:STD (F8)

@@
 
-\textbf{Fig.
-6} Contrast between the ablated model with only TA (shown in Figure \ref{fig5}) and the full model with both STD and TA.
-\textbf{a}, \textbf{b} Data shown in Figure \ref{fig5}ab, dev and msc columns, subtracted from the corresponding data for the full model.
-\textbf{c} Equivalent contrast (full - ablated model) in the response to A, in columns as above.
-\textbf{d} Network-wide mean $V_{TA}$ before A trials across networks.
-Left: Ablated and full model averages.
-Right: Difference between these model conditions.
-With the introduction of STD, $V_{TA}$ decreased more in dev than in msc (t = -3.49, p = 0.000342).
-\textbf{e} Mean number of spikes per target trial across networks.
-Left: Ablated and full model averages.
-Right: Difference between these model conditions.
-With the introduction of STD, the response decreased more in the std sequence than msc (t = -1.9, p = 0.0297), and more in msc than in dev (t = -2.99, p = 0.00168)
+This naturally raises the question of why, with the addition of STD, the non-target dev response was reduced more than other responses.
+We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.
+To show this in the sample network, we correlated the number of spikes in dev target trials with the spike-triggered depression variable $x_j$ at the start of later trials (cf.
+Equation \ref{eqn-xsyn}) across excitatory neurons.
+We found that, while the correlation was very clear in the trial immediately following target (Pearson's \rho = -0.984, p = 0), it was almost completely lost by the next trial (\rho = -0.0525, p = 0.138) under the influence of both decay and unrelated activity in the intervening trial.
+In fact, even without intervening activity, inputs to $x_j$ would decay to a negligible $e^\frac{1 s}{\tau_x} = 0.0013$ within two trials.
+Thus, STD could not have any stimulus-specific effects across dev target trials or msc trials.
 
-However, we noticed a systematic difference in the reduction of both response magnitude and $V_{TA}$ between the dev and msc conditions: The non-target response in the dev sequence was reduced more than that in the msc sequence, leading to a larger difference between the responses in these two contexts.
-$V_{TA}$ before A trials followed suit, leading to stronger de-adaptation and, as a consequence, a larger A trial response in dev in large parts of the network in the full model compared to TA-only.
-Across networks, only few cases of an increased dev response were seen; however, the response to the target stimulus (Figure \ref{fig6}e) was reduced least in dev, and most in std.
-This differential reduction was directly responsible for the larger DDI reported in Figure \ref{fig3}, but how did it arise?
+However, given the substantial correlation from one trial to the next, the frequently presented stimulus B could have experienced stimulus-specific adaptation as a result of STD.
+Referring back to Figure \ref{fig3}b, we see that there was indeed a significant, if small, effect of presentation frequency even in the ablated model with STD only, yielding a positive SSA index on average.
 
-Due to its relatively short time scale, the direct effect of STD across trials was likely limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.
-However, by reducing the standard (i.e., the dev non-target) response in the full model, we hypothesized that STD also affects adaptation levels, causing de-adaptation and thereby indirectly increasing the target dev response.
-
-In direct support of the notion that STD indeed carried over between (repeated) trials, we noted first that the response size in the STD-only condition was slightly less in std than in msc or dev (cf.
-Figure \ref{fig3}c), indicating some amount of SSA.
-To show that the across-trial effect of STD in the full model was limited to the dev non-target trials, however, we needed a more sophisticated tool, showing the level of depression experienced by each neuron.
-Since the effect of STD is to reduce excitatory postsynaptic potentials (EPSP), we estimated this reduction per synapse in a given trial type $S$ (e.g.
-A dev) as the product of the expected number of transmitted spikes $R^S$, the peak EPSP of the synapse ($k = 1.4 mV$, recall that all weights are 1), and its average depression coefficient $(1-x)$ at the start of the trial.
-Summing these estimates onto each postsynaptic neuron, we arrived at a rough estimate
-
+To show that STD carry-over was responsible for the reduction of non-target dev pulses relative to msc, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) is affected by changes in STD.
+Unlike $V_{TA}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects at the other side of a synapse.
+Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost.
+$D$ is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so must be calculated based on observed activity.
+Here, we used the average trial response $R_j^S$ of neuron $j$ to a given stimulus $S$, in number of spikes, as the presynaptic drive.
+Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at rest (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.
+Thus, ignoring added depression due to repeated spikes within a trial, and instead using the average depression at the start of trials $x_j^S$, we estimated the average peak EPSP evoked by a presynaptic neuron $j$ as $R_j^S k x_j^S$.
+Finally, with $x_j \in [0,1]$ and summing over presynaptic neurons, we could estimate the peak EPSP lost to depression in a given neuron $i$ as
 \begin{equation}
-    D^S = \sum_{i \in pre} R^S_i k (1-x_i^{t=0})
+    D_i^S = \sum_{j \in pre} R_j^S k (1-x_j^S)
 \end{equation}
-
-of the average depolarization lost due to STD.
+Notice that $D$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $V_{TA}$, albeit more approximate.
 
@@
     \centering
-    \includegraphics{fig7.pdf}
-    \caption{\textbf{a} Estimated average STD-derived postsynaptic depression $D$ in target (A) trials (top row) and B trials (bottom row).
-Note that the difference between A and B in msc (middle column) is due only to the network configuration.
-\textbf{b} Statistics across networks of the contrast (dev - msc) in $D$ (cf.
-\textbf{a}, right column), and in $V_{TA}$ at the start of A trials (cf.
-Figure \ref{fig6}d).
-In each network, averages were taken over neurons with at least 0.2 spikes per dev trial (B trials for $\Delta D^B$, A trials for $\Delta D^A$ and $\Delta V_{TA}$).
-Whereas the third box ($\Delta V_{TA}$ total) indicates the averages in the full model, the final box ($\Delta V_{TA}$ added) indicates the difference between the full and the ablated (TA-only) model.
-This additional amount of $\Delta V_{TA}$, caused indirectly by the addition of STD, was less (i.e., more negative) than the directly STD-related $\Delta D^A$ itself (t = -2.09, p = 0.0195)}
-    \label{fig7}
+    \includegraphics{fig8.pdf}
+    \caption{}
+    \label{fig8}
 \end{figure*}
+\textbf{Fig. 8} Estimated STD-mediated depression $D$ has small direct effects.
+\textbf{a} Average depression $D$ in target (A) trials (top row) and B trials (bottom row) in the dev (left) and msc sequence (center), and their contrast (right).
+The color scale is shared across columns (and between dev and msc) for direct comparability.
+\textbf{b} Statistics across networks and stimuli of the contrast (dev - msc) in $D$ ($D^B$ is shown first, since all other data in this panel relate directly to A), and in $V_{TA}$ at the start of A trials (cf.
+Figure \ref{fig7}d).
+In each network, averages were taken over neurons active (>0.2 spikes/trial) in either dev or msc trials (B trials for $\Delta D^B$, A trials for $\Delta D^A$ and $\Delta V_{TA}$).
+Whereas the third box ($\Delta V_{TA}$ total) indicates averages in the full model, the final box ($\Delta V_{TA}$ added) indicates the difference between the full and the ablated (TA only) model
 
-\textbf{Fig.
-7 a} Estimated average STD-derived postsynaptic depression $D$ in target (A) trials (top row) and B trials (bottom row).
-Note that the difference between A and B in msc (middle column) is due only to the network configuration.
-\textbf{b} Statistics across networks of the contrast (dev - msc) in $D$ (cf.
-\textbf{a}, right column), and in $V_{TA}$ at the start of A trials (cf.
-Figure \ref{fig6}d).
-In each network, averages were taken over neurons with at least 0.2 spikes per dev trial (B trials for $\Delta D^B$, A trials for $\Delta D^A$ and $\Delta V_{TA}$).
-Whereas the third box ($\Delta V_{TA}$ total) indicates the averages in the full model, the final box ($\Delta V_{TA}$ added) indicates the difference between the full and the ablated (TA-only) model.
-This additional amount of $\Delta V_{TA}$, caused indirectly by the addition of STD, was less (i.e., more negative) than the directly STD-related $\Delta D^A$ itself (t = -2.09, p = 0.0195)
+As noted above, we hypothesized that, in the full model, STD is directly responsible only for a reduction of non-target dev trials, relative to msc, whereas differences between target dev and msc trials are not a result of STD directly.
+To show this, we calculated $D$ for A and B trials, and contrasted it across dev and msc conditions as with $V_{TA}$ before.
+In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the network in both conditions.
+While there were differences between the two conditions, they were relatively small and did not appear to follow a readily identifiable pattern.
+Conversely, B trials (bottom) were spatially more clustered around the stimulation site in both conditions.
+In the msc sequence, we can attribute the difference between A and B primarily to the response size: The context was the same, but the B response tended not to spread out as much in this network.
+However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site (see also Figure \ref{fig6}a), with depression, of course, similarly limited.
+As a result, the contrast (dev - msc) shows slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicative of increased synaptic depression in dev.
 
-In Figure \ref{fig7}, we show that, in A dev and A msc trials, depression had relatively evenly distributed effects across the network.
-Neither sequence seemed to have a stronger overall effect, despite a slight trend towards weaker depression near the stimulus site in the dev sequence.
-In sharp contrast, in B dev trials, depression was very strong and very clearly localized to the stimulation site.
-This location dependence was closely mirrored in the B dev network response, which was largely confined to the area near the stimulation site (data not shown).
-To quantify this result while avoiding spurious events, we averaged the depression value $D$ over neurons with at least 1 spike per 5 trials of the corresponding type in the dev sequence.
-Across networks, we found that depression in the dev sequence was systematically stronger for B than for A (t = 7.5, p = 7.07e-12).
-More importantly, the level of depression in A trials changed considerably less between msc and dev sequences when compared to B trials (t = -6.7, p = 3.92e-10).
-Taken together, this shows that, indeed, the dominant direct effect of STD was to weaken the response to the frequently presented standard stimulus.
+In order to show that this was a robust effect, we turn to statistics across networks and stimuli.
+We calculated the trial-average depression values $D^A$ and $D^B$ in dev and msc conditions, subtracted them, then calculated the mean over neurons, excluding neurons with less than 0.2 spikes per trial in both conditions [because $\Delta D^B$ $\approx$ zero if I don't do this.
+Not sure how to justify...].
+The resulting contrasts are summarized in the left half of Figure \ref{fig8}b.
+We note first that, using the same exclusion criteria, we found that neurons received more depression in B trials than in A trials in the dev sequence (t = -4.4, p = 1.17e-05; one-sided t-test, n = 120), but not in the msc sequence (t = -1.14, p = 0.129).
+Furthermore, $D^B$ was greater in dev than msc (t = 2.94, p = 0.002).
+Both findings confirm that STD carried over between the frequent standard trials, but not, or less so, between control trials.
 
-Finally, to complete the picture and confirm our hypothesis in full, we compared the small, but significant relief from depression in A dev trials relative to msc (t = -3.2, p = 0.001) to the difference in $V_{TA}$, applying the same neuron exclusion criteria.
-We found that the relief from suppression, which was shown in the TA-only analysis above to be a key ingredient to deviance detection, was driven almost exclusively by lower $V_{TA}$ (t = 11, p = 4.06e-19).
-More importantly, adding STD to the model induced a much larger additional difference in $V_{TA}$ compared to the TA-only condition, than in directly STD-related $D$ (t = -2.09, p = 0.0195), indicating that, indeed, STD affected the deviant response primarily through indirect effects.
+However, we also found that $D^A$ was greater in msc than dev (t = -3.4, p = 0.000458), suggesting that STD did exert stimulus-independent suppression in the msc sequence, similar to what we described earlier for TA.
+The magnitude of the effect, however, was very different.
+In the right half of Figure \ref{fig8}b, we show the contrast of $V_{TA}$ before A trials, which was clearly much greater than the contrast in $D^A$.
+We wondered, however, how much $V_{TA}$ was affected by the addition of STD.
+To assess this, we subtracted the contrast (dev - msc) of $V_{TA}$ in the full model from the same contrast calculated in the ablated model with TA only.
+This additional decrease in $V_{TA}$ from msc to dev was substantially greater than the corresponding decrease in $D^A$ (t = -1.93, p = 0.0277), indicating that the direct effect of STD on the target dev response was less than indirect effects.
+
+To sum up, adding STD to the model had a number of direct and indirect effects.
+Firstly, it reduced responses in all conditions, likely by reducing the synaptic output of neurons at stimulus sites, which reliably fired several spikes immediately following stimulation.
+Secondly, it directly reduced the response to frequently presented dev B trials through stimulus-specific adaptation.
+Thirdly, STD was very slightly reduced in A trials in dev over msc, mirroring the direct stimulus-independent adaptation effect detailed in the previous section for TA.
+And finally, the decreased dev B trial response allowed TA in the wider network to recover more, which in turn led to increased responses in dev A trials.
+This indirect effect was substantially greater than the direct effect on target trials, as shown above, indicating that it was solely or largely responsible for the increased DDI.
 

### 2022-12-15 paper-1/manuscript.tex
Editing

@@
 
-% adding STD: What happens?
 As noted previously, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA.
@@
 
-% Why does dXA increase?:
 To understand the reason for this difference, we first focused on the target response and asked how and why the target response contrast (dev - msc) changed as STD was added to the model.
@@
 Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921).
-In other words, the reason for the larger target dev response, relative to msc, is the smaller non-target dev response, relative to msc.
+In other words, the reason for the larger target dev response, relative to msc, is the smaller non-target dev response, relative to msc; furthermore, since the relationships established in the previous section remained largely unchanged, this effect appeared to be mediated primarily through TA.
 

### 2022-12-15 paper-1/manuscript.tex
Rev title
V_rest caps consistency

@@
 
-\title{Short-term synaptic and neuronal plasticity act in synergy for deviance detection in spiking networks}
+\title{Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks}
 
@@
 \begin{equation}
-    \tau_m \frac{dV_m}{dt} = (v_{rest}-V_m) + I_{syn}(t)
+    \tau_m \frac{dV_m}{dt} = (V_{rest}-V_m) + I_{syn}(t)
 \end{equation}
-with $tau_m = 30$~ms the membrane time constant and $v_{rest} = -60$~mV the resting membrane potential. Synaptic currents were modeled in a conductance-based manner, following
+with $tau_m = 30$~ms the membrane time constant and $V_{rest} = -60$~mV the resting membrane potential. Synaptic currents were modeled in a conductance-based manner, following
 \begin{equation}
@@
 
-When a neuron's membrane potential reached the firing threshold $v_\theta$, a spike was emitted, and the potential was clamped to $V_m = v_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons).
+When a neuron's membrane potential reached the firing threshold $v_\theta$, a spike was emitted, and the potential was clamped to $V_m = V_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons).
 Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $v_\theta = \theta_0 = -54$~mV \refp{Mensi2012-au}.
@@
 \textbf{d} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
-\textbf{e} Network-wide mean contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\\Delta V_{TA}$ vs response size contrast across excitatory neurons).
+\textbf{e} Network-wide mean contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta V_{TA}$ vs response size contrast across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 \textbf{c} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation and the contrast in $V_{TA}$ at the start of target trials, and associated linear regression.
-\textbf{d} Network-wide mean contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\\Delta V_{TA}$ across excitatory neurons).
+\textbf{d} Network-wide mean contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\Delta V_{TA}$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.

### 2022-12-15 paper-1/manuscript.tex
Editing

@@
 2} Deviance detection in vitro and in silico.
-\textbf{a} Response of a 2 \times 4~mm neuronal cell culture on a multi-electrode array, adapted from \cite{Kubota2021-dx}.
+\textbf{a} Response of a 2 $\times$ 4~mm neuronal cell culture on a multi-electrode array, adapted from \cite{Kubota2021-dx}.
 Curves represent the average number of spikes per electrode in each 10~ms bin in response to target stimulation in the indicated sequence.
@@
 Equation \ref{eqn-xsyn}) across excitatory neurons.
-We found that, while the correlation was very clear in the trial immediately following target (Pearson's \rho = -0.984, p = 0), it was almost completely lost by the next trial (\rho = -0.0525, p = 0.138) under the influence of both decay and unrelated activity in the intervening trial.
+We found that, while the correlation was very clear in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the influence of both decay and unrelated activity in the intervening trial.
 In fact, even without intervening activity, inputs to $x_j$ would decay to a negligible $e^\frac{1 s}{\tau_x} = 0.0013$ within two trials.

### 2022-12-19 paper-1/manuscript.tex
Spaces around =

@@
 
-800 excitatory and 200 inhibitory neurons were placed randomly in a two-dimensional circular space with a radius of 4~mm, and formed synapses of weight $w=1$ with 50 randomly selected postsynaptic partners within a range of 2~mm (excitatory) or 1~mm (inhibitory), reflecting the notion that excitatory neurons project over longer distances, while inhibitory neurons mainly connect to their local neighborhood.
+800 excitatory and 200 inhibitory neurons were placed randomly in a two-dimensional circular space with a radius of 4~mm, and formed synapses of weight $w = 1$ with 50 randomly selected postsynaptic partners within a range of 2~mm (excitatory) or 1~mm (inhibitory), reflecting the notion that excitatory neurons project over longer distances, while inhibitory neurons mainly connect to their local neighborhood.
 Connectivity and weights remained fixed throughout.
@@
 We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response of each excitatory neuron in terms of spikes fired.
-In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (t = -17.3, p = 2.57e-57, median -0.166 mV; one-sided t-test, n=800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 3.1e-36).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Average $V_{TA}$ was lower in dev than msc (t = -7.32, p = 1.58e-11, median = -0.074 mV; one-sided t-test, n=120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (t = -23.1, p = 3.72e-46, median = -0.357).
+In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (t = -17.3, p = 2.57e-57, median -0.166 mV; one-sided t-test, n = 800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 3.1e-36).
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Average $V_{TA}$ was lower in dev than msc (t = -7.32, p = 1.58e-11, median = -0.074 mV; one-sided t-test, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (t = -23.1, p = 3.72e-46, median = -0.357).
 This indicates that networks were less adapted during oddball sequences, and that this lack of adaptation played an important role in the increased response to deviant inputs.
@@
 
-Correlating the non-target response and thresholds excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev (t = -7.61, p = 3.5e-12, median = -0.044; one-sided t-test, n=800) almost perfectly predicted the resulting reduction in threshold adaptation levels (Pearson's $\rho$ = 0.99, p = 0).
-This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.994 and a consistent reduction in the average non-target response between msc and dev sequences (t = -11.6, p = 1.46e-21, median = -0.058; one-sided t-test, n=120), clearly validating our notion that the non-target trial response is the primary driver of threshold adaptation.
+Correlating the non-target response and thresholds excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev (t = -7.61, p = 3.5e-12, median = -0.044; one-sided t-test, n = 800) almost perfectly predicted the resulting reduction in threshold adaptation levels (Pearson's $\rho$ = 0.99, p = 0).
+This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.994 and a consistent reduction in the average non-target response between msc and dev sequences (t = -11.6, p = 1.46e-21, median = -0.058; one-sided t-test, n = 120), clearly validating our notion that the non-target trial response is the primary driver of threshold adaptation.
 

### 2022-12-19 paper-1/manuscript.tex
Minor rev sec Results:STD, F7

@@
 On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (data not shown).
-However, this reduction was not uniform: Target responses were reduced most in the std sequence, and least in the dev sequence, this difference leading directly to the greater DD and SSA indices reported in Figure \ref{fig3}.
+However, this reduction was not uniform: Target responses were reduced most in the std sequence, and least in the dev sequence.
 
-To understand the reason for this difference, we first focused on the target response and asked how and why the target response contrast (dev - msc) changed as STD was added to the model.
-Firstly, as shown in Figure \ref{fig7}a, the target response was greater in dev than msc in both the full and the ablated model (indicating a positive DDI).
+To understand the reason for this asymmetry, we first focused on the target response and asked how and why the target response contrast (dev - msc) changed as STD was added to the model.
+As shown in Figure \ref{fig7}a, the target response was greater in dev than msc in both the full and the ablated model (indicating a positive DDI in both models, cf.
+also Figure \ref{fig3}).
 In most neurons, the contrast was greater in the full model, such that the difference (full - ablated) was positive, as indicated by the largely orange plot in the right column.
@@
 For clarity, the difference is therefore negative, as indicated by the largely blue plot in the right column.
-Lastly, we trace this one step further back in Figure \ref{fig7}c: Again as established in the previous section, a greater contrast in $V_{TA}$ may be explained by a similarly enhanced contrast (dev - msc) in the non-target response, which is exactly what we found: The non-target response was smaller in dev than msc by a greater margin in the full model, as indicated by the largely blue plot in the right column.
+Lastly, we trace this one step further back in Figure \ref{fig7}c: Again as established in the previous section, a greater contrast in $V_{TA}$ may be explained by a similarly enhanced contrast (dev - msc) in the non-target response.
+We found exactly this: The non-target response was smaller in dev than msc by a greater margin in the full model, as indicated by the largely blue plot in the right column.
 
-This pattern is perhaps simpler to grasp in the aggregate across networks and stimuli.
+This pattern was reflected also in the aggregate across networks and stimuli.
 Firstly, the contrast (dev - msc) in the target response (Figure \ref{fig7}d, left) was positive in both models, but the full model contrast was greater than the ablated model contrast (t = 2.99, p = 0.00168; one-sided t-test, n = 120).
-Referring to the previous section, we locate the cause for this in an enhanced contrast in $V_{TA}$, as shown in Figure \ref{fig7}e; the contrast (dev - msc) in the full model was more negative than that in the ablated model (t = -3.49, p = 0.000342).
-Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921).
-In other words, the reason for the larger target dev response, relative to msc, is the smaller non-target dev response, relative to msc; furthermore, since the relationships established in the previous section remained largely unchanged, this effect appeared to be mediated primarily through TA.
+Referring to the previous section, we identify the cause for this in an enhanced contrast in $V_{TA}$, as shown in Figure \ref{fig7}e; the contrast (dev - msc) in the full model was more negative than that in the ablated model (t = -3.49, p = 0.000342).
+Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely the non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921).
+To sum up, the reason for the larger target dev response, relative to msc, was the smaller non-target dev response, relative to msc; furthermore, since the relationships established in the previous section remained largely unchanged, this effect appeared to be mediated primarily through TA.
 

### 2022-12-19 paper-1/manuscript.tex
Rev sec Results:STD, F8 partial

@@
 We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.
-To show this in the sample network, we correlated the number of spikes in dev target trials with the spike-triggered depression variable $x_j$ at the start of later trials (cf.
+
+We had good reasons for this assumption: Firstly, notice that the ablated model with STD only produced only a positive SSA index (i.e., the target response in msc was greater than in std), but not a significantly positive DDI (i.e., no difference between target dev and msc responses).
+Secondly, we note that updates to the spike-triggered depression variable $x_j$ (cf.
+Equation \ref{eqn-xsyn}) would decay to a negligible $e^\frac{-1 s}{\tau_x} = 0.0013$ times the original value within two trials, making it unlikely that inputs separated by intervening trials could be "remembered" in the network.
+Conversely, the frequently presented stimulus B could lead to a build-up of depression strong enough to suppress its own response.
+Referring back to Figure \ref{fig3}b, we see that there was indeed a significant, if small, effect of presentation frequency even in the ablated model with STD only, yielding a positive SSA index on average.
+% To show this in the sample network, we correlated the number of spikes in dev target trials with the spike-triggered depression variable $x_j$ at the start of later trials (cf.
 Equation \ref{eqn-xsyn}) across excitatory neurons.
-We found that, while the correlation was very clear in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the influence of both decay and unrelated activity in the intervening trial.
-In fact, even without intervening activity, inputs to $x_j$ would decay to a negligible $e^\frac{1 s}{\tau_x} = 0.0013$ within two trials.
-Thus, STD could not have any stimulus-specific effects across dev target trials or msc trials.
+We found that, while the correlation was very clear in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the influence of both decay and unrelated activity in the intervening trial.
 
-However, given the substantial correlation from one trial to the next, the frequently presented stimulus B could have experienced stimulus-specific adaptation as a result of STD.
-Referring back to Figure \ref{fig3}b, we see that there was indeed a significant, if small, effect of presentation frequency even in the ablated model with STD only, yielding a positive SSA index on average.
-
-To show that STD carry-over was responsible for the reduction of non-target dev pulses relative to msc, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) is affected by changes in STD.
-Unlike $V_{TA}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects at the other side of a synapse.
-Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost.
-$D$ is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so must be calculated based on observed activity.
+To show that STD being carried over between trials was responsible for the reduction of non-target dev pulses relative to msc, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
+Unlike $V_{TA}$, which directly affected the neuron to which the value belongs, $x_j$ exerted its effects on the other side of a synapse.
+Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $V_{TA}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to bring a neuron to fire.
+A depression measure is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
 Here, we used the average trial response $R_j^S$ of neuron $j$ to a given stimulus $S$, in number of spikes, as the presynaptic drive.
 Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at rest (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.
-Thus, ignoring added depression due to repeated spikes within a trial, and instead using the average depression at the start of trials $x_j^S$, we estimated the average peak EPSP evoked by a presynaptic neuron $j$ as $R_j^S k x_j^S$.
-Finally, with $x_j \in [0,1]$ and summing over presynaptic neurons, we could estimate the peak EPSP lost to depression in a given neuron $i$ as
+Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $R_j^S k x_j^S$.
+For this estimate, we ignored added depression due to repeated spikes within a trial, instead defining $x_j^S$ as the average depression at the start of trials.
+Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
 \begin{equation}
@@
 Notice that $D$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $V_{TA}$, albeit more approximate.
+Lastly, since we contrasted this measure between conditions (see below), we aimed to avoid effects purely due to differences in presynaptic activity.
+Therefore, we calculated $R^S$ as the average response across relevant trials in all conditions, leaving the contrast entirely in the hands of $x^S$.
 
@@
 The color scale is shared across columns (and between dev and msc) for direct comparability.
-\textbf{b} Statistics across networks and stimuli of the contrast (dev - msc) in $D$ ($D^B$ is shown first, since all other data in this panel relate directly to A), and in $V_{TA}$ at the start of A trials (cf.
-Figure \ref{fig7}d).
-In each network, averages were taken over neurons active (>0.2 spikes/trial) in either dev or msc trials (B trials for $\Delta D^B$, A trials for $\Delta D^A$ and $\Delta V_{TA}$).
-Whereas the third box ($\Delta V_{TA}$ total) indicates averages in the full model, the final box ($\Delta V_{TA}$ added) indicates the difference between the full and the ablated (TA only) model
+\textbf{b} Statistics across networks and stimuli of the contrast (dev - msc) in $D$ in A trials (left) and B trials (right)
 
@@
 To show this, we calculated $D$ for A and B trials, and contrasted it across dev and msc conditions as with $V_{TA}$ before.
-In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the network in both conditions.
-While there were differences between the two conditions, they were relatively small and did not appear to follow a readily identifiable pattern.
-Conversely, B trials (bottom) were spatially more clustered around the stimulation site in both conditions.
-In the msc sequence, we can attribute the difference between A and B primarily to the response size: The context was the same, but the B response tended not to spread out as much in this network.
-However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site (see also Figure \ref{fig6}a), with depression, of course, similarly limited.
-As a result, the contrast (dev - msc) shows slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicative of increased synaptic depression in dev.
+In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both conditions.
+The msc sequence, however, clearly evoked greater depression, as borne out in the contrast (dev - msc), which was negative throughout.
+Conversely, depression in B trials (bottom) were spatially more clustered around the stimulation site in both conditions.
+In the msc sequence, we can attribute the difference between A and B primarily to the response size: The context was the same, but responses to B tended not to spread out as much in this network.
+However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site (see also Figure \ref{fig6}a), with depression by definition similarly limited.
+As a result, the contrast (dev - msc) showed slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicative of greatly increased synaptic depression in dev.
 

### 2022-12-20 plot_utils.py
Refactored inset_hist

@@ -3,6 +3,8 @@ import numpy as np
 from matplotlib.colors import LinearSegmentedColormap
 from matplotlib.colors import colorConverter
 from brian2.units import msecond
+import scipy.stats as stats
+import seaborn as sns
 import styling
 
 
@@ -110,3 +112,31 @@ def fill_ratios(*ratios, to=100):
     remainder = to-total
     ratios[ratios<0] = remainder / (ratios<0).sum()
     return ratios
+
+def inset_hist(ax, data, x=True, median_color='C1'):
+    if x:
+        y = ax.get_ylim()
+        ax.set_ylim(top=y[1] + .1*(y[1]-y[0]))
+        tx = ax.twinx()
+    else:
+        y = ax.get_xlim()
+        ax.set_xlim(right=y[1] + .1*(y[1]-y[0]))
+        tx = ax.twiny()
+    binned, bins, *_ = tx.hist(data, 20, color='grey', histtype='stepfilled', edgecolor='dimgrey', orientation='vertical' if x else 'horizontal')
+    ci = stats.bootstrap([data], np.median, n_resamples=10000)
+    lo, hi = ci.confidence_interval
+    bsize = np.diff(bins)[0]
+    mask = (bins > lo - bsize) & (bins < hi + bsize)
+    y = bins[mask]
+    y[0], y[-1] = lo, hi
+
+    if x:
+        tx.fill_between(y, binned[mask[1:]], step='pre', ec='dimgrey', fc=median_color)
+        tx.set_ylim(bottom=-10*binned.max())
+        tx.set_yticks([])
+    else:
+        tx.fill_betweenx(y, binned[mask[1:]], step='pre', ec='dimgrey', fc=median_color)
+        tx.set_xlim(left=-10*binned.max())
+        tx.set_xticks([])
+    sns.despine(ax=tx)
+    return tx

### 2022-12-20 paper-1/manuscript.tex
Rev sec Results:STD, F8 partial

@@
 \textbf{b} Statistics across networks and stimuli of the contrast (dev - msc) in $D$ in A trials (left) and B trials (right)
+\textbf{c} Relationship across neurons of the sample network between the contrast (dev - msc) in $D^S$ and the contrast in average number of spikes fired in $S$ trials, and associated linear regression.
+Left, target trials; right, B trials.
+\textbf{d} Network-wide mean contrast (dev - msc) in $D^S$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in panel c ($\Delta D^S$ vs response size contrast across neurons).
+Each marker corresponds to one network and target stimulus.
+Red markers indicate a significant negative correlation (one-sided t-test, significance level 0.05).
+Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
 
@@
 The msc sequence, however, clearly evoked greater depression, as borne out in the contrast (dev - msc), which was negative throughout.
-Conversely, depression in B trials (bottom) were spatially more clustered around the stimulation site in both conditions.
+Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both conditions.
 In the msc sequence, we can attribute the difference between A and B primarily to the response size: The context was the same, but responses to B tended not to spread out as much in this network.
-However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site (see also Figure \ref{fig6}a), with depression by definition similarly limited.
-As a result, the contrast (dev - msc) showed slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicative of greatly increased synaptic depression in dev.
+However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site (see also Figure \ref{fig6}a), with depression therefore similarly limited.
+As a result, the contrast (dev - msc) showed slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicating greatly increased synaptic depression in dev.
 
-In order to show that this was a robust effect, we turn to statistics across networks and stimuli.
-We calculated the trial-average depression values $D^A$ and $D^B$ in dev and msc conditions, subtracted them, then calculated the mean over neurons, excluding neurons with less than 0.2 spikes per trial in both conditions [because $\Delta D^B$ $\approx$ zero if I don't do this.
-Not sure how to justify...].
-The resulting contrasts are summarized in the left half of Figure \ref{fig8}b.
-We note first that, using the same exclusion criteria, we found that neurons received more depression in B trials than in A trials in the dev sequence (t = -4.4, p = 1.17e-05; one-sided t-test, n = 120), but not in the msc sequence (t = -1.14, p = 0.129).
-Furthermore, $D^B$ was greater in dev than msc (t = 2.94, p = 0.002).
-Both findings confirm that STD carried over between the frequent standard trials, but not, or less so, between control trials.
+To show how robust these effects were, we turn to statistics across networks and stimuli.
+We calculated the trial-average depression values $D^A$ and $D^B$ in dev and msc conditions, subtracted them, then calculated the mean over neurons.
+The resulting contrasts are summarized in Figure \ref{fig8}b.
+We note first that neurons were subject to more depression in B trials than in A trials in the dev sequence (-4.47, p = 9.02e-06; one-sided t-test, n = 120, data not shown), but not in the msc sequence (by definition).
+Furthermore, $D^B$ was greater in dev than msc (8.02, p = 4.05e-13).
+Both findings confirm that STD carried over between the frequent standard trials, but not, or less so, between the less frequent control and deviant trials.
+However, we also found that $D^A$ was greater in msc than dev (-7.54, p = 5.04e-12; one-sided t-test, n = 120), suggesting that STD did exert stimulus-independent suppression in the msc sequence, as is clearly seen in the sample network, and similar to what we described earlier for TA.
+
+Finally, to show that $D$ reflects suppression of a neuron's activity, we quantified the relationship between the contrasts in $D$ and the corresponding response size contrast.
+In the sample network (Figure \ref{fig8}c), we see that neurons with a greater decrease in $D$ from msc to dev had correspondingly greater response increase, and vice versa.
+Across both A and B trials, this relationship was captured well by a linear regression (A: Pearson's $\rho$ = -0.541, p = 2.28e-77, B: $\rho$ = -0.186, p = 1.61e-09; one-sided t-test, n = 1000).
+These findings were borne out across most networks and stimuli (Figure \ref{fig8}d), showing that $D$ captures the impact of STD not only on $V_m$, but also on the resulting output spikes.
+
+To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction of the response to the frequently presented standard stimuli (dev B) by adaptation, and an increase in the response to deviant stimuli (dev A) by relief from adaptation, relative to control.
+% ...
 

### 2022-12-20 paper-1/manuscript.tex
Rev figure titles

@@
 \end{figure*}
-\textbf{Fig. 4} Responses and threshold adaptation with TA only.
+\textbf{Fig. 4} Responses to dev target trials are larger due to lower $V_{TA}$.
 \textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
@@
 \end{figure*}
-\textbf{Fig. 5} Non-target response magnitude determines thresholds.
+\textbf{Fig. 5} Lower $V_{TA}$ in dev is a consequence of lower dev non-target response magnitude.
 \textbf{a} $V_{TA}$ at the start of A trials (mean over trials) in the sample network, with each neuron shown in its spatial location.
@@
 \end{figure*}
-\textbf{Fig. 6} Stimulus replacement and adaptation both lead to a reduced dev non-target response.
+\textbf{Fig. 6} Dev non-target responses are reduced due to adaptation.
 \textbf{a} Non-target response contrasts in the sample network: The replacement contrast (top right) compares the response to B in msc (top left) to all non-target responses in msc (top center).
@@
 \end{figure*}
-\textbf{Fig. 7} Adding STD to the model enhances the effect of TA.
+\textbf{Fig. 7} Adding STD to the model enhances the direct effect of TA.
 \textbf{a} Contrast (dev - msc) of the response to A, in terms of mean number of spikes per trial, in the sample network, with each neuron shown in its spatial location.
@@
 \end{figure*}
-\textbf{Fig. 8} Estimated STD-mediated depression $D$ has small direct effects.
+\textbf{Fig. 8} In the full model, STD reduces dev non-target and increases dev target responses relative to msc.
 \textbf{a} Average depression $D$ in target (A) trials (top row) and B trials (bottom row) in the dev (left) and msc sequence (center), and their contrast (right).

### 2022-12-20 paper-1/manuscript.tex
Editing

@@
 Each marker corresponds to one network and target stimulus.
-Red markers indicate a significant correlation (p < 0.05).
+Red markers indicate a significant correlation (two-sided t-test, significance level 0.05).
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
@@
 We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response of each excitatory neuron in terms of spikes fired.
-In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (t = -17.3, p = 2.57e-57, median -0.166 mV; one-sided t-test, n = 800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 3.1e-36).
+In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (t = -17.3, p = 2.57e-57, median -0.166 mV; one-sided t-test, n = 800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided t-test, n = 800).
 As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Average $V_{TA}$ was lower in dev than msc (t = -7.32, p = 1.58e-11, median = -0.074 mV; one-sided t-test, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (t = -23.1, p = 3.72e-46, median = -0.357).
@@
 Each marker corresponds to one network and target stimulus.
-Red markers indicate a significant correlation (p < 0.05).
+Red markers indicate a significant positive correlation (one-sided t-test, significance level 0.05).
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations
@@
 Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $V_{TA}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to bring a neuron to fire.
-A depression measure is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
+A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
 Here, we used the average trial response $R_j^S$ of neuron $j$ to a given stimulus $S$, in number of spikes, as the presynaptic drive.

### 2022-12-20 paper-1/manuscript.tex
Editing

@@
 Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely the non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921).
-To sum up, the reason for the larger target dev response, relative to msc, was the smaller non-target dev response, relative to msc; furthermore, since the relationships established in the previous section remained largely unchanged, this effect appeared to be mediated primarily through TA.
+To sum up, the reason for the larger target dev response relative to msc, was the smaller non-target dev response relative to msc; furthermore, since the relationships established in the previous section remained largely unchanged, this effect appeared to be mediated primarily through TA.
 

### 2022-12-20 paper-1/manuscript.tex
Rev sec Results:STD, F8 (done!)

@@
 To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction of the response to the frequently presented standard stimuli (dev B) by adaptation, and an increase in the response to deviant stimuli (dev A) by relief from adaptation, relative to control.
-% ...
-
-However, we also found that $D^A$ was greater in msc than dev (t = -3.4, p = 0.000458), suggesting that STD did exert stimulus-independent suppression in the msc sequence, similar to what we described earlier for TA.
-The magnitude of the effect, however, was very different.
-In the right half of Figure \ref{fig8}b, we show the contrast of $V_{TA}$ before A trials, which was clearly much greater than the contrast in $D^A$.
-We wondered, however, how much $V_{TA}$ was affected by the addition of STD.
-To assess this, we subtracted the contrast (dev - msc) of $V_{TA}$ in the full model from the same contrast calculated in the ablated model with TA only.
-This additional decrease in $V_{TA}$ from msc to dev was substantially greater than the corresponding decrease in $D^A$ (t = -1.93, p = 0.0277), indicating that the direct effect of STD on the target dev response was less than indirect effects.
-
-To sum up, adding STD to the model had a number of direct and indirect effects.
-Firstly, it reduced responses in all conditions, likely by reducing the synaptic output of neurons at stimulus sites, which reliably fired several spikes immediately following stimulation.
-Secondly, it directly reduced the response to frequently presented dev B trials through stimulus-specific adaptation.
-Thirdly, STD was very slightly reduced in A trials in dev over msc, mirroring the direct stimulus-independent adaptation effect detailed in the previous section for TA.
-And finally, the decreased dev B trial response allowed TA in the wider network to recover more, which in turn led to increased responses in dev A trials.
-This indirect effect was substantially greater than the direct effect on target trials, as shown above, indicating that it was solely or largely responsible for the increased DDI.
+Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in dev B responses also entailed lower $V_{TA}$, and higher dev A response by consequence.
+Thus, adding STD to the model increased the DDI both directly, as outlined at the start of this paragraph, and indirectly, by altering the pattern of TA by way of changed responses.
 

### 2022-12-21 paper-1/manuscript.tex
Editing

@@
 
-Across networks and stimuli (Figure \ref{fig2}c), we found that our model reliably produced mismatch responses (t = 11.3, p = 6.7e-21; one-sided t-test for mismatch index > 0, n = 120) while both SSA index (t = 7.84, p = 1.05e-12) and DDI (t = 6.49, p = 1.05e-09) were also positive.
+Across networks and stimuli (Figure \ref{fig2}c), we found that our model reliably produced positive mismatch responses (t = 11.3, p = 6.7e-21; one-sided t-test, n = 120) while both SSA index (t = 7.84, p = 1.05e-12) and DDI (t = 6.49, p = 1.05e-09) were also positive.
 The average DDI value reached around 0.1, which was less than the values reported by \cite{Kubota2021-dx}, but note that our target stimuli were presented more frequently (20\%) than theirs (10\%) in the dev and msc sequences.
@@
 \textbf{b} SSA indices across all networks and stimuli.
-Asterisks indicate data greater than 0 at a significance level of p<0.05, see main text for detailed statistics.
+Asterisks indicate data greater than 0, established by t-test with significance level 0.05, see main text for detailed statistics.
 \textbf{c} Deviance detection indices across all networks and stimuli.

### 2022-12-21 paper-1/manuscript.tex
Rev Fig6 caption

@@
 \end{figure*}
-\textbf{Fig. 6} Dev non-target responses are reduced due to adaptation.
-\textbf{a} Non-target response contrasts in the sample network: The replacement contrast (top right) compares the response to B in msc (top left) to all non-target responses in msc (top center).
+\textbf{Fig. 6} Dev non-target responses are reduced due to adaptation, which is caused by greater activity in the early responders to B.
+\textbf{a} Non-target response contrasts in the sample network: The replacement contrast (top right) compares the response to B in msc (top left) to all non-target responses, including B, in msc (top center).
 The adaptation contrast (bottom right) compares the response to B in dev (bottom left) to the response to B in msc (bottom center).
 Both contrasts together add up to the non-target contrast (dev - msc) shown in Figure \ref{fig5}b.
-Note that the contrast column (right) shares a color scale to allow a direct comparison.
-\textbf{b} Network-wide mean replacement (left) and adaptation (right) contrasts across networks and stimuli, as well as the mean replacement contrast filtered for neurons near B, i.e., neurons within one excitatory synapse of the neurons under direct stimulation of B (center).
-\textbf{c} Mean $V_{TA}$ contrast (dev - msc) at the start of A trials, filtered for neurons near B.
-Each marker corresponds to one network; the sample network is highlighted in red.
-The histogram in the background shows the distribution of the data, with the orange portion indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations
+Note that the contrast column (right) shares a color scale to allow direct comparison.
+\textbf{b} Network-wide mean replacement and adaptation contrasts across networks and stimuli.
+\textbf{c} Relationship across networks and stimuli between the replacement contrast and the contrast in $V_{TA}$ at the start of A trials, and associated linear regression.
+For both measures, only the first 100 neurons to respond to stimulation by B in any condition were included.
+Each marker corresponds to one network and target stimulus.
 

### 2022-12-21 paper-1/manuscript.tex
Editing

@@
 When a neuron's membrane potential reached the firing threshold $v_\theta$, a spike was emitted, and the potential was clamped to $V_m = V_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons).
-Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $v_\theta = \theta_0 = -54$~mV \refp{Mensi2012-au}.
+Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $v_\theta = \theta_0 = -54$~mV \citep{Mensi2012-au}.
 In contrast, excitatory neurons were modeled with threshold adaptation (TA) as in \cite{Teeter2018-iz}, following
@@
 \end{align}
-with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \refp{Pozzorini2015-ei}.
+with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \citep{Pozzorini2015-ei}.
 

### 2022-12-21 styling.py
Drop legend font size to 6

@@ -20,8 +20,8 @@ sns.set_theme(style='ticks', font_scale=0.75, rc={
     'grid.linewidth': 0.4,
     'lines.linewidth': 0.5,
     'patch.linewidth': 0.5,
-    'legend.fontsize': 8,
-    'legend.title_fontsize': 9,
+    'legend.fontsize': 6,
+    'legend.title_fontsize': 8,
     'xtick.labelsize': 6,
     'ytick.labelsize': 6,
     'xtick.major.size': 2,

### 2022-12-21 paper-1/manuscript.tex
Rev Fig1

@@
 
-When a neuron's membrane potential reached the firing threshold $v_\theta$, a spike was emitted, and the potential was clamped to $V_m = V_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons).
-Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $v_\theta = \theta_0 = -54$~mV \citep{Mensi2012-au}.
+When a neuron's membrane potential reached the firing threshold $V_\theta$, a spike was emitted, and the potential was clamped to $V_m = V_{reset} = -74$~mV for a fixed refractory period of 3~ms (excitatory neurons).
+Inhibitory neurons were modeled as fast-spiking cells with a refractory period of 2~ms and a constant firing threshold $V_\theta = \theta_0 = -54$~mV \citep{Mensi2012-au}.
 In contrast, excitatory neurons were modeled with threshold adaptation (TA) as in \cite{Teeter2018-iz}, following
 \begin{align}
-    v_\theta &= \theta_0 + V_{TA}(t) \nonumber \\
+    V_\theta &= \theta_0 + V_{TA}(t) \nonumber \\
     \tau_{\theta} \frac{dV_{TA}}{dt} &= -V_{TA} + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
@@
 \textbf{Fig.
-1}Model and paradigm.
-\textbf{a} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron, illustrated above the plots, firing pre-determined spikes.
+1} Model and paradigm.
+\textbf{a} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes, illustrated above the plots.
 The excitatory connection has weight $w = 3$ for demonstration purposes only; a single connection with $w = 1$ is normally unable to evoke postsynaptic firing.
-Top: Membrane voltage (black), firing threshold baseline (gray) and adaptive threshold (green) demonstrating TA.
-Middle: Excitatory ($g_e$, blue) and inhibitory ($g_i$, red) synaptic conductances, demonstrating excitatory STD.
-Bottom: Inputs to $g_e$ and $g_i$ (vertical bars in the appropriate colors), and the depression variable $x$ of the excitatory presynaptic neuron (dashed line).
-\textbf{b} Sample network layout, showing excitatory and inhibitory neurons and all outgoing synaptic connections from two neurons of each type.
-\textbf{c} Distance from stimulation site 0 in terms of minimum number of excitatory synapses.
-Stimulation is provided to the 10 neurons nearest to the center of each of 5 stimulation sites (round markers), which are regularly spaced on a circle with a radius of 2.5 mm.
+Top: Inputs to the (post-)synaptic conductances $g_e$ and $g_i$ (vertical bars), and the STD depression variable $x_{Exc}$ of the excitatory presynaptic neuron (dashed line).
+Note that excitatory inputs are equal to $U x_{Exc} w$ as per Equation \ref{eqn-gsyn}.
+Middle: Excitatory and inhibitory synaptic conductances ($g_e$ and $g_i$, respectively).
+Bottom: Membrane voltage, firing threshold baseline (gray) and adaptive threshold, demonstrating TA in the excitatory postsynaptic neuron.
+\textbf{b} Sample network layout, showing the spatial location of excitatory and inhibitory neurons, and all outgoing synaptic connections from two neurons of each type.
+\textbf{c} Distance from stimulation site 1 in terms of minimum number of excitatory synapses.
+Stimulation is provided to the 10 neurons nearest to the center of each of 5 stimulation sites (large circles), which are regularly spaced on a circle with a radius of 2.5 mm.
+\textbf{d} Illustration of the stimulation paradigm.
+Each marker represents a stimulation, with the stimulus identity indicated by color.
+The target stimulus (A) is represented as solid red, and non-target stimuli are represented with faded colors.
+Three randomized sequences were presented, containing 80\% A and 20\% B (std), 20\% A and 80\% B (dev), and 20\% of each of 5 stimuli, including A and B (msc).
+Note that, while the onset asynchrony of 500 ms is represented faithfully, the horizontal extent of the stimulation markers has no meaning; stimuli were presented instantaneously rather than over an extended period.
 

### 2022-12-22 paper-1/manuscript.tex
Rev Fig2-3, minus Fig2 text

@@
 \textbf{Fig.
-2} Deviance detection in vitro and in silico.
-\textbf{a} Response of a 2 $\times$ 4~mm neuronal cell culture on a multi-electrode array, adapted from \cite{Kubota2021-dx}.
-Curves represent the average number of spikes per electrode in each 10~ms bin in response to target stimulation in the indicated sequence.
-Unlike in our study, oddball sequences here were split 10\%/90\%, and 10 equally frequent stimuli were used for the msc sequence.
-\textbf{b} Sample model response, showing average network-wide spike counts after stimulation with target (A) in the indicated sequence (same color code as panel \textbf{a}).
-Both in vitro and in the model, the dev response is very large, followed by an intermediate msc response and a low std response.
-\textbf{c} Mismatch, SSA and deviance detection indices (see Equation \ref{eqn-ddi}) across networks and stimuli (n = 120).
+2} Deviance detection in our model.
+\textbf{a} Sample network target response intensity by trial type.
+Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step.
+Target trials are taken from the std (400), msc (100), and dev (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section).
+\textbf{b} Response summary, showing average network-wide spike counts in target trials in each sequence.
+The underlying data are the same as in panel a.
+Note how the dev response is very large, followed by an intermediate msc response and a low std response.
+\textbf{c} Mismatch, adaptation, and deviance detection indices (see Equation \ref{eqn-ddi}) across networks and stimuli (n = 120).
 In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points
@@
 
-Our first aim was to replicate deviance detection properties reported in dissociated neuronal cell culture by \cite{Kubota2021-dx}, reproduced in Figure \ref{fig2}a.
+% Our first aim was to replicate deviance detection properties reported in dissociated neuronal cell culture by \cite{Kubota2021-dx}, reproduced in Figure \ref{fig2}a.
 In their case, cell cultures -- like our model, randomly connected networks with two-dimensional extent -- were stimulated at 10 different sites in msc sequences, and at two sites with 10\%/90\% probability in dev and std sequences.
@@
 
-In our model, we could readily identify networks and stimuli whose responses developed in a qualitatively similar fashion, with high activity in response to target in the dev sequence, low activity in the std sequence, and intermediate activity in the msc sequence, as illustrated with an example in Figure \ref{fig2}b.
+% In our model, we could readily identify networks and stimuli whose responses developed in a qualitatively similar fashion, with high activity in response to target in the dev sequence, low activity in the std sequence, and intermediate activity in the msc sequence, as illustrated with an example in Figure \ref{fig2}b.
 The model responses developed on a compressed time scale compared to neuronal culture, which we tentatively attribute to unfaithful aspects of the model, including a lack of conduction delay or spontaneous activity, a much smaller number of participating neurons, and structural differences.
 
-Across networks and stimuli (Figure \ref{fig2}c), we found that our model reliably produced positive mismatch responses (t = 11.3, p = 6.7e-21; one-sided t-test, n = 120) while both SSA index (t = 7.84, p = 1.05e-12) and DDI (t = 6.49, p = 1.05e-09) were also positive.
+% Across networks and stimuli (Figure \ref{fig2}c), we found that our model reliably produced positive mismatch responses (T = 6.67e+03, p = 9.26e-16; one-sided Wilcoxon signed-rank test, n = 120) while both SSA index (T = 6.19e+03, p = 9.76e-12) and DDI (T = 5.97e+03, p = 4.75e-10) were also positive.
 The average DDI value reached around 0.1, which was less than the values reported by \cite{Kubota2021-dx}, but note that our target stimuli were presented more frequently (20\%) than theirs (10\%) in the dev and msc sequences.
@@
 \textbf{a} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
-The color code is the same as in Figure \ref{fig2}.
-\textbf{b} SSA indices across all networks and stimuli.
-Asterisks indicate data greater than 0, established by t-test with significance level 0.05, see main text for detailed statistics.
-\textbf{c} Deviance detection indices across all networks and stimuli.
+Lines show the average (trial mean +- SEM, left axis) network-wide spike counts in target trials in each sequence.
+Boxes summarize the spike count totals per trial (right axis), the means of which form the basis for index calculations (see panel c).
+Note the different axis scaling on the top row (without TA) and the bottom row (with TA).
+\textbf{b} Adaptation indices $<$msc - std$>$ across all networks and stimuli.
+Asterisks indicate data greater than 0, established with a Wilcoxon signed-rank test at a significance level 0.05.
+See main text for detailed statistics.
+\textbf{c} Deviance detection indices $<$dev - msc$>$ across all networks and stimuli
 
@@
 
-The SSA index (Figure \ref{fig3}b) was exactly 0 as expected in the ablation with no plasticity, and positive with any plasticity applied (STD only: t = 3.17, p = 0.000983, median 0.0039; TA only: t = 6.74, p = 3e-10, median 0.097; STD+TA: t = 7.84, p = 1.05e-12, median 0.135; one-sided t-tests, n = 120).
+The SSA index (Figure \ref{fig3}b) was exactly 0 as expected in the ablation with no plasticity, and positive with any plasticity applied (STD only: T = 4.97e+03, p = 0.000222, median 0.0039; TA only: T = 5.99e+03, p = 3.3e-10, median 0.097; STD+TA: T = 6.19e+03, p = 9.76e-12, median 0.135; one-sided Wilcoxon signed-rank tests, n = 120).
 Since the SSA index measures the degree of adaptation to frequent stimulus presentation, this indicates that both plasticity mechanisms were capable of imbuing the networks with some level of history dependence.
 
-Finally, across networks and stimuli, the DDI (Figure \ref{fig3}c) was exactly 0 with no plasticity, not significantly different from 0 with STD only (t = -1.89, p = 0.0609, median -2.1e-04; two-sided t-test, n = 120), but clearly positive for TA only (t = 5.52, p = 1.01e-07, median 0.077; one-sided t-test, n = 120) and the full model (t = 6.49, p = 1.05e-09, median 0.114).
-To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: t = 6.7, p = 4.53e-10; TA: t = 3.2, p = 0.00103), and greater too than the sum of the DDI values across both STD only and TA only models (t = 3.6, p = 0.000199).
+Finally, across networks and stimuli, the DDI (Figure \ref{fig3}c) was exactly 0 with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
+To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: T = 6.1e+03, p = 4.95e-11; TA: T = 4.87e+03, p = 0.000604), and greater too than the sum of the DDI values across both STD only and TA only models (T = 5.12e+03, p = 4.52e-05).
 This suggests that TA and STD act not as independent mechanisms in this paradigm, but rather interact constructively to enhance the network's ability to encode input regularities.

### 2022-12-22 paper-1/manuscript.tex
Rev Fig4

@@
 Left and middle column, target trials in dev sequence and msc sequence, respectively; right column, contrast between these two.
-\textbf{b} Target trial average of the threshold adaptation voltage $V_{TA}$, with axes, neuron order and columns as in panel a.
-\textbf{c} Target trial average of the membrane potential $V_m$, with axes, neuron order and columns as in panel a.
-The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the color mask.
+\textbf{b} Target trial average of the threshold adaptation voltage $V_{TA}$, with neuron order and columns as in panel a.
+\textbf{c} Target trial average of the membrane potential $V_m$, with neuron order and columns as in panel a.
+The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the colormap.
 \textbf{d} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
@@
 We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response of each excitatory neuron in terms of spikes fired.
-In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (t = -17.3, p = 2.57e-57, median -0.166 mV; one-sided t-test, n = 800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided t-test, n = 800).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Average $V_{TA}$ was lower in dev than msc (t = -7.32, p = 1.58e-11, median = -0.074 mV; one-sided t-test, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (t = -23.1, p = 3.72e-46, median = -0.357).
+In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Average $V_{TA}$ was lower in dev than msc (T = 1.24e+03, p = 2.07e-10, median = -0.074 mV; one-sided Wilcoxon, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
 This indicates that networks were less adapted during oddball sequences, and that this lack of adaptation played an important role in the increased response to deviant inputs.

### 2022-12-26 plot_utils.py
Improved inset_hist versatility

@@ -113,16 +113,20 @@ def fill_ratios(*ratios, to=100):
     ratios[ratios<0] = remainder / (ratios<0).sum()
     return ratios
 
-def inset_hist(ax, data, x=True, median_color='C1'):
+def inset_hist(ax, data, x=True, median_color='C1', rescale=True, **kwargs):
     if x:
         y = ax.get_ylim()
-        ax.set_ylim(top=y[1] + .1*(y[1]-y[0]))
+        if rescale:
+            ax.set_ylim(top=y[1] + .1*(y[1]-y[0]))
         tx = ax.twinx()
     else:
         y = ax.get_xlim()
-        ax.set_xlim(right=y[1] + .1*(y[1]-y[0]))
+        if rescale:
+            ax.set_xlim(right=y[1] + .1*(y[1]-y[0]))
         tx = ax.twiny()
-    binned, bins, *_ = tx.hist(data, 20, color='grey', histtype='stepfilled', edgecolor='dimgrey', orientation='vertical' if x else 'horizontal')
+    hargs = dict(color='grey', histtype='stepfilled', edgecolor='dimgrey', orientation='vertical' if x else 'horizontal')
+    hargs.update(**kwargs)
+    binned, bins, *_ = tx.hist(data, 20, **hargs)
     ci = stats.bootstrap([data], np.median, n_resamples=10000)
     lo, hi = ci.confidence_interval
     bsize = np.diff(bins)[0]

### 2022-12-26 paper-1/manuscript.tex
Added TODOs and discussion notes
Rev F6 caption

@@
 \end{figure*}
-\textbf{Fig. 6} Dev non-target responses are reduced due to adaptation, which is caused by greater activity in the early responders to B.
-\textbf{a} Non-target response contrasts in the sample network: The replacement contrast (top right) compares the response to B in msc (top left) to all non-target responses, including B, in msc (top center).
-The adaptation contrast (bottom right) compares the response to B in dev (bottom left) to the response to B in msc (bottom center).
-Both contrasts together add up to the non-target contrast (dev - msc) shown in Figure \ref{fig5}b.
-Note that the contrast column (right) shares a color scale to allow direct comparison.
-\textbf{b} Network-wide mean replacement and adaptation contrasts across networks and stimuli.
-\textbf{c} Relationship across networks and stimuli between the replacement contrast and the contrast in $V_{TA}$ at the start of A trials, and associated linear regression.
-For both measures, only the first 100 neurons to respond to stimulation by B in any condition were included.
-Each marker corresponds to one network and target stimulus.
+% TODO 5-8: Re-label stim locations once per figure
+\textbf{Fig. 6} Dev non-target responses are reduced due to adaptation in early responders to B.
+\textbf{a} The sample network, colored by latency rank, i.e.
+the rank order of the time to first spike, in response to stimulation at site B in any condition.
+The response develops asymmetrically due to structural properties, not due to interactions with A.
+Note that non-responding neurons are ranked last and appear dark blue.
+\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (dev - msc) and the contrast in $V_{TA}$ at the start of B trials.
+Neurons are colored by latency rank, in the sense that the earliest 100 responders are shown as red, and the later neurons as blue.
+Inset histograms show the distributions and estimated median of the corresponding values across these two categories.
+\textbf{c} Mean non-target response contrast (dev - msc) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labelled "late", across networks and stimuli.
+\textbf{c} Mean contrast (dev - msc) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli
 
+% TODO rev Fig6 text
 Finally, we examined the cause of the reduced response in non-target trials.
@@
 To show this, we first identified all neurons located within a single excitatory synapse of the stimulated neurons (cf.
-Figure \ref{1}c), and calculated the average replacement effect for these alone.
+Figure \ref{fig1}c), and calculated the average replacement effect for these alone.
 In this subpopulation, replacement yielded a clear response increase, shown in the central box of Figure \ref{fig6}b (t = 8.21, p = 1.5e-13; one-sided t-test).
@@
 Deviant detection then arises as a relief from this blanket suppression, or as a result of adaptation shifting from the entire network to a small region around the frequently presented standard.
+% END TODO
 
@@
 \end{figure*}
+% TODO 7: Add F6 conclusion to complete the story
+% Consider alternative ways of presenting the left side without using so much space/mental effort (stats?)
+% a-c: Add (dev-msc) to yaxis label
+% d,e: Mark significance
 \textbf{Fig. 7} Adding STD to the model enhances the direct effect of TA.
@@
 
+\begin{figure*}%
+    \centering
+    \includegraphics{fig9.pdf}
+    \caption{}
+    \label{fig9}
+\end{figure*}
+% TODO 9: Needs to be more generic, not specific to the task. Show the process, step by step.
+
+% No tuning required to get DDI!
+
+% Culture, rats, humans...
+
+% Response of a 2 $\times$ 4~mm neuronal cell culture on a multi-electrode array, adapted from \cite{Kubota2021-dx}.
+Curves represent the average number of spikes per electrode in each 10~ms bin in response to target stimulation in the indicated sequence.
+Unlike in our study, oddball sequences here were split 10\%/90\%, and 10 equally frequent stimuli were used for the msc sequence.
+
+% adaptation vs deviance detection, in the context of pred coding, means DD is derived from some pred/err interaction, cf Fong+20
+% deviant-alone always highest, cf. Taaseh2011-gg
+
+% Replacement vs adaptation is not a clean separation, since $V_{TA}$ is affected by replacement, but modulates adaptation
+
+% The prime effect of TA (reduce std by local suppression, relieve global suppression for a larger dev) is repeated with STD, in exactly the same way, just weaker.
+% What makes the two synergise is precisely this symmetry: STD reduces std by local suppression, which relieves not only STD-related global suppression, but also TA-related global suppression.
+And, of course, vice versa!
+% On the other hand, theoretical expectations lead us to expect competition to occur in the area of local suppression: Such suppression has diminishing returns; in the extreme case, if one mechanism reduces the response to zero, the other mechanism is left without room for further reductions.
+Or: If each mechanism independently cuts the response in half, then we should expect a quarter of the response to remain
+
+% Spatial overlap
+% Structure
+% Omission
+% Time (ISI <> tau)
+% Neurobiology
+% long-term plasticity
+% Spontaneous activity
+% Prediction signals that cancel expected inputs would reduce adaptation
+% Testable predictions. Eg: Block NMDA, spontaneous activity, TA-related channels; easier: early vs late subpopulations
+
 \backmatter

### 2022-12-27 paper-1/manuscript.tex
Rev sec Results:F6

@@
 
-Correlating the non-target response and thresholds excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev (t = -7.61, p = 3.5e-12, median = -0.044; one-sided t-test, n = 800) almost perfectly predicted the resulting reduction in threshold adaptation levels (Pearson's $\rho$ = 0.99, p = 0).
+Correlating the non-target response and thresholds excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev (t = -7.61, p = 3.5e-12, median = -0.044; one-sided t-test, n = 800) almost perfectly predicted the resulting reduction in threshold adaptation levels (Pearson's $\rho$ = 0.99).
 This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.994 and a consistent reduction in the average non-target response between msc and dev sequences (t = -11.6, p = 1.46e-21, median = -0.058; one-sided t-test, n = 120), clearly validating our notion that the non-target trial response is the primary driver of threshold adaptation.
@@
 \textbf{a} The sample network, colored by latency rank, i.e.
-the rank order of the time to first spike, in response to stimulation at site B in any condition.
+the rank order of the time to first spike in response to stimulation at site B in any condition.
 The response develops asymmetrically due to structural properties, not due to interactions with A.
 Note that non-responding neurons are ranked last and appear dark blue.
-\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (dev - msc) and the contrast in $V_{TA}$ at the start of B trials.
-Neurons are colored by latency rank, in the sense that the earliest 100 responders are shown as red, and the later neurons as blue.
-Inset histograms show the distributions and estimated median of the corresponding values across these two categories.
-\textbf{c} Mean non-target response contrast (dev - msc) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labelled "late", across networks and stimuli.
-\textbf{c} Mean contrast (dev - msc) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli
+\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (dev - msc) and the contrast in $V_{TA}$ at the start of B trials, and associated linear regression.
+Neurons are colored by latency rank as in panel a.
+Additionally, the earliest 100 responders are plotted with crosses, while later responders are plotted dots.
+Inset histograms show the distributions of the corresponding values across these two categories, and the confidence intervals of their medians, calculated by bootstrap with 10000 iterations.
+\textbf{c} Median non-target response contrast (dev - msc) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labelled "late", across networks and stimuli.
+\textbf{c} Median contrast (dev - msc) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli
 
-% TODO rev Fig6 text
 Finally, we examined the cause of the reduced response in non-target trials.
 Going from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation.
-Secondly, the more frequent presentation of B changes adaptation, likely reducing the average response to B.
-In order to separate these two, we split the non-target contrast shown in Figure \ref{fig5}b into two components: a replacement effect, contrasting the response to B in the msc sequence with the response to all non-target stimuli in the same sequence; and an adaptation effect, contrasting the response to B in dev with the response to B in msc.
-For the sample network, these contrasts are shown in Figure \ref{fig6}a.
-Unsurprisingly, replacement yielded a decreased response near the stimulation sites for C, D and E, and an increased response in areas near the stimulation site for B.
-By contrast, adaptation caused a decreased response in virtually the entire network, particularly in the vicinity of the stimulation site for B.
+Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and therefore decreasing the average response.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (a) responds very quickly to stimulation in B, (b) responds more strongly to non-target stimuli in dev than msc, and (c) has increased TA levels as a result, reducing the later response in the same way as demonstrated in Figure \ref{fig4} in reverse.
 
-Across networks and stimuli (Figure \ref{fig6}b), we saw a wide range of replacement effects, including both increased and decreased firing, likely as a result of differing sensitivities of the networks to the various non-target stimuli.
-On average, stimulus replacement did not have a significant effect (t = -0.255, p = 0.799; two-sided t-test).
-In contrast, adaptation yielded a reduced response on average (t = -9.01, p = 2.02e-15; one-sided t-test), indicating that the more frequent presentation of B indeed lowered the network's response to this stimulus.
+To see this in the sample network, consider first the latency of responses to B across all conditions, shown by way of rank order in Figure \ref{fig6}a.
+We found that the response developed in an approximately radial pattern, spreading outward from the stimulus site.
+We did notice that the early response appears to gravitate towards the stimulus site for A, but confirmed that this was the case in all conditions separately, indicating a structural rather than dynamic cause (data not shown).
+We then notionally split the network into an early portion, containing the first 100 neurons to respond to B in any condition, and a late portion, containing the remaining 900 neurons.
+In Figure \ref{fig6}b, we related the response contrast (dev - msc) to the contrast in $V_{TA}$ at the start of B trials.
+Across all neurons, a strong linear relationship (Pearson's $\rho$ = 0.98) echoed the related finding in Figure \ref{fig5}c.
+As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the dev sequence, while the network as a whole decreased.
+Correspondingly, the TA levels of this portion of the network also tended to increase, in clear contrast to the network as a whole.
 
-Finally, given the clear spatial pattern exhibited by the replacement effect in the sample network, we surmised that adaptation is driven mainly by neurons close to the stimulation site B responding more frequently due to the more frequent presentation of the stimulus.
-To show this, we first identified all neurons located within a single excitatory synapse of the stimulated neurons (cf.
-Figure \ref{fig1}c), and calculated the average replacement effect for these alone.
-In this subpopulation, replacement yielded a clear response increase, shown in the central box of Figure \ref{fig6}b (t = 8.21, p = 1.5e-13; one-sided t-test).
-As shown in Figure \ref{fig6}c, this response increase directly translated into elevated thresholds in dev trials compared to msc (t = 5.01, p = 9.59e-07).
+Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B.
+We first confirmed that the non-target response and $V_{TA}$ contrasts were robustly correlated (Median Pearson's $\rho$ 0.996, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon test for $\rho > 0$, n = 120).
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $V_{TA}$ at the start of B trials (Figure \ref{fig6}d) within these two portions.
+We found that the response in the early portion was consistently greater in dev than msc (T = 6.98e+03, p = 8.48e-19; one-sided Wilcoxon, n = 120), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
+The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in dev (T = 4.63e+03, p = 2.21e-17), while the late portion was more recovered (T = 26, p = 2.97e-17).
 
-Earlier, we showed that the early response is strongly influenced by TA (Figure \ref{fig4}h).
-Here, we showed that $V_{TA}$ near the stimulation site was increased in dev as a result of the more frequent presentation of B.
-We posit that this increase in TA affects the early response, reducing the response further downstream and freeing up the rest of the network to respond more strongly to the infrequent target stimulus.
+To sum up, we showed in Figure \ref{fig4} that greater $V_{TA}$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response.
+Here, we showed that the neurons responding earliest to B exhibit higher $V_{TA}$ in the dev sequence as a consequence of an increased response, in direct contrast to the response and TA levels of the network as a whole.
 
-To conclude, TA plays two complementary roles in deviance detection, affecting the standard and deviant response separately: In its first role, as we have shown with the adaptation contrast above, TA reduces the response to frequently presented stimuli regardless of context, leading to a stronger response to infrequently presented stimuli (SSA).
-Since this happens without the need for large-scale adaptation -- only a small fraction of the network near the stimulation site adapts strongly to frequently presented inputs -- the network as a whole is allowed to relax in the oddball condition.
-In its second role, TA causes network-wide adaptation in the control condition, where many different stimuli are presented, all of which are infrequently presented and therefore able to evoke large responses.
-Deviant detection then arises as a relief from this blanket suppression, or as a result of adaptation shifting from the entire network to a small region around the frequently presented standard.
-% END TODO
+Thus, we have finally traced the greater response to A in the dev sequence to its roots, and revealed two complementary roles of TA: The frequent presentation of non-target stimulus B in the dev sequence and concomitant increased activity in a subset of neurons responding early to B causes \emph{local adaptation} by TA in this subset, leading to smaller network responses.
+By contrast, stimuli in the msc sequence are presented more sporadically, allowing local adaptation to decay, thus leading to larger network responses.
+This difference in non-target response sizes -- small in dev, large in msc -- is then reflected in network-wide \emph{global adaptation}, such that the network is more excitable to stimulation with A (or other non-standard inputs) in the dev sequence.
 

### 2022-12-27 paper-1/manuscript.tex
Rev sec Results:F4 in light of better analysis

@@
 While thresholds were slighly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (Figure \ref{fig4}g), $V_{TA}$ constituted a major driver of increased firing only right at the onset of the response, with later contributions amounting to less than 20\% of the total voltage difference.
-This pattern was consistent across networks and stimuli, as shown in Figure \ref{fig4}h.
+In relative terms (Figure \ref{fig4}g), $V_{TA}$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+A similar pattern emerged across networks and stimuli, as shown in Figure \ref{fig4}h.
+Here, only the initial response was driven mainly by $V_{TA}$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
 This indicates that adaptation initiated and dominated the increase in firing in the beginning, whereas the late response was enlarged primarily as a result of increased presynaptic excitation.

### 2022-12-27 paper-1/manuscript.tex
s/mean/median/ for F4,5,7 stats

@@
 \textbf{d} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
-\textbf{e} Network-wide mean contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta V_{TA}$ vs response size contrast across excitatory neurons).
+\textbf{e} Network-wide median contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta V_{TA}$ vs response size contrast across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 \textbf{c} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation and the contrast in $V_{TA}$ at the start of target trials, and associated linear regression.
-\textbf{d} Network-wide mean contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\Delta V_{TA}$ across excitatory neurons).
+\textbf{d} Network-wide median contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\Delta V_{TA}$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 Neurons are colored by latency rank as in panel a.
-Additionally, the earliest 100 responders are plotted with crosses, while later responders are plotted dots.
+Additionally, the earliest 100 responders are plotted with crosses, while later responders are plotted with dots.
 Inset histograms show the distributions of the corresponding values across these two categories, and the confidence intervals of their medians, calculated by bootstrap with 10000 iterations.

### 2022-12-27 paper-1/manuscript.tex
Rev F7 legend

@@
 \end{figure*}
-% TODO 7: Add F6 conclusion to complete the story
-% Consider alternative ways of presenting the left side without using so much space/mental effort (stats?)
-% a-c: Add (dev-msc) to yaxis label
-% d,e: Mark significance
-\textbf{Fig. 7} Adding STD to the model enhances the direct effect of TA.
-\textbf{a} Contrast (dev - msc) of the response to A, in terms of mean number of spikes per trial, in the sample network, with each neuron shown in its spatial location.
-Left and middle column, full model with STD and TA, and ablated model with only TA, respectively.
-Right column, subtraction of these two.
-Stimulus locations are as shown in Figure \ref{fig5}a.
-\textbf{b} Contrast of mean $V_{TA}$ at the start of target trials, in columns as in panel a.
-\textbf{c} Contrast of the response to non-target stimuli, in columns as in panel a.
-\textbf{d} Network-wide mean contrast of the response to target (left) and non-target stimuli (right) across networks and stimuli, arranged equivalently to the columns in panels a-c.
-Blue and green boxes represent the full and ablated model, respectively, and white boxes the difference between these two.
-\textbf{e} Network-wide mean contrast of $V_{TA}$ before A trials across networks and stimuli, arranged as in panel d
+\textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
+\textbf{a} Median contrast (dev - msc) in the non-target response (top) and in $V_{TA}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+Compare to Figure \ref{fig6}c-d.
+\textbf{b} Top: Median contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the linear correlation between non-target response contrast and $\Delta V_{TA}$ at the start of target trials across excitatory neurons.
+Each marker corresponds to one network and target stimulus.
+Data from the full model with STD and TA, compare Figure \ref{fig5}d.
+Middle: Non-target response contrast in the full model, plotted against the same contrast in the TA-only model.
+The red line is the identity function; points above this line represent networks with greater response reduction (dev - msc) in the full model than in the ablated model.
+Bottom: Subtraction of the two response contrasts.
+\textbf{c} Top: Median contrast (dev - msc) in $V_{TA}$ at the start of target trials across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the linear correlation between $\Delta V_{TA}$ and the target response contrast across excitatory neurons.
+Each marker corresponds to one network and target stimulus.
+Data from the full model with STD and TA, compare Figure \ref{fig4}e.
+Middle: $\Delta V_{TA}$ in the full model, plotted against the same contrast in the TA-only model.
+The red line is the identity function; points above this line represent networks with greater reduction (dev - msc) in $V_{TA}$ in the full model than in the ablated model.
+Bottom: Subtraction of the two $V_{TA}$ contrasts.
 

### 2022-12-27 paper-1/manuscript.tex
Editing / mostly stats updates

@@
 Each marker corresponds to one network and target stimulus.
-Red markers indicate a significant correlation (two-sided t-test, significance level 0.05).
+Red markers indicate a significant negative correlation (two-sided Wald test, significance level 0.05).
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
@@
 In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Average $V_{TA}$ was lower in dev than msc (T = 1.24e+03, p = 2.07e-10, median = -0.074 mV; one-sided Wilcoxon, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Median $V_{TA}$ was lower in dev than msc (T = 514, p = 1.67e-16, median = -0.094 mV; one-sided Wilcoxon, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
 This indicates that networks were less adapted during oddball sequences, and that this lack of adaptation played an important role in the increased response to deviant inputs.
@@
 
-Correlating the non-target response and thresholds excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev (t = -7.61, p = 3.5e-12, median = -0.044; one-sided t-test, n = 800) almost perfectly predicted the resulting reduction in threshold adaptation levels (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.994 and a consistent reduction in the average non-target response between msc and dev sequences (t = -11.6, p = 1.46e-21, median = -0.058; one-sided t-test, n = 120), clearly validating our notion that the non-target trial response is the primary driver of threshold adaptation.
+Correlating the non-target response contrast (dev - msc) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev almost perfectly predicted the resulting reduction in threshold adaptation levels (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between msc and dev sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the non-target trial response is the primary driver of threshold adaptation.
 
@@
 Each marker corresponds to one network and target stimulus.
-Red markers indicate a significant positive correlation (one-sided t-test, significance level 0.05).
+Red markers indicate a significant positive correlation (one-sided Wald test, significance level 0.05).
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations
@@
 Going from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation.
-Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and therefore decreasing the average response.
+Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
 We hypothesized that this adaptation is mediated by a small subset of neurons that (a) responds very quickly to stimulation in B, (b) responds more strongly to non-target stimuli in dev than msc, and (c) has increased TA levels as a result, reducing the later response in the same way as demonstrated in Figure \ref{fig4} in reverse.
@@
 In Figure \ref{fig6}b, we related the response contrast (dev - msc) to the contrast in $V_{TA}$ at the start of B trials.
-Across all neurons, a strong linear relationship (Pearson's $\rho$ = 0.98) echoed the related finding in Figure \ref{fig5}c.
-As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the dev sequence, while the network as a whole decreased.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993) echoed the related finding in Figure \ref{fig5}c.
+As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the dev sequence, while the response of the network as a whole decreased.
 Correspondingly, the TA levels of this portion of the network also tended to increase, in clear contrast to the network as a whole.
@@
 Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B.
-We first confirmed that the non-target response and $V_{TA}$ contrasts were robustly correlated (Median Pearson's $\rho$ 0.996, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon test for $\rho > 0$, n = 120).
+We first confirmed that the non-target response and $V_{TA}$ contrasts were robustly correlated (Median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon test for $\rho > 0$, n = 120).
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $V_{TA}$ at the start of B trials (Figure \ref{fig6}d) within these two portions.
@@
 
+% TODO rev Fig7
 As noted previously, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA.
@@
 To sum up, the reason for the larger target dev response relative to msc, was the smaller non-target dev response relative to msc; furthermore, since the relationships established in the previous section remained largely unchanged, this effect appeared to be mediated primarily through TA.
+% end TODO
 

### 2022-12-28 paper-1/manuscript.tex
Rev sec Results:F7

@@
 
-% TODO rev Fig7
-As noted previously, short-term depression alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA.
-To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks with both TA and STD.
+As noted previously, STD alone exerted no deviance detection effect, but its addition did enhance the deviance detection effect established by TA.
+To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD.
 On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (data not shown).
@@
 
-To understand the reason for this asymmetry, we first focused on the target response and asked how and why the target response contrast (dev - msc) changed as STD was added to the model.
-As shown in Figure \ref{fig7}a, the target response was greater in dev than msc in both the full and the ablated model (indicating a positive DDI in both models, cf.
-also Figure \ref{fig3}).
-In most neurons, the contrast was greater in the full model, such that the difference (full - ablated) was positive, as indicated by the largely orange plot in the right column.
-Following the logic established in the previous section, we may expect a greater target response in dev to be caused by lower $V_{TA}$ relative to msc.
-Indeed, comparing the contrast (dev - msc) in $V_{TA}$ before target trials across models (Figure \ref{fig7}b), we see that, while the contrast is negative (i.e., lower in dev) in both models, the magnitude of the contrast is greater in the full model.
-For clarity, the difference is therefore negative, as indicated by the largely blue plot in the right column.
-Lastly, we trace this one step further back in Figure \ref{fig7}c: Again as established in the previous section, a greater contrast in $V_{TA}$ may be explained by a similarly enhanced contrast (dev - msc) in the non-target response.
-We found exactly this: The non-target response was smaller in dev than msc by a greater margin in the full model, as indicated by the largely blue plot in the right column.
+To understand how STD affects deviance detection, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
+Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}a the full model equivalent of Figure \ref{fig6}c and d.
+The contrasts (dev - msc) in the non-target response magnitude and in the resulting $V_{TA}$ levels at the start of B trials both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the dev sequence.
+Likewise, the correlation across excitatory neurons between non-target response contrast and the contrast in $V_{TA}$ before B trials remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992) This indicates that local adaptation to B mediated by TA remained a key driver of the reduced response to non-target trials.
 
-This pattern was reflected also in the aggregate across networks and stimuli.
-Firstly, the contrast (dev - msc) in the target response (Figure \ref{fig7}d, left) was positive in both models, but the full model contrast was greater than the ablated model contrast (t = 2.99, p = 0.00168; one-sided t-test, n = 120).
-Referring to the previous section, we identify the cause for this in an enhanced contrast in $V_{TA}$, as shown in Figure \ref{fig7}e; the contrast (dev - msc) in the full model was more negative than that in the ablated model (t = -3.49, p = 0.000342).
-Lastly, returning to Figure \ref{fig7}d (right), we see that the driver of $V_{TA}$, namely the non-target response size, also showed an enhanced contrast (dev - msc) in the full model relative to the ablation (t = -3.19, p = 0.000921).
-To sum up, the reason for the larger target dev response relative to msc, was the smaller non-target dev response relative to msc; furthermore, since the relationships established in the previous section remained largely unchanged, this effect appeared to be mediated primarily through TA.
-% end TODO
+Inspecting the non-target response reduction in the full model in Figure \ref{fig7}b, we found that it was on average larger than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), if only by a small amount on the order of 0.03 spikes per neuron.
+Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
+Indeed, the non-target response contrast remained tightly correlated with the contrast (dev - msc) in $V_{TA}$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
+Correspondingly, $V_{TA}$ was reduced in dev, as shown in Figure \ref{fig7}d, and this too by a slighly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+
+Finally, based on the ablated model analysis in Figure \ref{fig4}, we sought to confirm that lowered $V_{TA}$ before target trials was responsible for the increased target response.
+As shown in Figure \ref{fig7}e, the contrasts (dev - msc) in $V_{TA}$ at the start of target trials and in response magnitude to A were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response was clearly increased in dev over msc, and this increase was greater than in the ablated model (T = 1.94e+03, p = 0.000723), as shown in Figure \ref{fig7}f.
+This greater contrast, of course, directly reflects the greater DDI reported in Figure \ref{fig3}.
+
+Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local adaptation to B resulting in reduced B responses and reduced global adaptation, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
+The primary cause for the greater target response, therefore, appeared to be the smaller non-target response, which was where we noted the first clear shift towards increased deviance detection.
 
@@
 \textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
-\textbf{a} Median contrast (dev - msc) in the non-target response (top) and in $V_{TA}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+\textbf{a} Median contrast (dev - msc) in the non-target response (left) and in $V_{TA}$ at the start of B trials (right) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
 Compare to Figure \ref{fig6}c-d.
-\textbf{b} Top: Median contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the linear correlation between non-target response contrast and $\Delta V_{TA}$ at the start of target trials across excitatory neurons.
-Each marker corresponds to one network and target stimulus.
-Data from the full model with STD and TA, compare Figure \ref{fig5}d.
-Middle: Non-target response contrast in the full model, plotted against the same contrast in the TA-only model.
+\textbf{b} Median contrast (dev - msc) in the non-target response magnitude across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
 The red line is the identity function; points above this line represent networks with greater response reduction (dev - msc) in the full model than in the ablated model.
-Bottom: Subtraction of the two response contrasts.
-\textbf{c} Top: Median contrast (dev - msc) in $V_{TA}$ at the start of target trials across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the linear correlation between $\Delta V_{TA}$ and the target response contrast across excitatory neurons.
-Each marker corresponds to one network and target stimulus.
-Data from the full model with STD and TA, compare Figure \ref{fig4}e.
-Middle: $\Delta V_{TA}$ in the full model, plotted against the same contrast in the TA-only model.
-The red line is the identity function; points above this line represent networks with greater reduction (dev - msc) in $V_{TA}$ in the full model than in the ablated model.
-Bottom: Subtraction of the two $V_{TA}$ contrasts.
+The subtraction of the two quantities in the box plot below shows that the non-target response tends to be reduced more in the full model.
+\textbf{c} Non-target response contrast, plotted against its Pearson correlation coefficient with the corresponding contrast in $V_{TA}$ at the start of target trials, in the full model.
+Compare to Figure \ref{Fig5}d.
+\textbf{d} Median contrast (dev - msc) in $V_{TA}$ at the start of target trials across networks and stimuli in the full and ablated models, arranged as in panel b.
+The subtraction of the two quantities in the box plot below shows that $V_{TA}$ tends to be reduced more in the full model.
+\textbf{e} $\Delta V_{TA}$ at the start of target trials, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response magnitude, in the full model.
+Compare to Figure \ref{Fig4}e.
+\textbf{f} Median contrast (dev - msc) in the target response magnitude across networks and stimuli in the full and ablated models, arranged as in panel b.
+The subtraction of the two quantities in the box plot below shows that the target response tends to be increased more in the full model.
 
-This naturally raises the question of why, with the addition of STD, the non-target dev response was reduced more than other responses.
+This naturally raises the question of why, with the addition of STD, the non-target dev response was reduced more than the non-target msc responses.
 We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.

### 2022-12-28 paper-1/manuscript.tex
Added grand median comparisons to F7 stats

@@
 
-Inspecting the non-target response reduction in the full model in Figure \ref{fig7}b, we found that it was on average larger than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), if only by a small amount on the order of 0.03 spikes per neuron.
+Inspecting the non-target response reduction in the full model in Figure \ref{fig7}b, we found that it was on average larger than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians differing by 36\%.
 Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
@@
 As shown in Figure \ref{fig7}e, the contrasts (dev - msc) in $V_{TA}$ at the start of target trials and in response magnitude to A were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response was clearly increased in dev over msc, and this increase was greater than in the ablated model (T = 1.94e+03, p = 0.000723), as shown in Figure \ref{fig7}f.
+Despite the weaker correlation, however, the full model target response was clearly increased in dev over msc, as shown in Figure \ref{fig7}f, and this increase was greater than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians differing by 94\%.
 This greater contrast, of course, directly reflects the greater DDI reported in Figure \ref{fig3}.

### 2022-12-28 paper-1/manuscript.tex
local/global adaptation -> suppression

@@
 
-Thus, we have finally traced the greater response to A in the dev sequence to its roots, and revealed two complementary roles of TA: The frequent presentation of non-target stimulus B in the dev sequence and concomitant increased activity in a subset of neurons responding early to B causes \emph{local adaptation} by TA in this subset, leading to smaller network responses.
-By contrast, stimuli in the msc sequence are presented more sporadically, allowing local adaptation to decay, thus leading to larger network responses.
-This difference in non-target response sizes -- small in dev, large in msc -- is then reflected in network-wide \emph{global adaptation}, such that the network is more excitable to stimulation with A (or other non-standard inputs) in the dev sequence.
+Thus, we have finally traced the greater response to A in the dev sequence to its roots, and revealed two complementary roles of TA: The frequent presentation of non-target stimulus B in the dev sequence and concomitant increased activity in a subset of neurons responding early to B causes \emph{local suppression} by TA in this subset, leading to smaller network responses.
+By contrast, stimuli in the msc sequence are presented more sporadically, allowing local suppression to decay, thus leading to larger network responses.
+This difference in non-target response sizes -- small in dev, large in msc -- is then reflected in network-wide \emph{global suppression}, such that the network is more excitable to stimulation with A (or other non-standard inputs) in the dev sequence.
 
@@
 The contrasts (dev - msc) in the non-target response magnitude and in the resulting $V_{TA}$ levels at the start of B trials both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the dev sequence.
-Likewise, the correlation across excitatory neurons between non-target response contrast and the contrast in $V_{TA}$ before B trials remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992) This indicates that local adaptation to B mediated by TA remained a key driver of the reduced response to non-target trials.
+Likewise, the correlation across excitatory neurons between non-target response contrast and the contrast in $V_{TA}$ before B trials remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992) This indicates that local suppression to B mediated by TA remained a key driver of the reduced response to non-target trials.
 
@@
 
-Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local adaptation to B resulting in reduced B responses and reduced global adaptation, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
+Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
 The primary cause for the greater target response, therefore, appeared to be the smaller non-target response, which was where we noted the first clear shift towards increased deviance detection.
@@
 We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.
+This would mean that STD is capable only of local suppression, reducing the standard response, but not of global suppression, leaving networks with only STD unable to differentiate between dev and msc target trials.
 

### 2022-12-28 paper-1/manuscript.tex
Rev sec Results:F8

@@
 
-We had good reasons for this assumption: Firstly, notice that the ablated model with STD only produced only a positive SSA index (i.e., the target response in msc was greater than in std), but not a significantly positive DDI (i.e., no difference between target dev and msc responses).
+We had good reasons for this assumption: Firstly, referring back to Figure \ref{fig3}, notice that the ablated model with STD only produced only a positive SSA index (i.e., the target response in msc was greater than in std), but not a significant DDI (i.e., no difference between target dev and msc responses).
 Secondly, we note that updates to the spike-triggered depression variable $x_j$ (cf.
@@
 Conversely, the frequently presented stimulus B could lead to a build-up of depression strong enough to suppress its own response.
-Referring back to Figure \ref{fig3}b, we see that there was indeed a significant, if small, effect of presentation frequency even in the ablated model with STD only, yielding a positive SSA index on average.
-% To show this in the sample network, we correlated the number of spikes in dev target trials with the spike-triggered depression variable $x_j$ at the start of later trials (cf.
+To show that the network does not carry information beyond the most recent trial in the sample network, we correlated the number of spikes in dev target trials with the spike-triggered depression variable $x_j$ at the start of later trials (cf.
 Equation \ref{eqn-xsyn}) across excitatory neurons.
-We found that, while the correlation was very clear in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the influence of both decay and unrelated activity in the intervening trial.
+We found that, while the correlation was very clear in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
 
-To show that STD being carried over between trials was responsible for the reduction of non-target dev pulses relative to msc, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
-Unlike $V_{TA}$, which directly affected the neuron to which the value belongs, $x_j$ exerted its effects on the other side of a synapse.
-Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $V_{TA}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to bring a neuron to fire.
+To show that STD being carried over between repeated trials was responsible for the reduction of non-target dev pulses relative to msc, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
+Unlike $V_{TA}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
+Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $V_{TA}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
 A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
@@
 Lastly, since we contrasted this measure between conditions (see below), we aimed to avoid effects purely due to differences in presynaptic activity.
-Therefore, we calculated $R^S$ as the average response across relevant trials in all conditions, leaving the contrast entirely in the hands of $x^S$.
+Therefore, we calculated $R^S$ as the average response across $S$ trials in all conditions, leaving the contrast entirely in the hands of $x^S$.
 
@@
 \textbf{Fig. 8} In the full model, STD reduces dev non-target and increases dev target responses relative to msc.
-\textbf{a} Average depression $D$ in target (A) trials (top row) and B trials (bottom row) in the dev (left) and msc sequence (center), and their contrast (right).
+\textbf{a} Average depression $D$ in target (A) trials (top row) and B trials (bottom row) in the dev (left) and msc sequence (center), and their contrast (right) in the sample network.
 The color scale is shared across columns (and between dev and msc) for direct comparability.
-\textbf{b} Statistics across networks and stimuli of the contrast (dev - msc) in $D$ in A trials (left) and B trials (right)
-\textbf{c} Relationship across neurons of the sample network between the contrast (dev - msc) in $D^S$ and the contrast in average number of spikes fired in $S$ trials, and associated linear regression.
-Left, target trials; right, B trials.
-\textbf{d} Network-wide mean contrast (dev - msc) in $D^S$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in panel c ($\Delta D^S$ vs response size contrast across neurons).
-Each marker corresponds to one network and target stimulus.
-Red markers indicate a significant negative correlation (one-sided t-test, significance level 0.05).
-Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
+\textbf{b} Relationship across neurons of the sample network between the contrast (dev - msc) in $D^A$ and the contrast in the target response, and associated linear regression.
+\textbf{c} Median contrast (dev - msc) in $D^B$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
+\textbf{d} Relationship across networks and stimuli between the median contrast in $D^B$ among the early portion and the median contrast in the response to B among the entire network, and associated linear regression.
+\textbf{e} Network-wide median contrast in $D^A$, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response, across networks and stimuli.
+The plot shows both that the negative correlation shown in panel b is robust (most networks are plotted with red markers, i.e.
+significantly negatively correlated), and that $\Delta D^A$ is less than 0 on average (see the inset histogram on the horizontal axis)
 
@@
 Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both conditions.
-In the msc sequence, we can attribute the difference between A and B primarily to the response size: The context was the same, but responses to B tended not to spread out as much in this network.
-However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site (see also Figure \ref{fig6}a), with depression therefore similarly limited.
+In the msc sequence, we can attribute the difference between A and B primarily to the response pattern: The context was the same, but responses to B tended not to spread out as much in this network.
+However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly limited.
 As a result, the contrast (dev - msc) showed slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicating greatly increased synaptic depression in dev.
 
-To show how robust these effects were, we turn to statistics across networks and stimuli.
-We calculated the trial-average depression values $D^A$ and $D^B$ in dev and msc conditions, subtracted them, then calculated the mean over neurons.
-The resulting contrasts are summarized in Figure \ref{fig8}b.
-We note first that neurons were subject to more depression in B trials than in A trials in the dev sequence (-4.47, p = 9.02e-06; one-sided t-test, n = 120, data not shown), but not in the msc sequence (by definition).
-Furthermore, $D^B$ was greater in dev than msc (8.02, p = 4.05e-13).
-Both findings confirm that STD carried over between the frequent standard trials, but not, or less so, between the less frequent control and deviant trials.
-However, we also found that $D^A$ was greater in msc than dev (-7.54, p = 5.04e-12; one-sided t-test, n = 120), suggesting that STD did exert stimulus-independent suppression in the msc sequence, as is clearly seen in the sample network, and similar to what we described earlier for TA.
+To show that $D$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated the contrast (dev - msc) in $D^A$ against the target response contrast across neurons of the sample network.
+Given the relatively large responses in the contrasted conditions, and the infrequent presentation of these trials, allowing causality to mostly flow from $D$ to response, rather than also in reverse, we are confident that this relationship roughly reflects a causal influence.
+As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that $D$ reduces responses as expected.
 
-Finally, to show that $D$ reflects suppression of a neuron's activity, we quantified the relationship between the contrasts in $D$ and the corresponding response size contrast.
-In the sample network (Figure \ref{fig8}c), we see that neurons with a greater decrease in $D$ from msc to dev had correspondingly greater response increase, and vice versa.
-Across both A and B trials, this relationship was captured well by a linear regression (A: Pearson's $\rho$ = -0.541, p = 2.28e-77, B: $\rho$ = -0.186, p = 1.61e-09; one-sided t-test, n = 1000).
-These findings were borne out across most networks and stimuli (Figure \ref{fig8}d), showing that $D$ captures the impact of STD not only on $V_m$, but also on the resulting output spikes.
+Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
+First, in Figure \ref{fig8}c, we recapitulate the logic established in Figure \ref{fig6}: Given the spatial clustering of $D^B$, we surmised a similar local suppression effect as with TA.
+We found that, indeed, neurons responding early to stimulation in B showed greater $D^B$ in dev than msc (T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced $D^B$ (T = 2.24e+03, p = 0.000133).
+We can relate this directly to the non-target response contrasts in the early and late portions shown in Figure \ref{fig7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in dev causes local suppression in terms of $D$.
+% Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
 
-To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction of the response to the frequently presented standard stimuli (dev B) by adaptation, and an increase in the response to deviant stimuli (dev A) by relief from adaptation, relative to control.
-Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in dev B responses also entailed lower $V_{TA}$, and higher dev A response by consequence.
-Thus, adding STD to the model increased the DDI both directly, as outlined at the start of this paragraph, and indirectly, by altering the pattern of TA by way of changed responses.
+To show that increased local suppression by STD reduces the non-target response, we correlated the median contrast in $D^B$ among the early portion of networks with the median contrast of the total response to B.
+As we show in Figure \ref{fig8}d, higher $\Delta D^B$ was clearly predictive of a lower network response ($\rho$ = -0.581).
+This confirms again the integrity of the measure $D$, but also that STD-mediated local suppression played a role in reducing the non-target response in the dev sequence.
+% Possible concern: The equivalent hasn't been shown in TA (Fig6), but likely could be (in the same fashion; dTA ~ dXB)
+
+Finally, given the very clear contrast in $D^A$ in the sample network, we considered the possibility that, in the context of the full model, STD could also evoke global suppression in the msc sequence, contrary to our stated hypothesis.
+If so, we should expect to see less $D^A$ in dev than msc.
+To confirm this, we analysed the contrast (dev - msc) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120).
+In addition, as shown in Figure \ref{fig8}e, we correlated the contrast in $D^A$ with the target response contrast in each network, and found that weaker depression in dev predicted a greater response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
+We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).
+Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in msc than dev in terms of $D$, this difference appears to manifest weakly in activity differences.
+
+To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction of the response to the frequently presented standard stimuli (dev B) by local suppression, and an increase in the response to deviant stimuli (dev A) by relief from global suppression, relative to control.
+Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in dev B responses also entailed lower $V_{TA}$, and higher dev A responses by consequence.
+Thus, adding STD to the model increased the DDI both directly, by reducing global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
+
+We note that the direct and indirect effect of STD on global suppression appear to be on the same order of magnitude.
+To see this, compare the change in the $V_{TA}$ contrast at the start of A trials as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta D^A$ (Figure \ref{fig8}e), both of which are on the order of $\pm$ 0.1 mV.
+Conversely, notice that, since TA and STD work in similar fashion to increase the dev target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
 

### 2022-12-28 paper-1/manuscript.tex
Rev sec Methods:Model (better refs to Fig1)

@@
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-See Figure \ref{fig1}b and c for a representative example of the resulting network structure.
+Figure \ref{fig1}b shows a representative example of the resulting network structure, which we will be referring to as the sample network throughout this paper.
+In Figure \ref{fig1}c, we show the distance from the neurons stimulated at site 1 to each neuron in the network, in terms of the minimum number of excitatory synapses to reach the target, in order to give an intuition for how excitation may spread.
+In addition to the sample network, we generated additional networks for statistics purposes.
 Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response: Networks where fewer than 500 neurons responded to stimulation at any of the five sites after full recovery were discarded.

### 2022-12-28 paper-1/manuscript.tex
Rev sec Results:F2

@@
 
-% Our first aim was to replicate deviance detection properties reported in dissociated neuronal cell culture by \cite{Kubota2021-dx}, reproduced in Figure \ref{fig2}a.
-In their case, cell cultures -- like our model, randomly connected networks with two-dimensional extent -- were stimulated at 10 different sites in msc sequences, and at two sites with 10\%/90\% probability in dev and std sequences.
-Deviance detection was assessed analogously to our definition of DDI and reached an average level of around 0.3 across two cultures.
+To show that our model is capable of deviance detection, we first show the responses of the sample network in all conditions (dev, msc, and std sequences) in Figure \ref{fig2}a.
+Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color.
+All responses share an early activity pattern, which derives from the stimulated neurons and their immediate postsynaptic targets, but differences soon start to emerge.
+The dev sequence responses, shown at the top of the plot, exhibit a clear peak in activity around 15 ms after stimulation.
+The std sequence responses show a later, more diffuse activity peak around 20 ms, and the msc sequence responses lack any obvious activity cluster.
+We sum these patterns up in Figure \ref{fig2}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
+Clearly, the target response in the dev sequence is greater than either the msc or the std response.
 
-% In our model, we could readily identify networks and stimuli whose responses developed in a qualitatively similar fashion, with high activity in response to target in the dev sequence, low activity in the std sequence, and intermediate activity in the msc sequence, as illustrated with an example in Figure \ref{fig2}b.
-The model responses developed on a compressed time scale compared to neuronal culture, which we tentatively attribute to unfaithful aspects of the model, including a lack of conduction delay or spontaneous activity, a much smaller number of participating neurons, and structural differences.
-
-% Across networks and stimuli (Figure \ref{fig2}c), we found that our model reliably produced positive mismatch responses (T = 6.67e+03, p = 9.26e-16; one-sided Wilcoxon signed-rank test, n = 120) while both SSA index (T = 6.19e+03, p = 9.76e-12) and DDI (T = 5.97e+03, p = 4.75e-10) were also positive.
-The average DDI value reached around 0.1, which was less than the values reported by \cite{Kubota2021-dx}, but note that our target stimuli were presented more frequently (20\%) than theirs (10\%) in the dev and msc sequences.
+We could readily identify similar patterns emerging across most networks and stimulus assignments.
+To show that deviance detection is robust to network and stimulus identity, we calculated the DDI and related indices mentioned above for each network, and show them in a summary plot in Figure \ref{fig2}c.
+We found that our model reliably produced positive mismatch indices (T = 6.67e+03, p = 9.26e-16, median 0.253; one-sided Wilcoxon signed-rank test, n = 120) while both SSA index (T = 6.19e+03, p = 9.76e-12, median 0.135) and DDI (T = 5.97e+03, p = 4.75e-10, median 0.114) were also positive, despite some outliers.
+This shows that our model was well behaved and suitable for an investigation of deviance detection.
 

### 2022-12-29 paper-1/manuscript.tex
SSA disambiguation

@@
 Note how the dev response is very large, followed by an intermediate msc response and a low std response.
-\textbf{c} Mismatch, adaptation, and deviance detection indices (see Equation \ref{eqn-ddi}) across networks and stimuli (n = 120).
+\textbf{c} Mismatch, repetition suppression, and deviance detection indices (see Equation \ref{eqn-ddi}) across networks and stimuli (n = 120).
 In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points
@@
 \end{equation}
-We define stimulus-specific adaptation (SSA, $R^A_{msc} - R^A_{std}$) and mismatch ($R^A_{dev} - R^A_{std}$) indices analogously, capturing respectively the effect of adaptation, and of both adaptation and deviance detection together.
+We define related indices, contrasting the responses in msc against std, and in dev against std, analogously.
+Since contrasts of both kinds have been called "stimulus-specific adaptation" or SSA in the literature \citep{Kubota2021-dx,Taaseh2011-gg}, we chose to avoid such ambiguity, using instead the term "repetition suppression" (RSI, $R^A_{msc} - R^A_{std}$) to refer to the reduction in response induced mainly by a difference in presentation frequency while controlling for context, and the term "mismatch" (MMI, $R^A_{dev} - R^A_{std}$) to refer to the uncontrolled contrast, which captures a combination of repetition suppression and deviance detection.
 
@@
 To show that deviance detection is robust to network and stimulus identity, we calculated the DDI and related indices mentioned above for each network, and show them in a summary plot in Figure \ref{fig2}c.
-We found that our model reliably produced positive mismatch indices (T = 6.67e+03, p = 9.26e-16, median 0.253; one-sided Wilcoxon signed-rank test, n = 120) while both SSA index (T = 6.19e+03, p = 9.76e-12, median 0.135) and DDI (T = 5.97e+03, p = 4.75e-10, median 0.114) were also positive, despite some outliers.
+We found that our model reliably produced positive mismatch indices (T = 6.67e+03, p = 9.26e-16, median 0.253; one-sided Wilcoxon signed-rank test, n = 120) while both RSI (T = 6.19e+03, p = 9.76e-12, median 0.135) and DDI (T = 5.97e+03, p = 4.75e-10, median 0.114) were also positive, despite some outliers.
 This shows that our model was well behaved and suitable for an investigation of deviance detection.
@@
 Note the different axis scaling on the top row (without TA) and the bottom row (with TA).
-\textbf{b} Adaptation indices $<$msc - std$>$ across all networks and stimuli.
+\textbf{b} Repetition suppression indices $<$msc - std$>$ across all networks and stimuli.
 Asterisks indicate data greater than 0, established with a Wilcoxon signed-rank test at a significance level 0.05.
@@
 The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the msc sequence and calculated z-scores with respect to the population.
-We then focused on networks whose absolute z-score was less than 1 for both A and B responses, and hand-picked a sample that exhibited both an increased DDI in STD+TA over TA only, and positive SSA indices in both STD+TA and TA only conditions.
+We then focused on networks whose absolute z-score was less than 1 for both A and B responses, and hand-picked a sample that exhibited both an increased DDI in STD+TA over TA only, and positive RSI in both STD+TA and TA only conditions.
 
@@
 
-The SSA index (Figure \ref{fig3}b) was exactly 0 as expected in the ablation with no plasticity, and positive with any plasticity applied (STD only: T = 4.97e+03, p = 0.000222, median 0.0039; TA only: T = 5.99e+03, p = 3.3e-10, median 0.097; STD+TA: T = 6.19e+03, p = 9.76e-12, median 0.135; one-sided Wilcoxon signed-rank tests, n = 120).
-Since the SSA index measures the degree of adaptation to frequent stimulus presentation, this indicates that both plasticity mechanisms were capable of imbuing the networks with some level of history dependence.
+The RSI (Figure \ref{fig3}b) was exactly 0 as expected in the ablation with no plasticity, and positive with any plasticity applied (STD only: T = 4.97e+03, p = 0.000222, median 0.0039; TA only: T = 5.99e+03, p = 3.3e-10, median 0.097; STD+TA: T = 6.19e+03, p = 9.76e-12, median 0.135; one-sided Wilcoxon signed-rank tests, n = 120).
+Since the RSI measures the degree of adaptation to frequent stimulus presentation, this indicates that both plasticity mechanisms were capable of imbuing the networks with some level of history dependence.
 
@@
 
-We had good reasons for this assumption: Firstly, referring back to Figure \ref{fig3}, notice that the ablated model with STD only produced only a positive SSA index (i.e., the target response in msc was greater than in std), but not a significant DDI (i.e., no difference between target dev and msc responses).
+We had good reasons for this assumption: Firstly, referring back to Figure \ref{fig3}, notice that the ablated model with STD only produced only a positive RSI (i.e., the target response in msc was greater than in std), but not a significant DDI (i.e., no difference between target dev and msc responses).
 Secondly, we note that updates to the spike-triggered depression variable $x_j$ (cf.

### 2022-12-29 paper-1/manuscript.tex
Editing

@@
 Note the different axis scaling on the top row (without TA) and the bottom row (with TA).
-\textbf{b} Repetition suppression indices $<$msc - std$>$ across all networks and stimuli.
+\textbf{b} Repetition suppression indices $<$msc - std$>$ across networks and stimuli.
 Asterisks indicate data greater than 0, established with a Wilcoxon signed-rank test at a significance level 0.05.
 See main text for detailed statistics.
-\textbf{c} Deviance detection indices $<$dev - msc$>$ across all networks and stimuli
+\textbf{c} Deviance detection indices $<$dev - msc$>$ across networks and stimuli
 
-Having confirmed the presence of deviance detection in our model, we then turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-In the remainder of the paper, we will proceed to detail the processes leading to deviance detection on a sample network, then confirm the highlighted trends statistically with reference to the full set of 120 networks and stimuli.
+Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
+In the remainder of the paper, we will proceed to detail the processes leading to deviance detection on a sample network, then confirm the highlighted trends statistically with reference to the full set of 120 networks and stimulus assignments.
 The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the msc sequence and calculated z-scores with respect to the population.
-We then focused on networks whose absolute z-score was less than 1 for both A and B responses, and hand-picked a sample that exhibited both an increased DDI in STD+TA over TA only, and positive RSI in both STD+TA and TA only conditions.
+Considering only networks whose absolute z-score was less than 1 for both A and B responses, we hand-picked a sample network and stimulus assignment that exhibited both an increased DDI in STD+TA over TA only, and positive RSI in both STD+TA and TA only conditions.
 
@@
 
-In the following, we will use a sample network and target stimulus to first build an understanding of how TA leads to deviance detection when acting alone, i.e.
-in the ablated TA only model.
+In the following, we will first build an understanding of how TA leads to deviance detection when acting alone, i.e.
+in the ablated TA only model, working backwards from the increased target response in dev, identifying and tracing causal factors step by step.
 Then, we will build on this understanding to elucidate how the addition of STD, which was shown above to be ineffective for deviance detection on its own, can lead to an increased deviant response in the full model.
@@
 
-To understand why the same stimulus A evoked a higher response in dev than in msc with TA only, we first compared the average responses in the two conditions, sorting neurons by their response onset to permit a high-resolution view of the population activity.
+To understand why the same stimulus A evoked a higher response in dev than in msc with TA only, we first compared the average responses in the two conditions, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
 As shown in Figure \ref{fig4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both conditions.
@@
 In later parts of the response (top half of the panel), the dev response was larger, but not obviously shifted in time.
-Since the stimulus was identical in both conditions, and TA was the only plsaticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.
+Since the stimulus was identical in both conditions, and TA was the only plasticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.
 
@@
 As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target dev trials relative to msc, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
-In contrast, once the bulk of activity had occurred, many neurons showed higher thresholds in dev (right panel, later orange portions), since they spiked more often.
-Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $V_{TA}$.
+On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in dev (right panel, later orange portions), since they spiked more often.
+Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $V_{TA}$, leading to numerous horizontal breaks in the plot.
 
-To assess the difference in synaptic currents in a way that allowed direct comparison to $V_{TA}$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs (resets after spikes notwithstanding).
-As the direct cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
+To assess the difference in synaptic currents in a way that allowed direct comparison to $V_{TA}$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
+As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
 The contrast (right panel) shows not only a notable relative increase in membrane voltages in dev around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
@@
 We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response of each excitatory neuron in terms of spikes fired.
-In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800), and this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
 As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Median $V_{TA}$ was lower in dev than msc (T = 514, p = 1.67e-16, median = -0.094 mV; one-sided Wilcoxon, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
@@
 Here, only the initial response was driven mainly by $V_{TA}$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
-This indicates that adaptation initiated and dominated the increase in firing in the beginning, whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
+This indicates that adaptation initiated and dominated the increase in firing at the start of trials, whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
 
@@
 Conversely, in the dev sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
+
 Next, we turned to the likely cause of this difference, the non-target responses (Figure \ref{fig5}b), mapped into space as the average number of spikes fired in each neuron.
 As expected, the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds.
-Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (alongside 300 C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, most of these very close to the stimulation site of B.
+Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (alongside the same number of C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, most of these very close to the stimulation site of B.
 
-Correlating the non-target response contrast (dev - msc) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev almost perfectly predicted the resulting reduction in threshold adaptation levels (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between msc and dev sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the non-target trial response is the primary driver of threshold adaptation.
+Correlating the non-target response contrast (dev - msc) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev almost perfectly predicted the resulting reduction in TA levels (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between msc and dev sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the non-target trial response is the primary driver of TA.
 
@@
 \end{figure*}
-% TODO 5-8: Re-label stim locations once per figure
 \textbf{Fig. 6} Dev non-target responses are reduced due to adaptation in early responders to B.
@@
 Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
-We hypothesized that this adaptation is mediated by a small subset of neurons that (a) responds very quickly to stimulation in B, (b) responds more strongly to non-target stimuli in dev than msc, and (c) has increased TA levels as a result, reducing the later response in the same way as demonstrated in Figure \ref{fig4} in reverse.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and comparatively strongly to stimulation in B, and therefore more to non-target stimuli in dev than msc, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
 
 To see this in the sample network, consider first the latency of responses to B across all conditions, shown by way of rank order in Figure \ref{fig6}a.
-We found that the response developed in an approximately radial pattern, spreading outward from the stimulus site.
-We did notice that the early response appears to gravitate towards the stimulus site for A, but confirmed that this was the case in all conditions separately, indicating a structural rather than dynamic cause (data not shown).
+We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site.
+We did notice that the early response appeared to gravitate towards the stimulus site for A, but confirmed that this was the case in all conditions separately, indicating a structural rather than dynamic cause (data not shown).
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B in any condition, and a late portion, containing the remaining 900 neurons.
-In Figure \ref{fig6}b, we related the response contrast (dev - msc) to the contrast in $V_{TA}$ at the start of B trials.
+In Figure \ref{fig6}b, we related the non-target response contrast (dev - msc) to the contrast in $V_{TA}$ at the start of B trials.
 Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993) echoed the related finding in Figure \ref{fig5}c.
@@
 Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B.
-We first confirmed that the non-target response and $V_{TA}$ contrasts were robustly correlated (Median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon test for $\rho > 0$, n = 120).
+We first confirmed that the non-target response and $V_{TA}$ contrasts were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $V_{TA}$ at the start of B trials (Figure \ref{fig6}d) within these two portions.
-We found that the response in the early portion was consistently greater in dev than msc (T = 6.98e+03, p = 8.48e-19; one-sided Wilcoxon, n = 120), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
-The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in dev (T = 4.63e+03, p = 2.21e-17), while the late portion was more recovered (T = 26, p = 2.97e-17).
+We found that the response in the early portion was consistently greater in dev than msc (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
+The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in dev (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
 

### 2022-12-29 paper-1/manuscript.tex
Rev Fig6 legend & text (added panel e)

@@
 \textbf{c} Median contrast (dev - msc) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli
+\textbf{d} Relationship across networks and stimuli between the median contrast in $V_{TA}$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
 
@@
 The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in dev (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
+Finally, to show that early $V_{TA}$ was causal in reducing the response to B, we correlated the median contrast in $V_{TA}$ among the early portion of neurons with the network-wide median contrast of the response to stimulation in B, as shown in Figure \ref{fig6}e.
+Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
 
 To sum up, we showed in Figure \ref{fig4} that greater $V_{TA}$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response.
-Here, we showed that the neurons responding earliest to B exhibit higher $V_{TA}$ in the dev sequence as a consequence of an increased response, in direct contrast to the response and TA levels of the network as a whole.
+Here, we showed that the neurons responding earliest to B exhibit higher $V_{TA}$ in the dev sequence as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
 

### 2022-12-29 paper-1/manuscript.tex
Editing

@@
 By contrast, stimuli in the msc sequence are presented more sporadically, allowing local suppression to decay, thus leading to larger network responses.
-This difference in non-target response sizes -- small in dev, large in msc -- is then reflected in network-wide \emph{global suppression}, such that the network is more excitable to stimulation with A (or other non-standard inputs) in the dev sequence.
+This difference in non-target response sizes -- small in dev, large in msc -- is then reflected in network-wide \emph{global suppression}, such that the network is more excitable to stimulation with A (or other non-standard inputs) in the dev sequence than in the control context.
 
@@
 
-To understand how STD affects deviance detection, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
+To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
 Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}a the full model equivalent of Figure \ref{fig6}c and d.
 The contrasts (dev - msc) in the non-target response magnitude and in the resulting $V_{TA}$ levels at the start of B trials both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the dev sequence.
-Likewise, the correlation across excitatory neurons between non-target response contrast and the contrast in $V_{TA}$ before B trials remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992) This indicates that local suppression to B mediated by TA remained a key driver of the reduced response to non-target trials.
+Likewise, the correlation across excitatory neurons between these two contrasts remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992).
+This indicates that local suppression to B mediated by TA remained a key driver of the reduced response to non-target trials.
 
-Inspecting the non-target response reduction in the full model in Figure \ref{fig7}b, we found that it was on average larger than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians differing by 36\%.
+Comparing the non-target response reductions in Figure \ref{fig7}b, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians differing by 36\%.
 Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
@@
 As shown in Figure \ref{fig7}e, the contrasts (dev - msc) in $V_{TA}$ at the start of target trials and in response magnitude to A were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response was clearly increased in dev over msc, as shown in Figure \ref{fig7}f, and this increase was greater than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians differing by 94\%.
+Despite the weaker correlation, however, the full model target response was clearly increased in dev over msc, as shown in Figure \ref{fig7}f, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians differing by 94\%.
 This greater contrast, of course, directly reflects the greater DDI reported in Figure \ref{fig3}.
@@
 
-This naturally raises the question of why, with the addition of STD, the non-target dev response was reduced more than the non-target msc responses.
+This naturally raises the question of why, with the addition of STD, the non-target dev response was reduced more than non-target msc responses.
 We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.
@@
 
-We had good reasons for this assumption: Firstly, referring back to Figure \ref{fig3}, notice that the ablated model with STD only produced only a positive RSI (i.e., the target response in msc was greater than in std), but not a significant DDI (i.e., no difference between target dev and msc responses).
+We had good reasons for this assumption: Firstly, referring back to Figure \ref{fig3}, notice that the ablated model with STD only produced a positive RSI (i.e., the target response in msc was greater than in std), but not a significant DDI (i.e., no difference between target dev and msc responses).
 Secondly, we note that updates to the spike-triggered depression variable $x_j$ (cf.
 Equation \ref{eqn-xsyn}) would decay to a negligible $e^\frac{-1 s}{\tau_x} = 0.0013$ times the original value within two trials, making it unlikely that inputs separated by intervening trials could be "remembered" in the network.
-Conversely, the frequently presented stimulus B could lead to a build-up of depression strong enough to suppress its own response.
-To show that the network does not carry information beyond the most recent trial in the sample network, we correlated the number of spikes in dev target trials with the spike-triggered depression variable $x_j$ at the start of later trials (cf.
+Conversely, the frequently presented stimulus B may have been able to lead to a build-up of depression strong enough to suppress its own response.
+
+To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated the number of spikes in dev target trials with $x_j$ at the start of later trials (cf.
 Equation \ref{eqn-xsyn}) across excitatory neurons.
-We found that, while the correlation was very clear in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
+We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
 
@@
 Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $V_{TA}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
+
 A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
-Here, we used the average trial response $R_j^S$ of neuron $j$ to a given stimulus $S$, in number of spikes, as the presynaptic drive.
-Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at rest (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.
+Here, we used the average trial response $R_j^S$ of neuron $j$ to a given stimulus $S$, in number of spikes, as a measure of the presynaptic drive.
+Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.
 Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $R_j^S k x_j^S$.
-For this estimate, we ignored added depression due to repeated spikes within a trial, instead defining $x_j^S$ as the average depression at the start of trials.
+For this estimate, we ignored added depression due to repeated spikes within a trial, instead defining $x_j^S$ as the mean depression at the start of trials.
 Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
@@
 Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both conditions.
-In the msc sequence, we can attribute the difference between A and B primarily to the response pattern: The context was the same, but responses to B tended not to spread out as much in this network.
-However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly limited.
+In the msc sequence, we can attribute the difference between A and B primarily to different response patterns: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network.
+However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited.
 As a result, the contrast (dev - msc) showed slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicating greatly increased synaptic depression in dev.
@@
 
-To show that increased local suppression by STD reduces the non-target response, we correlated the median contrast in $D^B$ among the early portion of networks with the median contrast of the total response to B.
-As we show in Figure \ref{fig8}d, higher $\Delta D^B$ was clearly predictive of a lower network response ($\rho$ = -0.581).
+To show that increased local suppression by STD reduces the non-target response, we correlated the median contrast in $D^B$ among the early portion of networks with the median contrast of the total response to B, echoing Figure \ref{fig6}e.
+As we show in Figure \ref{fig8}d, higher $\Delta D^B$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000).
 This confirms again the integrity of the measure $D$, but also that STD-mediated local suppression played a role in reducing the non-target response in the dev sequence.
-% Possible concern: The equivalent hasn't been shown in TA (Fig6), but likely could be (in the same fashion; dTA ~ dXB)
 
@@
 To confirm this, we analysed the contrast (dev - msc) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120).
-In addition, as shown in Figure \ref{fig8}e, we correlated the contrast in $D^A$ with the target response contrast in each network, and found that weaker depression in dev predicted a greater response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
+In addition, as shown in Figure \ref{fig8}e, we correlated the contrast in $D^A$ with the target response contrast across neurons in each network, and found that weaker depression in dev predicted a greater response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
 We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).
-Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in msc than dev in terms of $D$, this difference appears to manifest weakly in activity differences.
+Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in msc than dev in terms of $D$, this difference appears to manifest only weakly in activity differences.
 
 To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction of the response to the frequently presented standard stimuli (dev B) by local suppression, and an increase in the response to deviant stimuli (dev A) by relief from global suppression, relative to control.
-Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in dev B responses also entailed lower $V_{TA}$, and higher dev A responses by consequence.
-Thus, adding STD to the model increased the DDI both directly, by reducing global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
+Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in dev B responses also entailed lower $V_{TA}$, which independently lead to higher dev A responses.
+Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
 
-We note that the direct and indirect effect of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in the $V_{TA}$ contrast at the start of A trials as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta D^A$ (Figure \ref{fig8}e), both of which are on the order of $\pm$ 0.1 mV.
+We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
+To see this, compare the change in the $V_{TA}$ contrast at the start of A trials as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta D^A$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the dev target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
+Thus, this synergy allows the individually weak STD to exert a significantly larger effect when paired with the stronger TA.
 

### 2023-01-06 paper-1/manuscript.tex
Moved figures to/near the start of their text
Added SI figures
+todo: Discussion
Bugfixes
Rev: remove RSI, MMI

@@
 
+We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs.
+Following related research \citep{Kubota2021-dx,Harms2014-ah,Jacobsen2001-sc}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the dev sequence to A in the std sequence).
+In other words, we used the msc sequence as a neutral baseline where the target stimulus was presented no more often than in the dev sequence, but no expectation of a competing stimulus B could be formed.
+To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the dev and msc sequences, denoted as $R^A_{dev}$ and $R^A_{msc}$, defining a deviance detection index (DDI) as
+\begin{equation}
+    DDI = \frac{R^A_{dev} - R^A_{msc}}{R^A_{dev} + R^A_{msc}} \label{eqn-ddi}
+\end{equation}
+
 \begin{figure*}%
@@
 Note how the dev response is very large, followed by an intermediate msc response and a low std response.
-\textbf{c} Mismatch, repetition suppression, and deviance detection indices (see Equation \ref{eqn-ddi}) across networks and stimuli (n = 120).
-In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points
-
-We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs.
-Following related research \citep{Kubota2021-dx,Harms2014-ah,Jacobsen2001-sc}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the dev sequence to A in the std sequence).
-In other words, we used the msc sequence as a neutral baseline where the target stimulus was presented no more often than in the dev sequence, but no expectation of a competing stimulus B could be formed.
-To quantify deviance detection, we compared the mean number of spikes $R^A$ fired across the network in response to target stimulus A in the dev and msc sequences, defining a deviance detection index (DDI) as
-\begin{equation}
-    DDI = \frac{R^A_{dev} - R^A_{msc}}{R^A_{dev} + R^A_{msc}} \label{eqn-ddi}
-\end{equation}
-We define related indices, contrasting the responses in msc against std, and in dev against std, analogously.
-Since contrasts of both kinds have been called "stimulus-specific adaptation" or SSA in the literature \citep{Kubota2021-dx,Taaseh2011-gg}, we chose to avoid such ambiguity, using instead the term "repetition suppression" (RSI, $R^A_{msc} - R^A_{std}$) to refer to the reduction in response induced mainly by a difference in presentation frequency while controlling for context, and the term "mismatch" (MMI, $R^A_{dev} - R^A_{std}$) to refer to the uncontrolled contrast, which captures a combination of repetition suppression and deviance detection.
+\textbf{c} Response sizes $R^A_{dev}$ and $R^A_{std}$, normalized by $R^A_{msc}$, across networks and stimuli (n = 120).
+In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
+\textbf{d} Deviance detection indices (DDI, see Equation {eqn-ddi}) across networks and stimuli
 
@@
 We could readily identify similar patterns emerging across most networks and stimulus assignments.
-To show that deviance detection is robust to network and stimulus identity, we calculated the DDI and related indices mentioned above for each network, and show them in a summary plot in Figure \ref{fig2}c.
-We found that our model reliably produced positive mismatch indices (T = 6.67e+03, p = 9.26e-16, median 0.253; one-sided Wilcoxon signed-rank test, n = 120) while both RSI (T = 6.19e+03, p = 9.76e-12, median 0.135) and DDI (T = 5.97e+03, p = 4.75e-10, median 0.114) were also positive, despite some outliers.
+To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across conditions.
+Figure \ref{fig2}c shows a summary of the average dev and std response sizes, normalized by the response in the msc sequence.
+The data clearly confirmed that dev responses were larger than msc (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and std responses smaller (T = 687, p = 6.42e-15).
+Accordingly, the DDI (Figure \ref{fig2}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
 This shows that our model was well behaved and suitable for an investigation of deviance detection.
+
+Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
+In the remainder of the paper, we will proceed to detail the processes leading to deviance detection on a sample network, then confirm the highlighted trends statistically with reference to the full set of 120 networks and stimulus assignments.
+The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the msc sequence and calculated z-scores with respect to the population.
+Considering only networks whose absolute z-score was less than 1 for both A and B responses, we hand-picked a sample network and stimulus assignment that exhibited both an increased DDI in STD+TA over TA only, and greater msc than std responses in both STD+TA and TA only conditions.
 
@@
 Note the different axis scaling on the top row (without TA) and the bottom row (with TA).
-\textbf{b} Repetition suppression indices $<$msc - std$>$ across networks and stimuli.
+\textbf{b} Deviance detection indices across networks and stimuli.
 Asterisks indicate data greater than 0, established with a Wilcoxon signed-rank test at a significance level 0.05.
-See main text for detailed statistics.
-\textbf{c} Deviance detection indices $<$dev - msc$>$ across networks and stimuli
-
-Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-In the remainder of the paper, we will proceed to detail the processes leading to deviance detection on a sample network, then confirm the highlighted trends statistically with reference to the full set of 120 networks and stimulus assignments.
-The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the msc sequence and calculated z-scores with respect to the population.
-Considering only networks whose absolute z-score was less than 1 for both A and B responses, we hand-picked a sample network and stimulus assignment that exhibited both an increased DDI in STD+TA over TA only, and positive RSI in both STD+TA and TA only conditions.
+See main text for detailed statistics
 
@@
 
-The RSI (Figure \ref{fig3}b) was exactly 0 as expected in the ablation with no plasticity, and positive with any plasticity applied (STD only: T = 4.97e+03, p = 0.000222, median 0.0039; TA only: T = 5.99e+03, p = 3.3e-10, median 0.097; STD+TA: T = 6.19e+03, p = 9.76e-12, median 0.135; one-sided Wilcoxon signed-rank tests, n = 120).
-Since the RSI measures the degree of adaptation to frequent stimulus presentation, this indicates that both plasticity mechanisms were capable of imbuing the networks with some level of history dependence.
-
-Finally, across networks and stimuli, the DDI (Figure \ref{fig3}c) was exactly 0 with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Finally, across networks and stimuli, the DDI (Figure \ref{fig3}c) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
 To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: T = 6.1e+03, p = 4.95e-11; TA: T = 4.87e+03, p = 0.000604), and greater too than the sum of the DDI values across both STD only and TA only models (T = 5.12e+03, p = 4.52e-05).
@@
 Since the stimulus was identical in both conditions, and TA was the only plasticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.
-
-We quantified the former with the strength of adaptation $V_{TA}$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
-As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target dev trials relative to msc, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
-On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in dev (right panel, later orange portions), since they spiked more often.
-Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $V_{TA}$, leading to numerous horizontal breaks in the plot.
-
-To assess the difference in synaptic currents in a way that allowed direct comparison to $V_{TA}$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
-As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
-The contrast (right panel) shows not only a notable relative increase in membrane voltages in dev around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
 
@@
 \textbf{h} Relative contribution of $V_{TA}$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
+
+We quantified the former with the strength of adaptation $V_{TA}$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
+As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target dev trials relative to msc, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in dev (right panel, later orange portions), since they spiked more often.
+Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $V_{TA}$, leading to numerous horizontal breaks in the plot.
+
+To assess the difference in synaptic currents in a way that allowed direct comparison to $V_{TA}$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
+As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
+The contrast (right panel) shows not only a notable relative increase in membrane voltages in dev around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
 
@@
 
-First, we mapped the average value of $V_{TA}$ at the start of target trials to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
-In the msc sequence, $V_{TA}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
-Conversely, in the dev sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
-
-Next, we turned to the likely cause of this difference, the non-target responses (Figure \ref{fig5}b), mapped into space as the average number of spikes fired in each neuron.
-As expected, the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds.
-Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (alongside the same number of C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, most of these very close to the stimulation site of B.
-
-Correlating the non-target response contrast (dev - msc) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev almost perfectly predicted the resulting reduction in TA levels (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between msc and dev sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the non-target trial response is the primary driver of TA.
-
 \begin{figure*}%
@@
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations
+
+First, we mapped the average value of $V_{TA}$ at the start of target trials to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
+In the msc sequence, $V_{TA}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
+Conversely, in the dev sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
+
+Next, we turned to the likely cause of this difference, the non-target responses (Figure \ref{fig5}b), mapped into space as the average number of spikes fired in each neuron.
+As expected, the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds.
+Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (alongside the same number of C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, most of these very close to the stimulation site of B.
+
+Correlating the non-target response contrast (dev - msc) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev almost perfectly predicted the resulting reduction in TA levels (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between msc and dev sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the non-target trial response is the primary driver of TA.
+
+Finally, we examined the cause of the reduced response in non-target trials.
+Going from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation.
+Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and comparatively strongly to stimulation in B, and therefore more to non-target stimuli in dev than msc, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
 
@@
 \textbf{d} Relationship across networks and stimuli between the median contrast in $V_{TA}$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
-
-Finally, we examined the cause of the reduced response in non-target trials.
-Going from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation.
-Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
-We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and comparatively strongly to stimulation in B, and therefore more to non-target stimuli in dev than msc, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
 
@@
 
+\begin{figure*}%
+    \centering
+    \includegraphics{fig7.pdf}
+    \caption{}
+    \label{fig7}
+\end{figure*}
+\textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
+\textbf{a} Median contrast (dev - msc) in the non-target response (left) and in $V_{TA}$ at the start of B trials (right) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+Compare to Figure \ref{fig6}c-d.
+\textbf{b} Median contrast (dev - msc) in the non-target response magnitude across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
+The red line is the identity function; points above this line represent networks with greater response reduction (dev - msc) in the full model than in the ablated model.
+The subtraction of the two quantities in the box plot below shows that the non-target response tends to be reduced more in the full model.
+\textbf{c} Non-target response contrast, plotted against its Pearson correlation coefficient with the corresponding contrast in $V_{TA}$ at the start of target trials, in the full model.
+Compare to Figure \ref{fig5}d.
+\textbf{d} Median contrast (dev - msc) in $V_{TA}$ at the start of target trials across networks and stimuli in the full and ablated models, arranged as in panel b.
+The subtraction of the two quantities in the box plot below shows that $V_{TA}$ tends to be reduced more in the full model.
+\textbf{e} $\Delta V_{TA}$ at the start of target trials, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response magnitude, in the full model.
+Compare to Figure \ref{fig4}e.
+\textbf{f} Median contrast (dev - msc) in the target response magnitude across networks and stimuli in the full and ablated models, arranged as in panel b.
+The subtraction of the two quantities in the box plot below shows that the target response tends to be increased more in the full model
+
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig7.pdf}
-    \caption{}
-    \label{fig7}
-\end{figure*}
-\textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
-\textbf{a} Median contrast (dev - msc) in the non-target response (left) and in $V_{TA}$ at the start of B trials (right) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
-Compare to Figure \ref{fig6}c-d.
-\textbf{b} Median contrast (dev - msc) in the non-target response magnitude across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
-The red line is the identity function; points above this line represent networks with greater response reduction (dev - msc) in the full model than in the ablated model.
-The subtraction of the two quantities in the box plot below shows that the non-target response tends to be reduced more in the full model.
-\textbf{c} Non-target response contrast, plotted against its Pearson correlation coefficient with the corresponding contrast in $V_{TA}$ at the start of target trials, in the full model.
-Compare to Figure \ref{Fig5}d.
-\textbf{d} Median contrast (dev - msc) in $V_{TA}$ at the start of target trials across networks and stimuli in the full and ablated models, arranged as in panel b.
-The subtraction of the two quantities in the box plot below shows that $V_{TA}$ tends to be reduced more in the full model.
-\textbf{e} $\Delta V_{TA}$ at the start of target trials, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response magnitude, in the full model.
-Compare to Figure \ref{Fig4}e.
-\textbf{f} Median contrast (dev - msc) in the target response magnitude across networks and stimuli in the full and ablated models, arranged as in panel b.
-The subtraction of the two quantities in the box plot below shows that the target response tends to be increased more in the full model.
-
 This naturally raises the question of why, with the addition of STD, the non-target dev response was reduced more than non-target msc responses.
@@
 
-We had good reasons for this assumption: Firstly, referring back to Figure \ref{fig3}, notice that the ablated model with STD only produced a positive RSI (i.e., the target response in msc was greater than in std), but not a significant DDI (i.e., no difference between target dev and msc responses).
-Secondly, we note that updates to the spike-triggered depression variable $x_j$ (cf.
+We had good reasons for this assumption: A priori, updates to the spike-triggered depression variable $x_j$ (cf.
 Equation \ref{eqn-xsyn}) would decay to a negligible $e^\frac{-1 s}{\tau_x} = 0.0013$ times the original value within two trials, making it unlikely that inputs separated by intervening trials could be "remembered" in the network.
@@
 
+TODO
+
 \begin{figure*}%
@@
 
-% \begin{appendices}
+\begin{appendices}
 
-% \section{Section title of first appendix}\label{secA1}
+\section{Supplementary infomation}\label{secA1}
+
+\begin{figure*}%
+    \centering
+    \includegraphics{fig4-withSTD.pdf}
+    \caption{}
+    \label{sfig4}
+\end{figure*}
+
+\begin{figure*}%
+    \centering
+    \includegraphics{fig5-withSTD.pdf}
+    \caption{}
+    \label{sfig5}
+\end{figure*}
+
+\begin{figure*}%
+    \centering
+    \includegraphics{fig6-withSTD.pdf}
+    \caption{}
+    \label{sfig6}
+\end{figure*}
+
+\begin{figure*}%
+    \centering
+    \includegraphics{fig8-noTA.pdf}
+    \caption{}
+    \label{sfig8}
+\end{figure*}
 
@@
 
-% \end{appendices}
+\end{appendices}
 

### 2023-01-06 paper-1/manuscript.tex
Italicize dev/msc/std

@@
 \newcommand{\orcidauthorB}{0000-0001-9402-0851}
+
+% Shorter italic dev/msc/std
+\newcommand{\dev}{\textit{dev}}
+\newcommand{\msc}{\textit{msc}}
+\newcommand{\std}{\textit{std}}
 
@@
 The target stimulus (A) is represented as solid red, and non-target stimuli are represented with faded colors.
-Three randomized sequences were presented, containing 80\% A and 20\% B (std), 20\% A and 80\% B (dev), and 20\% of each of 5 stimuli, including A and B (msc).
+Three randomized sequences were presented, containing 80\% A and 20\% B (\std{}), 20\% A and 80\% B (\dev{}), and 20\% of each of 5 stimuli, including A and B (\msc{}).
 Note that, while the onset asynchrony of 500 ms is represented faithfully, the horizontal extent of the stimulation markers has no meaning; stimuli were presented instantaneously rather than over an extended period.
@@
 To investigate deviance detection, we used a classical oddball paradigm as illustrated in Figure \ref{fig1}d, presenting a total of 500 stimuli at regular intervals of 500~ms.
-Target stimuli (labeled "A" or "target" hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (std), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (dev) with 100 A and 400 B, and the many-standards control sequence (msc) with 100 presentations of each of the stimuli A through E.
-In this way, A is presented as the predominant and therefore expected stimulus (std), as an infrequent violation of an expectation (of B, dev), and, to control for pure adaptation effects, as an equally infrequent stimulus with no strong expectation of any particular input (msc).
+Target stimuli (labeled "A" or "target" hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (\std{}), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (\dev{}) with 100 A and 400 B, and the many-standards control sequence (\msc{}) with 100 presentations of each of the stimuli A through E.
+In this way, A is presented as the predominant and therefore expected stimulus (\std{}), as an infrequent violation of an expectation (of B, \dev{}), and, to control for pure adaptation effects, as an equally infrequent stimulus with no strong expectation of any particular input (\msc{}).
 
-In each network, we simulated oddball sequences with two A/B stimulus pairs (sites 1 and 2 in one pair, and sites 3 and 5 in the other pair), where each stimulus site in a pair was used once as target (A) and once as non-target (B), and a single msc sequence involving all sites.
+In each network, we simulated oddball sequences with two A/B stimulus pairs (sites 1 and 2 in one pair, and sites 3 and 5 in the other pair), where each stimulus site in a pair was used once as target (A) and once as non-target (B), and a single \msc{} sequence involving all sites.
 Networks were reset to a fully recovered state between sequence presentations to guarantee their independence.
@@
 We define deviance detection as a network's ability to pick out unexpected inputs from an otherwise homogeneous or otherwise unsurprising stream of inputs.
-Following related research \citep{Kubota2021-dx,Harms2014-ah,Jacobsen2001-sc}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the dev sequence to A in the std sequence).
-In other words, we used the msc sequence as a neutral baseline where the target stimulus was presented no more often than in the dev sequence, but no expectation of a competing stimulus B could be formed.
-To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the dev and msc sequences, denoted as $R^A_{dev}$ and $R^A_{msc}$, defining a deviance detection index (DDI) as
+Following related research \citep{Kubota2021-dx,Harms2014-ah,Jacobsen2001-sc}, we were careful to exclude effects due to stimulus identity (i.e., avoiding direct comparisons between different stimuli) or due to adaptation to frequent stimulus presentation (i.e., avoiding direct comparisons of A in the \dev{} sequence to A in the \std{} sequence).
+In other words, we used the \msc{} sequence as a neutral baseline where the target stimulus was presented no more often than in the \dev{} sequence, but no expectation of a competing stimulus B could be formed.
+To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the \dev{} and \msc{} sequences, denoted as $R^A_{dev}$ and $R^A_{msc}$, defining a deviance detection index (DDI) as
 \begin{equation}
@@
 Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step.
-Target trials are taken from the std (400), msc (100), and dev (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section).
+Target trials are taken from the \std{} (400), \msc{} (100), and \dev{} (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section).
 \textbf{b} Response summary, showing average network-wide spike counts in target trials in each sequence.
 The underlying data are the same as in panel a.
-Note how the dev response is very large, followed by an intermediate msc response and a low std response.
+Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
 \textbf{c} Response sizes $R^A_{dev}$ and $R^A_{std}$, normalized by $R^A_{msc}$, across networks and stimuli (n = 120).
@@
 
-To show that our model is capable of deviance detection, we first show the responses of the sample network in all conditions (dev, msc, and std sequences) in Figure \ref{fig2}a.
+To show that our model is capable of deviance detection, we first show the responses of the sample network in all conditions (\dev{}, \msc{}, and \std{} sequences) in Figure \ref{fig2}a.
 Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color.
 All responses share an early activity pattern, which derives from the stimulated neurons and their immediate postsynaptic targets, but differences soon start to emerge.
-The dev sequence responses, shown at the top of the plot, exhibit a clear peak in activity around 15 ms after stimulation.
-The std sequence responses show a later, more diffuse activity peak around 20 ms, and the msc sequence responses lack any obvious activity cluster.
+The \dev{} sequence responses, shown at the top of the plot, exhibit a clear peak in activity around 15 ms after stimulation.
+The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster.
 We sum these patterns up in Figure \ref{fig2}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
-Clearly, the target response in the dev sequence is greater than either the msc or the std response.
+Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.
 
@@
 To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across conditions.
-Figure \ref{fig2}c shows a summary of the average dev and std response sizes, normalized by the response in the msc sequence.
-The data clearly confirmed that dev responses were larger than msc (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and std responses smaller (T = 687, p = 6.42e-15).
+Figure \ref{fig2}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
+The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15).
 Accordingly, the DDI (Figure \ref{fig2}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
@@
 In the remainder of the paper, we will proceed to detail the processes leading to deviance detection on a sample network, then confirm the highlighted trends statistically with reference to the full set of 120 networks and stimulus assignments.
-The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the msc sequence and calculated z-scores with respect to the population.
-Considering only networks whose absolute z-score was less than 1 for both A and B responses, we hand-picked a sample network and stimulus assignment that exhibited both an increased DDI in STD+TA over TA only, and greater msc than std responses in both STD+TA and TA only conditions.
+The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the \msc{} sequence and calculated z-scores with respect to the population.
+Considering only networks whose absolute z-score was less than 1 for both A and B responses, we hand-picked a sample network and stimulus assignment that exhibited both an increased DDI in STD+TA over TA only, and greater \msc{} than \std{} responses in both STD+TA and TA only conditions.
 
@@
 With no plasticity, the responses are indistinguishable between sequences, yielding a DDI of 0, and indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
-With STD only, the response in the dev sequence developed earlier and very slightly larger, yielding a DDI of 0.03.
-With TA only, the response became noticeably smaller, but the sequences became readily distinguishable, with high dev, intermediate msc, and low std responses, yielding a DDI of 0.19.
-Finally, in the full model, the dev response stood out more clearly, leaving much smaller msc and std responses and yielding a DDI of 0.4.
+With STD only, the response in the \dev{} sequence developed earlier and very slightly larger, yielding a DDI of 0.03.
+With TA only, the response became noticeably smaller, but the sequences became readily distinguishable, with high \dev{}, intermediate \msc{}, and low \std{} responses, yielding a DDI of 0.19.
+Finally, in the full model, the \dev{} response stood out more clearly, leaving much smaller \msc{} and \std{} responses and yielding a DDI of 0.4.
 
@@
 In the following, we will first build an understanding of how TA leads to deviance detection when acting alone, i.e.
-in the ablated TA only model, working backwards from the increased target response in dev, identifying and tracing causal factors step by step.
+in the ablated TA only model, working backwards from the increased target response in \dev{}, identifying and tracing causal factors step by step.
 Then, we will build on this understanding to elucidate how the addition of STD, which was shown above to be ineffective for deviance detection on its own, can lead to an increased deviant response in the full model.
@@
 
-To understand why the same stimulus A evoked a higher response in dev than in msc with TA only, we first compared the average responses in the two conditions, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
+To understand why the same stimulus A evoked a higher response in \dev{} than in \msc{} with TA only, we first compared the average responses in the two conditions, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
 As shown in Figure \ref{fig4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both conditions.
-The contrast (right) revealed a tendency of dev spikes to occur earlier, particularly shortly after stimulation, as shown by the succession of orange (higher dev activity) by blue (higher msc activity) in the lower parts of the panel.
-In later parts of the response (top half of the panel), the dev response was larger, but not obviously shifted in time.
+The contrast (right) revealed a tendency of \dev{} spikes to occur earlier, particularly shortly after stimulation, as shown by the succession of orange (higher \dev{} activity) by blue (higher \msc{} activity) in the lower parts of the panel.
+In later parts of the response (top half of the panel), the \dev{} response was larger, but not obviously shifted in time.
 Since the stimulus was identical in both conditions, and TA was the only plasticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.
@@
 \end{figure*}
-\textbf{Fig. 4} Responses to dev target trials are larger due to lower $V_{TA}$.
+\textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower $V_{TA}$.
 \textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
-Left and middle column, target trials in dev sequence and msc sequence, respectively; right column, contrast between these two.
+Left and middle column, target trials in \dev{} sequence and \msc{} sequence, respectively; right column, contrast between these two.
 \textbf{b} Target trial average of the threshold adaptation voltage $V_{TA}$, with neuron order and columns as in panel a.
@@
 The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the colormap.
-\textbf{d} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
-\textbf{e} Network-wide median contrast (dev - msc) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta V_{TA}$ vs response size contrast across excitatory neurons).
+\textbf{d} Relationship across excitatory neurons of the sample network between the contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
+\textbf{e} Network-wide median contrast (\dev{} - \msc{}) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta V_{TA}$ vs response size contrast across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
-\textbf{f} In the sample network, contribution of the contrast in $V_{TA}$ and $V_m$ to bins with increased firing in dev (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
+\textbf{f} In the sample network, contribution of the contrast in $V_{TA}$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
 \textbf{g} Contribution of $V_{TA}$ relative to the sum of the absolute values of the weighted contributions in panel f.
@@
 We quantified the former with the strength of adaptation $V_{TA}$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
-As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target dev trials relative to msc, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
-On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in dev (right panel, later orange portions), since they spiked more often.
+As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
 Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $V_{TA}$, leading to numerous horizontal breaks in the plot.
@@
 As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
-The contrast (right panel) shows not only a notable relative increase in membrane voltages in dev around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
+The contrast (right panel) shows not only a notable relative increase in membrane voltages in \dev{} around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
 
-Based on these measures, we then asked why there were more spikes in dev trials.
-We first focused on the contribution of $V_{TA}$ alone in an effort to confirm that thresholds are lower in dev, and that this reduction in thresholds causes increased firing.
+Based on these measures, we then asked why there were more spikes in \dev{} trials.
+We first focused on the contribution of $V_{TA}$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
 We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response of each excitatory neuron in terms of spikes fired.
-In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in dev than msc (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Median $V_{TA}$ was lower in dev than msc (T = 514, p = 1.67e-16, median = -0.094 mV; one-sided Wilcoxon, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
+In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Median $V_{TA}$ was lower in \dev{} than \msc{} (T = 514, p = 1.67e-16, median = -0.094 mV; one-sided Wilcoxon, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
 This indicates that networks were less adapted during oddball sequences, and that this lack of adaptation played an important role in the increased response to deviant inputs.
@@
 Yet, as we saw in Figure \ref{fig4}c, the membrane voltages also differed between the two conditions.
-In order to assess the relative contribution of $V_{TA}$ to the increased dev response, we chose to focus exclusively on time points with higher firing probability in the dev sequence, discarding data points with equal or lower firing probability relative to the msc sequence.
+In order to assess the relative contribution of $V_{TA}$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
 We then weighted the remaining contrasts in $V_{TA}$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
-The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased dev response.
+The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slighly lowered throughout the response, the membrane potential was dramatically increased.
@@
 
-Having established that the increased response to A in dev trials was driven by lower thresholds, we next turned to the cause of this reduction.
-A priori, since threshold adaptation is a direct reflection of activity, and since the response to target trials was larger in dev than in msc, the lower thresholds in dev must have been caused by lower activity in non-target trials, i.e.
-lower responses to B in dev than to B through E in msc.
-Yet, with activity modulated by thresholds as shown above, we might expect to observe higher dev responses regardless of stimulus.
+Having established that the increased response to A in \dev{} trials was driven by lower thresholds, we next turned to the cause of this reduction.
+A priori, since threshold adaptation is a direct reflection of activity, and since the response to target trials was larger in \dev{} than in \msc{}, the lower thresholds in \dev{} must have been caused by lower activity in non-target trials, i.e.
+lower responses to B in \dev{} than to B through E in \msc{}.
+Yet, with activity modulated by thresholds as shown above, we might expect to observe higher \dev{} responses regardless of stimulus.
 To resolve this apparent paradox, we will turn to an analysis of the spatial arrangement of activity and adaptation.
@@
 \end{figure*}
-\textbf{Fig. 5} Lower $V_{TA}$ in dev is a consequence of lower dev non-target response magnitude.
+\textbf{Fig. 5} Lower $V_{TA}$ in \dev{} is a consequence of lower \dev{} non-target response magnitude.
 \textbf{a} $V_{TA}$ at the start of A trials (mean over trials) in the sample network, with each neuron shown in its spatial location.
-Left and middle column, raw averages for dev and msc sequences, respectively; right column, contrast between dev and msc averages.
-A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the msc sequence.
-\textbf{b} Average response (in spikes per trial) to non-target trials (B in dev, B, C, D, and E in msc). Columns are arranged as in panel a.
-\textbf{c} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in average number of spikes fired in response to non-target stimulation and the contrast in $V_{TA}$ at the start of target trials, and associated linear regression.
-\textbf{d} Network-wide median contrast (dev - msc) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\Delta V_{TA}$ across excitatory neurons).
+Left and middle column, raw averages for \dev{} and \msc{} sequences, respectively; right column, contrast between \dev{} and \msc{} averages.
+A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the \msc{} sequence.
+\textbf{b} Average response (in spikes per trial) to non-target trials (B in \dev{}, B, C, D, and E in \msc{}). Columns are arranged as in panel a.
+\textbf{c} Relationship across excitatory neurons of the sample network between the contrast (\dev{} - \msc{}) in average number of spikes fired in response to non-target stimulation and the contrast in $V_{TA}$ at the start of target trials, and associated linear regression.
+\textbf{d} Network-wide median contrast (\dev{} - \msc{}) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\Delta V_{TA}$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 First, we mapped the average value of $V_{TA}$ at the start of target trials to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
-In the msc sequence, $V_{TA}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
-Conversely, in the dev sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
+In the \msc{} sequence, $V_{TA}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
+Conversely, in the \dev{} sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
 
@@
 As expected, the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds.
-Notably, although the comparison is between 400 B trials in dev, and only 100 B trials (alongside the same number of C, D and E trials) in msc, only a small handful of neurons responded more to non-target stimulation in dev, most of these very close to the stimulation site of B.
+Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.
 
-Correlating the non-target response contrast (dev - msc) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in dev almost perfectly predicted the resulting reduction in TA levels (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between msc and dev sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the non-target trial response is the primary driver of TA.
+Correlating the non-target response contrast (\dev{} - \msc{}) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the non-target trial response is the primary driver of TA.
 
 Finally, we examined the cause of the reduced response in non-target trials.
-Going from the msc sequence to the dev sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation.
+Going from the \msc{} sequence to the \dev{} sequence, two things change: Firstly, stimulations with non-target C, D and E are replaced with B, which would cause different responses even in the absence of adaptation.
 Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
-We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and comparatively strongly to stimulation in B, and therefore more to non-target stimuli in dev than msc, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and comparatively strongly to stimulation in B, and therefore more to non-target stimuli in \dev{} than \msc{}, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
 
@@
 Note that non-responding neurons are ranked last and appear dark blue.
-\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (dev - msc) and the contrast in $V_{TA}$ at the start of B trials, and associated linear regression.
+\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (\dev{} - \msc{}) and the contrast in $V_{TA}$ at the start of B trials, and associated linear regression.
 Neurons are colored by latency rank as in panel a.
@@
 Inset histograms show the distributions of the corresponding values across these two categories, and the confidence intervals of their medians, calculated by bootstrap with 10000 iterations.
-\textbf{c} Median non-target response contrast (dev - msc) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labelled "late", across networks and stimuli.
-\textbf{c} Median contrast (dev - msc) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli
+\textbf{c} Median non-target response contrast (\dev{} - \msc{}) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labelled "late", across networks and stimuli.
+\textbf{c} Median contrast (\dev{} - \msc{}) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli
 \textbf{d} Relationship across networks and stimuli between the median contrast in $V_{TA}$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B in any condition, and a late portion, containing the remaining 900 neurons.
-In Figure \ref{fig6}b, we related the non-target response contrast (dev - msc) to the contrast in $V_{TA}$ at the start of B trials.
+In Figure \ref{fig6}b, we related the non-target response contrast (\dev{} - \msc{}) to the contrast in $V_{TA}$ at the start of B trials.
 Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993) echoed the related finding in Figure \ref{fig5}c.
-As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the dev sequence, while the response of the network as a whole decreased.
+As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the \dev{} sequence, while the response of the network as a whole decreased.
 Correspondingly, the TA levels of this portion of the network also tended to increase, in clear contrast to the network as a whole.
@@
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $V_{TA}$ at the start of B trials (Figure \ref{fig6}d) within these two portions.
-We found that the response in the early portion was consistently greater in dev than msc (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
-The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in dev (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
+The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
 Finally, to show that early $V_{TA}$ was causal in reducing the response to B, we correlated the median contrast in $V_{TA}$ among the early portion of neurons with the network-wide median contrast of the response to stimulation in B, as shown in Figure \ref{fig6}e.
@@
 To sum up, we showed in Figure \ref{fig4} that greater $V_{TA}$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response.
-Here, we showed that the neurons responding earliest to B exhibit higher $V_{TA}$ in the dev sequence as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
+Here, we showed that the neurons responding earliest to B exhibit higher $V_{TA}$ in the \dev{} sequence as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
 
-Thus, we have finally traced the greater response to A in the dev sequence to its roots, and revealed two complementary roles of TA: The frequent presentation of non-target stimulus B in the dev sequence and concomitant increased activity in a subset of neurons responding early to B causes \emph{local suppression} by TA in this subset, leading to smaller network responses.
-By contrast, stimuli in the msc sequence are presented more sporadically, allowing local suppression to decay, thus leading to larger network responses.
-This difference in non-target response sizes -- small in dev, large in msc -- is then reflected in network-wide \emph{global suppression}, such that the network is more excitable to stimulation with A (or other non-standard inputs) in the dev sequence than in the control context.
+Thus, we have finally traced the greater response to A in the \dev{} sequence to its roots, and revealed two complementary roles of TA: The frequent presentation of non-target stimulus B in the \dev{} sequence and concomitant increased activity in a subset of neurons responding early to B causes \emph{local suppression} by TA in this subset, leading to smaller network responses.
+By contrast, stimuli in the \msc{} sequence are presented more sporadically, allowing local suppression to decay, thus leading to larger network responses.
+This difference in non-target response sizes -- small in \dev{}, large in \msc{} -- is then reflected in network-wide \emph{global suppression}, such that the network is more excitable to stimulation with A (or other non-standard inputs) in the \dev{} sequence than in the control context.
 
@@
 On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (data not shown).
-However, this reduction was not uniform: Target responses were reduced most in the std sequence, and least in the dev sequence.
+However, this reduction was not uniform: Target responses were reduced most in the \std{} sequence, and least in the \dev{} sequence.
 
@@
 \textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
-\textbf{a} Median contrast (dev - msc) in the non-target response (left) and in $V_{TA}$ at the start of B trials (right) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+\textbf{a} Median contrast (\dev{} - \msc{}) in the non-target response (left) and in $V_{TA}$ at the start of B trials (right) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
 Compare to Figure \ref{fig6}c-d.
-\textbf{b} Median contrast (dev - msc) in the non-target response magnitude across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
-The red line is the identity function; points above this line represent networks with greater response reduction (dev - msc) in the full model than in the ablated model.
+\textbf{b} Median contrast (\dev{} - \msc{}) in the non-target response magnitude across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
+The red line is the identity function; points above this line represent networks with greater response reduction (\dev{} - \msc{}) in the full model than in the ablated model.
 The subtraction of the two quantities in the box plot below shows that the non-target response tends to be reduced more in the full model.
@@
 Compare to Figure \ref{fig5}d.
-\textbf{d} Median contrast (dev - msc) in $V_{TA}$ at the start of target trials across networks and stimuli in the full and ablated models, arranged as in panel b.
+\textbf{d} Median contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials across networks and stimuli in the full and ablated models, arranged as in panel b.
 The subtraction of the two quantities in the box plot below shows that $V_{TA}$ tends to be reduced more in the full model.
@@
 Compare to Figure \ref{fig4}e.
-\textbf{f} Median contrast (dev - msc) in the target response magnitude across networks and stimuli in the full and ablated models, arranged as in panel b.
+\textbf{f} Median contrast (\dev{} - \msc{}) in the target response magnitude across networks and stimuli in the full and ablated models, arranged as in panel b.
 The subtraction of the two quantities in the box plot below shows that the target response tends to be increased more in the full model
@@
 Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}a the full model equivalent of Figure \ref{fig6}c and d.
-The contrasts (dev - msc) in the non-target response magnitude and in the resulting $V_{TA}$ levels at the start of B trials both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the dev sequence.
+The contrasts (\dev{} - \msc{}) in the non-target response magnitude and in the resulting $V_{TA}$ levels at the start of B trials both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
 Likewise, the correlation across excitatory neurons between these two contrasts remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992).
@@
 Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, the non-target response contrast remained tightly correlated with the contrast (dev - msc) in $V_{TA}$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
-Correspondingly, $V_{TA}$ was reduced in dev, as shown in Figure \ref{fig7}d, and this too by a slighly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Indeed, the non-target response contrast remained tightly correlated with the contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
+Correspondingly, $V_{TA}$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slighly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
 Finally, based on the ablated model analysis in Figure \ref{fig4}, we sought to confirm that lowered $V_{TA}$ before target trials was responsible for the increased target response.
-As shown in Figure \ref{fig7}e, the contrasts (dev - msc) in $V_{TA}$ at the start of target trials and in response magnitude to A were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response was clearly increased in dev over msc, as shown in Figure \ref{fig7}f, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians differing by 94\%.
+As shown in Figure \ref{fig7}e, the contrasts (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials and in response magnitude to A were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response was clearly increased in \dev{} over \msc{}, as shown in Figure \ref{fig7}f, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians differing by 94\%.
 This greater contrast, of course, directly reflects the greater DDI reported in Figure \ref{fig3}.
@@
 
-This naturally raises the question of why, with the addition of STD, the non-target dev response was reduced more than non-target msc responses.
+This naturally raises the question of why, with the addition of STD, the non-target \dev{} response was reduced more than non-target \msc{} responses.
 We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.
-This would mean that STD is capable only of local suppression, reducing the standard response, but not of global suppression, leaving networks with only STD unable to differentiate between dev and msc target trials.
+This would mean that STD is capable only of local suppression, reducing the standard response, but not of global suppression, leaving networks with only STD unable to differentiate between \dev{} and \msc{} target trials.
 
@@
 
-To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated the number of spikes in dev target trials with $x_j$ at the start of later trials (cf.
+To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated the number of spikes in \dev{} target trials with $x_j$ at the start of later trials (cf.
 Equation \ref{eqn-xsyn}) across excitatory neurons.
@@
 
-To show that STD being carried over between repeated trials was responsible for the reduction of non-target dev pulses relative to msc, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
+To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} pulses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
 Unlike $V_{TA}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
@@
 \end{figure*}
-\textbf{Fig. 8} In the full model, STD reduces dev non-target and increases dev target responses relative to msc.
-\textbf{a} Average depression $D$ in target (A) trials (top row) and B trials (bottom row) in the dev (left) and msc sequence (center), and their contrast (right) in the sample network.
-The color scale is shared across columns (and between dev and msc) for direct comparability.
-\textbf{b} Relationship across neurons of the sample network between the contrast (dev - msc) in $D^A$ and the contrast in the target response, and associated linear regression.
-\textbf{c} Median contrast (dev - msc) in $D^B$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
+\textbf{Fig. 8} In the full model, STD reduces \dev{} non-target and increases \dev{} target responses relative to \msc{}.
+\textbf{a} Average depression $D$ in target (A) trials (top row) and B trials (bottom row) in the \dev{} (left) and \msc{} sequence (center), and their contrast (right) in the sample network.
+The color scale is shared across columns (and between \dev{} and \msc{}) for direct comparability.
+\textbf{b} Relationship across neurons of the sample network between the contrast (\dev{} - \msc{}) in $D^A$ and the contrast in the target response, and associated linear regression.
+\textbf{c} Median contrast (\dev{} - \msc{}) in $D^B$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
 \textbf{d} Relationship across networks and stimuli between the median contrast in $D^B$ among the early portion and the median contrast in the response to B among the entire network, and associated linear regression.
@@
 
-As noted above, we hypothesized that, in the full model, STD is directly responsible only for a reduction of non-target dev trials, relative to msc, whereas differences between target dev and msc trials are not a result of STD directly.
-To show this, we calculated $D$ for A and B trials, and contrasted it across dev and msc conditions as with $V_{TA}$ before.
+As noted above, we hypothesized that, in the full model, STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly.
+To show this, we calculated $D$ for A and B trials, and contrasted it across \dev{} and \msc{} conditions as with $V_{TA}$ before.
 In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both conditions.
-The msc sequence, however, clearly evoked greater depression, as borne out in the contrast (dev - msc), which was negative throughout.
+The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast (\dev{} - \msc{}), which was negative throughout.
 Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both conditions.
-In the msc sequence, we can attribute the difference between A and B primarily to different response patterns: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network.
-However, in the dev sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited.
-As a result, the contrast (dev - msc) showed slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicating greatly increased synaptic depression in dev.
+In the \msc{} sequence, we can attribute the difference between A and B primarily to different response patterns: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network.
+However, in the \dev{} sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited.
+As a result, the contrast (\dev{} - \msc{}) showed slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicating greatly increased synaptic depression in \dev{}.
 
-To show that $D$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated the contrast (dev - msc) in $D^A$ against the target response contrast across neurons of the sample network.
+To show that $D$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated the contrast (\dev{} - \msc{}) in $D^A$ against the target response contrast across neurons of the sample network.
 Given the relatively large responses in the contrasted conditions, and the infrequent presentation of these trials, allowing causality to mostly flow from $D$ to response, rather than also in reverse, we are confident that this relationship roughly reflects a causal influence.
@@
 First, in Figure \ref{fig8}c, we recapitulate the logic established in Figure \ref{fig6}: Given the spatial clustering of $D^B$, we surmised a similar local suppression effect as with TA.
-We found that, indeed, neurons responding early to stimulation in B showed greater $D^B$ in dev than msc (T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced $D^B$ (T = 2.24e+03, p = 0.000133).
-We can relate this directly to the non-target response contrasts in the early and late portions shown in Figure \ref{fig7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in dev causes local suppression in terms of $D$.
+We found that, indeed, neurons responding early to stimulation in B showed greater $D^B$ in \dev{} than \msc{} (T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced $D^B$ (T = 2.24e+03, p = 0.000133).
+We can relate this directly to the non-target response contrasts in the early and late portions shown in Figure \ref{fig7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local suppression in terms of $D$.
 % Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
@@
 As we show in Figure \ref{fig8}d, higher $\Delta D^B$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000).
-This confirms again the integrity of the measure $D$, but also that STD-mediated local suppression played a role in reducing the non-target response in the dev sequence.
+This confirms again the integrity of the measure $D$, but also that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
 
-Finally, given the very clear contrast in $D^A$ in the sample network, we considered the possibility that, in the context of the full model, STD could also evoke global suppression in the msc sequence, contrary to our stated hypothesis.
-If so, we should expect to see less $D^A$ in dev than msc.
-To confirm this, we analysed the contrast (dev - msc) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120).
-In addition, as shown in Figure \ref{fig8}e, we correlated the contrast in $D^A$ with the target response contrast across neurons in each network, and found that weaker depression in dev predicted a greater response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
+Finally, given the very clear contrast in $D^A$ in the sample network, we considered the possibility that, in the context of the full model, STD could also evoke global suppression in the \msc{} sequence, contrary to our stated hypothesis.
+If so, we should expect to see less $D^A$ in \dev{} than \msc{}.
+To confirm this, we analysed the contrast (\dev{} - \msc{}) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120).
+In addition, as shown in Figure \ref{fig8}e, we correlated the contrast in $D^A$ with the target response contrast across neurons in each network, and found that weaker depression in \dev{} predicted a greater response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
 We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).
-Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in msc than dev in terms of $D$, this difference appears to manifest only weakly in activity differences.
+Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in \msc{} than \dev{} in terms of $D$, this difference appears to manifest only weakly in activity differences.
 
-To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction of the response to the frequently presented standard stimuli (dev B) by local suppression, and an increase in the response to deviant stimuli (dev A) by relief from global suppression, relative to control.
-Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in dev B responses also entailed lower $V_{TA}$, which independently lead to higher dev A responses.
+To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction of the response to the frequently presented standard stimuli (\dev{} B) by local suppression, and an increase in the response to deviant stimuli (\dev{} A) by relief from global suppression, relative to control.
+Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in \dev{} B responses also entailed lower $V_{TA}$, which independently lead to higher \dev{} A responses.
 Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
@@
 To see this, compare the change in the $V_{TA}$ contrast at the start of A trials as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta D^A$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
-Conversely, notice that, since TA and STD work in similar fashion to increase the dev target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
+Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
 Thus, this synergy allows the individually weak STD to exert a significantly larger effect when paired with the stronger TA.

### 2023-01-06 paper-1/manuscript.tex
Rev: Introduce sample network in methods

@@
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-Figure \ref{fig1}b shows a representative example of the resulting network structure, which we will be referring to as the sample network throughout this paper.
-In Figure \ref{fig1}c, we show the distance from the neurons stimulated at site 1 to each neuron in the network, in terms of the minimum number of excitatory synapses to reach the target, in order to give an intuition for how excitation may spread.
-In addition to the sample network, we generated additional networks for statistics purposes.
+Figure \ref{fig1}b shows a representative example of the resulting network structure.
+In Figure \ref{fig1}c, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
 Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response: Networks where fewer than 500 neurons responded to stimulation at any of the five sites after full recovery were discarded.
@@
 To turn off TA, we replaced Equation \ref{eqn-TA} with a constant $V_{TA} = 0$, making the firing threshold completely unadaptive.
+
+\subsection{Sample network}\label{sec-sample}
+
+We present most of our analysis at two levels: We first detail the processes leading to deviance detection on a sample network, then confirm the highlighted trends statistically with reference to the full set of 120 networks and stimulus assignments.
+The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the \msc{} sequence and calculated z-scores with respect to the population.
+We considered only networks whose absolute z-score was less than 1 for both A and B responses, then hand-picked a sample network and stimulus assignment that exhibited both an increased DDI (see Equation \ref{eqn-ddi} below) in STD+TA over TA only, and greater \msc{} than \std{} responses in both STD+TA and TA only conditions.
 
@@
 
-Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-In the remainder of the paper, we will proceed to detail the processes leading to deviance detection on a sample network, then confirm the highlighted trends statistically with reference to the full set of 120 networks and stimulus assignments.
-The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the \msc{} sequence and calculated z-scores with respect to the population.
-Considering only networks whose absolute z-score was less than 1 for both A and B responses, we hand-picked a sample network and stimulus assignment that exhibited both an increased DDI in STD+TA over TA only, and greater \msc{} than \std{} responses in both STD+TA and TA only conditions.
-
 \begin{figure*}%
@@
 
+Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
 Figure \ref{fig3}a shows the responses to target stimulation of the sample network under ablation.

### 2023-01-06 paper-1/manuscript.tex
Rev: Drop F4 mention of temporal shift of response

@@
 As shown in Figure \ref{fig4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both conditions.
-The contrast (right) revealed a tendency of \dev{} spikes to occur earlier, particularly shortly after stimulation, as shown by the succession of orange (higher \dev{} activity) by blue (higher \msc{} activity) in the lower parts of the panel.
-In later parts of the response (top half of the panel), the \dev{} response was larger, but not obviously shifted in time.
+The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, indicating higher \dev{} activity.
 Since the stimulus was identical in both conditions, and TA was the only plasticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.

### 2023-01-06 paper-1/manuscript.tex
Rev: Clarify F4d conclusion

@@
 We first focused on the contribution of $V_{TA}$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response of each excitatory neuron in terms of spikes fired.
-In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: Median $V_{TA}$ was lower in \dev{} than \msc{} (T = 514, p = 1.67e-16, median = -0.094 mV; one-sided Wilcoxon, n = 120), and the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
-This indicates that networks were less adapted during oddball sequences, and that this lack of adaptation played an important role in the increased response to deviant inputs.
+We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response in terms of spikes fired across excitatory neurons.
+In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: The median $V_{TA}$ among excitatory neurons was lower in \dev{} than \msc{} (T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+In addition, the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
+This indicates that neurons with a greater decrease in $V_{TA}$ in \dev{} tended to respond more strongly to target stimuli.
 
@@
 The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased \dev{} response.
-While thresholds were slighly lowered throughout the response, the membrane potential was dramatically increased.
+While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
 In relative terms (Figure \ref{fig4}g), $V_{TA}$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.

### 2023-01-06 paper-1/manuscript.tex
Rev: Clarify dev low TA = high A != low B paradox

@@
 
-Having established that the increased response to A in \dev{} trials was driven by lower thresholds, we next turned to the cause of this reduction.
+Having established that the increased response to A in \dev{} trials was driven by lower thresholds before stimulation, we next turned to the cause of this reduction.
 A priori, since threshold adaptation is a direct reflection of activity, and since the response to target trials was larger in \dev{} than in \msc{}, the lower thresholds in \dev{} must have been caused by lower activity in non-target trials, i.e.
 lower responses to B in \dev{} than to B through E in \msc{}.
-Yet, with activity modulated by thresholds as shown above, we might expect to observe higher \dev{} responses regardless of stimulus.
+Yet, since activity is modulated by thresholds as shown above, and since the dev sequence was random, meaning that thresholds before stimulation could carry no information about the upcoming stimulus, we might expect the lowered thresholds in \dev{} to yield not only higher target responses, but also higher non-target responses.
 To resolve this apparent paradox, we will turn to an analysis of the spatial arrangement of activity and adaptation.

### 2023-01-06 paper-1/manuscript.tex
Rev: Stronger F5d conclusion

@@
 
-Correlating the non-target response contrast (\dev{} - \msc{}) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the non-target trial response is the primary driver of TA.
+Correlating the non-target response contrast (\dev{} - \msc{}) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
 

### 2023-01-06 paper-1/manuscript.tex
Rev: F6 caption and text

@@
 Note that non-responding neurons are ranked last and appear dark blue.
-\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (\dev{} - \msc{}) and the contrast in $V_{TA}$ at the start of B trials, and associated linear regression.
+\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (\dev{} - \msc{}) and the contrast in $V_{TA}$ at the start of B trials, and associated linear regression across all excitatory neurons.
 Neurons are colored by latency rank as in panel a.
-Additionally, the earliest 100 responders are plotted with crosses, while later responders are plotted with dots.
-Inset histograms show the distributions of the corresponding values across these two categories, and the confidence intervals of their medians, calculated by bootstrap with 10000 iterations.
-\textbf{c} Median non-target response contrast (\dev{} - \msc{}) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labelled "late", across networks and stimuli.
-\textbf{c} Median contrast (\dev{} - \msc{}) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli
+Additionally, the earliest 100 responders are shown in the left plot, while later responders are shown in the right plot.
+Inset histograms show the distributions of the plotted values, with the confidence intervals of their medians, calculated by bootstrap with 10000 iterations, in darker color.
+\textbf{c} Median non-target response contrast (\dev{} - \msc{}) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labeled "late", across networks and stimuli.
+\textbf{c} Median contrast (\dev{} - \msc{}) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli.
 \textbf{d} Relationship across networks and stimuli between the median contrast in $V_{TA}$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
 
-To see this in the sample network, consider first the latency of responses to B across all conditions, shown by way of rank order in Figure \ref{fig6}a.
+To show this, we sought to isolate this subset of early responders to B.
+In a given network and stimulus assignment, we found, for each neuron, the minimum response time (i.e., time to first spike) across B trials in \dev{}, \msc{} and \std{} sequences, and ranked neurons by this response time.
+The ranking in the sample network is shown in Figure \ref{fig6}a.
 We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site.
 We did notice that the early response appeared to gravitate towards the stimulus site for A, but confirmed that this was the case in all conditions separately, indicating a structural rather than dynamic cause (data not shown).
-We then notionally split the network into an early portion, containing the first 100 neurons to respond to B in any condition, and a late portion, containing the remaining 900 neurons.
+We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
 In Figure \ref{fig6}b, we related the non-target response contrast (\dev{} - \msc{}) to the contrast in $V_{TA}$ at the start of B trials.
-Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993) echoed the related finding in Figure \ref{fig5}c.
-As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the \dev{} sequence, while the response of the network as a whole decreased.
-Correspondingly, the TA levels of this portion of the network also tended to increase, in clear contrast to the network as a whole.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated the same response contrast to $\Delta V_{TA}$ before A.
+As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the \dev{} sequence, while the response of the later portion of the network decreased.
+Correspondingly, the TA levels of the early portion also tended to increase, in clear contrast to the later portion.
 
@@
 The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
+
 Finally, to show that early $V_{TA}$ was causal in reducing the response to B, we correlated the median contrast in $V_{TA}$ among the early portion of neurons with the network-wide median contrast of the response to stimulation in B, as shown in Figure \ref{fig6}e.

### 2023-01-06 paper-1/manuscript.tex
Rev: F7 text minor

@@
 Indeed, the non-target response contrast remained tightly correlated with the contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
-Correspondingly, $V_{TA}$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slighly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Correspondingly, $V_{TA}$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
@@
 We hypothesized that, due to its relatively short time constant, the direct effect of STD across trials was limited to repeated trials of the same stimulus, such as the frequent stimulus in oddball sequences.
-This would mean that STD is capable only of local suppression, reducing the standard response, but not of global suppression, leaving networks with only STD unable to differentiate between \dev{} and \msc{} target trials.
+This would mean that STD is capable only of local suppression, reducing the non-target response in \dev{}, but not of global suppression, leaving networks with only STD unable to differentiate between \dev{} and \msc{} target trials.
 
@@
 
-To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} pulses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
+To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} responses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
 Unlike $V_{TA}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.

### 2023-01-06 paper-1/manuscript.tex
Rev: Tighten explanation of mean R for D

@@
 A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
-Here, we used the average trial response $R_j^S$ of neuron $j$ to a given stimulus $S$, in number of spikes, as a measure of the presynaptic drive.
+In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $R_j^S$ of a neuron $j$ to a given stimulus $S$ across all sequences as a measure of the presynaptic drive.
 Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.
-Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $R_j^S k x_j^S$.
-For this estimate, we ignored added depression due to repeated spikes within a trial, instead defining $x_j^S$ as the mean depression at the start of trials.
+Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $R_j^S k x_j^S$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials.
 Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
@@
 Notice that $D$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $V_{TA}$, albeit more approximate.
-Lastly, since we contrasted this measure between conditions (see below), we aimed to avoid effects purely due to differences in presynaptic activity.
-Therefore, we calculated $R^S$ as the average response across $S$ trials in all conditions, leaving the contrast entirely in the hands of $x^S$.
 

### 2023-01-06 paper-1/manuscript.tex
Rev F8

@@
 \textbf{b} Relationship across neurons of the sample network between the contrast (\dev{} - \msc{}) in $D^A$ and the contrast in the target response, and associated linear regression.
-\textbf{c} Median contrast (\dev{} - \msc{}) in $D^B$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
-\textbf{d} Relationship across networks and stimuli between the median contrast in $D^B$ among the early portion and the median contrast in the response to B among the entire network, and associated linear regression.
-\textbf{e} Network-wide median contrast in $D^A$, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response, across networks and stimuli.
-The plot shows both that the negative correlation shown in panel b is robust (most networks are plotted with red markers, i.e.
-significantly negatively correlated), and that $\Delta D^A$ is less than 0 on average (see the inset histogram on the horizontal axis)
+\textbf{c} Network-wide median contrast in $D^A$, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response, across networks and stimuli.
+\textbf{d} Median contrast (\dev{} - \msc{}) in $D^B$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
+\textbf{e} Relationship across networks and stimuli between the median contrast in $D^B$ among the early portion and the median contrast in the response to B among the entire network, and associated linear regression
 
@@
 To show that $D$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated the contrast (\dev{} - \msc{}) in $D^A$ against the target response contrast across neurons of the sample network.
-Given the relatively large responses in the contrasted conditions, and the infrequent presentation of these trials, allowing causality to mostly flow from $D$ to response, rather than also in reverse, we are confident that this relationship roughly reflects a causal influence.
-As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that $D$ reduces responses as expected.
+Notice that this mirrors the correlation shown in Figure \ref{fig4} between $\Delta V_{TA}$ before A and the response contrast.
+As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $D$ causes reduced responses as expected.
+We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
+We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).
 
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in Figure \ref{fig8}c, we recapitulate the logic established in Figure \ref{fig6}: Given the spatial clustering of $D^B$, we surmised a similar local suppression effect as with TA.
+First, in Figure \ref{fig8}d, we recapitulate the logic established in Figure \ref{fig6}: Given the spatial clustering of $D^B$, we surmised a similar local suppression effect as with TA.
 We found that, indeed, neurons responding early to stimulation in B showed greater $D^B$ in \dev{} than \msc{} (T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced $D^B$ (T = 2.24e+03, p = 0.000133).
@@
 To show that increased local suppression by STD reduces the non-target response, we correlated the median contrast in $D^B$ among the early portion of networks with the median contrast of the total response to B, echoing Figure \ref{fig6}e.
-As we show in Figure \ref{fig8}d, higher $\Delta D^B$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000).
+As we show in Figure \ref{fig8}e, higher $\Delta D^B$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000).
 This confirms again the integrity of the measure $D$, but also that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
@@
 If so, we should expect to see less $D^A$ in \dev{} than \msc{}.
-To confirm this, we analysed the contrast (\dev{} - \msc{}) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120).
-In addition, as shown in Figure \ref{fig8}e, we correlated the contrast in $D^A$ with the target response contrast across neurons in each network, and found that weaker depression in \dev{} predicted a greater response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
-We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).
+To confirm this, we analysed the contrast (\dev{} - \msc{}) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
 Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in \msc{} than \dev{} in terms of $D$, this difference appears to manifest only weakly in activity differences.

### 2023-01-06 paper-1/manuscript.tex
Rev: Concluding paragraph

@@
 
-To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction of the response to the frequently presented standard stimuli (\dev{} B) by local suppression, and an increase in the response to deviant stimuli (\dev{} A) by relief from global suppression, relative to control.
-Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in \dev{} B responses also entailed lower $V_{TA}$, which independently lead to higher \dev{} A responses.
+To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli by local suppression, and an increase in the response to target stimuli by relief from global suppression, relative to \msc{}.
+Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in \dev{} B responses also entailed lower $V_{TA}$, which independently led to higher \dev{} A responses.
 Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.

### 2023-01-06 paper-1/manuscript.tex
Typo

@@
 If so, we should expect to see less $D^A$ in \dev{} than \msc{}.
-To confirm this, we analysed the contrast (\dev{} - \msc{}) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
+To confirm this, we analyzed the contrast (\dev{} - \msc{}) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
 Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in \msc{} than \dev{} in terms of $D$, this difference appears to manifest only weakly in activity differences.

### 2023-01-12 paper-1/manuscript.tex
Fix undefined abbrev DDI

@@
 The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the \msc{} sequence and calculated z-scores with respect to the population.
-We considered only networks whose absolute z-score was less than 1 for both A and B responses, then hand-picked a sample network and stimulus assignment that exhibited both an increased DDI (see Equation \ref{eqn-ddi} below) in STD+TA over TA only, and greater \msc{} than \std{} responses in both STD+TA and TA only conditions.
+We considered only networks whose absolute z-score was less than 1 for both A and B responses, then hand-picked a sample network and stimulus assignment that exhibited both an increased deviance detection index (see Equation \ref{eqn-ddi} below) in STD+TA over TA only, and greater \msc{} than \std{} responses in both STD+TA and TA only conditions.
 

### 2023-01-12 paper-1/manuscript.tex
condition -> sequence for dev/msc/std

@@
 
-To show that our model is capable of deviance detection, we first show the responses of the sample network in all conditions (\dev{}, \msc{}, and \std{} sequences) in Figure \ref{fig2}a.
+To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in Figure \ref{fig2}a.
 Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color.
@@
 We could readily identify similar patterns emerging across most networks and stimulus assignments.
-To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across conditions.
+To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences.
 Figure \ref{fig2}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
@@
 
-To understand why the same stimulus A evoked a higher response in \dev{} than in \msc{} with TA only, we first compared the average responses in the two conditions, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
-As shown in Figure \ref{fig4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both conditions.
+To understand why the same stimulus A evoked a higher response in \dev{} than in \msc{} with TA only, we first compared the average responses in the two sequences, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
+As shown in Figure \ref{fig4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
 The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, indicating higher \dev{} activity.
-Since the stimulus was identical in both conditions, and TA was the only plasticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.
+Since the stimulus was identical in both sequences, and TA was the only plasticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.
 
@@
 \textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower $V_{TA}$.
-\textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and conditions.
+\textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
 Left and middle column, target trials in \dev{} sequence and \msc{} sequence, respectively; right column, contrast between these two.
@@
 
-Yet, as we saw in Figure \ref{fig4}c, the membrane voltages also differed between the two conditions.
+Yet, as we saw in Figure \ref{fig4}c, the membrane voltages also differed between the two sequences.
 In order to assess the relative contribution of $V_{TA}$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
@@
 \textbf{a} The sample network, colored by latency rank, i.e.
-the rank order of the time to first spike in response to stimulation at site B in any condition.
+the rank order of the time to first spike in response to stimulation at site B in any sequence.
 The response develops asymmetrically due to structural properties, not due to interactions with A.
@@
 We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site.
-We did notice that the early response appeared to gravitate towards the stimulus site for A, but confirmed that this was the case in all conditions separately, indicating a structural rather than dynamic cause (data not shown).
+We did notice that the early response appeared to gravitate towards the stimulus site for A, but confirmed that this was the case in all sequences separately, indicating a structural rather than dynamic cause (data not shown).
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
@@
 As noted above, we hypothesized that, in the full model, STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly.
-To show this, we calculated $D$ for A and B trials, and contrasted it across \dev{} and \msc{} conditions as with $V_{TA}$ before.
-In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both conditions.
+To show this, we calculated $D$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $V_{TA}$ before.
+In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
 The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast (\dev{} - \msc{}), which was negative throughout.
-Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both conditions.
+Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both sequences.
 In the \msc{} sequence, we can attribute the difference between A and B primarily to different response patterns: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network.

### 2023-01-12 paper-1/manuscript.tex
Fixed inaccurate description of F4d result

@@
 We first focused on the contribution of $V_{TA}$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response in terms of spikes fired across excitatory neurons.
+We correlated the contrast (dev - msc) of $V_{TA}$ at the start of target trials (i.e., at the left edge of the contrast histogram in Figure \ref{fig4}b) with the target trial response size contrast, across excitatory neurons.
 In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).

### 2023-01-12 paper-1/manuscript.tex
Replaced V_{TA} with \theta (text)

--- 2023-01-12 paper-1/manuscript.tex
@@
 \begin{align}
-    V_\theta &= \theta_0 + V_{TA}(t) \nonumber \\
-    \tau_{\theta} \frac{dV_{TA}}{dt} &= -V_{TA} + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
+    V_\theta &= \theta_0 + \theta(t) \nonumber \\
+    \tau_{\theta} \frac{d\theta}{dt} &= -\theta + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
 \end{align}
@@
 This corresponds to immediate recovery ($\tau_x = 0$) and was done to maintain the magnitude of the excitatory postsynaptic potential (EPSP) in the recovered state regardless of whether STD was turned on or off.
-To turn off TA, we replaced Equation \ref{eqn-TA} with a constant $V_{TA} = 0$, making the firing threshold completely unadaptive.
+To turn off TA, we replaced Equation \ref{eqn-TA} with a constant $\theta = 0$, making the firing threshold completely unadaptive.
 
@@
 \end{figure*}
-\textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower $V_{TA}$.
+\textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower $\theta$.
 \textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
 Left and middle column, target trials in \dev{} sequence and \msc{} sequence, respectively; right column, contrast between these two.
-\textbf{b} Target trial average of the threshold adaptation voltage $V_{TA}$, with neuron order and columns as in panel a.
+\textbf{b} Target trial average of the threshold adaptation voltage $\theta$, with neuron order and columns as in panel a.
 \textbf{c} Target trial average of the membrane potential $V_m$, with neuron order and columns as in panel a.
 The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the colormap.
-\textbf{d} Relationship across excitatory neurons of the sample network between the contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
-\textbf{e} Network-wide median contrast (\dev{} - \msc{}) in $V_{TA}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta V_{TA}$ vs response size contrast across excitatory neurons).
+\textbf{d} Relationship across excitatory neurons of the sample network between the contrast (\dev{} - \msc{}) in $\theta$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
+\textbf{e} Network-wide median contrast (\dev{} - \msc{}) in $\theta$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \theta$ vs response size contrast across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
-\textbf{f} In the sample network, contribution of the contrast in $V_{TA}$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
-\textbf{g} Contribution of $V_{TA}$ relative to the sum of the absolute values of the weighted contributions in panel f.
-\textbf{h} Relative contribution of $V_{TA}$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
+\textbf{f} In the sample network, contribution of the contrast in $\theta$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
+\textbf{g} Contribution of $\theta$ relative to the sum of the absolute values of the weighted contributions in panel f.
+\textbf{h} Relative contribution of $\theta$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
 
-We quantified the former with the strength of adaptation $V_{TA}$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
+We quantified the former with the strength of adaptation $\theta$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
 As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
-Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $V_{TA}$, leading to numerous horizontal breaks in the plot.
+Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $\theta$, leading to numerous horizontal breaks in the plot.
 
-To assess the difference in synaptic currents in a way that allowed direct comparison to $V_{TA}$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
+To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
 As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
@@
 Based on these measures, we then asked why there were more spikes in \dev{} trials.
-We first focused on the contribution of $V_{TA}$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We correlated the level of $V_{TA}$ at the start of target trials (i.e., at the left edge of the histograms in Figure \ref{fig4}b) with the average target trial response in terms of spikes fired across excitatory neurons.
+We first focused on the contribution of $\theta$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
+We correlated the contrast (dev - msc) of $\theta$ at the start of target trials (i.e., at the left edge of the contrast histogram in Figure \ref{fig4}b) with the target trial response size contrast, across excitatory neurons.
 In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: The median $V_{TA}$ among excitatory neurons was lower in \dev{} than \msc{} (T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
-In addition, the neuron-level correlation between the contrasts in $V_{TA}$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
-This indicates that neurons with a greater decrease in $V_{TA}$ in \dev{} tended to respond more strongly to target stimuli.
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: The median $\theta$ among excitatory neurons was lower in \dev{} than \msc{} (T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+In addition, the neuron-level correlation between the contrasts in $\theta$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
+This indicates that neurons with a greater decrease in $\theta$ in \dev{} tended to respond more strongly to target stimuli.
 
 Yet, as we saw in Figure \ref{fig4}c, the membrane voltages also differed between the two sequences.
-In order to assess the relative contribution of $V_{TA}$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
-We then weighted the remaining contrasts in $V_{TA}$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
+In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
+We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
 The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (Figure \ref{fig4}g), $V_{TA}$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+In relative terms (Figure \ref{fig4}g), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
 A similar pattern emerged across networks and stimuli, as shown in Figure \ref{fig4}h.
-Here, only the initial response was driven mainly by $V_{TA}$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
+Here, only the initial response was driven mainly by $\theta$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
 This indicates that adaptation initiated and dominated the increase in firing at the start of trials, whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
@@
 \end{figure*}
-\textbf{Fig. 5} Lower $V_{TA}$ in \dev{} is a consequence of lower \dev{} non-target response magnitude.
-\textbf{a} $V_{TA}$ at the start of A trials (mean over trials) in the sample network, with each neuron shown in its spatial location.
+\textbf{Fig. 5} Lower $\theta$ in \dev{} is a consequence of lower \dev{} non-target response magnitude.
+\textbf{a} $\theta$ at the start of A trials (mean over trials) in the sample network, with each neuron shown in its spatial location.
 Left and middle column, raw averages for \dev{} and \msc{} sequences, respectively; right column, contrast between \dev{} and \msc{} averages.
@@
 \textbf{b} Average response (in spikes per trial) to non-target trials (B in \dev{}, B, C, D, and E in \msc{}). Columns are arranged as in panel a.
-\textbf{c} Relationship across excitatory neurons of the sample network between the contrast (\dev{} - \msc{}) in average number of spikes fired in response to non-target stimulation and the contrast in $V_{TA}$ at the start of target trials, and associated linear regression.
-\textbf{d} Network-wide median contrast (\dev{} - \msc{}) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\Delta V_{TA}$ across excitatory neurons).
+\textbf{c} Relationship across excitatory neurons of the sample network between the contrast (\dev{} - \msc{}) in average number of spikes fired in response to non-target stimulation and the contrast in $\theta$ at the start of target trials, and associated linear regression.
+\textbf{d} Network-wide median contrast (\dev{} - \msc{}) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\Delta \theta$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 
-First, we mapped the average value of $V_{TA}$ at the start of target trials to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
-In the \msc{} sequence, $V_{TA}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
+First, we mapped the average value of $\theta$ at the start of target trials to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
+In the \msc{} sequence, $\theta$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
 Conversely, in the \dev{} sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
@@
 
-Correlating the non-target response contrast (\dev{} - \msc{}) with the contrast in $V_{TA}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
+Correlating the non-target response contrast (\dev{} - \msc{}) with the contrast in $\theta$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
 This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
@@
 Note that non-responding neurons are ranked last and appear dark blue.
-\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (\dev{} - \msc{}) and the contrast in $V_{TA}$ at the start of B trials, and associated linear regression across all excitatory neurons.
+\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (\dev{} - \msc{}) and the contrast in $\theta$ at the start of B trials, and associated linear regression across all excitatory neurons.
 Neurons are colored by latency rank as in panel a.
@@
 \textbf{c} Median non-target response contrast (\dev{} - \msc{}) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labeled "late", across networks and stimuli.
-\textbf{c} Median contrast (\dev{} - \msc{}) in $V_{TA}$ among the early and late neurons, respectively, across networks and stimuli.
-\textbf{d} Relationship across networks and stimuli between the median contrast in $V_{TA}$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
+\textbf{c} Median contrast (\dev{} - \msc{}) in $\theta$ among the early and late neurons, respectively, across networks and stimuli.
+\textbf{d} Relationship across networks and stimuli between the median contrast in $\theta$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
 
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
-In Figure \ref{fig6}b, we related the non-target response contrast (\dev{} - \msc{}) to the contrast in $V_{TA}$ at the start of B trials.
-Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated the same response contrast to $\Delta V_{TA}$ before A.
+In Figure \ref{fig6}b, we related the non-target response contrast (\dev{} - \msc{}) to the contrast in $\theta$ at the start of B trials.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated the same response contrast to $\Delta \theta$ before A.
 As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the \dev{} sequence, while the response of the later portion of the network decreased.
@@
 Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B.
-We first confirmed that the non-target response and $V_{TA}$ contrasts were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
-In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $V_{TA}$ at the start of B trials (Figure \ref{fig6}d) within these two portions.
+We first confirmed that the non-target response and $\theta$ contrasts were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $\theta$ at the start of B trials (Figure \ref{fig6}d) within these two portions.
 We found that the response in the early portion was consistently greater in \dev{} than \msc{} (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
-The contrast in $V_{TA}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
+The contrast in $\theta$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
 
-Finally, to show that early $V_{TA}$ was causal in reducing the response to B, we correlated the median contrast in $V_{TA}$ among the early portion of neurons with the network-wide median contrast of the response to stimulation in B, as shown in Figure \ref{fig6}e.
+Finally, to show that early $\theta$ was causal in reducing the response to B, we correlated the median contrast in $\theta$ among the early portion of neurons with the network-wide median contrast of the response to stimulation in B, as shown in Figure \ref{fig6}e.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
 
-To sum up, we showed in Figure \ref{fig4} that greater $V_{TA}$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response.
-Here, we showed that the neurons responding earliest to B exhibit higher $V_{TA}$ in the \dev{} sequence as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
+To sum up, we showed in Figure \ref{fig4} that greater $\theta$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response.
+Here, we showed that the neurons responding earliest to B exhibit higher $\theta$ in the \dev{} sequence as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
 
@@
 \textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
-\textbf{a} Median contrast (\dev{} - \msc{}) in the non-target response (left) and in $V_{TA}$ at the start of B trials (right) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+\textbf{a} Median contrast (\dev{} - \msc{}) in the non-target response (left) and in $\theta$ at the start of B trials (right) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
 Compare to Figure \ref{fig6}c-d.
@@
 The subtraction of the two quantities in the box plot below shows that the non-target response tends to be reduced more in the full model.
-\textbf{c} Non-target response contrast, plotted against its Pearson correlation coefficient with the corresponding contrast in $V_{TA}$ at the start of target trials, in the full model.
+\textbf{c} Non-target response contrast, plotted against its Pearson correlation coefficient with the corresponding contrast in $\theta$ at the start of target trials, in the full model.
 Compare to Figure \ref{fig5}d.
-\textbf{d} Median contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials across networks and stimuli in the full and ablated models, arranged as in panel b.
-The subtraction of the two quantities in the box plot below shows that $V_{TA}$ tends to be reduced more in the full model.
-\textbf{e} $\Delta V_{TA}$ at the start of target trials, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response magnitude, in the full model.
+\textbf{d} Median contrast (\dev{} - \msc{}) in $\theta$ at the start of target trials across networks and stimuli in the full and ablated models, arranged as in panel b.
+The subtraction of the two quantities in the box plot below shows that $\theta$ tends to be reduced more in the full model.
+\textbf{e} $\Delta \theta$ at the start of target trials, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response magnitude, in the full model.
 Compare to Figure \ref{fig4}e.
@@
 Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}a the full model equivalent of Figure \ref{fig6}c and d.
-The contrasts (\dev{} - \msc{}) in the non-target response magnitude and in the resulting $V_{TA}$ levels at the start of B trials both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
+The contrasts (\dev{} - \msc{}) in the non-target response magnitude and in the resulting $\theta$ levels at the start of B trials both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
 Likewise, the correlation across excitatory neurons between these two contrasts remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992).
@@
 Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, the non-target response contrast remained tightly correlated with the contrast (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
-Correspondingly, $V_{TA}$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Indeed, the non-target response contrast remained tightly correlated with the contrast (\dev{} - \msc{}) in $\theta$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
+Correspondingly, $\theta$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
-Finally, based on the ablated model analysis in Figure \ref{fig4}, we sought to confirm that lowered $V_{TA}$ before target trials was responsible for the increased target response.
-As shown in Figure \ref{fig7}e, the contrasts (\dev{} - \msc{}) in $V_{TA}$ at the start of target trials and in response magnitude to A were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Finally, based on the ablated model analysis in Figure \ref{fig4}, we sought to confirm that lowered $\theta$ before target trials was responsible for the increased target response.
+As shown in Figure \ref{fig7}e, the contrasts (\dev{} - \msc{}) in $\theta$ at the start of target trials and in response magnitude to A were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
 Despite the weaker correlation, however, the full model target response was clearly increased in \dev{} over \msc{}, as shown in Figure \ref{fig7}f, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians differing by 94\%.
@@
 To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} responses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
-Unlike $V_{TA}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
-Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $V_{TA}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
+Unlike $\theta$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
+Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $\theta$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
 
@@
 \end{equation}
-Notice that $D$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $V_{TA}$, albeit more approximate.
+Notice that $D$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $\theta$, albeit more approximate.
 
@@
 As noted above, we hypothesized that, in the full model, STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly.
-To show this, we calculated $D$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $V_{TA}$ before.
+To show this, we calculated $D$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $\theta$ before.
 In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
@@
 To show that $D$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated the contrast (\dev{} - \msc{}) in $D^A$ against the target response contrast across neurons of the sample network.
-Notice that this mirrors the correlation shown in Figure \ref{fig4} between $\Delta V_{TA}$ before A and the response contrast.
+Notice that this mirrors the correlation shown in Figure \ref{fig4} between $\Delta \theta$ before A and the response contrast.
 As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $D$ causes reduced responses as expected.
@@
 To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli by local suppression, and an increase in the response to target stimuli by relief from global suppression, relative to \msc{}.
-Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in \dev{} B responses also entailed lower $V_{TA}$, which independently led to higher \dev{} A responses.
+Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in \dev{} B responses also entailed lower $\theta$, which independently led to higher \dev{} A responses.
 Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in the $V_{TA}$ contrast at the start of A trials as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta D^A$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in the $\theta$ contrast at the start of A trials as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta D^A$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
@@
 
-% Replacement vs adaptation is not a clean separation, since $V_{TA}$ is affected by replacement, but modulates adaptation
+% Replacement vs adaptation is not a clean separation, since $\theta$ is affected by replacement, but modulates adaptation
 

### 2023-01-12 paper-1/manuscript.tex
Fixed F6 caption panel labels

@@
 \textbf{c} Median non-target response contrast (\dev{} - \msc{}) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labeled "late", across networks and stimuli.
-\textbf{c} Median contrast (\dev{} - \msc{}) in $\theta$ among the early and late neurons, respectively, across networks and stimuli.
-\textbf{d} Relationship across networks and stimuli between the median contrast in $\theta$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
+\textbf{d} Median contrast (\dev{} - \msc{}) in $\theta$ among the early and late neurons, respectively, across networks and stimuli.
+\textbf{e} Relationship across networks and stimuli between the median contrast in $\theta$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
 

### 2023-01-12 paper-1/manuscript.tex
Editing

@@
 To show this, we sought to isolate this subset of early responders to B.
-In a given network and stimulus assignment, we found, for each neuron, the minimum response time (i.e., time to first spike) across B trials in \dev{}, \msc{} and \std{} sequences, and ranked neurons by this response time.
+In a given network and stimulus assignment, we found, for each neuron, the minimum response time (i.e., time to first spike) across B trials in all sequences, and ranked neurons by this response time.
 The ranking in the sample network is shown in Figure \ref{fig6}a.

### 2023-01-12 paper-1/manuscript.tex
Edit F7 layout (early/late horizontal like in F6)

@@
 \textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
-\textbf{a} Median contrast (\dev{} - \msc{}) in the non-target response (left) and in $\theta$ at the start of B trials (right) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+\textbf{a} Median contrast (\dev{} - \msc{}) in the non-target response (top) and in $\theta$ at the start of B trials (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
 Compare to Figure \ref{fig6}c-d.

### 2023-01-12 paper-1/manuscript.tex
Removed unclear/unnecessary/unfounded conclusion
Editing

@@
 This corresponds to immediate recovery ($\tau_x = 0$) and was done to maintain the magnitude of the excitatory postsynaptic potential (EPSP) in the recovered state regardless of whether STD was turned on or off.
-To turn off TA, we replaced Equation \ref{eqn-TA} with a constant $\theta = 0$, making the firing threshold completely unadaptive.
+To turn off TA, we replaced Equation \ref{eqn-TA} with a constant $V_{\theta} = \theta_0$, making the firing threshold completely unadaptive.
 
@@
 As we show in Figure \ref{fig8}e, higher $\Delta D^B$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000).
-This confirms again the integrity of the measure $D$, but also that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
+This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
 

### 2023-01-12 paper-1/manuscript.tex
Improved response/TA/D notation
Some possible overreach here, and the notation needs to be introduced.

@@
 \newcommand{\std}{\textit{std}}
+
+% Common notation
+\renewcommand{\R}[3][]{{}^{#1}_{}\!R^{#2}_{#3}}
+\renewcommand{\T}[3][]{{}^{#1}_{}T^{#2}_{#3}}
+\newcommand{\D}[3][]{{}^{#1}_{}\!D^{#2}_{#3}}
 
@@
 \begin{align}
-    V_\theta &= \theta_0 + \theta(t) \nonumber \\
-    \tau_{\theta} \frac{d\theta}{dt} &= -\theta + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
+    V_\theta &= \theta_0 + \T{}{}(t) \nonumber \\
+    \tau_{\theta} \frac{d\T{}{}}{dt} &= -\T{}{} + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
 \end{align}
@@
 \end{figure*}
-\textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower $\theta$.
+\textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower $\T{A}{}$.
 \textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
-Left and middle column, target trials in \dev{} sequence and \msc{} sequence, respectively; right column, contrast between these two.
-\textbf{b} Target trial average of the threshold adaptation voltage $\theta$, with neuron order and columns as in panel a.
+Left and middle column, target trials in \dev{} sequence ($\R[dev]{A}{i}$) and \msc{} sequence ($\R[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \R{A}{i}$).
+\textbf{b} Target trial average of the threshold adaptation voltage $\T{}{}$, with neuron order and columns as in panel a.
 \textbf{c} Target trial average of the membrane potential $V_m$, with neuron order and columns as in panel a.
 The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the colormap.
-\textbf{d} Relationship across excitatory neurons of the sample network between the contrast (\dev{} - \msc{}) in $\theta$ at the start of target trials and the contrast in average number of spikes fired in target trials, and associated linear regression.
-\textbf{e} Network-wide median contrast (\dev{} - \msc{}) in $\theta$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \theta$ vs response size contrast across excitatory neurons).
+\textbf{d} Relationship across excitatory neurons of the sample network between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$, and associated linear regression.
+\textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \T{A}{i}$ vs $\Delta \R{A}{i}$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
-Red markers indicate a significant negative correlation (two-sided Wald test, significance level 0.05).
+Red markers indicate a significant negative correlation (one-sided Wald test, significance level 0.05).
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
-\textbf{f} In the sample network, contribution of the contrast in $\theta$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
-\textbf{g} Contribution of $\theta$ relative to the sum of the absolute values of the weighted contributions in panel f.
-\textbf{h} Relative contribution of $\theta$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
+\textbf{f} In the sample network, contribution of the contrast in $\T{}{}$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
+\textbf{g} Contribution of $\T{}{}$ relative to the sum of the absolute values of the weighted contributions in panel f.
+\textbf{h} Relative contribution of $\T{}{}$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
 
-We quantified the former with the strength of adaptation $\theta$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
+We quantified the former with the strength of adaptation $\T{}{}$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
 As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
-Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $\theta$, leading to numerous horizontal breaks in the plot.
+Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $\T{}{}$, leading to numerous horizontal breaks in the plot.
 
-To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
+To assess the difference in synaptic currents in a way that allowed direct comparison to $\T{}{}$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
 As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
@@
 Based on these measures, we then asked why there were more spikes in \dev{} trials.
-We first focused on the contribution of $\theta$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We correlated the contrast (dev - msc) of $\theta$ at the start of target trials (i.e., at the left edge of the contrast histogram in Figure \ref{fig4}b) with the target trial response size contrast, across excitatory neurons.
+We first focused on the contribution of $\T{}{}$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
+We correlated the contrast in $\T{}{}$ at the start of target trials, i.e., $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$ corresponding to the leftmost datapoints of the contrast histogram in Figure \ref{fig4}b, with the target trial response size contrast $\Delta \R{A}{i}$, across excitatory neurons.
 In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: The median $\theta$ among excitatory neurons was lower in \dev{} than \msc{} (T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
-In addition, the neuron-level correlation between the contrasts in $\theta$ and target response was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
-This indicates that neurons with a greater decrease in $\theta$ in \dev{} tended to respond more strongly to target stimuli.
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} (T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+In addition, the correlation across excitatory neurons between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
+This indicates that neurons with a greater decrease in $\T[dev]{A}{}$ tended to respond more strongly to target stimuli.
 
 Yet, as we saw in Figure \ref{fig4}c, the membrane voltages also differed between the two sequences.
-In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
-We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
+In order to assess the relative contribution of $\T{}{}$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
+We then weighted the remaining contrasts in $\T{}{}$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
 The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (Figure \ref{fig4}g), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+In relative terms (Figure \ref{fig4}g), $\T{}{}$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
 A similar pattern emerged across networks and stimuli, as shown in Figure \ref{fig4}h.
-Here, only the initial response was driven mainly by $\theta$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
+Here, only the initial response was driven mainly by $\T{}{}$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
 This indicates that adaptation initiated and dominated the increase in firing at the start of trials, whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
@@
 \end{figure*}
-\textbf{Fig. 5} Lower $\theta$ in \dev{} is a consequence of lower \dev{} non-target response magnitude.
-\textbf{a} $\theta$ at the start of A trials (mean over trials) in the sample network, with each neuron shown in its spatial location.
-Left and middle column, raw averages for \dev{} and \msc{} sequences, respectively; right column, contrast between \dev{} and \msc{} averages.
+\textbf{Fig. 5} Lower $\T[dev]{}{}$ is a consequence of lower \dev{} non-target response magnitude.
+\textbf{a} $\T{}{}$ at the start of A trials (mean over trials) in the sample network, with each neuron shown in its spatial location.
+Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
 A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the \msc{} sequence.
-\textbf{b} Average response (in spikes per trial) to non-target trials (B in \dev{}, B, C, D, and E in \msc{}). Columns are arranged as in panel a.
-\textbf{c} Relationship across excitatory neurons of the sample network between the contrast (\dev{} - \msc{}) in average number of spikes fired in response to non-target stimulation and the contrast in $\theta$ at the start of target trials, and associated linear regression.
-\textbf{d} Network-wide median contrast (\dev{} - \msc{}) in the non-target response magnitude across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel (non-target response size contrast vs $\Delta \theta$ across excitatory neurons).
+\textbf{b} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel a.
+\textbf{c} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$, and associated linear regression.
+\textbf{d} $\Delta \R{N}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \R{N}{i}$ vs $\Delta \T{A}{i}$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 
-First, we mapped the average value of $\theta$ at the start of target trials to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
-In the \msc{} sequence, $\theta$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
+First, we mapped the average value of $\T{}{}$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
+In the \msc{} sequence, $\T[msc]{A}{i}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
 Conversely, in the \dev{} sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
 
-Next, we turned to the likely cause of this difference, the non-target responses (Figure \ref{fig5}b), mapped into space as the average number of spikes fired in each neuron.
+Next, we turned to the likely cause of this difference, the non-target responses (Figure \ref{fig5}b), mapped into space as the average number of spikes fired in each neuron, $\R{N}{i}$.
 As expected, the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds.
@@
 
-Correlating the non-target response contrast (\dev{} - \msc{}) with the contrast in $\theta$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
+Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
 
@@
 Note that non-responding neurons are ranked last and appear dark blue.
-\textbf{b} Relationship across excitatory neurons of the sample network between the non-target response contrast (\dev{} - \msc{}) and the contrast in $\theta$ at the start of B trials, and associated linear regression across all excitatory neurons.
+\textbf{b} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$, and associated linear regression across all excitatory neurons.
 Neurons are colored by latency rank as in panel a.
@@
 Inset histograms show the distributions of the plotted values, with the confidence intervals of their medians, calculated by bootstrap with 10000 iterations, in darker color.
-\textbf{c} Median non-target response contrast (\dev{} - \msc{}) among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labeled "late", across networks and stimuli.
-\textbf{d} Median contrast (\dev{} - \msc{}) in $\theta$ among the early and late neurons, respectively, across networks and stimuli.
-\textbf{e} Relationship across networks and stimuli between the median contrast in $\theta$ among the early portion and the median contrast in the response to stimulation in B, and associated linear regression
+\textbf{c} Median non-target response contrast $\Delta \R{N}{}$ among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labeled "late", across networks and stimuli.
+\textbf{d} Median contrast in $\T{B}{}$ among the early and late neurons, respectively, across networks and stimuli.
+\textbf{e} Relationship across networks and stimuli between $\Delta \T{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression
 
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
-In Figure \ref{fig6}b, we related the non-target response contrast (\dev{} - \msc{}) to the contrast in $\theta$ at the start of B trials.
-Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated the same response contrast to $\Delta \theta$ before A.
+In Figure \ref{fig6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\T{}{}$ at the start of B trials, $\Delta \T{B}{i}$.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
 As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the \dev{} sequence, while the response of the later portion of the network decreased.
@@
 Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B.
-We first confirmed that the non-target response and $\theta$ contrasts were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
-In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast (Figure \ref{fig6}c) and the median contrast in $\theta$ at the start of B trials (Figure \ref{fig6}d) within these two portions.
+We first confirmed that $\Delta \R{N}{i}$ and $\Delta T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, Figure \ref{fig6}c) and the median contrast in $\T{}{}$ at the start of B trials ($\Delta \T{B}{}$, Figure \ref{fig6}d) within these two portions.
 We found that the response in the early portion was consistently greater in \dev{} than \msc{} (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
-The contrast in $\theta$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
+$\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
 
-Finally, to show that early $\theta$ was causal in reducing the response to B, we correlated the median contrast in $\theta$ among the early portion of neurons with the network-wide median contrast of the response to stimulation in B, as shown in Figure \ref{fig6}e.
+Finally, to show that early $\T{B}{}$ was causal in reducing the response to B, we correlated the median contrast in $\T{}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in Figure \ref{fig6}e.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
 
-To sum up, we showed in Figure \ref{fig4} that greater $\theta$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response.
-Here, we showed that the neurons responding earliest to B exhibit higher $\theta$ in the \dev{} sequence as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
+To sum up, we showed in Figure \ref{fig4} that greater $\T[msc]{A}{}$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response.
+Here, we showed that the neurons responding earliest to B exhibit higher $\T[dev]{B}{}$ as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
 
@@
 \textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
-\textbf{a} Median contrast (\dev{} - \msc{}) in the non-target response (top) and in $\theta$ at the start of B trials (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+\textbf{a} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
 Compare to Figure \ref{fig6}c-d.
-\textbf{b} Median contrast (\dev{} - \msc{}) in the non-target response magnitude across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
+\textbf{b} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
 The red line is the identity function; points above this line represent networks with greater response reduction (\dev{} - \msc{}) in the full model than in the ablated model.
 The subtraction of the two quantities in the box plot below shows that the non-target response tends to be reduced more in the full model.
-\textbf{c} Non-target response contrast, plotted against its Pearson correlation coefficient with the corresponding contrast in $\theta$ at the start of target trials, in the full model.
+\textbf{c} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
 Compare to Figure \ref{fig5}d.
-\textbf{d} Median contrast (\dev{} - \msc{}) in $\theta$ at the start of target trials across networks and stimuli in the full and ablated models, arranged as in panel b.
-The subtraction of the two quantities in the box plot below shows that $\theta$ tends to be reduced more in the full model.
-\textbf{e} $\Delta \theta$ at the start of target trials, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response magnitude, in the full model.
+\textbf{d} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel b.
+The subtraction of the two quantities in the box plot below shows that $\T{A}$ tends to be reduced more in the full model.
+\textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
 Compare to Figure \ref{fig4}e.
-\textbf{f} Median contrast (\dev{} - \msc{}) in the target response magnitude across networks and stimuli in the full and ablated models, arranged as in panel b.
+\textbf{f} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel b.
 The subtraction of the two quantities in the box plot below shows that the target response tends to be increased more in the full model
@@
 Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}a the full model equivalent of Figure \ref{fig6}c and d.
-The contrasts (\dev{} - \msc{}) in the non-target response magnitude and in the resulting $\theta$ levels at the start of B trials both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
+The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
 Likewise, the correlation across excitatory neurons between these two contrasts remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992).
@@
 
-Comparing the non-target response reductions in Figure \ref{fig7}b, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians differing by 36\%.
+Comparing the non-target response reductions in Figure \ref{fig7}b, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
 Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, the non-target response contrast remained tightly correlated with the contrast (\dev{} - \msc{}) in $\theta$ at the start of target trials (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
-Correspondingly, $\theta$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
+Correspondingly, $\T{A}{exc}$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
-Finally, based on the ablated model analysis in Figure \ref{fig4}, we sought to confirm that lowered $\theta$ before target trials was responsible for the increased target response.
-As shown in Figure \ref{fig7}e, the contrasts (\dev{} - \msc{}) in $\theta$ at the start of target trials and in response magnitude to A were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response was clearly increased in \dev{} over \msc{}, as shown in Figure \ref{fig7}f, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians differing by 94\%.
+Finally, based on the ablated model analysis in Figure \ref{fig4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
+As shown in Figure \ref{fig7}e, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in Figure \ref{fig7}f, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
 This greater contrast, of course, directly reflects the greater DDI reported in Figure \ref{fig3}.
@@
 To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} responses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
-Unlike $\theta$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
-Therefore, we devised a measure $D$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $\theta$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
+Unlike $\T{}{}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
+Therefore, we devised a measure $\D{}{}$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $\T{}{}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
 
 A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
-In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $R_j^S$ of a neuron $j$ to a given stimulus $S$ across all sequences as a measure of the presynaptic drive.
+In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $\R{S}{j} = \frac{1}{3} \sum_{seq \in (std, dev, msc)} \R[seq]{S}{j}$ of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive.
 Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.
-Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $R_j^S k x_j^S$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials.
+Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $\R{S}{j} k x_j^S$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials.
 Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
 \begin{equation}
-    D_i^S = \sum_{j \in pre} R_j^S k (1-x_j^S)
+    \D{S}{i} = \sum_{j \in pre} \R{S}{j} k (1-x_j^S)
 \end{equation}
-Notice that $D$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $\theta$, albeit more approximate.
+Notice that $\D{}{}$ is measured in volts due to scaling with $k$, and is therefore roughly comparable to $\T{}{}$, albeit more approximate.
 
@@
 \textbf{Fig. 8} In the full model, STD reduces \dev{} non-target and increases \dev{} target responses relative to \msc{}.
-\textbf{a} Average depression $D$ in target (A) trials (top row) and B trials (bottom row) in the \dev{} (left) and \msc{} sequence (center), and their contrast (right) in the sample network.
+\textbf{a} Average depression $\D{}{i}$ in target (A) trials (top row) and B trials (bottom row) in the \dev{} (left) and \msc{} sequence (center), and their contrast (right) in the sample network.
 The color scale is shared across columns (and between \dev{} and \msc{}) for direct comparability.
-\textbf{b} Relationship across neurons of the sample network between the contrast (\dev{} - \msc{}) in $D^A$ and the contrast in the target response, and associated linear regression.
-\textbf{c} Network-wide median contrast in $D^A$, plotted against its Pearson correlation coefficient with the corresponding contrast in the target response, across networks and stimuli.
-\textbf{d} Median contrast (\dev{} - \msc{}) in $D^B$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
-\textbf{e} Relationship across networks and stimuli between the median contrast in $D^B$ among the early portion and the median contrast in the response to B among the entire network, and associated linear regression
+\textbf{b} Relationship across neurons of the sample network between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$, and associated linear regression.
+\textbf{c} $\Delta \D{A}{}$ across networks and stimuli, plotted against the corresponding Pearson coefficient between $\Delta \D{A}{}$ and $\Delta \R{A}{}$.
+\textbf{d} $\Delta \D{B}{}$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
+\textbf{e} Relationship across networks and stimuli between $\Delta \D{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression
 
 As noted above, we hypothesized that, in the full model, STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly.
-To show this, we calculated $D$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $\theta$ before.
+To show this, we calculated $\D{}{}$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
 In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
-The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast (\dev{} - \msc{}), which was negative throughout.
+The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast $\Delta \D{A}{i}$, which was negative throughout.
 Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both sequences.
@@
 However, in the \dev{} sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited.
-As a result, the contrast (\dev{} - \msc{}) showed slightly reduced $D$ in the periphery, but strikingly elevated $D$ near the stimulation site, indicating greatly increased synaptic depression in \dev{}.
+As a result, the contrast $\Delta \D{B}{i}$ was slightly reduced in the periphery, but strikingly elevated near the stimulation site, indicating greatly increased synaptic depression in \dev{}.
 
-To show that $D$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated the contrast (\dev{} - \msc{}) in $D^A$ against the target response contrast across neurons of the sample network.
-Notice that this mirrors the correlation shown in Figure \ref{fig4} between $\Delta \theta$ before A and the response contrast.
-As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $D$ causes reduced responses as expected.
+To show that $\D{}{}$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated $\Delta \D{A}{i}$ with $\Delta \R{A}{i}$ across neurons of the sample network.
+Notice that this mirrors the correlation shown in Figure \ref{fig4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
+As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
 We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
@@
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in Figure \ref{fig8}d, we recapitulate the logic established in Figure \ref{fig6}: Given the spatial clustering of $D^B$, we surmised a similar local suppression effect as with TA.
-We found that, indeed, neurons responding early to stimulation in B showed greater $D^B$ in \dev{} than \msc{} (T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced $D^B$ (T = 2.24e+03, p = 0.000133).
-We can relate this directly to the non-target response contrasts in the early and late portions shown in Figure \ref{fig7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local suppression in terms of $D$.
+First, in Figure \ref{fig8}d, we recapitulate the logic established in Figure \ref{fig6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
+We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
+We can relate this directly to the non-target response contrasts in the early and late portions shown in Figure \ref{fig7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local suppression in terms of $\D{}{}$.
 % Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
 
-To show that increased local suppression by STD reduces the non-target response, we correlated the median contrast in $D^B$ among the early portion of networks with the median contrast of the total response to B, echoing Figure \ref{fig6}e.
-As we show in Figure \ref{fig8}e, higher $\Delta D^B$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000).
+To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing Figure \ref{fig6}e.
+As we show in Figure \ref{fig8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000).
 This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
 
-Finally, given the very clear contrast in $D^A$ in the sample network, we considered the possibility that, in the context of the full model, STD could also evoke global suppression in the \msc{} sequence, contrary to our stated hypothesis.
-If so, we should expect to see less $D^A$ in \dev{} than \msc{}.
-To confirm this, we analyzed the contrast (\dev{} - \msc{}) in $D^A$ across networks and stimuli, and found its network median to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
-Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in \msc{} than \dev{} in terms of $D$, this difference appears to manifest only weakly in activity differences.
+Finally, given the very clear contrast in $\D{A}{i}$ in the sample network, we considered the possibility that, in the context of the full model, STD could also evoke global suppression in the \msc{} sequence, contrary to our stated hypothesis.
+If so, we should expect to see less $\D{A}{}$ in \dev{} than \msc{}.
+To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli, and found it to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
+Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in \msc{} than \dev{} in terms of $\D{}{}$, this difference appears to manifest only weakly in activity differences.
 
-To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli by local suppression, and an increase in the response to target stimuli by relief from global suppression, relative to \msc{}.
-Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in \dev{} B responses also entailed lower $\theta$, which independently led to higher \dev{} A responses.
+To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli (lower $\R[dev]{B}{}$) by local suppression, and an increase in the response to target stimuli (higher $\R[dev]{A}{}$) by relief from global suppression, relative to \msc{}.
+Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
 Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in the $\theta$ contrast at the start of A trials as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta D^A$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta \D{A}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.

### 2023-01-13 paper-1/manuscript.tex
TODOs
Added sec Methods:Notation

@@
 \newcommand{\D}[3][]{{}^{#1}_{}\!D^{#2}_{#3}}
+\newcommand{\mean}[1]{\overline{#1}}
 
@@
 
+\subsection{Notation}\label{sec-notation}
+
+Throughout this paper, we refer to three main quantities of interest: the stimulus response $\R{}{}$ defined below, the threshold adaptation value $\T{}{}$ defined in Equation \ref{eqn-TA}, and the short-term depression effect $\D{}{}$, which is introduced later in Equation \ref{eqn-D}.
+We employ sub- and superscripts to identify the precise quantity as follows, using $\R{}{}$ for illustration: $\R[seq]{X}{p}$ refers to the response of neuron(s) $p$, averaged across all trials of type $X$ in the sequence $seq$.
+Where $p$ does not identify individual neurons (e.g., $i$ or $j$), but rather a subpopulation (e.g., $exc$) or the whole network (no index), this refers to the median (of the individual quantity $\R[seq]{X}{i}$) across the indicated set of neurons.
+The trial type $X$ is either $A$ or $B$, referring to trials with stimulation at these sites, or $N$, indicating all \textbf{n}on-target trials B through E.
+Finally, we use the notation $\Delta \R{}{}$ to refer to the sequence contrast $\R[dev]{}{} - \R[msc]{}{}$, and $\mean{\R{}{p}}$ to indicate the mean, rather than median, of the individual quantities in a population $p$.
+
+We define the response $\R{X}{i}$ as the mean spike count of neuron $i$ across time bins in trials of type $X$.
+Almost equivalently, we define $\T{X}{i}$ (and $\D{}{}$ analogously) as the mean across trials of type $X$ of $\T{}{i}(t=0)$, i.e., the average TA voltage at the beginning of trials, immediately before stimulation.
+
 \section{Results}\label{sec-results}
@@
 This indicates that neurons with a greater decrease in $\T[dev]{A}{}$ tended to respond more strongly to target stimuli.
+% TODO: edit final sentence to end with ... tended to show a greater increase in their response to target stimuli.
 
@@
 A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the \msc{} sequence.
+% TODO: Drop "for this .. figures" since they're everywhere now.
 \textbf{b} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel a.
@@
 \textbf{e} Relationship across networks and stimuli between $\Delta \T{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression
+% TODO: Amend to correctly identify early in panel b as the first 100 EXC. Consider making this consistent across sample and stats; currently, stats is indiscriminate.
 
@@
 Here, we showed that the neurons responding earliest to B exhibit higher $\T[dev]{B}{}$ as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
+% TODO replace with: To sum up, we showed in Figure \ref{fig4} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
+Here, we showed that the neurons responding earliest to B exhibited higher $\T[dev]{B}{}$, which analogously reduced the response of the network as a whole to B.
+We also showed that this increase in $\T[dev]{B}{early}$ was a consequence of an increased response to non-target stimulation in the early portion, or in other words, of the more frequent presentation of B.
 
@@
 However, this reduction was not uniform: Target responses were reduced most in the \std{} sequence, and least in the \dev{} sequence.
+% TODO: Show this, somewhere?
 
@@
 We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
+% TODO: SFig for this.
 
@@
 \begin{equation}
-    \D{S}{i} = \sum_{j \in pre} \R{S}{j} k (1-x_j^S)
+    \D{S}{i} = \sum_{j \in pre} \R{S}{j} k (1-x_j^S) \label{eqn-D}
 \end{equation}
@@
 As a result, the contrast $\Delta \D{B}{i}$ was slightly reduced in the periphery, but strikingly elevated near the stimulation site, indicating greatly increased synaptic depression in \dev{}.
+% TODO: After "different response patterns", clarify that these are caused by structural, not dynamic, differences
 
@@
 Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in \msc{} than \dev{} in terms of $\D{}{}$, this difference appears to manifest only weakly in activity differences.
+% TODO: In hindsight, I'm not sure where this "weakly" claim comes from.
+Perhaps from Fig7, lower panels?
+If so, look at the scales in b and f and reconsider.
+Note also how the next paragraph seems to require a "not weak" conclusion here.
 

### 2023-01-13 paper-1/manuscript.tex
Rev sec Results:DD (F2) for notation consistency

@@
 In other words, we used the \msc{} sequence as a neutral baseline where the target stimulus was presented no more often than in the \dev{} sequence, but no expectation of a competing stimulus B could be formed.
-To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the \dev{} and \msc{} sequences, denoted as $R^A_{dev}$ and $R^A_{msc}$, defining a deviance detection index (DDI) as
+To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the \dev{} and \msc{} sequences, defining a deviance detection index (DDI) as
 \begin{equation}
-    DDI = \frac{R^A_{dev} - R^A_{msc}}{R^A_{dev} + R^A_{msc}} \label{eqn-ddi}
+    DDI = \frac{\mean{\R[dev]{A}{}} - \mean{\R[msc]{A}{}}}{\mean{\R[dev]{A}{}} + \mean{\R[msc]{A}{}}} \label{eqn-ddi}
 \end{equation}
@@
 Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
-\textbf{c} Response sizes $R^A_{dev}$ and $R^A_{std}$, normalized by $R^A_{msc}$, across networks and stimuli (n = 120).
+\textbf{c} Mean response sizes $\mean{\R[dev]{A}{}}$ and $\mean{\R[std]{A}{}}$, normalized by $\mean{\R[msc]{A}{}}$, across networks and stimuli (n = 120).
 In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.

### 2023-01-13 paper-1/manuscript.tex
Fixed a bug in F8 caption

@@
 \textbf{b} Relationship across neurons of the sample network between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$, and associated linear regression.
-\textbf{c} $\Delta \D{A}{}$ across networks and stimuli, plotted against the corresponding Pearson coefficient between $\Delta \D{A}{}$ and $\Delta \R{A}{}$.
+\textbf{c} $\Delta \D{A}{}$ across networks and stimuli, plotted against the corresponding Pearson coefficient between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$.
 \textbf{d} $\Delta \D{B}{}$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.

### 2023-01-13 paper-1/manuscript.tex
Fixed missing test indicator

@@
 As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
-We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14), though the median correlation coefficient $\rho$ was only -0.282.
+We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120), though the median correlation coefficient $\rho$ was only -0.282.
 We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).

### 2023-01-13 plot_utils.py
Incorporate notation into all figures

@@ -5,6 +5,7 @@ from matplotlib.colors import colorConverter
 from brian2.units import msecond
 import scipy.stats as stats
 import seaborn as sns
+import string
 import styling
 
 
@@ -19,10 +20,27 @@ colors = {
     'Vm': 'orange'
 }
 
+def _quantity_label(core, seq, stim, index, mean):
+    template = string.Template(r'{}^{$seq}_{}$core^{$stim}_{$index}')
+    label = template.substitute(core=core, seq=seq, stim=stim, index=index)
+    if mean:
+        label = string.Template(r'\overline{$label}').substitute(label=label)
+    return label
+
+# Note: Negative space before R and D omitted for better kerning on \Delta \[RD] in the font used in mpl.
+def R(seq='', stim='', index='', mean=False):
+    return _quantity_label('R', seq, stim, index, mean)
+
+def T(seq='', stim='', index='', mean=False):
+    return _quantity_label('T', seq, stim, index, mean)
+
+def D(seq='', stim='', index='', mean=False):
+    return _quantity_label('D', seq, stim, index, mean)
+
 labels = {
     'pspike': 'Spike probability',
     'Depression': '$V_{STD}$ (mV)',
-    'Threshold': '$V_{TA}$ (mV)',
+    'Threshold': f'${T()}$ (mV)',
     'Synapses': '$V_{syn}$ (mV)',
     'Reset': '$V_{reset}$ (mV)',
     'Vm': '$V_m$ (mV)'

### 2023-01-13 paper-1/manuscript.tex
Rev F8c and stats to use median, not mean, D

@@
 If so, we should expect to see less $\D{A}{}$ in \dev{} than \msc{}.
-To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli, and found it to be robustly negative (T = 955, p = 1.23e-12; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
+To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli, and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
 Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in \msc{} than \dev{} in terms of $\D{}{}$, this difference appears to manifest only weakly in activity differences.

### 2023-01-13 paper-1/manuscript.tex
Editing

@@
 In addition, the correlation across excitatory neurons between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
-This indicates that neurons with a greater decrease in $\T[dev]{A}{}$ tended to respond more strongly to target stimuli.
-% TODO: edit final sentence to end with ... tended to show a greater increase in their response to target stimuli.
+This indicates that neurons with a greater decrease in $\T[dev]{A}{i}$ tended to show a greater increase in their response to target stimuli.
 
@@
 Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
-A and B stimulus locations for this and following figures are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the \msc{} sequence.
-% TODO: Drop "for this .. figures" since they're everywhere now.
+A and B stimulus locations are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the \msc{} sequence.
 \textbf{b} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel a.

### 2023-01-13 paper-1/manuscript.tex
Clarified F6 conclusion

@@
 
-To sum up, we showed in Figure \ref{fig4} that greater $\T[msc]{A}{}$ in the portion of the network responding earliest to a stimulus plays an important role in reducing the overall response.
-Here, we showed that the neurons responding earliest to B exhibit higher $\T[dev]{B}{}$ as a consequence of an increased response, and that this increase in TA was directly responsible for reducing the response size and TA levels of the network as a whole.
-% TODO replace with: To sum up, we showed in Figure \ref{fig4} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
+To sum up, we showed in Figure \ref{fig4} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
 Here, we showed that the neurons responding earliest to B exhibited higher $\T[dev]{B}{}$, which analogously reduced the response of the network as a whole to B.
-We also showed that this increase in $\T[dev]{B}{early}$ was a consequence of an increased response to non-target stimulation in the early portion, or in other words, of the more frequent presentation of B.
+We also showed that this increase in $\T[dev]{B}{early}$ was a consequence of an increased response to non-target stimulation in the early portion, or in other words, a direct consequence of the more frequent presentation of B.
 

### 2023-01-13 paper-1/manuscript.tex
Editing

@@
 By contrast, stimuli in the \msc{} sequence are presented more sporadically, allowing local suppression to decay, thus leading to larger network responses.
-This difference in non-target response sizes -- small in \dev{}, large in \msc{} -- is then reflected in network-wide \emph{global suppression}, such that the network is more excitable to stimulation with A (or other non-standard inputs) in the \dev{} sequence than in the control context.
+This difference in non-target response sizes -- small in \dev{}, large in \msc{} -- is then reflected in network-wide \emph{global suppression}, such that the network is more excitable to stimulation with A (or other infrequent inputs) in the \dev{} sequence than in the control context.
 

### 2023-01-13 paper-1/manuscript.tex
Rev Fig8 conclusion

@@
 To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli, and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
-Taken together, we cautiously conclude that, while there is clear evidence for greater global suppression in \msc{} than \dev{} in terms of $\D{}{}$, this difference appears to manifest only weakly in activity differences.
-% TODO: In hindsight, I'm not sure where this "weakly" claim comes from.
-Perhaps from Fig7, lower panels?
-If so, look at the scales in b and f and reconsider.
-Note also how the next paragraph seems to require a "not weak" conclusion here.
+We conclude that there is clear evidence for greater STD-mediated global suppression in \msc{} than \dev{} in the full model.
 

### 2023-01-13 paper-1/manuscript.tex
Editing

@@
 Conversely, depression in B trials (bottom) was spatially more clustered around the stimulation site in both sequences.
-In the \msc{} sequence, we can attribute the difference between A and B primarily to different response patterns: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network.
+In the \msc{} sequence, we can attribute the difference between A and B primarily to different response patterns due to network structure, not dynamics: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network.
 However, in the \dev{} sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited.
 As a result, the contrast $\Delta \D{B}{i}$ was slightly reduced in the periphery, but strikingly elevated near the stimulation site, indicating greatly increased synaptic depression in \dev{}.
-% TODO: After "different response patterns", clarify that these are caused by structural, not dynamic, differences
 

### 2023-01-13 paper-1/manuscript.tex
Killed a todo by looking closely
Added F7a: response size difference with/out STD

@@
 Neurons are colored by latency rank as in panel a.
-Additionally, the earliest 100 responders are shown in the left plot, while later responders are shown in the right plot.
+Additionally, excitatory neurons among the earliest 100 responders are shown in the left plot, while later responders are shown in the right plot.
 Inset histograms show the distributions of the plotted values, with the confidence intervals of their medians, calculated by bootstrap with 10000 iterations, in darker color.
@@
 \textbf{e} Relationship across networks and stimuli between $\Delta \T{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression
-% TODO: Amend to correctly identify early in panel b as the first 100 EXC. Consider making this consistent across sample and stats; currently, stats is indiscriminate.
 
@@
 To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD.
-On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (data not shown).
-However, this reduction was not uniform: Target responses were reduced most in the \std{} sequence, and least in the \dev{} sequence.
-% TODO: Show this, somewhere?
+On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (see Figure \ref{fig7}a).
+However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (T = 2.66e+03, p = 0.00554).
 
@@
 \end{figure*}
-\textbf{Fig. 7} Adding STD to the model maintains or slighly enhances the effects of TA.
-\textbf{a} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+\textbf{Fig. 7} Adding STD to the model slightly enhances the effects of TA.
+\textbf{a} Reductions in mean response size $\mean{\R{A}}$ in each sequence as a consequence of adding STD to the model, across networks and stimuli.
+\textbf{b} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
 Compare to Figure \ref{fig6}c-d.
-\textbf{b} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
+\textbf{c} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
 The red line is the identity function; points above this line represent networks with greater response reduction (\dev{} - \msc{}) in the full model than in the ablated model.
 The subtraction of the two quantities in the box plot below shows that the non-target response tends to be reduced more in the full model.
-\textbf{c} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
+\textbf{d} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
 Compare to Figure \ref{fig5}d.
-\textbf{d} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel b.
+\textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
 The subtraction of the two quantities in the box plot below shows that $\T{A}$ tends to be reduced more in the full model.
-\textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
+\textbf{f} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
 Compare to Figure \ref{fig4}e.
-\textbf{f} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel b.
+\textbf{g} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
 The subtraction of the two quantities in the box plot below shows that the target response tends to be increased more in the full model
@@
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
-Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}a the full model equivalent of Figure \ref{fig6}c and d.
+Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}b the full model equivalent of Figure \ref{fig6}c and d.
 The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
@@
 
-Comparing the non-target response reductions in Figure \ref{fig7}b, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
+Comparing the non-target response reductions in Figure \ref{fig7}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
 Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}c.
-Correspondingly, $\T{A}{exc}$ was reduced in \dev{}, as shown in Figure \ref{fig7}d, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}d.
+Correspondingly, $\T{A}{exc}$ was reduced in \dev{}, as shown in Figure \ref{fig7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
 Finally, based on the ablated model analysis in Figure \ref{fig4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
-As shown in Figure \ref{fig7}e, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in Figure \ref{fig7}f, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
+As shown in Figure \ref{fig7}f, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in Figure \ref{fig7}g, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
 This greater contrast, of course, directly reflects the greater DDI reported in Figure \ref{fig3}.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}d, bottom) with the corresponding $\Delta \D{A}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}e, bottom) with the corresponding $\Delta \D{A}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.

### 2023-01-16 paper-1/manuscript.tex
Added SFig to show lost STD correlation

@@
 We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
-% TODO: SFig for this.
+Similar behavior was also seen in the \msc{} sequence, as shown in Figure \ref{sfig-STD-correlation-decay}.
 
@@
 
+\begin{figure}%
+    \centering
+    \includegraphics{sfig-STD-correlation-decay.pdf}
+    \caption{}
+    \label{sfig-STD-correlation-decay}
+\end{figure*}
+\textbf{Figure S?} The STD variable $x_i$ only correlates with the activity of the immediately preceding trial.
+Each marker represents the coefficient $\rho$ of the Pearson correlation between the spike counts in a target trial and the corresponding values $x_i$ at the beginning of the immediately following trial ($\Delta$ trials = 1) or after one intervening trial ($\Delta$ trials = 2).
+Recall that $x$ is reduced upon spiking, which implies that an effective correlation is negative.
+Color represents the p-value of the alternative hypothesis that $\rho < 0$.
+Horizontal lines indicate the median $\rho$ and p.
+Horizontal jitter is introduced to aid visual separability of the markers.
+
 \begin{figure*}%

### 2023-01-17 paper-1/manuscript.tex
Latex fixes & minor edits

@@
 Three randomized sequences were presented, containing 80\% A and 20\% B (\std{}), 20\% A and 80\% B (\dev{}), and 20\% of each of 5 stimuli, including A and B (\msc{}).
-Note that, while the onset asynchrony of 500 ms is represented faithfully, the horizontal extent of the stimulation markers has no meaning; stimuli were presented instantaneously rather than over an extended period.
+Note that, while the onset asynchrony of 500 ms is represented faithfully, the horizontal extent of the stimulation markers has no meaning; stimuli were presented instantaneously rather than over an extended period
 
@@
 
-Throughout this paper, we refer to three main quantities of interest: the stimulus response $\R{}{}$ defined below, the threshold adaptation value $\T{}{}$ defined in Equation \ref{eqn-TA}, and the short-term depression effect $\D{}{}$, which is introduced later in Equation \ref{eqn-D}.
+Throughout this paper, we refer to three main quantities of interest: the stimulus response $\R{}{}$ defined in the next paragraph, the threshold adaptation value $\T{}{}$ defined in Equation \ref{eqn-TA}, and the short-term depression effect $\D{}{}$, which will be introduced later in Equation \ref{eqn-D}.
 We employ sub- and superscripts to identify the precise quantity as follows, using $\R{}{}$ for illustration: $\R[seq]{X}{p}$ refers to the response of neuron(s) $p$, averaged across all trials of type $X$ in the sequence $seq$.
@@
 We define the response $\R{X}{i}$ as the mean spike count of neuron $i$ across time bins in trials of type $X$.
-Almost equivalently, we define $\T{X}{i}$ (and $\D{}{}$ analogously) as the mean across trials of type $X$ of $\T{}{i}(t=0)$, i.e., the average TA voltage at the beginning of trials, immediately before stimulation.
+Almost equivalently, we define $\T{X}{i}$ (and $\D{X}{i}$ analogously) as the mean across trials of type $X$ of $\T{}{i}(t=0)$, i.e., the average TA voltage at the beginning of trials, immediately before stimulation.
 
@@
 \textbf{Fig. 7} Adding STD to the model slightly enhances the effects of TA.
-\textbf{a} Reductions in mean response size $\mean{\R{A}}$ in each sequence as a consequence of adding STD to the model, across networks and stimuli.
+\textbf{a} Reductions in mean response size $\mean{\R{A}{}}$ in each sequence as a consequence of adding STD to the model, across networks and stimuli.
 \textbf{b} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
@@
 \textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
-The subtraction of the two quantities in the box plot below shows that $\T{A}$ tends to be reduced more in the full model.
+The subtraction of the two quantities in the box plot below shows that $\T{A}{}$ tends to be reduced more in the full model.
 \textbf{f} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}e, bottom) with the corresponding $\Delta \D{A}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}e, bottom) with the corresponding $\Delta \D{A}{}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
@@
 
-\begin{figure}%
+\begin{figure*}%
     \centering

### 2023-01-17 paper-1/manuscript.tex
Editing

@@
 lower responses to B in \dev{} than to B through E in \msc{}.
-Yet, since activity is modulated by thresholds as shown above, and since the dev sequence was random, meaning that thresholds before stimulation could carry no information about the upcoming stimulus, we might expect the lowered thresholds in \dev{} to yield not only higher target responses, but also higher non-target responses.
+Yet, since activity is modulated by thresholds as shown above, and since the dev sequence was random, implying that $\T[dev]{A}{} \approx \T[dev]{B}{}$, we might expect the lowered thresholds in \dev{} to yield not only higher target responses, but also higher non-target responses.
 To resolve this apparent paradox, we will turn to an analysis of the spatial arrangement of activity and adaptation.

### 2023-01-17 paper-1/manuscript.tex
Editing

@@
 
-Next, we turned to the likely cause of this difference, the non-target responses (Figure \ref{fig5}b), mapped into space as the average number of spikes fired in each neuron, $\R{N}{i}$.
-As expected, the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds.
+Visualizing the likely cause of this difference, the non-target responses, in the same fashion (Figure \ref{fig5}b), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
 Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.
@@
 Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
-We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and comparatively strongly to stimulation in B, and therefore more to non-target stimuli in \dev{} than \msc{}, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B, and therefore more to non-target stimuli in \dev{} than \msc{}, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
 
@@
 Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
-As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red) increasing their response in the \dev{} sequence, while the response of the later portion of the network decreased.
+As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red, left plot) increasing their response in the \dev{} sequence, while the response of the later portion of the network (blue, right plot) decreased.
 Correspondingly, the TA levels of the early portion also tended to increase, in clear contrast to the later portion.
@@
 Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B.
-We first confirmed that $\Delta \R{N}{i}$ and $\Delta T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
+We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, Figure \ref{fig6}c) and the median contrast in $\T{}{}$ at the start of B trials ($\Delta \T{B}{}$, Figure \ref{fig6}d) within these two portions.
@@
 
-Finally, to show that early $\T{B}{}$ was causal in reducing the response to B, we correlated the median contrast in $\T{}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in Figure \ref{fig6}e.
+Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in Figure \ref{fig6}e.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
@@
 To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD.
-On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (see Figure \ref{fig7}a).
-However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (T = 2.66e+03, p = 0.00554).
+On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (Figure \ref{fig7}a).
+However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (more than \msc{}: T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (less than \msc{}: T = 2.66e+03, p = 0.00554).
 
@@
 Likewise, the correlation across excitatory neurons between these two contrasts remained positive (T = 7.26e+03, p = 9.86e-22) and very strong (median Pearson's $\rho$ = 0.992).
-This indicates that local suppression to B mediated by TA remained a key driver of the reduced response to non-target trials.
+This indicates that local suppression to B mediated by TA in the early responders remained a key driver of the reduced response to non-target trials.
 
@@
 Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}d.
-Correspondingly, $\T{A}{exc}$ was reduced in \dev{}, as shown in Figure \ref{fig7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in Figure \ref{fig7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
@@
 Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
-The primary cause for the greater target response, therefore, appeared to be the smaller non-target response, which was where we noted the first clear shift towards increased deviance detection.
+The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (Figure \ref{fig7}c), which was where we noted the first clear shift towards increased deviance detection.
 
@@
 
-To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated the number of spikes in \dev{} target trials with $x_j$ at the start of later trials (cf.
-Equation \ref{eqn-xsyn}) across excitatory neurons.
+To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated the number of spikes in \dev{} target trials with $x_j$ at the start of later trials across excitatory neurons.
 We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
@@
 To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} responses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
-Unlike $\T{}{}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
+Unlike $\T{}{j}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
 Therefore, we devised a measure $\D{}{}$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $\T{}{}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
@@
 A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
-In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $\R{S}{j} = \frac{1}{3} \sum_{seq \in (std, dev, msc)} \R[seq]{S}{j}$ of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive.
+In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $\R{S}{j} = \frac{1}{3} (\R[dev]{S}{j} + \R[msc]{S}{j} + \R[std]{S}{j})$ of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive.
 Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.
-Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $\R{S}{j} k x_j^S$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials.
+Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $\R{S}{j} x_j^S k$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials.
 Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
 \begin{equation}
-    \D{S}{i} = \sum_{j \in pre} \R{S}{j} k (1-x_j^S) \label{eqn-D}
+    \D{S}{i} = \sum_{j \in pre} \R{S}{j} (1-x_j^S) k \label{eqn-D}
 \end{equation}
@@
 
-As noted above, we hypothesized that, in the full model, STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly.
+As noted above, we hypothesized that STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly.
 To show this, we calculated $\D{}{}$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
@@
 In the \msc{} sequence, we can attribute the difference between A and B primarily to different response patterns due to network structure, not dynamics: The context, and thus $x$, was approximately the same, but responses to B tended not to spread out as much in this network.
-However, in the \dev{} sequence, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited.
+In the \dev{} sequence, however, the B response was limited almost exclusively to a very local patch around the stimulation site, with depression therefore similarly spatially limited.
 As a result, the contrast $\Delta \D{B}{i}$ was slightly reduced in the periphery, but strikingly elevated near the stimulation site, indicating greatly increased synaptic depression in \dev{}.
@@
 As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
-We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120), though the median correlation coefficient $\rho$ was only -0.282.
+We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
 We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).
@@
 To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing Figure \ref{fig6}e.
-As we show in Figure \ref{fig8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 1000).
+As we show in Figure \ref{fig8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
 This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
@@
 To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli (lower $\R[dev]{B}{}$) by local suppression, and an increase in the response to target stimuli (higher $\R[dev]{A}{}$) by relief from global suppression, relative to \msc{}.
-Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
+Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower thresholds $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
 Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}e, bottom) with the corresponding $\Delta \D{A}{}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.

### 2023-01-18 paper-1/manuscript.tex
Separate T and theta

@@
 \begin{align}
-    V_\theta &= \theta_0 + \T{}{}(t) \nonumber \\
-    \tau_{\theta} \frac{d\T{}{}}{dt} &= -\T{}{} + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
+    V_\theta &= \theta_0 + \theta(t) \nonumber \\
+    \tau_{\theta} \frac{d\theta}{dt} &= -\theta + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
 \end{align}
@@
 
-Throughout this paper, we refer to three main quantities of interest: the stimulus response $\R{}{}$ defined in the next paragraph, the threshold adaptation value $\T{}{}$ defined in Equation \ref{eqn-TA}, and the short-term depression effect $\D{}{}$, which will be introduced later in Equation \ref{eqn-D}.
+Throughout this paper, we refer to three main quantities of interest: the stimulus response magnitude $\R{}{}$, the threshold adaptation value $\T{}{}$, and the short-term depression value $\D{}{}$.
+We define the response $\R{X}{i}$ as the mean number of spikes fired by neuron $i$ across all trials of type $X$, i.e., the average response to $X$.
+We define $\T{X}{i}$ as the mean across trials of type $X$ of $\theta_i(t=0)$, i.e., the average TA voltage at the beginning of trials, immediately before stimulation.
+Lastly, the STD quantity $\D{X}{i}$ will be introduced later in Equation \ref{eqn-D}.
+
 We employ sub- and superscripts to identify the precise quantity as follows, using $\R{}{}$ for illustration: $\R[seq]{X}{p}$ refers to the response of neuron(s) $p$, averaged across all trials of type $X$ in the sequence $seq$.
-Where $p$ does not identify individual neurons (e.g., $i$ or $j$), but rather a subpopulation (e.g., $exc$) or the whole network (no index), this refers to the median (of the individual quantity $\R[seq]{X}{i}$) across the indicated set of neurons.
-The trial type $X$ is either $A$ or $B$, referring to trials with stimulation at these sites, or $N$, indicating all \textbf{n}on-target trials B through E.
+We use $p$ to either identify individual neurons (e.g., $i$ or $j$), or a subpopulation (e.g., $exc$) or the whole network (no index), in which case the designated quantity is the median (of the individual quantities $\R[seq]{X}{i}$) across the indicated set of neurons.
+The trial type $X$ is either $A$ or $B$, referring to trials with stimulation at these sites, or $N$, indicating all \textit{n}on-target trials B through E.
 Finally, we use the notation $\Delta \R{}{}$ to refer to the sequence contrast $\R[dev]{}{} - \R[msc]{}{}$, and $\mean{\R{}{p}}$ to indicate the mean, rather than median, of the individual quantities in a population $p$.
-
-We define the response $\R{X}{i}$ as the mean spike count of neuron $i$ across time bins in trials of type $X$.
-Almost equivalently, we define $\T{X}{i}$ (and $\D{X}{i}$ analogously) as the mean across trials of type $X$ of $\T{}{i}(t=0)$, i.e., the average TA voltage at the beginning of trials, immediately before stimulation.
 
@@
 \end{figure*}
-\textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower $\T{A}{}$.
+\textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower initial $\theta$.
 \textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
 Left and middle column, target trials in \dev{} sequence ($\R[dev]{A}{i}$) and \msc{} sequence ($\R[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \R{A}{i}$).
-\textbf{b} Target trial average of the threshold adaptation voltage $\T{}{}$, with neuron order and columns as in panel a.
+\textbf{b} Target trial average of the threshold adaptation voltage $\theta$, with neuron order and columns as in panel a.
 \textbf{c} Target trial average of the membrane potential $V_m$, with neuron order and columns as in panel a.
@@
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
-\textbf{f} In the sample network, contribution of the contrast in $\T{}{}$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
-\textbf{g} Contribution of $\T{}{}$ relative to the sum of the absolute values of the weighted contributions in panel f.
-\textbf{h} Relative contribution of $\T{}{}$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
+\textbf{f} In the sample network, contribution of the contrast in $\theta$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
+\textbf{g} Contribution of $\theta$ relative to the sum of the absolute values of the weighted contributions in panel f.
+\textbf{h} Relative contribution of $\theta$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
 
-We quantified the former with the strength of adaptation $\T{}{}$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
+We quantified the former with the strength of adaptation $\theta$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
 As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
-Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $\T{}{}$, leading to numerous horizontal breaks in the plot.
+Finally, note that inhibitory neurons, which were modeled without TA, naturally showed no difference in $\theta$, leading to numerous horizontal breaks in the plot.
 
-To assess the difference in synaptic currents in a way that allowed direct comparison to $\T{}{}$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
+To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
 As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
@@
 Based on these measures, we then asked why there were more spikes in \dev{} trials.
-We first focused on the contribution of $\T{}{}$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We correlated the contrast in $\T{}{}$ at the start of target trials, i.e., $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$ corresponding to the leftmost datapoints of the contrast histogram in Figure \ref{fig4}b, with the target trial response size contrast $\Delta \R{A}{i}$, across excitatory neurons.
+We first focused on the contribution of $\theta$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
+We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in Figure \ref{fig4}b, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$, across excitatory neurons.
 In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
@@
 Yet, as we saw in Figure \ref{fig4}c, the membrane voltages also differed between the two sequences.
-In order to assess the relative contribution of $\T{}{}$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
-We then weighted the remaining contrasts in $\T{}{}$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
+In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
+We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
 The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (Figure \ref{fig4}g), $\T{}{}$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+In relative terms (Figure \ref{fig4}g), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
 A similar pattern emerged across networks and stimuli, as shown in Figure \ref{fig4}h.
-Here, only the initial response was driven mainly by $\T{}{}$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
+Here, only the initial response was driven mainly by $\theta$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
 This indicates that adaptation initiated and dominated the increase in firing at the start of trials, whereas the late response was enlarged primarily as a result of increased presynaptic excitation.
@@
 \textbf{Fig. 5} Lower $\T[dev]{}{}$ is a consequence of lower \dev{} non-target response magnitude.
-\textbf{a} $\T{}{}$ at the start of A trials (mean over trials) in the sample network, with each neuron shown in its spatial location.
+\textbf{a} $\T{A}{i}$, i.e., $\theta_i$ at the start of A trials (mean over trials), in the sample network, with each neuron shown in its spatial location.
 Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
@@
 
-First, we mapped the average value of $\T{}{}$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
+First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
 In the \msc{} sequence, $\T[msc]{A}{i}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
-In Figure \ref{fig6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\T{}{}$ at the start of B trials, $\Delta \T{B}{i}$.
+In Figure \ref{fig6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
 Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
@@
 We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
-In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, Figure \ref{fig6}c) and the median contrast in $\T{}{}$ at the start of B trials ($\Delta \T{B}{}$, Figure \ref{fig6}d) within these two portions.
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, Figure \ref{fig6}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, Figure \ref{fig6}d) within these two portions.
 We found that the response in the early portion was consistently greater in \dev{} than \msc{} (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).

### 2023-01-18 plot_utils.py
Bugfix

@@ -151,13 +151,15 @@ def inset_hist(ax, data, x=True, median_color='C1', rescale=True, **kwargs):
     mask = (bins > lo - bsize) & (bins < hi + bsize)
     y = bins[mask]
     y[0], y[-1] = lo, hi
+    x_in_ci = binned[mask[1:]].copy()
+    x_in_ci[0] = x_in_ci[1]
 
     if x:
-        tx.fill_between(y, binned[mask[1:]], step='pre', ec='dimgrey', fc=median_color)
+        tx.fill_between(y, x_in_ci, step='pre', ec='dimgrey', fc=median_color)
         tx.set_ylim(bottom=-10*binned.max())
         tx.set_yticks([])
     else:
-        tx.fill_betweenx(y, binned[mask[1:]], step='pre', ec='dimgrey', fc=median_color)
+        tx.fill_betweenx(y, x_in_ci, step='pre', ec='dimgrey', fc=median_color)
         tx.set_xlim(left=-10*binned.max())
         tx.set_xticks([])
     sns.despine(ax=tx)

### 2023-01-18 paper-1/manuscript.tex
Disambiguate T (TA) from T (stats)
Minor tweak for script compat

@@
 % Common notation
-\renewcommand{\R}[3][]{{}^{#1}_{}\!R^{#2}_{#3}}
-\renewcommand{\T}[3][]{{}^{#1}_{}T^{#2}_{#3}}
-\newcommand{\D}[3][]{{}^{#1}_{}\!D^{#2}_{#3}}
+\renewcommand{\R}[3][]{{}^{#1}_{}\!\mathbf{R}^{#2}_{#3}}
+\renewcommand{\T}[3][]{{}^{#1}_{}\mathbf{T}^{#2}_{#3}}
+\newcommand{\D}[3][]{{}^{#1}_{}\!\mathbf{D}^{#2}_{#3}}
 \newcommand{\mean}[1]{\overline{#1}}
@@
 Finally, we use the notation $\Delta \R{}{}$ to refer to the sequence contrast $\R[dev]{}{} - \R[msc]{}{}$, and $\mean{\R{}{p}}$ to indicate the mean, rather than median, of the individual quantities in a population $p$.
+
+For disambiguation, note that the quantity $\T{X}{i}$, regardless of sub- and superscripts, is always in bold math font, whereas the Wilcoxon statistic T is typeset in normal font.
 
@@
 \end{figure*}
-\textbf{Figure S?} The STD variable $x_i$ only correlates with the activity of the immediately preceding trial.
+\textbf{Fig.
+S?} The STD variable $x_i$ only correlates with the activity of the immediately preceding trial.
 Each marker represents the coefficient $\rho$ of the Pearson correlation between the spike counts in a target trial and the corresponding values $x_i$ at the beginning of the immediately following trial ($\Delta$ trials = 1) or after one intervening trial ($\Delta$ trials = 2).

### 2023-01-19 paper-1/manuscript.tex
Undo notation spacing tweaks
Consistent newline use in fig legends
Editing

@@
 % Common notation
-\renewcommand{\R}[3][]{{}^{#1}_{}\!\mathbf{R}^{#2}_{#3}}
+\renewcommand{\R}[3][]{{}^{#1}_{}\mathbf{R}^{#2}_{#3}}
 \renewcommand{\T}[3][]{{}^{#1}_{}\mathbf{T}^{#2}_{#3}}
-\newcommand{\D}[3][]{{}^{#1}_{}\!\mathbf{D}^{#2}_{#3}}
+\newcommand{\D}[3][]{{}^{#1}_{}\mathbf{D}^{#2}_{#3}}
 \newcommand{\mean}[1]{\overline{#1}}
+\newcommand{\Rhat}[3][]{{}^{#1}_{}\mathbf{\widehat{R}}^{#2}_{#3}}
 
@@
 \end{figure*}
-\textbf{Fig.
-1} Model and paradigm.
+\textbf{Fig. 1} Model and paradigm.
 \textbf{a} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes, illustrated above the plots.
@@
 In other words, we used the \msc{} sequence as a neutral baseline where the target stimulus was presented no more often than in the \dev{} sequence, but no expectation of a competing stimulus B could be formed.
-To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the \dev{} and \msc{} sequences, defining a deviance detection index (DDI) as
+To quantify deviance detection, we compared the mean number of spikes fired per neuron in response to target stimulus A in the \dev{} and \msc{} sequences, denoted as $\mean{\R[dev]{A}{}}$ and $\mean{\R[msc]{A}{}}$, respectively, and define a deviance detection index (DDI) as
 \begin{equation}
@@
 \end{figure*}
-\textbf{Fig.
-2} Deviance detection in our model.
+\textbf{Fig. 2} Deviance detection in our model.
 \textbf{a} Sample network target response intensity by trial type.
@@
 \end{figure*}
-\textbf{Fig.
-3} Deviance detection under model ablation.
+\textbf{Fig. 3} Deviance detection under model ablation.
 \textbf{a} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
@@
 As shown in Figure \ref{fig4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
-The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, indicating higher \dev{} activity.
+The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, which indicate higher \dev{} activity.
 Since the stimulus was identical in both sequences, and TA was the only plasticity mechanism in play, these differences had to be due to (1) different firing thresholds, and/or (2) different synaptic currents as a consequence of different prior activity within the trial.
@@
 \textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower initial $\theta$.
-\textbf{a} Target trial average post-stimulus spike histogram in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
-Left and middle column, target trials in \dev{} sequence ($\R[dev]{A}{i}$) and \msc{} sequence ($\R[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \R{A}{i}$).
+\textbf{a} Post-stimulus spike histogram across target trials in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
+Left and middle column, target trials in the \dev{} sequence and \msc{} sequence, respectively; right column, contrast between these two.
 \textbf{b} Target trial average of the threshold adaptation voltage $\theta$, with neuron order and columns as in panel a.
@@
 The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the colormap.
-\textbf{d} Relationship across excitatory neurons of the sample network between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$, and associated linear regression.
+\textbf{d} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $\theta$ at the start of target trials ($\Delta \T{A}{i}$) and the contrast in the average number of spikes fired in target trials ($\Delta \R{A}{i}$), and associated linear regression.
 \textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \T{A}{i}$ vs $\Delta \R{A}{i}$ across excitatory neurons).
@@
 
-To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we referred to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
+To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we refer to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
 As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
@@
 Based on these measures, we then asked why there were more spikes in \dev{} trials.
-We first focused on the contribution of $\theta$ alone in an effort to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in Figure \ref{fig4}b, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$, across excitatory neurons.
+We first focused on the contribution of $\theta$ alone to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
+We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in Figure \ref{fig4}b, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
 In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} (T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
 In addition, the correlation across excitatory neurons between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
@@
 In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
-We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference, and summed across neurons.
+We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference and summed across neurons.
 The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased \dev{} response.
@@
 Having established that the increased response to A in \dev{} trials was driven by lower thresholds before stimulation, we next turned to the cause of this reduction.
-A priori, since threshold adaptation is a direct reflection of activity, and since the response to target trials was larger in \dev{} than in \msc{}, the lower thresholds in \dev{} must have been caused by lower activity in non-target trials, i.e.
-lower responses to B in \dev{} than to B through E in \msc{}.
+A priori, since threshold adaptation is a direct reflection of activity, and since the response to target trials was larger in \dev{} than in \msc{}, the lower thresholds in \dev{} must have been caused by lower activity in non-target trials, i.e., by lower responses to B in \dev{} than to B through E in \msc{}.
 Yet, since activity is modulated by thresholds as shown above, and since the dev sequence was random, implying that $\T[dev]{A}{} \approx \T[dev]{B}{}$, we might expect the lowered thresholds in \dev{} to yield not only higher target responses, but also higher non-target responses.
@@
 \end{figure*}
-\textbf{Fig. 5} Lower $\T[dev]{}{}$ is a consequence of lower \dev{} non-target response magnitude.
-\textbf{a} $\T{A}{i}$, i.e., $\theta_i$ at the start of A trials (mean over trials), in the sample network, with each neuron shown in its spatial location.
+\textbf{Fig. 5} Lower thresholds in \dev{} are a consequence of lower \dev{} non-target response magnitude.
+\textbf{a} $\T{A}{i}$ in the sample network, with each neuron shown in its spatial location.
 Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
@@
 In the \msc{} sequence, $\T[msc]{A}{i}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
-Conversely, in the \dev{} sequence, neurons near the non-target stimulation site (B) were strongly adapted, whereas the majority of the network was less adapted, consistent with the data shown above.
+Conversely, in the \dev{} sequence, a small number of neurons near the non-target stimulation site (B) were strongly adapted, whereas most of the network was less adapted, consistent with the data shown above.
 
@@
 Note that non-responding neurons are ranked last and appear dark blue.
-\textbf{b} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$, and associated linear regression across all excitatory neurons.
+\textbf{b} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$, and associated linear regression.
 Neurons are colored by latency rank as in panel a.
@@
 To show this, we sought to isolate this subset of early responders to B.
-In a given network and stimulus assignment, we found, for each neuron, the minimum response time (i.e., time to first spike) across B trials in all sequences, and ranked neurons by this response time.
+In a given network and stimulus assignment, we ranked each neuron by the minimum response time (i.e., time to first spike) across B trials in all sequences.
 The ranking in the sample network is shown in Figure \ref{fig6}a.
@@
 Across networks and stimuli, we found even stronger evidence that the early portion of the network drove adaptation to B.
-We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992, T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
+We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992: T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, Figure \ref{fig6}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, Figure \ref{fig6}d) within these two portions.
-We found that the response in the early portion was consistently greater in \dev{} than \msc{} (T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased (T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
 $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
 
-Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in Figure \ref{fig6}e.
+Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in Figure \ref{fig6}e.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
@@
 
-Thus, we have finally traced the greater response to A in the \dev{} sequence to its roots, and revealed two complementary roles of TA: The frequent presentation of non-target stimulus B in the \dev{} sequence and concomitant increased activity in a subset of neurons responding early to B causes \emph{local suppression} by TA in this subset, leading to smaller network responses.
+Thus, we have finally traced the greater response to A in the \dev{} sequence to its roots and revealed two complementary roles of TA: The frequent presentation of non-target stimulus B in the \dev{} sequence and concomitant increased activity in a subset of neurons responding early to B causes \emph{local suppression} by TA in this subset, leading to smaller network responses.
 By contrast, stimuli in the \msc{} sequence are presented more sporadically, allowing local suppression to decay, thus leading to larger network responses.
@@
 
-We had good reasons for this assumption: A priori, updates to the spike-triggered depression variable $x_j$ (cf.
+We had good reasons for this assumption: Updates to the spike-triggered depression variable $x_j$ (cf.
 Equation \ref{eqn-xsyn}) would decay to a negligible $e^\frac{-1 s}{\tau_x} = 0.0013$ times the original value within two trials, making it unlikely that inputs separated by intervening trials could be "remembered" in the network.
@@
 
-To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated the number of spikes in \dev{} target trials with $x_j$ at the start of later trials across excitatory neurons.
+To show in the sample network that the model does not maintain a history beyond the most recent trial, we correlated $\R[dev]{A}{j}$ with $x_j$ at the start of later trials across excitatory neurons.
 We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
@@
 To show that STD being carried over between repeated trials was responsible for the reduction of non-target \dev{} responses relative to \msc{}, however, we needed to estimate the extent to which activity of a given neuron (or, alternatively, its membrane potential $V_m$) was affected by changes in STD.
-Unlike $\T{}{j}$, which directly affects the neuron to which the value belongs, $x_j$ exerts its effects on the other side of a synapse.
-Therefore, we devised a measure $\D{}{}$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring the effect of $\T{}{}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
+Unlike $\theta_j$, which directly affects the activity of neuron $j$, $x_j$ only exerts its effects postsynaptically.
+Therefore, we devised a measure $\D{}{}$ to estimate the postsynaptic impact of STD in terms of membrane potential lost, mirroring $\T{}{}$ in the sense that decreasing $V_m$ and increasing the firing threshold both increase the depolarization necessary to make a neuron fire.
 
 A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
-In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $\R{S}{j} = \frac{1}{3} (\R[dev]{S}{j} + \R[msc]{S}{j} + \R[std]{S}{j})$ of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive.
+In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $\Rhat{S}{j} = \frac{1}{3} (\R[dev]{S}{j} + \R[msc]{S}{j} + \R[std]{S}{j})$ of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive.
 Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.
-Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $\R{S}{j} x_j^S k$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials.
-Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in a given neuron $i$ as
+Thus, we estimated the average peak EPSP evoked by a spike in presynaptic neuron $j$ as $\Rhat{S}{j} x_j^S k$, where we ignored added depression due to repeated spikes within a trial and instead defined $x_j^S$ as the mean depression at the start of trials.
+Finally, summing over presynaptic neurons, we estimated the peak EPSP lost to depression in each neuron $i$ as
 \begin{equation}
-    \D{S}{i} = \sum_{j \in pre} \R{S}{j} (1-x_j^S) k \label{eqn-D}
+    \D{S}{i} = \sum_{j \in pre} \Rhat{S}{j} (1-x_j^S) k \label{eqn-D}
 \end{equation}
@@
 As noted above, we hypothesized that STD is directly responsible only for a reduction of non-target \dev{} trials, relative to \msc{}, whereas differences between target \dev{} and \msc{} trials are not a result of STD directly.
-To show this, we calculated $\D{}{}$ for A and B trials, and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
+To show this, we calculated $\D{}{}$ for A and B trials and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
 In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
@@
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
-We can relate this directly to the non-target response contrasts in the early and late portions shown in Figure \ref{fig7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local suppression in terms of $\D{}{}$.
+We can relate this directly to the non-target response contrasts in the early and late portions shown in Figure \ref{fig7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
 % Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
@@
 If so, we should expect to see less $\D{A}{}$ in \dev{} than \msc{}.
-To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli, and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
+To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
 We conclude that there is clear evidence for greater STD-mediated global suppression in \msc{} than \dev{} in the full model.
@@
 
-\section{Supplementary infomation}\label{secA1}
+\section{Supplementary information}\label{secA1}
 

### 2023-01-20 styling.py
Drop label, annotation font size to 7

@@ -11,9 +11,9 @@ sns.set_theme(style='ticks', font_scale=0.75, rc={
     'text.usetex': False,
     'pdf.fonttype': 42,
     'ps.fonttype': 42,
-    'font.size': 8,
-    'axes.labelsize': 8,
-    'axes.titlesize': 9,
+    'font.size': 7,
+    'axes.labelsize': 7,
+    'axes.titlesize': 8,
     'axes.labelpad': 2,
     'axes.linewidth': 0.5,
     'axes.titlepad': 4,
@@ -21,7 +21,7 @@ sns.set_theme(style='ticks', font_scale=0.75, rc={
     'lines.linewidth': 0.5,
     'patch.linewidth': 0.5,
     'legend.fontsize': 6,
-    'legend.title_fontsize': 8,
+    'legend.title_fontsize': 7,
     'xtick.labelsize': 6,
     'ytick.labelsize': 6,
     'xtick.major.size': 2,
@@ -61,4 +61,4 @@ sns.set_theme(style='ticks', font_scale=0.75, rc={
 })
 
 label_kwargs = {'fontsize': 12, 'fontweight': 'bold'}
-annotation_kwargs = {'fontsize': 8}
+annotation_kwargs = {'fontsize': 7}

### 2023-01-25 paper-1/manuscript.tex
Editing

@@
 
-In the following, we will first build an understanding of how TA leads to deviance detection when acting alone, i.e.
-in the ablated TA only model, working backwards from the increased target response in \dev{}, identifying and tracing causal factors step by step.
+In the following, we will first build an understanding of how TA leads to deviance detection when acting alone, i.e., in the ablated TA only model, working backwards from the increased target response in \dev{}, identifying and tracing causal factors step by step.
 Then, we will build on this understanding to elucidate how the addition of STD, which was shown above to be ineffective for deviance detection on its own, can lead to an increased deviant response in the full model.

### 2023-01-27 filterrefs.py
Added bib minimizer script

@@ -1,0 +1,49 @@
+import sys
+import re
+
+
+if __name__ == '__main__':
+    texfile = sys.argv[1]
+    bibfile = sys.argv[2]
+    if not texfile.endswith('.tex') or not bibfile.endswith('.bib'):
+        print('Use: filterrefs.py document.tex bibliography.bib')
+        exit()
+
+    keys = []
+    pattern = r'\\citep?\{([^\}]+)\}'
+    with open(texfile) as tex:
+        doc = tex.read()
+    for match in re.finditer(pattern, doc, re.MULTILINE):
+        for key in match.group(1).split(','):
+            key = key.strip()
+            if key not in keys:
+                keys.append(key)
+
+    minbib = []
+    visited_keys = []
+    drop = False
+    print('Removed references from bib:\n')
+    with open(bibfile) as bib:
+        for line in bib.readlines():
+            if line.startswith('@') and '{' in line:
+                key = line[line.find('{') + 1:].rstrip().rstrip(',')
+                if key not in keys:
+                    drop = True
+                    print(key)
+                elif key not in visited_keys:
+                    visited_keys.append(key)
+            
+            if not drop:
+                minbib.append(line)
+            
+            if drop and line.startswith('}'):
+                drop = False
+    
+    with open(bibfile, 'w') as bib:
+        bib.writelines(minbib)
+    
+    undefined_refs = set(keys) - set(visited_keys)
+    if undefined_refs:
+        print('\nUndefined references in tex:\n')
+        for key in undefined_refs:
+            print(key)

### 2023-01-30 paper-1/manuscript.tex
Bugfix
Added introduction

@@
 
+Deviance detection (DD) is the ability of neural systems to identify statistically surprising sensory inputs or violations of established regularities and to respond to this surprise with an increased signal.
+Deviance detection is a critical component of any signal processing cascade, and is closely related to the concept of error signals in the predictive processing framework \citep{Bendixen2012-nx, Khouri2015-gr, Fong2020-em}.
+DD has been studied particularly in auditory processing with approaches from single-unit recordings \citep{Taaseh2011-gg} to EEG {Winkler2012-pr}.
+Impairments in DD have been reported in various disorders \citep{Fong2020-em}, including autism \citep{Schwartz2018-bc} and schizophrenia \citep{Tada2019-lj}, highlighting the need to better understand its mechanistic underpinnings.
+
+Existing models of DD focus on synaptic plasticity, either short-term \citep{Mill2011-ah, May2015-lt, Yarden2017-eh} or long-term \citep{Hertag2020-kc, Wacongne2012-ah}, and most models make use of strongly regimented network structure to harness lateral inhibition \citep{Garagnani2011-eu, Hertag2020-kc}, feedback inhibition \citep{Wacongne2012-ah}, or depression in feedforward pathways \citep{Mill2011-ah} as additional mechanisms to support DD.
+However, recent evidence from dissociated cell culture \citep{Kubota2021-dx} suggests that even randomly connected networks of cortical neurons are capable of deviance detection, implying that detailed network structure is not a requirement.
+
+Neuronal adaptation, i.e., changes in intrinsic neuronal sensitivity as opposed to changes to synaptic function, is a common phenomenon in excitatory neurons \citep{Sanchez-Vives2000-df, OLeary2010-ar, Sanchez-Aguilera2014-fd}.
+While neuronal rate adaptation has been successfully used to model SSA \citep{Garagnani2011-eu}, it has been suggested that this mechanism is incapable of producing true deviance detection due to a lack of stimulus sensitivity \citep{Yarden2017-eh}.
+However, to the extent that individual neurons are sensitive to different stimuli, we argue that neuronal plasticity is equally well placed to produce DD as synaptic plasticity.
+
+Here, we investigate how neuronal short-term plasticity, expressed as spike-triggered threshold adaptation, can lead to deviance detection.
+Specifically, we use a spatially constrained, randomly connected network of point neurons with threshold adaptation, and show that deviance detection emerges from this mechanism alone, requiring neither synaptic plasticity nor a tightly defined network structure.
+Furthermore, we model the conjunction of both neuronal and synaptic short-term plasticity, and study how these mechanisms interact within the network as it processes regular and irregular stimulus sequences.
+
+% Interactions between forms of plasticity:
+% Single-cell level, in physiology: Buonomano1998-hf
+% Single-cell modelled TA/STD interaction => oscillations, Mejias2011-rv
+% At a single synapse: Tsodyks1997-qt, Varela1997-nr, Hennig2008-sc, Mondal2022-av
+% Short / long, in physiology: Davies1991-kf, Iezzi2011-vb; in modelling Deperrois2020-sp
+% Sequence processing = different responses to repeats of the same input with STD/STF: cortical circuit model Buonomano1995-gq, hippocampus physiology Buonomano1997-ms
+% Gating/Filtering by balance of depression and facilitation: Anwar2017-qg, Mondal2022-av
+
+% Short-term plasticity as a memory mechanism: Jaaskelainen2011-cb
+
 \section{Methods}\label{sec-methods}
@@
 \end{equation}
-with $tau_m = 30$~ms the membrane time constant and $V_{rest} = -60$~mV the resting membrane potential. Synaptic currents were modeled in a conductance-based manner, following
+with $\tau_m = 30$~ms the membrane time constant and $V_{rest} = -60$~mV the resting membrane potential. Synaptic currents were modeled in a conductance-based manner, following
 \begin{equation}

### 2023-01-30 filterrefs.py
Improved bib filter

@@ -21,23 +21,24 @@ if __name__ == '__main__':
 
     minbib = []
     visited_keys = []
-    drop = False
-    print('Removed references from bib:\n')
+    drop = 0
     with open(bibfile) as bib:
         for line in bib.readlines():
             if line.startswith('@') and '{' in line:
                 key = line[line.find('{') + 1:].rstrip().rstrip(',')
                 if key not in keys:
-                    drop = True
-                    print(key)
+                    drop = -1
                 elif key not in visited_keys:
                     visited_keys.append(key)
             
             if not drop:
                 minbib.append(line)
             
-            if drop and line.startswith('}'):
-                drop = False
+            if drop:
+                if line.startswith('}'):
+                    drop = 1
+                elif drop > 0:
+                    drop -= 1
     
     with open(bibfile, 'w') as bib:
         bib.writelines(minbib)

### 2023-01-30 paper-1/manuscript.tex
+keywords
+abstract

@@
 
-\abstract{TODO}
+\abstract{
+    Sensory areas of cortex respond more strongly to infrequent stimuli when these violate previously established regularities, a phenomenon known as deviance detection (DD).
+Previous modeling work has mainly attempted to explain DD on the basis of synaptic plasticity and tightly constrained network structure.
+However, a large fraction of cortical neurons also exhibit intrinsic plasticity, an underexplored potential mechanism, and recent results show DD in dissociated neuronal culture, raising doubts about the necessity of network structure.
+    Here, we investigate DD in a spiking neuronal model with spatially constrained connectivity and two types of short-term plasticity, fast synaptic short-term depression (STD) and slower threshold adaptation (TA).
+We probe the model with oddball sequences at an intermediate inter-stimulus interval and assess DD.
+    We find that TA alone is sufficient to generate DD signals, and identify an interplay between suppression and its recovery as the underlying mechanism.
+    Further, we find that the full model with both STD and TA yields a DD signal that exceeds the sum of the DD signals in ablated models with either STD or TA turned off.
+We identify the cause of this in an interaction where the fast-decaying STD increases local suppression near the frequently stimulated neurons, reducing the stimulus response, and thereby indirectly promotes global recovery of TA.
+    We suggest that similar interactions between various forms of plasticity are likely to shape how sensory networks process regularities.
+}
 
-\keywords{keyword1, Keyword2, Keyword3, Keyword4}
+\keywords{deviance detection, adaptation, short-term plasticity, neuronal plasticity, synaptic plasticity}
 

### 2023-01-30 paper-1/manuscript.tex
Editing

@@
 A priori, since threshold adaptation is a direct reflection of activity, and since the response to target trials was larger in \dev{} than in \msc{}, the lower thresholds in \dev{} must have been caused by lower activity in non-target trials, i.e., by lower responses to B in \dev{} than to B through E in \msc{}.
-Yet, since activity is modulated by thresholds as shown above, and since the dev sequence was random, implying that $\T[dev]{A}{} \approx \T[dev]{B}{}$, we might expect the lowered thresholds in \dev{} to yield not only higher target responses, but also higher non-target responses.
+Yet, since activity was modulated by thresholds as shown above, and since the dev sequence was random, implying that $\T[dev]{A}{} \approx \T[dev]{B}{}$, we might expect the lowered thresholds in \dev{} to yield not only higher target responses, but also higher non-target responses.
 To resolve this apparent paradox, we will turn to an analysis of the spatial arrangement of activity and adaptation.
@@
 Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
-We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B, and therefore more to non-target stimuli in \dev{} than \msc{}, and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B (and therefore more to non-target stimuli in \dev{} than \msc{}), and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
 
@@
 A measure of STD is necessarily activity-dependent -- a postsynaptic neuron's membrane potential is driven only by presynaptic partners that fire -- so it must be calculated based on observed activity.
-In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response $\Rhat{S}{j} = \frac{1}{3} (\R[dev]{S}{j} + \R[msc]{S}{j} + \R[std]{S}{j})$ of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive.
+In order to respect this notion while at the same time minimizing the influence of activity differences between sequences, we used the average response
+\begin{equation}
+    \Rhat{S}{j} = \frac{1}{3} (\R[dev]{S}{j} + \R[msc]{S}{j} + \R[std]{S}{j}) \label{eqn-Rhat}
+\end{equation}
+of a neuron $j$ to a given stimulus $S$ as a measure of the presynaptic drive.
 Before depression, a single excitatory presynaptic spike evokes a postsynaptic potential (EPSP) peaking at approximately $k = 1.4$ mV at $V_m = V_{rest}$ (recall that $w = 1$ for all synapses), and STD linearly scales this EPSP by presynaptic $x_j$.

### 2023-01-30 paper-1/manuscript.tex
Rev sec Notation

@@
 Throughout this paper, we refer to three main quantities of interest: the stimulus response magnitude $\R{}{}$, the threshold adaptation value $\T{}{}$, and the short-term depression value $\D{}{}$.
-We define the response $\R{X}{i}$ as the mean number of spikes fired by neuron $i$ across all trials of type $X$, i.e., the average response to $X$.
-We define $\T{X}{i}$ as the mean across trials of type $X$ of $\theta_i(t=0)$, i.e., the average TA voltage at the beginning of trials, immediately before stimulation.
-Lastly, the STD quantity $\D{X}{i}$ will be introduced later in Equation \ref{eqn-D}.
+We define the response $\R{}{}$ as the mean number of spikes fired across all indicated trials (see below), i.e., the average trial response.
+We define $\T{}{}$ as the mean across trials of $\theta(t=0)$, i.e., the average TA voltage at the beginning of trials, immediately before stimulation.
+Lastly, the related STD quantity $\D{}{}$ will be introduced later in Equation \ref{eqn-D}.
 
-We employ sub- and superscripts to identify the precise quantity as follows, using $\R{}{}$ for illustration: $\R[seq]{X}{p}$ refers to the response of neuron(s) $p$, averaged across all trials of type $X$ in the sequence $seq$.
-We use $p$ to either identify individual neurons (e.g., $i$ or $j$), or a subpopulation (e.g., $exc$) or the whole network (no index), in which case the designated quantity is the median (of the individual quantities $\R[seq]{X}{i}$) across the indicated set of neurons.
+We employ sub- and superscripts to indicate the precise quantity as follows, using $\R{}{}$ for illustration: $\R[seq]{X}{p}$ refers to the response of neuron(s) $p$, averaged across all trials of type $X$ in the sequence $seq$.
+We use $p$ to either identify individual neurons (e.g., $i$ or $j$), or a subpopulation (e.g., $exc$) or the whole network (no index), in which case the designated quantity is the median of the individual quantities $\R[seq]{X}{i}$ across the indicated set of neurons.
 The trial type $X$ is either $A$ or $B$, referring to trials with stimulation at these sites, or $N$, indicating all \textit{n}on-target trials B through E.

### 2023-01-30 paper-1/manuscript.tex
Editing

@@
 We probe the model with oddball sequences at an intermediate inter-stimulus interval and assess DD.
-    We find that TA alone is sufficient to generate DD signals, and identify an interplay between suppression and its recovery as the underlying mechanism.
+    We find that TA alone is sufficient to generate DD signals and identify an interplay between suppression and its recovery as the underlying mechanism.
     Further, we find that the full model with both STD and TA yields a DD signal that exceeds the sum of the DD signals in ablated models with either STD or TA turned off.
@@
 Deviance detection (DD) is the ability of neural systems to identify statistically surprising sensory inputs or violations of established regularities and to respond to this surprise with an increased signal.
-Deviance detection is a critical component of any signal processing cascade, and is closely related to the concept of error signals in the predictive processing framework \citep{Bendixen2012-nx, Khouri2015-gr, Fong2020-em}.
+Deviance detection is a critical component of any signal processing cascade and is closely related to the concept of error signals in the predictive processing framework \citep{Bendixen2012-nx, Khouri2015-gr, Fong2020-em}.
 DD has been studied particularly in auditory processing with approaches from single-unit recordings \citep{Taaseh2011-gg} to EEG {Winkler2012-pr}.

### 2023-02-01 plot_utils.py
Refactored panel labelling
This allows rapid change of labels (abc to ABC, e.g.)
It may also make labelling more robust to figure width changes.

@@ -46,6 +46,14 @@ labels = {
     'Vm': '$V_m$ (mV)'
 }
 
+def label_panel(ax, ord, lower=False, left=1):
+    lb = ax.set_title(styling.panel_labels[ord], loc='left', y=1, va='top' if lower else 'baseline', **styling.label_kwargs, ha='left')
+    bb_plotonly = ax.get_window_extent()
+    bb_withdeco = ax.get_tightbbox()
+    x = left*(bb_withdeco.xmin - bb_plotonly.xmin) / bb_plotonly.bounds[2]
+    lb.set_position((x, 1))
+    return lb
+
 
 # From http://www.ccctool.com/html_v_0_9_0_3/CCC_Tool/cccTool.html
 ccc_divergent = LinearSegmentedColormap.from_list(

### 2023-02-01 styling.py

@@ -62,3 +62,6 @@ sns.set_theme(style='ticks', font_scale=0.75, rc={
 
 label_kwargs = {'fontsize': 12, 'fontweight': 'bold'}
 annotation_kwargs = {'fontsize': 7}
+
+# panel_labels = 'ABCDEFGHIJK'
+panel_labels = 'abcdefghijk'

### 2023-02-06 paper-1/manuscript.tex
Housekeeping
Rev abstract

@@
 
-%%%%%=============================================================================%%%%
-%%%%  Remarks: This template is provided to aid authors with the preparation
-%%%%  of original research articles intended for submission to journals published 
-%%%%  by Springer Nature. The guidance has been prepared in partnership with 
-%%%%  production teams to conform to Springer Nature technical requirements. 
-%%%%  Editorial and presentation requirements differ among journal portfolios and 
-%%%%  research disciplines. You may find sections in this template are irrelevant 
-%%%%  to your work and are empowered to omit any such section if allowed by the 
-%%%%  journal you intend to submit to. The submission guidelines and policies 
-%%%%  of the journal take precedence. A detailed User Manual is available in the 
-%%%%  template package for technical guidance.
-%%%%%=============================================================================%%%%
-
 \jyear{2021}%
-
-%% as per the requirement new theorem styles can be included as shown below
-\theoremstyle{thmstyleone}%
-\newtheorem{theorem}{Theorem}%  meant for continuous numbers
-%%\newtheorem{theorem}{Theorem}[section]% meant for sectionwise numbers
-%% optional argument [theorem] produces theorem numbering sequence instead of independent numbers for Proposition
-\newtheorem{proposition}[theorem]{Proposition}% 
-%%\newtheorem{proposition}{Proposition}% to get separate numbers for theorem and proposition etc.
-
-\theoremstyle{thmstyletwo}%
-\newtheorem{example}{Example}%
-\newtheorem{remark}{Remark}%
-
-\theoremstyle{thmstylethree}%
-\newtheorem{definition}{Definition}%
 
@@
 
-%%=============================================================%%
-%% Prefix	-> \pfx{Dr}
-%% GivenName	-> \fnm{Joergen W.}
-%% Particle	-> \spfx{van der} -> surname prefix
-%% FamilyName	-> \sur{Ploeg}
-%% Suffix	-> \sfx{IV}
-%% NatureName	-> \tanm{Poet Laureate} -> Title after name
-%% Degrees	-> \dgr{MSc, PhD}
-%% \author*[1,2]{\pfx{Dr} \fnm{Joergen W.} \spfx{van der} \sur{Ploeg} \sfx{IV} \tanm{Poet Laureate} 
-%%                 \dgr{MSc, PhD}}\email{iauthor@gmail.com}
-%%=============================================================%%
-
 \author[1]{\fnm{Felix Benjamin} \sur{Kern} \orcidA{}}\email{kernfel@gmail.com}
@@
 
-%%==================================%%
-%% sample for unstructured abstract %%
-%%==================================%%
-
 \abstract{
@@
 Previous modeling work has mainly attempted to explain DD on the basis of synaptic plasticity and tightly constrained network structure.
-However, a large fraction of cortical neurons also exhibit intrinsic plasticity, an underexplored potential mechanism, and recent results show DD in dissociated neuronal culture, raising doubts about the necessity of network structure.
-    Here, we investigate DD in a spiking neuronal model with spatially constrained connectivity and two types of short-term plasticity, fast synaptic short-term depression (STD) and slower threshold adaptation (TA).
+However, a large fraction of cortical neurons also exhibit intrinsic plasticity, an underexplored potential mechanism.
+    Here, we investigate DD in a spiking neuronal model with simple spatially constrained connectivity and two types of short-term plasticity, fast synaptic short-term depression (STD) and slower threshold adaptation (TA).
 We probe the model with oddball sequences at an intermediate inter-stimulus interval and assess DD.
@@
     Further, we find that the full model with both STD and TA yields a DD signal that exceeds the sum of the DD signals in ablated models with either STD or TA turned off.
-We identify the cause of this in an interaction where the fast-decaying STD increases local suppression near the frequently stimulated neurons, reducing the stimulus response, and thereby indirectly promotes global recovery of TA.
+We identify the cause of this in an interaction between the contribution of STD to the local suppression of neurons sensitive to the frequently presented stimulus, and the recovery of TA-derived global suppression across the network.
     We suggest that similar interactions between various forms of plasticity are likely to shape how sensory networks process regularities.

### 2023-02-06 paper-1/manuscript.tex
Rev intro

@@
 
-Existing models of DD focus on synaptic plasticity, either short-term \citep{Mill2011-ah, May2015-lt, Yarden2017-eh} or long-term \citep{Hertag2020-kc, Wacongne2012-ah}, and most models make use of strongly regimented network structure to harness lateral inhibition \citep{Garagnani2011-eu, Hertag2020-kc}, feedback inhibition \citep{Wacongne2012-ah}, or depression in feedforward pathways \citep{Mill2011-ah} as additional mechanisms to support DD.
-However, recent evidence from dissociated cell culture \citep{Kubota2021-dx} suggests that even randomly connected networks of cortical neurons are capable of deviance detection, implying that detailed network structure is not a requirement.
+Existing models of DD place their primary focus on synaptic plasticity.
+While models using long-term plasticity have been proposed \citep{Wacongne2012-ah, Hertag2020-kc}, the rapid establishment of DD in vivo \citep{Taaseh2011-gg} suggests that fast-acting short-term plasticity mechanisms play a key role.
+Several models using synaptic short-term plasticity have been proposed \citep{Mill2011-ah, May2015-lt, Yarden2017-eh}, finding good agreement with physiological data.
 
-Neuronal adaptation, i.e., changes in intrinsic neuronal sensitivity as opposed to changes to synaptic function, is a common phenomenon in excitatory neurons \citep{Sanchez-Vives2000-df, OLeary2010-ar, Sanchez-Aguilera2014-fd}.
-While neuronal rate adaptation has been successfully used to model SSA \citep{Garagnani2011-eu}, it has been suggested that this mechanism is incapable of producing true deviance detection due to a lack of stimulus sensitivity \citep{Yarden2017-eh}.
-However, to the extent that individual neurons are sensitive to different stimuli, we argue that neuronal plasticity is equally well placed to produce DD as synaptic plasticity.
+Neuronal plasticity, on the other hand, has largely been neglected in previous work, used as a mere afterthought in the neuron model \citep{Mill2011-ah} or rejected a priori as incapable of producing DD \citep{Yarden2017-eh}.
+However, neuronal adaptation, i.e., an experience-dependent reduction in intrinsic neuronal sensitivity independent of synaptic function, is a common phenomenon in excitatory neurons \citep{Sanchez-Vives2000-df, Henze2001-xd, Sanchez-Aguilera2014-fd, Teeter2018-iz} that has been successfully used to model stimulus-specific adaptation (SSA), a precursor to and prerequisite for DD \citep{Garagnani2011-eu}.
+Neuronal adaptation is well placed to play a role in DD: It establishes quickly and recovers over the course of up to several seconds \citep{Henze2001-xd, Mensi2012-au}, allowing it to act as a memory of recent activity in a neuron.
 
-Here, we investigate how neuronal short-term plasticity, expressed as spike-triggered threshold adaptation, can lead to deviance detection.
-Specifically, we use a spatially constrained, randomly connected network of point neurons with threshold adaptation, and show that deviance detection emerges from this mechanism alone, requiring neither synaptic plasticity nor a tightly defined network structure.
+Here, we investigate how neuronal threshold adaptation can lead to deviance detection.
+Specifically, we use a spatially constrained, randomly connected network of point neurons with spike-triggered threshold adaptation, and show that deviance detection emerges from this mechanism alone, requiring neither synaptic plasticity nor a tightly defined network structure.
 Furthermore, we model the conjunction of both neuronal and synaptic short-term plasticity, and study how these mechanisms interact within the network as it processes regular and irregular stimulus sequences.

### 2023-02-06 paper-1/manuscript.tex
Fixed TA refs

@@
 However, neuronal adaptation, i.e., an experience-dependent reduction in intrinsic neuronal sensitivity independent of synaptic function, is a common phenomenon in excitatory neurons \citep{Sanchez-Vives2000-df, Henze2001-xd, Sanchez-Aguilera2014-fd, Teeter2018-iz} that has been successfully used to model stimulus-specific adaptation (SSA), a precursor to and prerequisite for DD \citep{Garagnani2011-eu}.
-Neuronal adaptation is well placed to play a role in DD: It establishes quickly and recovers over the course of up to several seconds \citep{Henze2001-xd, Mensi2012-au}, allowing it to act as a memory of recent activity in a neuron.
+Neuronal adaptation is well placed to play a role in DD: It establishes quickly and recovers over the course of up to several seconds \citep{Henze2001-xd, Pozzorini2015-ei}, allowing it to act as a memory of recent activity in a neuron.
 
@@
 \end{align}
-with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \citep{Pozzorini2015-ei}.
+with increment $\hat{\theta} = 1$~mV and decay time constant $\tau_{\theta} = 1$~s \citep{Henze2001-xd, Pozzorini2015-ei}.
 

### 2023-02-06 paper-1/manuscript.tex
editing

@@
 Here, we investigate how neuronal threshold adaptation can lead to deviance detection.
-Specifically, we use a spatially constrained, randomly connected network of point neurons with spike-triggered threshold adaptation, and show that deviance detection emerges from this mechanism alone, requiring neither synaptic plasticity nor a tightly defined network structure.
+Specifically, we use a spatially constrained, randomly connected network of point neurons with spike-triggered threshold adaptation and show that deviance detection emerges from this mechanism alone, requiring neither synaptic plasticity nor a tightly defined network structure.
 Furthermore, we model the conjunction of both neuronal and synaptic short-term plasticity, and study how these mechanisms interact within the network as it processes regular and irregular stimulus sequences.

### 2023-02-06 paper-1/manuscript.tex
Missed citation
command-based fig/panel refs

@@
 
+% Figure references
+\newcommand{\reffig}[1]{Figure \ref{fig#1}}
+\newcommand{\refpanel}[2]{Figure \ref{fig#1}\lowercase{#2}}
+
 \jyear{2021}%
@@
 Deviance detection is a critical component of any signal processing cascade and is closely related to the concept of error signals in the predictive processing framework \citep{Bendixen2012-nx, Khouri2015-gr, Fong2020-em}.
-DD has been studied particularly in auditory processing with approaches from single-unit recordings \citep{Taaseh2011-gg} to EEG {Winkler2012-pr}.
+DD has been studied particularly in auditory processing with approaches from single-unit recordings \citep{Taaseh2011-gg} to EEG \citep{Winkler2012-pr}.
 Impairments in DD have been reported in various disorders \citep{Fong2020-em}, including autism \citep{Schwartz2018-bc} and schizophrenia \citep{Tada2019-lj}, highlighting the need to better understand its mechanistic underpinnings.
@@
 
-Figure \ref{fig1}a shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
+\refpanel{1}{a} shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
 The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \citep{Stimberg2020-go}, and simulated with an integration time step of 1~ms.
@@
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-Figure \ref{fig1}b shows a representative example of the resulting network structure.
-In Figure \ref{fig1}c, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
+\refpanel{1}{b} shows a representative example of the resulting network structure.
+In \refpanel{1}{c}, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
 Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response: Networks where fewer than 500 neurons responded to stimulation at any of the five sites after full recovery were discarded.
@@
 
-To investigate deviance detection, we used a classical oddball paradigm as illustrated in Figure \ref{fig1}d, presenting a total of 500 stimuli at regular intervals of 500~ms.
+To investigate deviance detection, we used a classical oddball paradigm as illustrated in \refpanel{1}{d}, presenting a total of 500 stimuli at regular intervals of 500~ms.
 Target stimuli (labeled "A" or "target" hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (\std{}), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (\dev{}) with 100 A and 400 B, and the many-standards control sequence (\msc{}) with 100 presentations of each of the stimuli A through E.
@@
 
-To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in Figure \ref{fig2}a.
+To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \refpanel{2}{a}.
 Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color.
@@
 The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster.
-We sum these patterns up in Figure \ref{fig2}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
+We sum these patterns up in \refpanel{2}{b}, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
 Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.
@@
 To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences.
-Figure \ref{fig2}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
+\refpanel{2}{c} shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
 The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15).
-Accordingly, the DDI (Figure \ref{fig2}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Accordingly, the DDI (\refpanel{2}{d}) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
 This shows that our model was well behaved and suitable for an investigation of deviance detection.
@@
 Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-Figure \ref{fig3}a shows the responses to target stimulation of the sample network under ablation.
+\refpanel{3}{a} shows the responses to target stimulation of the sample network under ablation.
 With no plasticity, the responses are indistinguishable between sequences, yielding a DDI of 0, and indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
@@
 
-Finally, across networks and stimuli, the DDI (Figure \ref{fig3}c) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Finally, across networks and stimuli, the DDI (\refpanel{3}{c}) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
 To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: T = 6.1e+03, p = 4.95e-11; TA: T = 4.87e+03, p = 0.000604), and greater too than the sum of the DDI values across both STD only and TA only models (T = 5.12e+03, p = 4.52e-05).
@@
 To understand why the same stimulus A evoked a higher response in \dev{} than in \msc{} with TA only, we first compared the average responses in the two sequences, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
-As shown in Figure \ref{fig4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
+As shown in \refpanel{4}{a} (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
 The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, which indicate higher \dev{} activity.
@@
 We quantified the former with the strength of adaptation $\theta$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
-As shown in Figure \ref{fig4}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+As shown in \refpanel{4}{b}, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
@@
 To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we refer to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
-As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (Figure \ref{fig4}c) largely reflected that seen in spiking activity.
+As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\refpanel{4}{c}) largely reflected that seen in spiking activity.
 The contrast (right panel) shows not only a notable relative increase in membrane voltages in \dev{} around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
@@
 We first focused on the contribution of $\theta$ alone to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in Figure \ref{fig4}b, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
-In the sample network (Figure \ref{fig4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in Figure \ref{fig4}e, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \refpanel{4}{b}, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
+In the sample network (\refpanel{4}{d}), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+As shown in \refpanel{4}{e}, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
 In addition, the correlation across excitatory neurons between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
@@
 
-Yet, as we saw in Figure \ref{fig4}c, the membrane voltages also differed between the two sequences.
+Yet, as we saw in \refpanel{4}{c}, the membrane voltages also differed between the two sequences.
 In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
 We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference and summed across neurons.
-The resulting time courses, shown in Figure \ref{fig4}f, show how much these two measures contributed to the increased \dev{} response.
+The resulting time courses, shown in \refpanel{4}{f}, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (Figure \ref{fig4}g), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
-A similar pattern emerged across networks and stimuli, as shown in Figure \ref{fig4}h.
+In relative terms (\refpanel{4}{g}), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+A similar pattern emerged across networks and stimuli, as shown in \refpanel{4}{h}.
 Here, only the initial response was driven mainly by $\theta$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
@@
 Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
-A and B stimulus locations are highlighted in the left plot; see Figure \ref{fig1} for the locations of the remaining stimuli used in the \msc{} sequence.
+A and B stimulus locations are highlighted in the left plot; see \reffig{1} for the locations of the remaining stimuli used in the \msc{} sequence.
 \textbf{b} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel a.
@@
 
-First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in Figure \ref{fig5}a.
+First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \refpanel{5}{a}.
 In the \msc{} sequence, $\T[msc]{A}{i}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
@@
 
-Visualizing the likely cause of this difference, the non-target responses, in the same fashion (Figure \ref{fig5}b), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
+Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\refpanel{5}{b}), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
 Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.
 
-Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (Figure \ref{fig5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in Figure \ref{fig5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
+Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\refpanel{5}{c}), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in \refpanel{5}{d}, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
 
@@
 Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
-We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B (and therefore more to non-target stimuli in \dev{} than \msc{}), and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in Figure \ref{fig4}.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B (and therefore more to non-target stimuli in \dev{} than \msc{}), and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in \reffig{4}.
 
@@
 In a given network and stimulus assignment, we ranked each neuron by the minimum response time (i.e., time to first spike) across B trials in all sequences.
-The ranking in the sample network is shown in Figure \ref{fig6}a.
+The ranking in the sample network is shown in \refpanel{6}{a}.
 We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site.
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
-In Figure \ref{fig6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
-Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in Figure \ref{fig5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
+In \refpanel{6}{b}, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \refpanel{5}{c}, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
 As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red, left plot) increasing their response in the \dev{} sequence, while the response of the later portion of the network (blue, right plot) decreased.
@@
 We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992: T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
-In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, Figure \ref{fig6}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, Figure \ref{fig6}d) within these two portions.
-We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also Figure \ref{fig5}d).
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \refpanel{6}{c}) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \refpanel{6}{d}) within these two portions.
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \refpanel{5}{d}).
 $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
 
-Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in Figure \ref{fig6}e.
+Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \refpanel{6}{e}.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
 
-To sum up, we showed in Figure \ref{fig4} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
+To sum up, we showed in \reffig{4} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
 Here, we showed that the neurons responding earliest to B exhibited higher $\T[dev]{B}{}$, which analogously reduced the response of the network as a whole to B.
@@
 To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD.
-On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (Figure \ref{fig7}a).
+On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\refpanel{7}{a}).
 However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (more than \msc{}: T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (less than \msc{}: T = 2.66e+03, p = 0.00554).
@@
 \textbf{b} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
-Compare to Figure \ref{fig6}c-d.
+Compare to \refpanel{6}{c-d}.
 \textbf{c} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
@@
 \textbf{d} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
-Compare to Figure \ref{fig5}d.
+Compare to \refpanel{5}{d}.
 \textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
@@
 \textbf{f} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
-Compare to Figure \ref{fig4}e.
+Compare to \refpanel{4}{e}.
 \textbf{g} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
@@
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
-Starting at the cause of the reduced response in non-target trials, we show in Figure \ref{fig7}b the full model equivalent of Figure \ref{fig6}c and d.
+Starting at the cause of the reduced response in non-target trials, we show in \refpanel{7}{b} the full model equivalent of \refpanel{6}{c-d}.
 The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
@@
 
-Comparing the non-target response reductions in Figure \ref{fig7}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
-Based on the ablated model analysis in Figure \ref{fig5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in Figure \ref{fig7}d.
-Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in Figure \ref{fig7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Comparing the non-target response reductions in \refpanel{7}{c}, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
+Based on the ablated model analysis in \reffig{5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
+Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \refpanel{7}{d}.
+Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \refpanel{7}{e}, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
-Finally, based on the ablated model analysis in Figure \ref{fig4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
-As shown in Figure \ref{fig7}f, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in Figure \ref{fig7}g, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
-This greater contrast, of course, directly reflects the greater DDI reported in Figure \ref{fig3}.
+Finally, based on the ablated model analysis in \reffig{4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
+As shown in \refpanel{7}{f}, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \refpanel{7}{g}, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
+This greater contrast, of course, directly reflects the greater DDI reported in \reffig{3}.
 
 Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
-The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (Figure \ref{fig7}c), which was where we noted the first clear shift towards increased deviance detection.
+The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\refpanel{7}{c}), which was where we noted the first clear shift towards increased deviance detection.
 
@@
 We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
-Similar behavior was also seen in the \msc{} sequence, as shown in Figure \ref{sfig-STD-correlation-decay}.
+Similar behavior was also seen in the \msc{} sequence, as shown in \ref{sfig-STD-correlation-decay}.
 
@@
 To show this, we calculated $\D{}{}$ for A and B trials and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
-In Figure \ref{fig8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
+In \refpanel{8}{a}, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
 The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast $\Delta \D{A}{i}$, which was negative throughout.
@@
 To show that $\D{}{}$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated $\Delta \D{A}{i}$ with $\Delta \R{A}{i}$ across neurons of the sample network.
-Notice that this mirrors the correlation shown in Figure \ref{fig4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
-As shown in Figure \ref{fig8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
-We confirmed this relationship across networks and stimuli, see Figure \ref{fig8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
-We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see Figure \ref{sfig8}).
+Notice that this mirrors the correlation shown in \reffig{4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
+As shown in \refpanel{8}{b}, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
+We confirmed this relationship across networks and stimuli, see \refpanel{8}{c}, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
+We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \ref{sfig8}).
 
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in Figure \ref{fig8}d, we recapitulate the logic established in Figure \ref{fig6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
+First, in \refpanel{8}{d}, we recapitulate the logic established in \reffig{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
-We can relate this directly to the non-target response contrasts in the early and late portions shown in Figure \ref{fig7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
+We can relate this directly to the non-target response contrasts in the early and late portions shown in \refpanel{7}{a}, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
 % Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
 
-To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing Figure \ref{fig6}e.
-As we show in Figure \ref{fig8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
+To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \refpanel{6}{e}.
+As we show in \refpanel{8}{e}, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
 This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
@@
 If so, we should expect to see less $\D{A}{}$ in \dev{} than \msc{}.
-To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in Figure \ref{fig8}c.
+To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \refpanel{8}{c}.
 We conclude that there is clear evidence for greater STD-mediated global suppression in \msc{} than \dev{} in the full model.
@@
 To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli (lower $\R[dev]{B}{}$) by local suppression, and an increase in the response to target stimuli (higher $\R[dev]{A}{}$) by relief from global suppression, relative to \msc{}.
-Yet, as we showed in Figure \ref{fig7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower thresholds $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
+Yet, as we showed in \reffig{7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower thresholds $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
 Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (Figure \ref{fig7}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (Figure \ref{fig8}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\refpanel{7}{e}, bottom) with the corresponding value of $\Delta \D{A}{}$ (\refpanel{8}{e}), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.

### 2023-03-05 paper-1/manuscript.tex
More comments for discussion
Cleanup / import prep

@@
 \abstract{
-    Sensory areas of cortex respond more strongly to infrequent stimuli when these violate previously established regularities, a phenomenon known as deviance detection (DD).
-Previous modeling work has mainly attempted to explain DD on the basis of synaptic plasticity and tightly constrained network structure.
-However, a large fraction of cortical neurons also exhibit intrinsic plasticity, an underexplored potential mechanism.
-    Here, we investigate DD in a spiking neuronal model with simple spatially constrained connectivity and two types of short-term plasticity, fast synaptic short-term depression (STD) and slower threshold adaptation (TA).
-We probe the model with oddball sequences at an intermediate inter-stimulus interval and assess DD.
-    We find that TA alone is sufficient to generate DD signals and identify an interplay between suppression and its recovery as the underlying mechanism.
-    Further, we find that the full model with both STD and TA yields a DD signal that exceeds the sum of the DD signals in ablated models with either STD or TA turned off.
-We identify the cause of this in an interaction between the contribution of STD to the local suppression of neurons sensitive to the frequently presented stimulus, and the recovery of TA-derived global suppression across the network.
-    We suggest that similar interactions between various forms of plasticity are likely to shape how sensory networks process regularities.
+
 }
@@
 \section{Introduction}\label{sec-intro}
-
-Deviance detection (DD) is the ability of neural systems to identify statistically surprising sensory inputs or violations of established regularities and to respond to this surprise with an increased signal.
-Deviance detection is a critical component of any signal processing cascade and is closely related to the concept of error signals in the predictive processing framework \citep{Bendixen2012-nx, Khouri2015-gr, Fong2020-em}.
-DD has been studied particularly in auditory processing with approaches from single-unit recordings \citep{Taaseh2011-gg} to EEG \citep{Winkler2012-pr}.
-Impairments in DD have been reported in various disorders \citep{Fong2020-em}, including autism \citep{Schwartz2018-bc} and schizophrenia \citep{Tada2019-lj}, highlighting the need to better understand its mechanistic underpinnings.
-
-Existing models of DD place their primary focus on synaptic plasticity.
-While models using long-term plasticity have been proposed \citep{Wacongne2012-ah, Hertag2020-kc}, the rapid establishment of DD in vivo \citep{Taaseh2011-gg} suggests that fast-acting short-term plasticity mechanisms play a key role.
-Several models using synaptic short-term plasticity have been proposed \citep{Mill2011-ah, May2015-lt, Yarden2017-eh}, finding good agreement with physiological data.
-
-Neuronal plasticity, on the other hand, has largely been neglected in previous work, used as a mere afterthought in the neuron model \citep{Mill2011-ah} or rejected a priori as incapable of producing DD \citep{Yarden2017-eh}.
-However, neuronal adaptation, i.e., an experience-dependent reduction in intrinsic neuronal sensitivity independent of synaptic function, is a common phenomenon in excitatory neurons \citep{Sanchez-Vives2000-df, Henze2001-xd, Sanchez-Aguilera2014-fd, Teeter2018-iz} that has been successfully used to model stimulus-specific adaptation (SSA), a precursor to and prerequisite for DD \citep{Garagnani2011-eu}.
-Neuronal adaptation is well placed to play a role in DD: It establishes quickly and recovers over the course of up to several seconds \citep{Henze2001-xd, Pozzorini2015-ei}, allowing it to act as a memory of recent activity in a neuron.
-
-Here, we investigate how neuronal threshold adaptation can lead to deviance detection.
-Specifically, we use a spatially constrained, randomly connected network of point neurons with spike-triggered threshold adaptation and show that deviance detection emerges from this mechanism alone, requiring neither synaptic plasticity nor a tightly defined network structure.
-Furthermore, we model the conjunction of both neuronal and synaptic short-term plasticity, and study how these mechanisms interact within the network as it processes regular and irregular stimulus sequences.
-
-% Interactions between forms of plasticity:
-% Single-cell level, in physiology: Buonomano1998-hf
-% Single-cell modelled TA/STD interaction => oscillations, Mejias2011-rv
-% At a single synapse: Tsodyks1997-qt, Varela1997-nr, Hennig2008-sc, Mondal2022-av
-% Short / long, in physiology: Davies1991-kf, Iezzi2011-vb; in modelling Deperrois2020-sp
-% Sequence processing = different responses to repeats of the same input with STD/STF: cortical circuit model Buonomano1995-gq, hippocampus physiology Buonomano1997-ms
-% Gating/Filtering by balance of depression and facilitation: Anwar2017-qg, Mondal2022-av
-
-% Short-term plasticity as a memory mechanism: Jaaskelainen2011-cb
 
@@
 
-TODO
-
 \begin{figure*}%
@@
 \end{figure*}
-% TODO 9: Needs to be more generic, not specific to the task. Show the process, step by step.
-
-% No tuning required to get DDI!
-
-% Culture, rats, humans...
-
-% Response of a 2 $\times$ 4~mm neuronal cell culture on a multi-electrode array, adapted from \cite{Kubota2021-dx}.
-Curves represent the average number of spikes per electrode in each 10~ms bin in response to target stimulation in the indicated sequence.
-Unlike in our study, oddball sequences here were split 10\%/90\%, and 10 equally frequent stimuli were used for the msc sequence.
-
-% adaptation vs deviance detection, in the context of pred coding, means DD is derived from some pred/err interaction, cf Fong+20
-% deviant-alone always highest, cf. Taaseh2011-gg
-
-% Replacement vs adaptation is not a clean separation, since $\theta$ is affected by replacement, but modulates adaptation
-
-% The prime effect of TA (reduce std by local suppression, relieve global suppression for a larger dev) is repeated with STD, in exactly the same way, just weaker.
-% What makes the two synergise is precisely this symmetry: STD reduces std by local suppression, which relieves not only STD-related global suppression, but also TA-related global suppression.
-And, of course, vice versa!
-% On the other hand, theoretical expectations lead us to expect competition to occur in the area of local suppression: Such suppression has diminishing returns; in the extreme case, if one mechanism reduces the response to zero, the other mechanism is left without room for further reductions.
-Or: If each mechanism independently cuts the response in half, then we should expect a quarter of the response to remain
-
-% Spatial overlap
-% Structure
-% Omission
-% Time (ISI <> tau)
-% Neurobiology
-% long-term plasticity
-% Spontaneous activity
-% Prediction signals that cancel expected inputs would reduce adaptation
-% Testable predictions. Eg: Block NMDA, spontaneous activity, TA-related channels; easier: early vs late subpopulations
+\textbf{Fig. 8} TODO
 

### 2023-03-05 paper-1/manuscript.tex
Added mini-subsec about space

@@
 
+\subsection{Spatial basis of local suppression}
+
+Finally, we sought to confirm whether "local" and "global" suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network.
+To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated \R{A}{}, \T{A}{} and \D{A}{} as a function of this distance.
+Figure \refpanel{10}{a-c} shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (Figure \refpanel{10}{a}), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, \T[std]{A}{} (Figure \refpanel{10}{b}) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with \D{A}{} (Figure \refpanel{10}{c}), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+
+To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
+We then assessed the difference of the mean values of \T{A}{} and \D{A}{} across the two subsets in each sequence, as shown in Figure \refpanel{10}{d}.
+In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for \T{A}{} (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for \D{A}{} (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.
+Similarly, we found that global suppression was greater in \msc{} than in \dev{}, again for both \T{A}{} (T = 1.09e+03, p = 1.37e-11) and \D{A}{} (T = 927, p = 7.27e-13), indicating that the \msc{} sequence produced more suppression across the wider network, leading to smaller target responses than in the \dev{} sequence.
+Incidentally, this greater global suppression, caused by the varied stimulation paradigm, was also reflected in greater local suppression around the target stimulus site for \T{A}{} (T = 831, p = 1.15e-13) and \D{A}{} (T = 1.46e+03, p = 6.13e-09), providing an additional pathway for a comparatively large response in \dev{}.
+
 \section{Discussion}\label{sec-discussion}

### 2023-03-08 paper-1/manuscript.tex
Math mode for T/D/A

--- 2023-03-05 paper-1/manuscript.tex
@@
 
+\subsection{Spatial basis of local suppression}
+
+Finally, we sought to confirm whether "local" and "global" suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network.
+To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.
+Figure \refpanel{10}{a-c} shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (Figure \refpanel{10}{a}), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, $\T[std]{A}{}$ (Figure \refpanel{10}{b}) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with $\D{A}{}$ (Figure \refpanel{10}{c}), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+
+To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
+We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in Figure \refpanel{10}{d}.
+In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.
+Similarly, we found that global suppression was greater in \msc{} than in \dev{}, again for both $\T{A}{}$ (T = 1.09e+03, p = 1.37e-11) and $\D{A}{}$ (T = 927, p = 7.27e-13), indicating that the \msc{} sequence produced more suppression across the wider network, leading to smaller target responses than in the \dev{} sequence.
+Incidentally, this greater global suppression, caused by the varied stimulation paradigm, was also reflected in greater local suppression around the target stimulus site for $\T{A}{}$ (T = 831, p = 1.15e-13) and $\D{A}{}$ (T = 1.46e+03, p = 6.13e-09), providing an additional pathway for a comparatively large response in \dev{}.
+
 \section{Discussion}\label{sec-discussion}

### 2023-03-08 paper-1/manuscript.tex
Added F10 to manuscript

@@
 
+\begin{figure*}%
+    \centering
+    \includegraphics{fig10.pdf}
+    \caption{}
+    \label{fig10}
+\end{figure*}
+\textbf{Fig. 9} Suppression acts in spatially distinguishable subpopulations.
+\textbf{a} Response size $\R{A}{}$ as a function of distance from the center of the stimulation site, evaluated across all networks and stimuli in 200 bins between 0 mm and 6.5 mm, and normalized by the response size in recovered networks (initial trial).
+Each of the 200 data points represents the mean response across all relevant trials of all neurons in the corresponding distance bin.
+Left, spatial illustration, using all data with a fictive stimulus location.
+Right, same data plotted as lines.
+The vertical line indicates the chosen local/global divider.
+\textbf{b} As in panel a, but showing $\T{A}{}$ without normalization.
+\textbf{c} As in panel a, but showing $\D{A}{}$ without normalization.
+\textbf{d} Mean $\T{A}{}$ in neurons <2 mm from the stimulation site (local) and more distant neurons (global).
+Data are averaged across relevant neurons and trials in each network and stimulus, leaving 120 data points to be summarized by box plot.
+Comparisons of interest are between \std{} and \dev{} in the local subset, and between \msc{} and \dev{} in the global subset.
+\textbf{e} Corresponding mean $\D{A}{}$ in local and global portions of the network.
+
 Finally, we sought to confirm whether "local" and "global" suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network.

### 2023-03-08 paper-1/manuscript.tex
Added F9 legend

@@
 \end{figure*}
-\textbf{Fig. 8} TODO
+\textbf{Fig. 10} Illustration of the origin of DD in our model.
+\textbf{a} Left: In an oddball context, the regular inputs are encoded in the network by high local suppression, which reduces their overall response, leading to low global suppression.
+Deviant inputs encounter low local and global suppression and therefore evoke a large response.
+Right: In an irregular context, the network frequently encounters rare stimuli with low local suppression, leading to large overall responses and high global suppression, with these two (global response and suppression) counter-balancing each other.
+Deviant inputs encounter this high global suppression and therefore evoke smaller responses.
+\textbf{b} Adding additional suppression (e.g., STD) affects these situations differently.
+Left: In an oddball context, the additional suppression further reduces the regular response, permitting even greater recovery of the wider network.
+Right: In an irregular context, global suppression is instead increased, and all responses reduced.
 

### 2023-03-08 paper-1/manuscript.tex
Fixed figure refs

@@
 To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.
-Figure \refpanel{10}{a-c} shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
-Responses, normalized by the response to stimulation in the recovered networks (Figure \refpanel{10}{a}), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
-For thresholds, $\T[std]{A}{}$ (Figure \refpanel{10}{b}) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
-A similar pattern emerged with $\D{A}{}$ (Figure \refpanel{10}{c}), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+\refpanel{10}{a-c} shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (\refpanel{10}{a}), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, $\T[std]{A}{}$ (\refpanel{10}{b}) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with $\D{A}{}$ (\refpanel{10}{c}), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
 
 To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
-We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in Figure \refpanel{10}{d}.
+We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \refpanel{10}{d}.
 In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.

### 2023-03-16 paper-1/manuscript.tex
Inserted final framing sections

@@
 \abstract{
-
+    Sensory areas of cortex respond more strongly to infrequent stimuli when these violate previously established regularities, a phenomenon known as deviance detection (DD).
+Previous modeling work has mainly attempted to explain DD on the basis of synaptic plasticity.
+However, a large fraction of cortical neurons also exhibit firing rate adaptation, an underexplored potential mechanism.
+Here, we investigate DD in a spiking neuronal network model with two types of short-term plasticity, fast synaptic short-term depression (STD) and slower threshold adaptation (TA).
+We probe the model with an oddball stimulation paradigm and assess DD by evaluating the network responses.
+We find that TA is sufficient to elicit DD.
+It achieves this by suppressing neurons near the stimulation site that respond earliest to the frequently presented standard stimulus (local suppression), which diminishes the response and promotes the recovery (global suppression) of the wider network.
+Further, we find a synergy effect between STD and TA, where they interact with each other to achieve greater DD than the sum of their individual effects.
+We show that this synergy is caused by the local suppression added by STD, which inhibits the global response to the frequently presented stimulus, further reducing global suppression by TA and making the network more responsive to the deviant stimulus.
+We conclude that highly predictable information can be encoded in strong local suppression, which allows greater global recovery and subsequent heightened sensitivity for DD.
 }
@@
 \section{Introduction}\label{sec-intro}
+
+Deviance detection (DD) is the ability of neural systems to identify statistically surprising sensory inputs or violations of established regularities.
+It is a critical component of any signal processing cascade and is closely related to the concept of prediction errors in the predictive processing framework (Bendixen, SanMiguel, and Schröger 2012; Khouri and Nelken 2015; Carbajal and Malmierca 2018; Fong et al.
+2020), which has been suggested to underlie most of cognition (Friston 2005; Clark 2015).
+DD has been studied widely across sensory systems including the auditory (Nelken 2014; Escera, Leung, and Grimm 2014; Carbajal and Malmierca 2018), visual (Winkler and Czigler 2012; Pazo-Alvarez, Cadaveira, and Amenedo 2003), somatosensory (Näätänen 2009) domains and beyond.
+Impairments in DD have been reported in various disorders (Fong et al.
+2020), including autism (Schwartz, Shinn-Cunningham, and Tager-Flusberg 2018; Goris et al.
+2018; Hudac et al.
+2018; Vlaskamp et al.
+2017; Chen et al.
+2020), schizophrenia (Koshiyama et al.
+2020; H.
+K.
+Kim, Blumberger, and Daskalakis 2020; Salisbury, Coffman, and Haigh 2020; Tada et al.
+2019), ADHD (S.
+Kim et al.
+2021; Hsieh, Chien, and Gau 2021; Lee et al.
+2020) and others, highlighting the need to better understand its mechanistic underpinnings.
+
+Studies of DD must carefully distinguish between repetition suppression and true deviance detection (Ross and Hamm 2020).
+The former refers to the reduction of the response to a frequently presented stimulus and can be explained by any adaptive mechanism (May and Tiitinen 2010; Garagnani and Pulvermüller 2011).
+The latter refers to the enhancement of the response to an infrequent stimulus when it is statistically surprising, e.g., due to interrupting a series of regular stimuli, compared to an unsurprising control condition.
+Here, we study the latter, true DD, which has been suggested to rely on a comparison between prediction signals and received inputs (Parras et al.
+2017; Carbajal and Malmierca 2018; Ross and Hamm 2020).
+
+Existing computational models of DD place their primary focus on synaptic plasticity as the causal mechanism.
+While models using long-term plasticity have been proposed (Wacongne, Changeux, and Dehaene 2012; Hertäg and Sprekeler 2020), the rapid establishment of DD in vivo (Taaseh, Yaron, and Nelken 2011) suggests that fast-acting short-term plasticity mechanisms play a key role.
+Several models using short-term synaptic plasticity have been proposed (Mill et al.
+2011; May, Westö, and Tiitinen 2015; Yarden and Nelken 2017), finding good agreement with physiological data.
+However, short-term plasticity also occurs in neuronal excitability, with many neurons exhibiting an experience-dependent transient reduction in intrinsic neuronal sensitivity independent of synaptic function (Sanchez-Vives, Nowak, and McCormick 2000; Henze and Buzsáki 2001; Sánchez-Aguilera et al.
+2014).
+Despite operating at similar time scales to short-term synaptic plasticity, the effects of intrinsic neuronal plasticity on DD is unexplored.
+Previous modeling work has only used neuronal plasticity as a side effect of the chosen neuron model (Mill et al.
+2011) or rejected it a priori as incapable of producing true DD (Yarden and Nelken 2017).
+One exception is the work by Garagnani and Pulvermüller (2011), which showed that neuronal plasticity in a multi-layer firing rate model could give rise to greater responses to infrequent, surprising stimuli, but did not distinguish between true DD and repetition suppression.
+Furthermore, it remains unknown how short-term synaptic and neuronal plasticity, which modulate the input to and the output from a neuron, respectively, work together in a network for DD.
+
+Here, we model DD using a network model of spiking neurons with threshold adaptation and short-term synaptic depression, pursuing two aims: Firstly, we aim to show that neuronal adaptation can induce true DD in much the same way as synaptic depression and investigate the mechanistic basis of this effect in detail.
+Secondly, we aim to investigate the interactions between neuronal and synaptic short-term plasticity in the context of DD.
+We show that the two mechanisms work similarly and can synergize to enhance their effect.
+Our results widen the scope of possible mechanisms of DD and highlight the need to consider interactions between various forms of plasticity.
 
@@
 
+\subsection{Deviance detection emerges from learning regularities}
+
+In this study, we showed that non-synaptic neuronal adaptation can induce true deviance detection in a randomly wired, but spatially organized network of simple leaky integrate-and-fire neurons.
+Neuronal adaptation achieves this, as illustrated in Figure 10a, by suppressing the response of neurons local to the frequently stimulated site.
+This drastically reduces the response to frequent stimulations at an early stage of signal propagation through the network, allowing the more distant parts of the network to recover from adaptation and respond vigorously to infrequent deviant stimuli.
+By contrast, a more varied paradigm including inputs at multiple stimulation sites allows local suppression to partly recover between stimulus presentations, resulting in a greater response through greater recruitment of the wider network.
+As a result, the network becomes more adapted, diminishing the response to each individual stimulus.
+Deviance detection, then, is a side effect of learning and adapting to the regularity of the frequently presented stimulus, which leaves the wider network less adapted and more responsive than usual.
+
+\subsection{Biological basis of neuronal threshold adaptation}
+
+Adaptation of the neuronal firing threshold to past activity or membrane depolarization has been demonstrated at a wide variety of timescales ranging from milliseconds (Henze and Buzsáki 2001; Badel et al.
+2008; Higgs and Spain 2011), supporting precise spike timing (Azouz and Gray 2000; Huang et al.
+2016), to days (O’Leary, van Rossum, and Wyllie 2010), supporting activity homeostasis.
+History dependence on a timescale of hundreds of milliseconds to several seconds, as used in our work, has been described in various systems, including early sensory areas (Baccus and Meister 2002; Wark, Fairhall, and Rieke 2009) and later processing stages up to primary sensory cortices and beyond (Spain, Schwindt, and Crill 1991; Sanchez-Vives, Nowak, and McCormick 2000b; Henze and Buzsáki 2001; La Camera et al.
+2006; Abolafia et al.
+2011).
+Commonly described as spike-frequency adaptation (SFA), this intrinsic feature of neurons supports contrast adaptation (Sanchez-Vives, Nowak, and McCormick 2000b; Wang et al.
+2003), enhances the signal-to-noise ratio (Wark, Lundstrom, and Fairhall 2007; Chacron, Lindner, and Longtin 2007), and improves signal coding properties of the wider network (Prescott and Sejnowski 2008; Bohte 2012; H.
+Lee et al.
+2023).
+SFA can be induced by a number of mutually independent mechanisms, including a buildup of inactivation in voltage-gated sodium channels (Fleidervish, Friedman, and Gutnick 1996; Miles, Dai, and Brownstone 2005), and increased gating of calcium- and sodium-dependent potassium channels (Madison and Nicoll 1984; Schwindt, Spain, and Crill 1989; Sanchez-Vives, Nowak, and McCormick 2000a; Wang et al.
+2003; Faber and Sah 2005) and calcium-gated chloride channels (Ha and Cheong 2017; Ha et al.
+2016).
+Perhaps due to this variety in underlying mechanisms, how to model SFA in integrate-and-fire neurons has been debated, and models have been proposed using both hyperpolarizing currents (Izhikevich 2003; Brette and Gerstner 2005) and adaptive thresholds (Segundo et al.
+1968; Chacron, Pakdaman, and Longtin 2003; Kobayashi, Tsubo, and Shinomoto 2009).
+While there are subtle differences between current- and threshold-based models, and theoretical work has shown current-based approaches to be somewhat superior in capturing firing rate dynamics (Liu and Wang 2001; Benda, Maler, and Longtin 2010), recent large-scale model fitting studies have done very well with adaptive threshold models similar to that used here (Pozzorini et al.
+2015; Teeter et al.
+2018).
+
+\subsection{Threshold adaptation supports deviance detection}
+
+It has been suggested that non-synaptic adaptation such as TA lacks stimulus specificity and is therefore unable to support DD (Duque and Malmierca 2015; Yarden and Nelken 2017).
+In addition, a frequent argument made is that the various timescales involved in stimulus-specific adaptation (SSA) in an oddball context (Ulanovsky et al.
+2004) is best mirrored in the similar range of synaptic plasticity timescales (Tsodyks and Markram 1997; Varela et al.
+1997).
+Yet, in the auditory system, where DD has been studied most intensely, there is intrinsic adaptation at all stages, from the auditory nerve (Yates, Cody, and Johnstone 1983; Westerman and Smith 1984) through subcortical stations (Ingham and McAlpine 2004) to auditory cortex (Abolafia et al.
+2011).
+The timescales involved exhibit a hierarchy, going from tens to hundreds of milliseconds in the sensory and subcortical neurons to several seconds in cortex, a range likewise very similar to that seen in oddball SSA (Ulanovsky et al.
+2004).
+Furthermore, this hierarchy is accompanied by a similar hierarchy in SSA, with thalamus and inferior colliculus being sensitive to shorter stimulus intervals than cortex (Pérez-González and Malmierca 2014).
+Finally, it has been shown that combining intrinsic adaptation with STD imbues networks with exquisite sensitivity to changes in the input signal (Puccini, Sanchez-Vives, and Compte 2006).
+Having shown that TA can support true DD in an oddball context, we anticipate that future models of DD and its analogues will consider including TA or other forms of intrinsic plasticity in addition to synaptic depression.
+
+\subsection{Comparison to other models of DD}
+
+In many ways, our model closely resembles the more elaborate and explicitly cortical model proposed by Yarden and Nelken (2017), which uses short-term synaptic depression as its adaptation mechanism.
+Local suppression is implemented in their model in synaptic depletion in the cortical column sensitive to the target stimulus and is greatest for the standard and least for the deviant, with the many-standards paradigm consistently evoking significant activity across the wider network and inducing an elevated background suppression.
+In addition, both their model and earlier experimental results (Hershenhoren et al.
+2014) show occasional large responses to standard, which we have also observed anecdotally.
+Notably, however, our model operates without explicit compartmentalization into columns.
+Furthermore, Yarden and Nelken (2017) report a fine balance of parameter tuning, outside of which their model ceases to operate as intended and does not detect deviant inputs.
+While we did not explore parameter sensitivity, we take the very minimal tuning and exclusion criteria necessary in our model to indicate a certain level of robustness.
+With looser connectivity requirements and a weak dependence on parameter settings, our model is applicable to a wider range of neural architectures, including non-cortical circuitry.
+On the flipside, the nature of the structural constraint imposed makes it difficult to judge whether DD arises as a result of lateral inhibition or adaptation of lateral excitation as in (Yarden and Nelken 2017; Hertäg and Sprekeler 2020; Park and Geffen 2020) or, conversely, as a result of adaptation in feed-forward connections as in Mill et al.
+(2011).
+We speculate that the robustness of our model derives at least in part from this ambiguity: Like in the model by May and colleagues (2015), TA reduces all activity propagation, regardless of its destination, thus reducing the impact of structural properties.
+
+In contrast to the predominant thinking in the field (see e.g.
+Ross and Hamm (2020) for an extensive review), our work suggests that repetition suppression, which Ross and Hamm call SSA, and true DD can be mediated by a single mechanism, and that DD in particular does not require complex connectivity, as noted above.
+Indeed, while increased responses to deviant stimuli have been reported in cortex of many mammals, including humans (Tada et al.
+2019), other primates (Takaura and Fujii 2016), and rats (Taaseh, Yaron, and Nelken 2011; Shiramatsu, Kanzaki, and Takahashi 2013; Lauren Harms et al.
+2014), recent work has shown true DD even in dissociated neuronal culture (Kubota et al.
+2021), where connectivity is assumed to be comparatively simple, much like in our model.
+Rather than invalidating work dissecting the circuit-level mechanisms of cortical DD (Hamm and Yuste 2016; Attinger, Wang, and Keller 2017; Yarden, Mizrahi, and Nelken 2022), our model provides a possible explanation for the demonstrated presence of DD in subcortical sensory areas (Parras et al.
+2017) and suggests that what Parras and colleagues identify as prediction errors may in fact persist in the absence of top-down predictions.
+
+\subsection{Adaptation supports subcortical DD}
+
+There is indeed some evidence for this.
+If feedback signals within cortex and from cortex to midbrain sensory areas carry predictions that cancel out incoming activation, as is commonly assumed in the predictive coding framework (Rao and Ballard 1999; Friston 2005; Carbajal and Malmierca 2018), then we should expect these signals to interfere with adaptation-dependent processing, since neurons silenced by top-down predictions are unable to build up a stimulus history with adaptation or depression.
+Thus, knocking out predictive feedback should facilitate such adaptation-driven DD, while any DD based on predictive inputs would disappear.
+Indeed, such differential effects have been reported by Harms and colleagues (2018).
+Under the influence of NMDA blockers, which inhibit long-term plasticity in cortex and therefore prevent the formation of accurate predictive models, late components of DD in an oddball paradigm are reduced, while components within the first 10-20 ms after stimulus onset are enhanced.
+This suggests that early DD, likely driven by subcortical processing (Grimm and Escera 2012; Parras et al.
+2017; Font-Alaminos et al.
+2020), may be driven by adaptive mechanisms such as that demonstrated in our work, whereas later, cortical DD derives largely from more sophisticated computations involving explicit prediction and error signals (Schlossmacher et al.
+2022; Grimm, Escera, and Nelken 2016).
+It would be very interesting to explore this interaction in a hierarchical model that combines the adaptation-based approach in early, subcortical areas with a predictive approach (Wacongne, Changeux, and Dehaene 2012; Wacongne 2016; Hertäg and Sprekeler 2020) in higher, cortical areas.
+
+\subsection{TA and STD synergize to yield supralinear scaling}
+
+Our second main finding is that synaptic short-term depression, which acts in similar fashion to neuronal adaptation both in terms of how it is induced and in how it affects activity flow in the network, enhances the DD effect of neuronal adaptation, and vice versa.
+We demonstrated that this synergy is a result of increased local suppression of the frequently presented stimulus by STD, which decreases global suppression by both mechanisms in oddball sequences and enhances the deviant response, as illustrated in Figure 10b.
+Conversely, without strong local suppression in control sequences, global suppression is increased, leading to smaller individual responses.
+
+This synergy can be understood as a simple nonlinear effect of adaptation.
+While the effect on local suppression of adding STD is direct and mediated only by STD, the resulting reduction of global suppression in the oddball sequence depends entirely on the pre-existing suppression levels mediated by TA and is indeed counteracted to some extent by STD.
+We speculate that increasing the strength of TA, rather than adding STD, would have a similar effect.
+Thus, prior to saturation, each additional unit of adaptation from any sufficiently similar source would lead to increasing gains in DD.
+While it is a priori unsurprising that short-term plasticity should give rise to nonlinearities, the direction of the nonlinearity typically reflects the direction of plasticity: Presynaptic facilitation leads to supralinear activation of the post-synaptic side, and vice versa, depression leads to sublinear activation (Tsodyks and Markram 1997; Varela et al.
+1997), which is typically reflected also in population-level activity (Tsodyks, Pawelzik, and Markram 1998).
+The reverse has been shown in the context of inhibition-stabilized networks (Wu and Gjorgjieva 2022) as a result of the so-called paradoxical effect that increasing excitatory drive decreases activity (Sanzeni et al.
+2020; Tsodyks et al.
+1997).
+However, we have no reason to believe that our model is inhibition-stabilized and are not aware of other work showing supralinear scaling from depression.
+It would be interesting to confirm our intuition in the model presented here, as well as in a context of more realistic neural and network dynamics.
+
+In addition, while interactions between different forms of short-term plasticity have been investigated in detail at the level of synaptic processes (Tsodyks and Markram 1997; Varela et al.
+1997; Buonomano and Merzenich 1998; Hennig et al.
+2008; Anwar et al.
+2017; Mondal, Pena, and Rotstein 2022), we are aware of only two studies on the interaction between synaptic and neuronal short-term plasticity, investigating oscillations (Mejias and Torres 2011) and spike train irregularities (Ileri et al.
+2015) caused by such interactions, respectively.
+A principled investigation of the interactions between short-term plasticity mechanisms is missing and would seem to be an important addition to our understanding of neural circuit dynamics.
+
+\subsection{Differences between TA and STD}
+
+In the present work, we chose parameters such that TA had a much stronger effect than STD, particularly with the chosen time scales.
+We believe that the chosen values fall within a physiologically plausible range, and it was not our intention to skew the results in favor of TA.
+Indeed, we consider it likely that a more even tuning would yield a model capable of evoking comparable levels of DD with either mechanism, working individually and jointly as demonstrated here.
+
+In one aspect, however, neuronal and synaptic plasticity clearly differ: While TA throttles the output of the affected neurons, STD throttles the input to downstream targets.
+A re-examination of Figure 10b-c, for example, shows that STD affects more neurons as a direct result.
+The greater reach of the immediate effect of STD perhaps explains why TA has often been overlooked for deviance detection, besides the mistaken claim of stimulus specificity.
+We also note that this greater reach may result in greater overlap between the neurons experiencing suppression from several stimulus classes.
+This, in turn, would mean that STD may facilitate the kinds of interactions between different stimuli that give rise to DD, suggesting its applicability to a wider range of neural architectures and of stimuli than TA.
+
+\subsection{Impact of spontaneous activity}
+
+In comparison to biological sensory processing systems, our model is, of course, very simple.
+One key limitation is in the activity dynamics: Unlike living systems, our model does not produce any spontaneous activity, and evoked responses die out quickly.
+In contrast, even dissociated cell cultures show some amount of ongoing activity in absence of stimulation.
+Increased activity would likely substantially affect our results by preferentially following less adapted pathways and therefore reducing the contrast between adapted and recovered parts of the network.
+Greater spontaneous activity has indeed been linked to reduced DD in rodent inferior colliculus (Duque et al.
+2012; Duque and Malmierca 2015) and auditory cortex (Klein, von der Behrens, and Gaese 2014), though sensitivity appears to be considerably less than we would expect in our model.
+In addition, Abolafia and colleagues (2011) showed that increased background activity also attenuates the level of adaptation to stimuli in individual neurons.
+
+Relatedly, the lack of spontaneous activity also implies that no explicit predictive signals could possibly be generated in our model.
+Such signals would be necessary for the network to respond to the omission of expected stimuli, as has been demonstrated in cortex of humans (Lehmann, Arias, and Schönwiesner 2016; Demarchi, Sanchez, and Weisz 2019), marmosets (Jiang et al.
+2022), and rats (Auksztulewicz et al.
+2022).
+However, we are unaware of any work showing conclusively the presence or absence of omission responses in subcortical areas.
+Evidence from EEG presented by Lehmann and colleagues (2016) suggests that no such responses may occur, but a finer-grained analysis in animal models would help to further narrow down to what extent the early, subcortical DD processes are driven by prediction and adaptation (Schlossmacher et al.
+2022).
+Along with other considerations outlined above, the absence of evidence of omission responses seems to suggest that our adaptation-based model is a better description of subcortical processing.
+
+\subsection{Early and late response in culture}
+
+In dissociated culture (Kubota et al.
+2021), the stimulus response is biphasic, with a peak of directly evoked activity in the first 10 ms, and a longer response lasting up to 200 ms.
+In the results reported by Kubota and colleagues (Figure 2), DD appears to become evident only in the later response.
+Since we do not simulate conduction delays, the response in our model is compressed over a much shorter time period.
+However, we also see a larger response difference in the later stage of the response, as activity passes from the stimulated neurons and their immediate neighborhood into the wider network.
+It remains to be confirmed whether the distinction between early and late responses in culture is tied to separate neuron populations, as in our model, indicating some amount of linearity of activity propagation, or whether the early and late responses both develop in a distributed manner throughout the network.
+
+\subsection{Conclusion}
+
+As alluded to above, we anticipate that future models of DD and its attendant signatures (SSA, mismatch negativity, etc.) will rely on hierarchical structure, including not just different cortical areas (May, Westö, and Tiitinen 2015; Park and Geffen 2020), but also subcortical areas with their own processing dynamics.
+Given the evidence presented here, we strongly believe that neuronal adaptation will play a key role in such future models, working alongside synaptic short-term plasticity (Yarden and Nelken 2017) and long-term plasticity (Hertäg and Sprekeler 2020) to better explain the full richness of temporal sensory processing.
+
 \backmatter

### 2023-03-16 paper-1/manuscript.tex
Converted citations

--- 2023-03-08 paper-1/manuscript.tex
@@
 \section{Introduction}\label{sec-intro}
+
+Deviance detection (DD) is the ability of neural systems to identify statistically surprising sensory inputs or violations of established regularities.
+It is a critical component of any signal processing cascade and is closely related to the concept of prediction errors in the predictive processing framework \citep{Bendixen2012-nx, Khouri2015-gr, Carbajal2018-sd, Fong2020-em}, which has been suggested to underlie most of cognition \citep{Friston2005-jz, Clark2015-gl}.
+DD has been studied widely across sensory systems including the auditory \citep{Nelken2014-wr, Escera2014-tv, Carbajal2018-sd}, visual \citep{Winkler2012-pr, Pazo-Alvarez2003-kv}, somatosensory \citep{Naatanen2009-jx} domains and beyond.
+Impairments in DD have been reported in various disorders \citep{Fong2020-em}, including autism \citep{Schwartz2018-bc, Goris2018-bv, Hudac2018-jl, Vlaskamp2017-vs, Chen2020-lu}, schizophrenia \citep{Koshiyama2020-ca, Kim2020-fb, Salisbury2020-sd, Tada2019-lj}, ADHD \citep{Kim2021-uj, Hsieh2021-ti, Lee2020-np} and others, highlighting the need to better understand its mechanistic underpinnings.
+
+Studies of DD must carefully distinguish between repetition suppression and true deviance detection \citep{Ross2020-qf}.
+The former refers to the reduction of the response to a frequently presented stimulus and can be explained by any adaptive mechanism \citep{May2010-qn, Garagnani2011-eu}.
+The latter refers to the enhancement of the response to an infrequent stimulus when it is statistically surprising, e.g., due to interrupting a series of regular stimuli, compared to an unsurprising control condition.
+Here, we study the latter, true DD, which has been suggested to rely on a comparison between prediction signals and received inputs \citep{Parras2017-fp, Carbajal2018-sd, Ross2020-qf}.
+
+Existing computational models of DD place their primary focus on synaptic plasticity as the causal mechanism.
+While models using long-term plasticity have been proposed \citep{Wacongne2012-ah, Hertag2020-kc}, the rapid establishment of DD in vivo \citep{Taaseh2011-gg} suggests that fast-acting short-term plasticity mechanisms play a key role.
+Several models using short-term synaptic plasticity have been proposed \citep{Mill2011-ah, May2015-lt, Yarden2017-eh}, finding good agreement with physiological data.
+However, short-term plasticity also occurs in neuronal excitability, with many neurons exhibiting an experience-dependent transient reduction in intrinsic neuronal sensitivity independent of synaptic function \citep{Sanchez-Vives2000-df, Henze2001-xd, Sanchez-Aguilera2014-fd}.
+Despite operating at similar time scales to short-term synaptic plasticity, the effects of intrinsic neuronal plasticity on DD is unexplored.
+Previous modeling work has only used neuronal plasticity as a side effect of the chosen neuron model \citep{Mill2011-ah} or rejected it a priori as incapable of producing true DD \citep{Yarden2017-eh}.
+One exception is the work by \cite{Garagnani2011-eu}, which showed that neuronal plasticity in a multi-layer firing rate model could give rise to greater responses to infrequent, surprising stimuli, but did not distinguish between true DD and repetition suppression.
+Furthermore, it remains unknown how short-term synaptic and neuronal plasticity, which modulate the input to and the output from a neuron, respectively, work together in a network for DD.
+
+Here, we model DD using a network model of spiking neurons with threshold adaptation and short-term synaptic depression, pursuing two aims: Firstly, we aim to show that neuronal adaptation can induce true DD in much the same way as synaptic depression and investigate the mechanistic basis of this effect in detail.
+Secondly, we aim to investigate the interactions between neuronal and synaptic short-term plasticity in the context of DD.
+We show that the two mechanisms work similarly and can synergize to enhance their effect.
+Our results widen the scope of possible mechanisms of DD and highlight the need to consider interactions between various forms of plasticity.
 
@@
 
+\subsection{Deviance detection emerges from learning regularities}
+
+In this study, we showed that non-synaptic neuronal adaptation can induce true deviance detection in a randomly wired, but spatially organized network of simple leaky integrate-and-fire neurons.
+Neuronal adaptation achieves this, as illustrated in Figure 10a, by suppressing the response of neurons local to the frequently stimulated site.
+This drastically reduces the response to frequent stimulations at an early stage of signal propagation through the network, allowing the more distant parts of the network to recover from adaptation and respond vigorously to infrequent deviant stimuli.
+By contrast, a more varied paradigm including inputs at multiple stimulation sites allows local suppression to partly recover between stimulus presentations, resulting in a greater response through greater recruitment of the wider network.
+As a result, the network becomes more adapted, diminishing the response to each individual stimulus.
+Deviance detection, then, is a side effect of learning and adapting to the regularity of the frequently presented stimulus, which leaves the wider network less adapted and more responsive than usual.
+
+\subsection{Biological basis of neuronal threshold adaptation}
+
+Adaptation of the neuronal firing threshold to past activity or membrane depolarization has been demonstrated at a wide variety of timescales ranging from milliseconds \citep{Henze2001-xd, Badel2008-ld, Higgs2011-cy}, supporting precise spike timing \citep{Azouz2000-vh, Huang2016-ch}, to days \citep{OLeary2010-ar}, supporting activity homeostasis.
+History dependence on a timescale of hundreds of milliseconds to several seconds, as used in our work, has been described in various systems, including early sensory areas \citep{Baccus2002-jb, Wark2009-sf} and later processing stages up to primary sensory cortices and beyond \citep{Spain1991-fg, Sanchez-Vives2000-df, Henze2001-xd, La_Camera2006-dz, Abolafia2011-ig}.
+Commonly described as spike-frequency adaptation (SFA), this intrinsic feature of neurons supports contrast adaptation \citep{Sanchez-Vives2000-df, Wang2003-vv}, enhances the signal-to-noise ratio \citep{Wark2007-rx, Chacron2007-ak}, and improves signal coding properties of the wider network \citep{Prescott2008-oa, Bohte2012-gk, Lee2023-bu}.
+SFA can be induced by a number of mutually independent mechanisms, including a buildup of inactivation in voltage-gated sodium channels \citep{Fleidervish1996-ly, Miles2005-nz}, and increased gating of calcium- and sodium-dependent potassium channels \citep{Madison1984-zc, Schwindt1989-lv, Sanchez-Vives2000-rv, Wang2003-vv, Faber2005-vc} and calcium-gated chloride channels \citep{Ha2016-lc, Ha2017-bh}.
+Perhaps due to this variety in underlying mechanisms, how to model SFA in integrate-and-fire neurons has been debated, and models have been proposed using both hyperpolarizing currents \citep{Izhikevich2003-ty, Brette2005-ci} and adaptive thresholds \citep{Segundo1968-eu, Chacron2003-oz, Kobayashi2009-av}.
+While there are subtle differences between current- and threshold-based models, and theoretical work has shown current-based approaches to be somewhat superior in capturing firing rate dynamics \citep{Liu2001-yw, Benda2010-tz}, recent large-scale model fitting studies have done very well with adaptive threshold models similar to that used here \citep{Pozzorini2015-ei, Teeter2018-iz}.
+
+\subsection{Threshold adaptation supports deviance detection}
+
+It has been suggested that non-synaptic adaptation such as TA lacks stimulus specificity and is therefore unable to support DD \citep{Duque2015-eu, Yarden2017-eh}.
+In addition, a frequent argument made is that the various timescales involved in stimulus-specific adaptation (SSA) in an oddball context \citep{Ulanovsky2004-nf} is best mirrored in the similar range of synaptic plasticity timescales \citep{Tsodyks1997-qt, Varela1997-nr}.
+Yet, in the auditory system, where DD has been studied most intensely, there is intrinsic adaptation at all stages, from the auditory nerve \citep{Yates1983-uu, Westerman1984-ge} through subcortical stations \citep{Ingham2004-wc} to auditory cortex \citep{Abolafia2011-ig}.
+The timescales involved exhibit a hierarchy, going from tens to hundreds of milliseconds in the sensory and subcortical neurons to several seconds in cortex, a range likewise very similar to that seen in oddball SSA \citep{Ulanovsky2004-nf}.
+Furthermore, this hierarchy is accompanied by a similar hierarchy in SSA, with thalamus and inferior colliculus being sensitive to shorter stimulus intervals than cortex \citep{Perez-Gonzalez2014-pm}.
+Finally, it has been shown that combining intrinsic adaptation with STD imbues networks with exquisite sensitivity to changes in the input signal \citep{Puccini2006-tm}.
+Having shown that TA can support true DD in an oddball context, we anticipate that future models of DD and its analogues will consider including TA or other forms of intrinsic plasticity in addition to synaptic depression.
+
+\subsection{Comparison to other models of DD}
+
+In many ways, our model closely resembles the more elaborate and explicitly cortical model proposed by \cite{Yarden2017-eh}, which uses short-term synaptic depression as its adaptation mechanism.
+Local suppression is implemented in their model in synaptic depletion in the cortical column sensitive to the target stimulus and is greatest for the standard and least for the deviant, with the many-standards paradigm consistently evoking significant activity across the wider network and inducing an elevated background suppression.
+In addition, both their model and earlier experimental results \citep{Hershenhoren2014-bb} show occasional large responses to standard, which we have also observed anecdotally.
+Notably, however, our model operates without explicit compartmentalization into columns.
+Furthermore, \cite{Yarden2017-eh} report a fine balance of parameter tuning, outside of which their model ceases to operate as intended and does not detect deviant inputs.
+While we did not explore parameter sensitivity, we take the very minimal tuning and exclusion criteria necessary in our model to indicate a certain level of robustness.
+With looser connectivity requirements and a weak dependence on parameter settings, our model is applicable to a wider range of neural architectures, including non-cortical circuitry.
+On the flipside, the nature of the structural constraint imposed makes it difficult to judge whether DD arises as a result of lateral inhibition or adaptation of lateral excitation as in \citep{Yarden2017-eh, Hertag2020-kc, Park2020-oa} or, conversely, as a result of adaptation in feed-forward connections as in \cite{Mill2011-ah}.
+We speculate that the robustness of our model derives at least in part from this ambiguity: Like in the model by \cite{May2015-lt}, TA reduces all activity propagation, regardless of its destination, thus reducing the impact of structural properties.
+
+In contrast to the predominant thinking in the field (see e.g.
+\cite{Ross2020-qf} for an extensive review), our work suggests that repetition suppression, which Ross and Hamm call SSA, and true DD can be mediated by a single mechanism, and that DD in particular does not require complex connectivity, as noted above.
+Indeed, while increased responses to deviant stimuli have been reported in cortex of many mammals, including humans \citep{Tada2019-lj}, other primates \citep{Takaura2016-eh}, and rats \citep{Taaseh2011-gg, Shiramatsu2013-qz, Harms2014-ah}, recent work has shown true DD even in dissociated neuronal culture \citep{Kubota2021-dx}, where connectivity is assumed to be comparatively simple, much like in our model.
+Rather than invalidating work dissecting the circuit-level mechanisms of cortical DD \citep{Hamm2016-lm, Attinger2017-dw, Yarden2022-yy}, our model provides a possible explanation for the demonstrated presence of DD in subcortical sensory areas \citep{Parras2017-fp} and suggests that what Parras and colleagues identify as prediction errors may in fact persist in the absence of top-down predictions.
+
+\subsection{Adaptation supports subcortical DD}
+
+There is indeed some evidence for this.
+If feedback signals within cortex and from cortex to midbrain sensory areas carry predictions that cancel out incoming activation, as is commonly assumed in the predictive coding framework \citep{Rao1999-xc, Friston2005-jz, Carbajal2018-sd}, then we should expect these signals to interfere with adaptation-dependent processing, since neurons silenced by top-down predictions are unable to build up a stimulus history with adaptation or depression.
+Thus, knocking out predictive feedback should facilitate such adaptation-driven DD, while any DD based on predictive inputs would disappear.
+Indeed, such differential effects have been reported by \cite{Harms2018-bg}.
+Under the influence of NMDA blockers, which inhibit long-term plasticity in cortex and therefore prevent the formation of accurate predictive models, late components of DD in an oddball paradigm are reduced, while components within the first 10-20 ms after stimulus onset are enhanced.
+This suggests that early DD, likely driven by subcortical processing \citep{Grimm2012-tb, Parras2017-fp, Font-Alaminos2020-fs}, may be driven by adaptive mechanisms such as that demonstrated in our work, whereas later, cortical DD derives largely from more sophisticated computations involving explicit prediction and error signals \citep{Schlossmacher2022-zv, Grimm2016-fr}.
+It would be very interesting to explore this interaction in a hierarchical model that combines the adaptation-based approach in early, subcortical areas with a predictive approach \citep{Wacongne2012-ah, Wacongne2016-ht, Hertag2020-kc} in higher, cortical areas.
+
+\subsection{TA and STD synergize to yield supralinear scaling}
+
+Our second main finding is that synaptic short-term depression, which acts in similar fashion to neuronal adaptation both in terms of how it is induced and in how it affects activity flow in the network, enhances the DD effect of neuronal adaptation, and vice versa.
+We demonstrated that this synergy is a result of increased local suppression of the frequently presented stimulus by STD, which decreases global suppression by both mechanisms in oddball sequences and enhances the deviant response, as illustrated in Figure 10b.
+Conversely, without strong local suppression in control sequences, global suppression is increased, leading to smaller individual responses.
+
+This synergy can be understood as a simple nonlinear effect of adaptation.
+While the effect on local suppression of adding STD is direct and mediated only by STD, the resulting reduction of global suppression in the oddball sequence depends entirely on the pre-existing suppression levels mediated by TA and is indeed counteracted to some extent by STD.
+We speculate that increasing the strength of TA, rather than adding STD, would have a similar effect.
+Thus, prior to saturation, each additional unit of adaptation from any sufficiently similar source would lead to increasing gains in DD.
+While it is a priori unsurprising that short-term plasticity should give rise to nonlinearities, the direction of the nonlinearity typically reflects the direction of plasticity: Presynaptic facilitation leads to supralinear activation of the post-synaptic side, and vice versa, depression leads to sublinear activation \citep{Tsodyks1997-qt, Varela1997-nr}, which is typically reflected also in population-level activity \citep{Tsodyks1998-bq}.
+The reverse has been shown in the context of inhibition-stabilized networks \citep{Wu2022-qx} as a result of the so-called paradoxical effect that increasing excitatory drive decreases activity \citep{Tsodyks1997-hr, Sanzeni2020-su}.
+However, we have no reason to believe that our model is inhibition-stabilized and are not aware of other work showing supralinear scaling from depression.
+It would be interesting to confirm our intuition in the model presented here, as well as in a context of more realistic neural and network dynamics.
+
+In addition, while interactions between different forms of short-term plasticity have been investigated in detail at the level of synaptic processes \citep{Tsodyks1997-qt, Varela1997-nr, Buonomano1998-hf, Hennig2008-sc, Anwar2017-qg, Mondal2022-av}, we are aware of only two studies on the interaction between synaptic and neuronal short-term plasticity, investigating oscillations \citep{Mejias2011-rv} and spike train irregularities \citep{Ileri2015-eu} caused by such interactions, respectively.
+A principled investigation of the interactions between short-term plasticity mechanisms is missing and would seem to be an important addition to our understanding of neural circuit dynamics.
+
+\subsection{Differences between TA and STD}
+
+In the present work, we chose parameters such that TA had a much stronger effect than STD, particularly with the chosen time scales.
+We believe that the chosen values fall within a physiologically plausible range, and it was not our intention to skew the results in favor of TA.
+Indeed, we consider it likely that a more even tuning would yield a model capable of evoking comparable levels of DD with either mechanism, working individually and jointly as demonstrated here.
+
+In one aspect, however, neuronal and synaptic plasticity clearly differ: While TA throttles the output of the affected neurons, STD throttles the input to downstream targets.
+A re-examination of Figure 10b-c, for example, shows that STD affects more neurons as a direct result.
+The greater reach of the immediate effect of STD perhaps explains why TA has often been overlooked for deviance detection, besides the mistaken claim of stimulus specificity.
+We also note that this greater reach may result in greater overlap between the neurons experiencing suppression from several stimulus classes.
+This, in turn, would mean that STD may facilitate the kinds of interactions between different stimuli that give rise to DD, suggesting its applicability to a wider range of neural architectures and of stimuli than TA.
+
+\subsection{Impact of spontaneous activity}
+
+In comparison to biological sensory processing systems, our model is, of course, very simple.
+One key limitation is in the activity dynamics: Unlike living systems, our model does not produce any spontaneous activity, and evoked responses die out quickly.
+In contrast, even dissociated cell cultures show some amount of ongoing activity in absence of stimulation.
+Increased activity would likely substantially affect our results by preferentially following less adapted pathways and therefore reducing the contrast between adapted and recovered parts of the network.
+Greater spontaneous activity has indeed been linked to reduced DD in rodent inferior colliculus \citep{Duque2012-vx, Duque2015-eu} and auditory cortex \citep{Klein2014-op}, though sensitivity appears to be considerably less than we would expect in our model.
+In addition, \cite{Abolafia2011-ig} showed that increased background activity also attenuates the level of adaptation to stimuli in individual neurons.
+
+Relatedly, the lack of spontaneous activity also implies that no explicit predictive signals could possibly be generated in our model.
+Such signals would be necessary for the network to respond to the omission of expected stimuli, as has been demonstrated in cortex of humans \citep{Lehmann2016-ol, Demarchi2019-ch}, marmosets \citep{Jiang2022-ak}, and rats \citep{Auksztulewicz2022-gq}.
+However, we are unaware of any work showing conclusively the presence or absence of omission responses in subcortical areas.
+Evidence from EEG presented by \cite{Lehmann2016-ol} suggests that no such responses may occur, but a finer-grained analysis in animal models would help to further narrow down to what extent the early, subcortical DD processes are driven by prediction and adaptation \citep{Schlossmacher2022-zv}.
+Along with other considerations outlined above, the absence of evidence of omission responses seems to suggest that our adaptation-based model is a better description of subcortical processing.
+
+\subsection{Early and late response in culture}
+
+In dissociated culture \citep{Kubota2021-dx}, the stimulus response is biphasic, with a peak of directly evoked activity in the first 10 ms, and a longer response lasting up to 200 ms.
+In the results reported by Kubota and colleagues (Figure 2), DD appears to become evident only in the later response.
+Since we do not simulate conduction delays, the response in our model is compressed over a much shorter time period.
+However, we also see a larger response difference in the later stage of the response, as activity passes from the stimulated neurons and their immediate neighborhood into the wider network.
+It remains to be confirmed whether the distinction between early and late responses in culture is tied to separate neuron populations, as in our model, indicating some amount of linearity of activity propagation, or whether the early and late responses both develop in a distributed manner throughout the network.
+
+\subsection{Conclusion}
+
+As alluded to above, we anticipate that future models of DD and its attendant signatures (SSA, mismatch negativity, etc.) will rely on hierarchical structure, including not just different cortical areas \citep{May2015-lt, Park2020-oa}, but also subcortical areas with their own processing dynamics.
+Given the evidence presented here, we strongly believe that neuronal adaptation will play a key role in such future models, working alongside synaptic short-term plasticity \citep{Yarden2017-eh} and long-term plasticity \citep{Hertag2020-kc} to better explain the full richness of temporal sensory processing.
+
 \backmatter

### 2023-03-17 plot_utils.py
Added figwidth utils, changed to capital panels

@@ -55,6 +55,12 @@ def label_panel(ax, ord, lower=False, left=1):
     return lb
 
 
+def fullwidth(height):
+    return styling.fig_width, height * styling.fig_width / 6.29
+def halfwidth(height):
+    return styling.fig_width/2, height * styling.fig_width/2 / 3
+
+
 # From http://www.ccctool.com/html_v_0_9_0_3/CCC_Tool/cccTool.html
 ccc_divergent = LinearSegmentedColormap.from_list(
     'ccc_divergent', list(zip([0, .16, .35, .5, .62, .8, 1],

### 2023-03-17 styling.py

@@ -63,5 +63,6 @@ sns.set_theme(style='ticks', font_scale=0.75, rc={
 label_kwargs = {'fontsize': 12, 'fontweight': 'bold'}
 annotation_kwargs = {'fontsize': 7}
 
-# panel_labels = 'ABCDEFGHIJK'
-panel_labels = 'abcdefghijk'
+panel_labels = 'ABCDEFGHIJK'
+# panel_labels = 'abcdefghijk'
+fig_width = 8

### 2023-03-17 paper-1/manuscript.tex
Adjust to ``normal'' latex quoting

@@
 To investigate deviance detection, we used a classical oddball paradigm as illustrated in \refpanel{1}{d}, presenting a total of 500 stimuli at regular intervals of 500~ms.
-Target stimuli (labeled "A" or "target" hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The "standard" sequence (\std{}), consisting of 400 presentations of A and 100 presentations of B; the "deviant" sequence (\dev{}) with 100 A and 400 B, and the many-standards control sequence (\msc{}) with 100 presentations of each of the stimuli A through E.
+Target stimuli (labeled ``A'' or ``target'' hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The ``standard'' sequence (\std{}), consisting of 400 presentations of A and 100 presentations of B; the ``deviant'' sequence (\dev{}) with 100 A and 400 B, and the many-standards control sequence (\msc{}) with 100 presentations of each of the stimuli A through E.
 In this way, A is presented as the predominant and therefore expected stimulus (\std{}), as an infrequent violation of an expectation (of B, \dev{}), and, to control for pure adaptation effects, as an equally infrequent stimulus with no strong expectation of any particular input (\msc{}).
@@
 Inset histograms show the distributions of the plotted values, with the confidence intervals of their medians, calculated by bootstrap with 10000 iterations, in darker color.
-\textbf{c} Median non-target response contrast $\Delta \R{N}{}$ among the first 100 neurons to respond to stimulation in B, labeled "early", and among the later neurons, labeled "late", across networks and stimuli.
+\textbf{c} Median non-target response contrast $\Delta \R{N}{}$ among the first 100 neurons to respond to stimulation in B, labeled ``early'', and among the later neurons, labeled ``late'', across networks and stimuli.
 \textbf{d} Median contrast in $\T{B}{}$ among the early and late neurons, respectively, across networks and stimuli.
@@
 We had good reasons for this assumption: Updates to the spike-triggered depression variable $x_j$ (cf.
-Equation \ref{eqn-xsyn}) would decay to a negligible $e^\frac{-1 s}{\tau_x} = 0.0013$ times the original value within two trials, making it unlikely that inputs separated by intervening trials could be "remembered" in the network.
+Equation \ref{eqn-xsyn}) would decay to a negligible $e^\frac{-1 s}{\tau_x} = 0.0013$ times the original value within two trials, making it unlikely that inputs separated by intervening trials could be ``remembered'' in the network.
 Conversely, the frequently presented stimulus B may have been able to lead to a build-up of depression strong enough to suppress its own response.
@@
 
-Finally, we sought to confirm whether "local" and "global" suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network.
+Finally, we sought to confirm whether ``local'' and ``global'' suppression, which we defined so far by the latency of response, did in fact reflect a spatial property of the network.
 To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.

### 2023-03-17 paper-1/manuscript.tex
Align manuscript somewhat with elife

@@
 \usepackage{xcolor,hyperref}
-\usepackage[autostyle=false, style=english]{csquotes}
 %%%%
-
-% Avoid having to ``quote'' verbosely (csquotes package)
-\MakeOuterQuote{"}
 
@@
 
-% Figure references
-\newcommand{\reffig}[1]{Figure \ref{fig#1}}
-\newcommand{\refpanel}[2]{Figure \ref{fig#1}\lowercase{#2}}
+% Crossreferences
+\newcommand{\FIG}[1]{Figure \ref{fig:#1}}
+\newcommand{\EQ}[1]{Equation \ref{eq:#1}}
 
@@
 
-\section{Introduction}\label{sec-intro}
+\section{Introduction}\label{sec:intro}
 
@@
 
-\section{Methods}\label{sec-methods}
+\section{Methods}\label{sec:methods}
 
-\subsection{Model}\label{sec-model}
+\subsection{Model}\label{sec:model}
 
@@
     \tau_e \frac{dg_e}{dt} &= -g_e + \sum_{j \in \boldsymbol E} U x_j w \delta(t - \hat{t_j}) \nonumber \\
-    \tau_i \frac{dg_i}{dt} &= -g_i + \sum_{j \in \boldsymbol I} w \delta(t - \hat{t_j}) \label{eqn-gsyn}
+    \tau_i \frac{dg_i}{dt} &= -g_i + \sum_{j \in \boldsymbol I} w \delta(t - \hat{t_j}) \label{eq:gsyn}
 \end{align}
@@
 \begin{equation}
-    \tau_x \frac{dx_j}{dt} = (1-x_j) - U x_j \delta(t - \hat{t}) \label{eqn-xsyn}
+    \tau_x \frac{dx_j}{dt} = (1-x_j) - U x_j \delta(t - \hat{t}) \label{eq:xsyn}
 \end{equation}
@@
     V_\theta &= \theta_0 + \theta(t) \nonumber \\
-    \tau_{\theta} \frac{d\theta}{dt} &= -\theta + \hat{\theta} \delta(t - \hat{t}) \label{eqn-TA}
+    \tau_{\theta} \frac{d\theta}{dt} &= -\theta + \hat{\theta} \delta(t - \hat{t}) \label{eq:TA}
 \end{align}
@@
     \caption{}
-    \label{fig1}
+    \label{fig:1}
 \end{figure*}
@@
 Top: Inputs to the (post-)synaptic conductances $g_e$ and $g_i$ (vertical bars), and the STD depression variable $x_{Exc}$ of the excitatory presynaptic neuron (dashed line).
-Note that excitatory inputs are equal to $U x_{Exc} w$ as per Equation \ref{eqn-gsyn}.
+Note that excitatory inputs are equal to $U x_{Exc} w$ as per \EQ{gsyn}.
 Middle: Excitatory and inhibitory synaptic conductances ($g_e$ and $g_i$, respectively).
@@
 
-\refpanel{1}{a} shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
+\FIG{1}a shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
 The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \citep{Stimberg2020-go}, and simulated with an integration time step of 1~ms.
@@
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-\refpanel{1}{b} shows a representative example of the resulting network structure.
-In \refpanel{1}{c}, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
+\FIG{1}b shows a representative example of the resulting network structure.
+In \FIG{1}c, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
 Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response: Networks where fewer than 500 neurons responded to stimulation at any of the five sites after full recovery were discarded.
@@
 
-\subsection{Paradigm}\label{sec-paradigm}
+\subsection{Paradigm}\label{sec:paradigm}
 
-To investigate deviance detection, we used a classical oddball paradigm as illustrated in \refpanel{1}{d}, presenting a total of 500 stimuli at regular intervals of 500~ms.
+To investigate deviance detection, we used a classical oddball paradigm as illustrated in \FIG{1}d, presenting a total of 500 stimuli at regular intervals of 500~ms.
 Target stimuli (labeled ``A'' or ``target'' hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The ``standard'' sequence (\std{}), consisting of 400 presentations of A and 100 presentations of B; the ``deviant'' sequence (\dev{}) with 100 A and 400 B, and the many-standards control sequence (\msc{}) with 100 presentations of each of the stimuli A through E.
@@
 In order to disentangle the effects of STD and TA, we ran all simulations in four conditions: Without any short-term plasticity, with either TA only or STD only, and with both STD and TA (labeled STD+TA in the following, and corresponding to the full model as described above).
-To turn off STD, we replaced Equation \ref{eqn-xsyn} with a constant $x = 1$, but retained the scaling of inputs to $g_e$ with $U$ in Equation \ref{eqn-gsyn}.
+To turn off STD, we replaced \EQ{xsyn} with a constant $x = 1$, but retained the scaling of inputs to $g_e$ with $U$ in \EQ{gsyn}.
 This corresponds to immediate recovery ($\tau_x = 0$) and was done to maintain the magnitude of the excitatory postsynaptic potential (EPSP) in the recovered state regardless of whether STD was turned on or off.
-To turn off TA, we replaced Equation \ref{eqn-TA} with a constant $V_{\theta} = \theta_0$, making the firing threshold completely unadaptive.
+To turn off TA, we replaced \EQ{TA} with a constant $V_{\theta} = \theta_0$, making the firing threshold completely unadaptive.
 
-\subsection{Sample network}\label{sec-sample}
+\subsection{Sample network}\label{sec:sample}
 
@@
 The sample network was chosen to be roughly representative of the population as follows: We assessed each network's average response to stimulation with A and B in the \msc{} sequence and calculated z-scores with respect to the population.
-We considered only networks whose absolute z-score was less than 1 for both A and B responses, then hand-picked a sample network and stimulus assignment that exhibited both an increased deviance detection index (see Equation \ref{eqn-ddi} below) in STD+TA over TA only, and greater \msc{} than \std{} responses in both STD+TA and TA only conditions.
+We considered only networks whose absolute z-score was less than 1 for both A and B responses, then hand-picked a sample network and stimulus assignment that exhibited both an increased deviance detection index (see \EQ{ddi} below) in STD+TA over TA only, and greater \msc{} than \std{} responses in both STD+TA and TA only conditions.
 
-\subsection{Notation}\label{sec-notation}
+\subsection{Notation}\label{sec:notation}
 
@@
 We define $\T{}{}$ as the mean across trials of $\theta(t=0)$, i.e., the average TA voltage at the beginning of trials, immediately before stimulation.
-Lastly, the related STD quantity $\D{}{}$ will be introduced later in Equation \ref{eqn-D}.
+Lastly, the related STD quantity $\D{}{}$ will be introduced later in \EQ{D}.
 
@@
 
-\section{Results}\label{sec-results}
+\section{Results}\label{sec:results}
 
-\subsection{Deviance detection}\label{sec-dd}
+\subsection{Deviance detection}\label{sec:dd}
 
@@
 \begin{equation}
-    DDI = \frac{\mean{\R[dev]{A}{}} - \mean{\R[msc]{A}{}}}{\mean{\R[dev]{A}{}} + \mean{\R[msc]{A}{}}} \label{eqn-ddi}
+    DDI = \frac{\mean{\R[dev]{A}{}} - \mean{\R[msc]{A}{}}}{\mean{\R[dev]{A}{}} + \mean{\R[msc]{A}{}}} \label{eq:ddi}
 \end{equation}
@@
     \caption{}
-    \label{fig2}
+    \label{fig:2}
 \end{figure*}
@@
 In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
-\textbf{d} Deviance detection indices (DDI, see Equation {eqn-ddi}) across networks and stimuli
+\textbf{d} Deviance detection indices (DDI, see Equation {eq:ddi}) across networks and stimuli
 
-To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \refpanel{2}{a}.
+To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{2}a.
 Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color.
@@
 The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster.
-We sum these patterns up in \refpanel{2}{b}, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
+We sum these patterns up in \FIG{2}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
 Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.
@@
 To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences.
-\refpanel{2}{c} shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
+\FIG{2}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
 The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15).
-Accordingly, the DDI (\refpanel{2}{d}) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Accordingly, the DDI (\FIG{2}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
 This shows that our model was well behaved and suitable for an investigation of deviance detection.
@@
     \caption{}
-    \label{fig3}
+    \label{fig:3}
 \end{figure*}
@@
 Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-\refpanel{3}{a} shows the responses to target stimulation of the sample network under ablation.
+\FIG{3}a shows the responses to target stimulation of the sample network under ablation.
 With no plasticity, the responses are indistinguishable between sequences, yielding a DDI of 0, and indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
@@
 
-Finally, across networks and stimuli, the DDI (\refpanel{3}{c}) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Finally, across networks and stimuli, the DDI (\FIG{3}c) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
 To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: T = 6.1e+03, p = 4.95e-11; TA: T = 4.87e+03, p = 0.000604), and greater too than the sum of the DDI values across both STD only and TA only models (T = 5.12e+03, p = 4.52e-05).
@@
 
-\subsection{The role of TA in deviance detection}\label{sec-ta}
+\subsection{The role of TA in deviance detection}\label{sec:ta}
 
 To understand why the same stimulus A evoked a higher response in \dev{} than in \msc{} with TA only, we first compared the average responses in the two sequences, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
-As shown in \refpanel{4}{a} (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
+As shown in \FIG{4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
 The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, which indicate higher \dev{} activity.
@@
     \caption{}
-    \label{fig4}
+    \label{fig:4}
 \end{figure*}
@@
 
-We quantified the former with the strength of adaptation $\theta$ (Equation \ref{eqn-TA}), i.e., the amount by which the firing threshold was raised due to TA.
-As shown in \refpanel{4}{b}, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+We quantified the former with the strength of adaptation $\theta$ (\EQ{TA}), i.e., the amount by which the firing threshold was raised due to TA.
+As shown in \FIG{4}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
@@
 To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we refer to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
-As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\refpanel{4}{c}) largely reflected that seen in spiking activity.
+As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\FIG{4}c) largely reflected that seen in spiking activity.
 The contrast (right panel) shows not only a notable relative increase in membrane voltages in \dev{} around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
@@
 We first focused on the contribution of $\theta$ alone to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \refpanel{4}{b}, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
-In the sample network (\refpanel{4}{d}), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in \refpanel{4}{e}, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \FIG{4}b, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
+In the sample network (\FIG{4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+As shown in \FIG{4}e, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
 In addition, the correlation across excitatory neurons between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
@@
 
-Yet, as we saw in \refpanel{4}{c}, the membrane voltages also differed between the two sequences.
+Yet, as we saw in \FIG{4}c, the membrane voltages also differed between the two sequences.
 In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
 We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference and summed across neurons.
-The resulting time courses, shown in \refpanel{4}{f}, show how much these two measures contributed to the increased \dev{} response.
+The resulting time courses, shown in \FIG{4}f, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (\refpanel{4}{g}), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
-A similar pattern emerged across networks and stimuli, as shown in \refpanel{4}{h}.
+In relative terms (\FIG{4}g), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+A similar pattern emerged across networks and stimuli, as shown in \FIG{4}h.
 Here, only the initial response was driven mainly by $\theta$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
@@
     \caption{}
-    \label{fig5}
+    \label{fig:5}
 \end{figure*}
@@
 Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
-A and B stimulus locations are highlighted in the left plot; see \reffig{1} for the locations of the remaining stimuli used in the \msc{} sequence.
+A and B stimulus locations are highlighted in the left plot; see \FIG{1} for the locations of the remaining stimuli used in the \msc{} sequence.
 \textbf{b} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel a.
@@
 
-First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \refpanel{5}{a}.
+First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \FIG{5}a.
 In the \msc{} sequence, $\T[msc]{A}{i}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
@@
 
-Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\refpanel{5}{b}), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
+Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\FIG{5}b), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
 Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.
 
-Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\refpanel{5}{c}), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in \refpanel{5}{d}, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
+Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\FIG{5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in \FIG{5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
 
@@
 Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
-We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B (and therefore more to non-target stimuli in \dev{} than \msc{}), and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in \reffig{4}.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B (and therefore more to non-target stimuli in \dev{} than \msc{}), and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in \FIG{4}.
 
@@
     \caption{}
-    \label{fig6}
+    \label{fig:6}
 \end{figure*}
@@
 In a given network and stimulus assignment, we ranked each neuron by the minimum response time (i.e., time to first spike) across B trials in all sequences.
-The ranking in the sample network is shown in \refpanel{6}{a}.
+The ranking in the sample network is shown in \FIG{6}a.
 We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site.
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
-In \refpanel{6}{b}, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
-Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \refpanel{5}{c}, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
+In \FIG{6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
 As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red, left plot) increasing their response in the \dev{} sequence, while the response of the later portion of the network (blue, right plot) decreased.
@@
 We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992: T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
-In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \refpanel{6}{c}) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \refpanel{6}{d}) within these two portions.
-We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \refpanel{5}{d}).
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{6}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{6}d) within these two portions.
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{5}d).
 $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
 
-Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \refpanel{6}{e}.
+Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \FIG{6}e.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
 
-To sum up, we showed in \reffig{4} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
+To sum up, we showed in \FIG{4} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
 Here, we showed that the neurons responding earliest to B exhibited higher $\T[dev]{B}{}$, which analogously reduced the response of the network as a whole to B.
@@
 
-\subsection{The role of STD in deviance detection}\label{sec-std}
+\subsection{The role of STD in deviance detection}\label{sec:std}
 
@@
 To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD.
-On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\refpanel{7}{a}).
+On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\FIG{7}a).
 However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (more than \msc{}: T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (less than \msc{}: T = 2.66e+03, p = 0.00554).
@@
     \caption{}
-    \label{fig7}
+    \label{fig:7}
 \end{figure*}
@@
 \textbf{a} Reductions in mean response size $\mean{\R{A}{}}$ in each sequence as a consequence of adding STD to the model, across networks and stimuli.
-\textbf{b} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
-Compare to \refpanel{6}{c-d}.
+\textbf{b} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA. Compare to \FIG{6}c-d.
 \textbf{c} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
@@
 \textbf{d} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
-Compare to \refpanel{5}{d}.
+Compare to \FIG{5}d.
 \textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
@@
 \textbf{f} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
-Compare to \refpanel{4}{e}.
+Compare to \FIG{4}e.
 \textbf{g} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
@@
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
-Starting at the cause of the reduced response in non-target trials, we show in \refpanel{7}{b} the full model equivalent of \refpanel{6}{c-d}.
+Starting at the cause of the reduced response in non-target trials, we show in \FIG{7}b the full model equivalent of \FIG{6}c-d.
 The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
@@
 
-Comparing the non-target response reductions in \refpanel{7}{c}, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
-Based on the ablated model analysis in \reffig{5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \refpanel{7}{d}.
-Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \refpanel{7}{e}, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Comparing the non-target response reductions in \FIG{7}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
+Based on the ablated model analysis in \FIG{5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
+Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{7}d.
+Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \FIG{7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
-Finally, based on the ablated model analysis in \reffig{4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
-As shown in \refpanel{7}{f}, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \refpanel{7}{g}, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
-This greater contrast, of course, directly reflects the greater DDI reported in \reffig{3}.
+Finally, based on the ablated model analysis in \FIG{4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
+As shown in \FIG{7}f, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{7}g, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
+This greater contrast, of course, directly reflects the greater DDI reported in \FIG{3}.
 
 Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
-The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\refpanel{7}{c}), which was where we noted the first clear shift towards increased deviance detection.
+The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\FIG{7}c), which was where we noted the first clear shift towards increased deviance detection.
 
@@
 We had good reasons for this assumption: Updates to the spike-triggered depression variable $x_j$ (cf.
-Equation \ref{eqn-xsyn}) would decay to a negligible $e^\frac{-1 s}{\tau_x} = 0.0013$ times the original value within two trials, making it unlikely that inputs separated by intervening trials could be ``remembered'' in the network.
+\EQ{xsyn}) would decay to a negligible $e^\frac{-1 s}{\tau_x} = 0.0013$ times the original value within two trials, making it unlikely that inputs separated by intervening trials could be ``remembered'' in the network.
 Conversely, the frequently presented stimulus B may have been able to lead to a build-up of depression strong enough to suppress its own response.
@@
 \begin{equation}
-    \Rhat{S}{j} = \frac{1}{3} (\R[dev]{S}{j} + \R[msc]{S}{j} + \R[std]{S}{j}) \label{eqn-Rhat}
+    \Rhat{S}{j} = \frac{1}{3} (\R[dev]{S}{j} + \R[msc]{S}{j} + \R[std]{S}{j}) \label{eq:Rhat}
 \end{equation}
@@
 \begin{equation}
-    \D{S}{i} = \sum_{j \in pre} \Rhat{S}{j} (1-x_j^S) k \label{eqn-D}
+    \D{S}{i} = \sum_{j \in pre} \Rhat{S}{j} (1-x_j^S) k \label{eq:D}
 \end{equation}
@@
     \caption{}
-    \label{fig8}
+    \label{fig:8}
 \end{figure*}
@@
 To show this, we calculated $\D{}{}$ for A and B trials and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
-In \refpanel{8}{a}, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
+In \FIG{8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
 The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast $\Delta \D{A}{i}$, which was negative throughout.
@@
 To show that $\D{}{}$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated $\Delta \D{A}{i}$ with $\Delta \R{A}{i}$ across neurons of the sample network.
-Notice that this mirrors the correlation shown in \reffig{4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
-As shown in \refpanel{8}{b}, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
-We confirmed this relationship across networks and stimuli, see \refpanel{8}{c}, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
+Notice that this mirrors the correlation shown in \FIG{4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
+As shown in \FIG{8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
+We confirmed this relationship across networks and stimuli, see \FIG{8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
 We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \ref{sfig8}).
@@
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in \refpanel{8}{d}, we recapitulate the logic established in \reffig{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
+First, in \FIG{8}d, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
-We can relate this directly to the non-target response contrasts in the early and late portions shown in \refpanel{7}{a}, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
+We can relate this directly to the non-target response contrasts in the early and late portions shown in \FIG{7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
 % Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
 
-To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \refpanel{6}{e}.
-As we show in \refpanel{8}{e}, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
+To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{6}e.
+As we show in \FIG{8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
 This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
@@
 If so, we should expect to see less $\D{A}{}$ in \dev{} than \msc{}.
-To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \refpanel{8}{c}.
+To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \FIG{8}c.
 We conclude that there is clear evidence for greater STD-mediated global suppression in \msc{} than \dev{} in the full model.
@@
 To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli (lower $\R[dev]{B}{}$) by local suppression, and an increase in the response to target stimuli (higher $\R[dev]{A}{}$) by relief from global suppression, relative to \msc{}.
-Yet, as we showed in \reffig{7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower thresholds $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
+Yet, as we showed in \FIG{7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower thresholds $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
 Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\refpanel{7}{e}, bottom) with the corresponding value of $\Delta \D{A}{}$ (\refpanel{8}{e}), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{8}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
@@
     \caption{}
-    \label{fig10}
+    \label{fig:10}
 \end{figure*}
@@
 To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.
-\refpanel{10}{a-c} shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
-Responses, normalized by the response to stimulation in the recovered networks (\refpanel{10}{a}), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
-For thresholds, $\T[std]{A}{}$ (\refpanel{10}{b}) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
-A similar pattern emerged with $\D{A}{}$ (\refpanel{10}{c}), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+\FIG{10}a-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (\FIG{10}a), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, $\T[std]{A}{}$ (\FIG{10}b) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with $\D{A}{}$ (\FIG{10}c), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
 
 To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
-We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \refpanel{10}{d}.
+We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{10}d.
 In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.
@@
 
-\section{Discussion}\label{sec-discussion}
+\section{Discussion}\label{sec:discussion}
 
@@
     \caption{}
-    \label{fig9}
+    \label{fig:9}
 \end{figure*}

### 2023-03-17 paper-1/elife.tex
Added elife version of the manuscript

--- 2023-03-17 paper-1/manuscript.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\RequirePackage{tikz}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% ELIFE ARTICLE TEMPLATE
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% PREAMBLE 
+\documentclass[9pt,lineno]{elife}
+% Use the onehalfspacing option for 1.5 line spacing
+% Use the doublespacing option for 2.0 line spacing
+% Please note that these options may affect formatting.
+% Additionally, the use of the \newcommand function should be limited.
 
-\documentclass[pdflatex,referee,iicol,sn-basic]{sn-jnl}
+\usepackage{lipsum} % Required to insert dummy text
+\usepackage[version=4]{mhchem}
+\usepackage{siunitx}
+\DeclareSIUnit\Molar{M}
 
-%%%% Standard Packages
-\usepackage{xcolor,hyperref}
-%%%%
-
-% Make Orcid icon
-\definecolor{lime}{HTML}{A6CE39}
-\DeclareRobustCommand{\orcidicon}{%
-	\begin{tikzpicture}
-	\draw[lime, fill=lime] (0,0) 
-	circle [radius=0.16] 
-	node[white] {{\fontfamily{qag}\selectfont \tiny ID}};
-	\draw[white, fill=white] (-0.0625,0.095) 
-	circle [radius=0.007];
-	\end{tikzpicture}
-	\hspace{-2mm}
-}
-
-\foreach \x in {A, ..., Z}{%
-	\expandafter\xdef\csname orcid\x\endcsname{\noexpand\href{https://orcid.org/\csname orcidauthor\x\endcsname}{\noexpand\orcidicon}}
-}
-
-% Define the ORCID iD command for each author separately.
-\newcommand{\orcidauthorA}{0000-0002-3420-4576}
-\newcommand{\orcidauthorB}{0000-0001-9402-0851}
-
+% FK: Noted, but these \newcommands are really very useful here:
 % Shorter italic dev/msc/std
@@
 % Common notation
-\renewcommand{\R}[3][]{{}^{#1}_{}\mathbf{R}^{#2}_{#3}}
-\renewcommand{\T}[3][]{{}^{#1}_{}\mathbf{T}^{#2}_{#3}}
-\newcommand{\D}[3][]{{}^{#1}_{}\mathbf{D}^{#2}_{#3}}
+\newcommand{\R}[3][]{{}^{#1}_{}\boldsymbol R^{#2}_{#3}}
+\newcommand{\T}[3][]{{}^{#1}_{}\boldsymbol T^{#2}_{#3}}
+\newcommand{\D}[3][]{{}^{#1}_{}\boldsymbol D^{#2}_{#3}}
 \newcommand{\mean}[1]{\overline{#1}}
-\newcommand{\Rhat}[3][]{{}^{#1}_{}\mathbf{\widehat{R}}^{#2}_{#3}}
+\newcommand{\Rhat}[3][]{{}^{#1}_{}\widehat{\boldsymbol R}^{#2}_{#3}}
 
-% Crossreferences
-\newcommand{\FIG}[1]{Figure \ref{fig:#1}}
-\newcommand{\EQ}[1]{Equation \ref{eq:#1}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% ARTICLE SETUP
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks}
 
-\jyear{2021}%
+\author[1]{Felix Benjamin Kern}
+\author[1*]{Zenas C. Chao}
+\affil[1]{International Research Center for Neurointelligence (WPI-IRCN), The University of Tokyo}
 
-\raggedbottom
-%%\unnumbered% uncomment this for unnumbered level heads
+\corr{zenas.c.chao@gmail.com}{ZCC}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%% ARTICLE START
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\title{Short-term neuronal and synaptic plasticity act in synergy for deviance detection in spiking networks}
+\maketitle
 
-\author[1]{\fnm{Felix Benjamin} \sur{Kern} \orcidA{}}\email{kernfel@gmail.com}
-
-\author*[1]{\fnm{Zenas C.} \sur{Chao} \orcidB{}}\email{zenas.c.chao@gmail.com}
-
-\affil[1]{\orgdiv{International Research Center for Neurointelligence (WPI-IRCN)}, \orgname{The University of Tokyo}, \orgaddress{\city{Tokyo}, \country{Japan}}}
-
-\abstract{
+\begin{abstract}
 Sensory areas of cortex respond more strongly to infrequent stimuli when these violate previously established regularities, a phenomenon known as deviance detection (DD).
@@
 We conclude that highly predictable information can be encoded in strong local suppression, which allows greater global recovery and subsequent heightened sensitivity for DD.
-}
-
-\keywords{deviance detection, adaptation, short-term plasticity, neuronal plasticity, synaptic plasticity}
-
-\maketitle
+\end{abstract}
 
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig1.pdf}
-    \caption{}
-    \label{fig:1}
-\end{figure*}
-\textbf{Fig. 1} Model and paradigm.
-\textbf{a} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes, illustrated above the plots.
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig1}
+    \caption{%
+        Model and paradigm.
+        \textbf{A} Membrane and synaptic dynamics of a neuron innervated by an excitatory and an inhibitory neuron firing pre-determined spikes, illustrated above the plots.
 The excitatory connection has weight $w = 3$ for demonstration purposes only; a single connection with $w = 1$ is normally unable to evoke postsynaptic firing.
@@
 Bottom: Membrane voltage, firing threshold baseline (gray) and adaptive threshold, demonstrating TA in the excitatory postsynaptic neuron.
-\textbf{b} Sample network layout, showing the spatial location of excitatory and inhibitory neurons, and all outgoing synaptic connections from two neurons of each type.
-\textbf{c} Distance from stimulation site 1 in terms of minimum number of excitatory synapses.
+        \textbf{B} Sample network layout, showing the spatial location of excitatory and inhibitory neurons, and all outgoing synaptic connections from two neurons of each type.
+        \textbf{C} Distance from stimulation site 1 in terms of minimum number of excitatory synapses.
 Stimulation is provided to the 10 neurons nearest to the center of each of 5 stimulation sites (large circles), which are regularly spaced on a circle with a radius of 2.5 mm.
-\textbf{d} Illustration of the stimulation paradigm.
+        \textbf{D} Illustration of the stimulation paradigm.
 Each marker represents a stimulation, with the stimulus identity indicated by color.
@@
 Three randomized sequences were presented, containing 80\% A and 20\% B (\std{}), 20\% A and 80\% B (\dev{}), and 20\% of each of 5 stimuli, including A and B (\msc{}).
-Note that, while the onset asynchrony of 500 ms is represented faithfully, the horizontal extent of the stimulation markers has no meaning; stimuli were presented instantaneously rather than over an extended period
+Note that, while the onset asynchrony of 500 ms is represented faithfully, the horizontal extent of the stimulation markers has no meaning; stimuli were presented instantaneously rather than over an extended period.
+    }
+    \label{fig:1}
+\end{figure}
 
-\FIG{1}a shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
+\FIG{1}A shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
 The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \citep{Stimberg2020-go}, and simulated with an integration time step of 1~ms.
@@
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-\FIG{1}b shows a representative example of the resulting network structure.
-In \FIG{1}c, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
+\FIG{1}B shows a representative example of the resulting network structure.
+In \FIG{1}C, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
 Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response: Networks where fewer than 500 neurons responded to stimulation at any of the five sites after full recovery were discarded.
@@
 
-To investigate deviance detection, we used a classical oddball paradigm as illustrated in \FIG{1}d, presenting a total of 500 stimuli at regular intervals of 500~ms.
+To investigate deviance detection, we used a classical oddball paradigm as illustrated in \FIG{1}D, presenting a total of 500 stimuli at regular intervals of 500~ms.
 Target stimuli (labeled ``A'' or ``target'' hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The ``standard'' sequence (\std{}), consisting of 400 presentations of A and 100 presentations of B; the ``deviant'' sequence (\dev{}) with 100 A and 400 B, and the many-standards control sequence (\msc{}) with 100 presentations of each of the stimuli A through E.
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig2.pdf}
-    \caption{}
-    \label{fig:2}
-\end{figure*}
-\textbf{Fig. 2} Deviance detection in our model.
-\textbf{a} Sample network target response intensity by trial type.
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig2}
+    \caption{%
+        Deviance detection in our model.
+        \textbf{A} Sample network target response intensity by trial type.
 Each horizontal line of the plot summarizes the network response in one target trial in terms of spikes per integration time step.
 Target trials are taken from the \std{} (400), \msc{} (100), and \dev{} (100) sequences, and sorted by sequence time (earliest trials are at the bottom of each section).
-\textbf{b} Response summary, showing average network-wide spike counts in target trials in each sequence.
-The underlying data are the same as in panel a.
+        \textbf{B} Response summary, showing average network-wide spike counts in target trials in each sequence.
+The underlying data are the same as in panel A.
 Note how the \dev{} response is very large, followed by an intermediate \msc{} response and a low \std{} response.
-\textbf{c} Mean response sizes $\mean{\R[dev]{A}{}}$ and $\mean{\R[std]{A}{}}$, normalized by $\mean{\R[msc]{A}{}}$, across networks and stimuli (n = 120).
+        \textbf{C} Mean response sizes $\mean{\R[dev]{A}{}}$ and $\mean{\R[std]{A}{}}$, normalized by $\mean{\R[msc]{A}{}}$, across networks and stimuli (n = 120).
 In this and later boxplots, the median is indicated with an orange line, notches indicate the 95\% confidence interval for the median calculated by bootstrap with 10000 iterations, boxes indicate the inter-quartile range (IQR), whiskers extend to 1.5 times the IQR or to the data extrema, whichever is less, and fliers indicate individual data points beyond the whisker end points.
-\textbf{d} Deviance detection indices (DDI, see Equation {eq:ddi}) across networks and stimuli
+        \textbf{D} Deviance detection indices (DDI, see Equation {eq:ddi}) across networks and stimuli.
+    }
+    \label{fig:2}
+\end{figure}
 
-To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{2}a.
+To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{2}A.
 Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color.
@@
 The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster.
-We sum these patterns up in \FIG{2}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
+We sum these patterns up in \FIG{2}B, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
 Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.
@@
 To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences.
-\FIG{2}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
+\FIG{2}C shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
 The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15).
-Accordingly, the DDI (\FIG{2}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Accordingly, the DDI (\FIG{2}D) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
 This shows that our model was well behaved and suitable for an investigation of deviance detection.
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig3.pdf}
-    \caption{}
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig3}
+    \caption{%
+        Deviance detection under model ablation.
+        \textbf{A} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
+Lines show the average (trial mean +- SEM, left axis) network-wide spike counts in target trials in each sequence.
+Boxes summarize the spike count totals per trial (right axis), the means of which form the basis for index calculations (see panel C).
+Note the different axis scaling on the top row (without TA) and the bottom row (with TA).
+        \textbf{B} Deviance detection indices across networks and stimuli.
+Asterisks indicate data greater than 0, established with a Wilcoxon signed-rank test at a significance level 0.05.
+See main text for detailed statistics.
+    }
     \label{fig:3}
-\end{figure*}
-\textbf{Fig. 3} Deviance detection under model ablation.
-\textbf{a} Sample network responses to target stimulation, with the deviance detection index in each instance noted in the plot titles.
-Lines show the average (trial mean +- SEM, left axis) network-wide spike counts in target trials in each sequence.
-Boxes summarize the spike count totals per trial (right axis), the means of which form the basis for index calculations (see panel c).
-Note the different axis scaling on the top row (without TA) and the bottom row (with TA).
-\textbf{b} Deviance detection indices across networks and stimuli.
-Asterisks indicate data greater than 0, established with a Wilcoxon signed-rank test at a significance level 0.05.
-See main text for detailed statistics
+\end{figure}
 
 Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-\FIG{3}a shows the responses to target stimulation of the sample network under ablation.
+\FIG{3}A shows the responses to target stimulation of the sample network under ablation.
 With no plasticity, the responses are indistinguishable between sequences, yielding a DDI of 0, and indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
@@
 
-Finally, across networks and stimuli, the DDI (\FIG{3}c) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Finally, across networks and stimuli, the DDI (\FIG{3}C) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
 To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: T = 6.1e+03, p = 4.95e-11; TA: T = 4.87e+03, p = 0.000604), and greater too than the sum of the DDI values across both STD only and TA only models (T = 5.12e+03, p = 4.52e-05).
@@
 To understand why the same stimulus A evoked a higher response in \dev{} than in \msc{} with TA only, we first compared the average responses in the two sequences, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
-As shown in \FIG{4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
+As shown in \FIG{4}A (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
 The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, which indicate higher \dev{} activity.
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig4.pdf}
-    \caption{}
-    \label{fig:4}
-\end{figure*}
-\textbf{Fig. 4} Responses to \dev{} target trials are larger due to lower initial $\theta$.
-\textbf{a} Post-stimulus spike histogram across target trials in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig4}
+    \caption{%
+        Responses to \dev{} target trials are larger due to lower initial $\theta$.
+        \textbf{A} Post-stimulus spike histogram across target trials in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
 Left and middle column, target trials in the \dev{} sequence and \msc{} sequence, respectively; right column, contrast between these two.
-\textbf{b} Target trial average of the threshold adaptation voltage $\theta$, with neuron order and columns as in panel a.
-\textbf{c} Target trial average of the membrane potential $V_m$, with neuron order and columns as in panel a.
+        \textbf{B} Target trial average of the threshold adaptation voltage $\theta$, with neuron order and columns as in panel A.
+        \textbf{C} Target trial average of the membrane potential $V_m$, with neuron order and columns as in panel A.
 The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the colormap.
-\textbf{d} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $\theta$ at the start of target trials ($\Delta \T{A}{i}$) and the contrast in the average number of spikes fired in target trials ($\Delta \R{A}{i}$), and associated linear regression.
-\textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \T{A}{i}$ vs $\Delta \R{A}{i}$ across excitatory neurons).
+        \textbf{D} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $\theta$ at the start of target trials ($\Delta \T{A}{i}$) and the contrast in the average number of spikes fired in target trials ($\Delta \R{A}{i}$), and associated linear regression.
+        \textbf{E} $\Delta \T{A}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \T{A}{i}$ vs $\Delta \R{A}{i}$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
@@
 Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
-\textbf{f} In the sample network, contribution of the contrast in $\theta$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel a), weighted by that same contrast.
-\textbf{g} Contribution of $\theta$ relative to the sum of the absolute values of the weighted contributions in panel f.
-\textbf{h} Relative contribution of $\theta$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area)
+        \textbf{F} In the sample network, contribution of the contrast in $\theta$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel A), weighted by that same contrast.
+        \textbf{G} Contribution of $\theta$ relative to the sum of the absolute values of the weighted contributions in panel F.
+        \textbf{H} Relative contribution of $\theta$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area).
+    }
+    \label{fig:4}
+    % 
+    \figsupp[The equivalent figure in the full model with both TA and STD.]{%
+        The equivalent of \FIG{4} in the full model with both TA and STD.
+        \textbf{A} Post-stimulus spike histogram across target trials in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
+Left and middle column, target trials in the \dev{} sequence and \msc{} sequence, respectively; right column, contrast between these two.
+        \textbf{B} Target trial average of the threshold adaptation voltage $\theta$, with neuron order and columns as in panel A.
+        \textbf{C} Target trial average of the membrane potential $V_m$, with neuron order and columns as in panel A.
+The first 10 ms of the neurons under direct stimulation are masked out to avoid saturation of the colormap.
+        \textbf{D} Relationship across excitatory neurons of the sample network between the contrast (dev - msc) in $\theta$ at the start of target trials ($\Delta \T{A}{i}$) and the contrast in the average number of spikes fired in target trials ($\Delta \R{A}{i}$), and associated linear regression.
+        \textbf{E} $\Delta \T{A}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \T{A}{i}$ vs $\Delta \R{A}{i}$ across excitatory neurons).
+Each marker corresponds to one network and target stimulus.
+Red markers indicate a significant negative correlation (one-sided Wald test, significance level 0.05).
+Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
+        \textbf{F} In the sample network, contribution of the contrast in $\theta$ and $V_m$ to bins with increased firing in \dev{} (i.e., positive/red bins in the spike probability contrast, panel A), weighted by that same contrast.
+        \textbf{G} Contribution of $\theta$ relative to the sum of the absolute values of the weighted contributions in panel F.
+        \textbf{H} Relative contribution of $\theta$ across networks and stimuli, showing median (solid line) and inter-quartile range (shaded area).
+    }
+    {\includegraphics[width=\linewidth]{fig4-withSTD}}\label{figsupp:withSTD}
+\end{figure}
 
 We quantified the former with the strength of adaptation $\theta$ (\EQ{TA}), i.e., the amount by which the firing threshold was raised due to TA.
-As shown in \FIG{4}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+As shown in \FIG{4}B, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
@@
 To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we refer to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
-As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\FIG{4}c) largely reflected that seen in spiking activity.
+As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\FIG{4}C) largely reflected that seen in spiking activity.
 The contrast (right panel) shows not only a notable relative increase in membrane voltages in \dev{} around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
@@
 We first focused on the contribution of $\theta$ alone to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \FIG{4}b, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
-In the sample network (\FIG{4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in \FIG{4}e, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \FIG{4}B, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
+In the sample network (\FIG{4}D), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+As shown in \FIG{4}E, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
 In addition, the correlation across excitatory neurons between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
@@
 
-Yet, as we saw in \FIG{4}c, the membrane voltages also differed between the two sequences.
+Yet, as we saw in \FIG{4}C, the membrane voltages also differed between the two sequences.
 In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
 We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference and summed across neurons.
-The resulting time courses, shown in \FIG{4}f, show how much these two measures contributed to the increased \dev{} response.
+The resulting time courses, shown in \FIG{4}F, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (\FIG{4}g), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
-A similar pattern emerged across networks and stimuli, as shown in \FIG{4}h.
+In relative terms (\FIG{4}G), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+A similar pattern emerged across networks and stimuli, as shown in \FIG{4}H.
 Here, only the initial response was driven mainly by $\theta$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig5.pdf}
-    \caption{}
-    \label{fig:5}
-\end{figure*}
-\textbf{Fig. 5} Lower thresholds in \dev{} are a consequence of lower \dev{} non-target response magnitude.
-\textbf{a} $\T{A}{i}$ in the sample network, with each neuron shown in its spatial location.
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig5}
+    \caption{%
+        Lower thresholds in \dev{} are a consequence of lower \dev{} non-target response magnitude.
+        \textbf{A} $\T{A}{i}$ in the sample network, with each neuron shown in its spatial location.
 Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
 A and B stimulus locations are highlighted in the left plot; see \FIG{1} for the locations of the remaining stimuli used in the \msc{} sequence.
-\textbf{b} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel a.
-\textbf{c} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$, and associated linear regression.
-\textbf{d} $\Delta \R{N}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \R{N}{i}$ vs $\Delta \T{A}{i}$ across excitatory neurons).
+        \textbf{B} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel A.
+        \textbf{C} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$, and associated linear regression.
+        \textbf{D} $\Delta \R{N}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \R{N}{i}$ vs $\Delta \T{A}{i}$ across excitatory neurons).
 Each marker corresponds to one network and target stimulus.
 Red markers indicate a significant positive correlation (one-sided Wald test, significance level 0.05).
-Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations
+Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
+    }
+    \label{fig:5}
+    % 
+    \figsupp[The equivalent figure in the full model with both TA and STD.]{%
+        The equivalent of \FIG{5} in the full model with both TA and STD.
+        \textbf{A} $\T{A}{i}$ in the sample network, with each neuron shown in its spatial location.
+Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
+A and B stimulus locations are highlighted in the left plot; see \FIG{1} for the locations of the remaining stimuli used in the \msc{} sequence.
+        \textbf{B} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel A.
+        \textbf{C} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$, and associated linear regression.
+        \textbf{D} $\Delta \R{N}{exc}$ across networks and stimuli, plotted against the Pearson correlation coefficients $\rho$ of the relationship shown in the previous panel ($\Delta \R{N}{i}$ vs $\Delta \T{A}{i}$ across excitatory neurons).
+Each marker corresponds to one network and target stimulus.
+Red markers indicate a significant positive correlation (one-sided Wald test, significance level 0.05).
+Inset histograms show the distribution of the corresponding values, with orange portions indicating the 95\% confidence interval of the median, calculated by bootstrap with 10000 iterations.
+    }
+    {\includegraphics[width=\linewidth]{fig5-withSTD}}\label{figsupp:withSTD}
+\end{figure}
 
-First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \FIG{5}a.
+First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \FIG{5}A.
 In the \msc{} sequence, $\T[msc]{A}{i}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
@@
 
-Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\FIG{5}b), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
+Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\FIG{5}B), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
 Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.
 
-Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\FIG{5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in \FIG{5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
+Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\FIG{5}C), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in \FIG{5}D, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
 
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig6.pdf}
-    \caption{}
-    \label{fig:6}
-\end{figure*}
-\textbf{Fig. 6} Dev non-target responses are reduced due to adaptation in early responders to B.
-\textbf{a} The sample network, colored by latency rank, i.e.
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig6}
+    \caption{%
+        Dev non-target responses are reduced due to adaptation in early responders to B.
+        \textbf{A} The sample network, colored by latency rank, i.e.
 the rank order of the time to first spike in response to stimulation at site B in any sequence.
@@
 Note that non-responding neurons are ranked last and appear dark blue.
-\textbf{b} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$, and associated linear regression.
-Neurons are colored by latency rank as in panel a.
+        \textbf{B} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$, and associated linear regression.
+Neurons are colored by latency rank as in panel A.
 Additionally, excitatory neurons among the earliest 100 responders are shown in the left plot, while later responders are shown in the right plot.
 Inset histograms show the distributions of the plotted values, with the confidence intervals of their medians, calculated by bootstrap with 10000 iterations, in darker color.
-\textbf{c} Median non-target response contrast $\Delta \R{N}{}$ among the first 100 neurons to respond to stimulation in B, labeled ``early'', and among the later neurons, labeled ``late'', across networks and stimuli.
-\textbf{d} Median contrast in $\T{B}{}$ among the early and late neurons, respectively, across networks and stimuli.
-\textbf{e} Relationship across networks and stimuli between $\Delta \T{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression
+        \textbf{C} Median non-target response contrast $\Delta \R{N}{}$ among the first 100 neurons to respond to stimulation in B, labeled ``early'', and among the later neurons, labeled ``late'', across networks and stimuli.
+        \textbf{D} Median contrast in $\T{B}{}$ among the early and late neurons, respectively, across networks and stimuli.
+        \textbf{E} Relationship across networks and stimuli between $\Delta \T{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression.
+    }
+    \label{fig:6}
+    % 
+    \figsupp[The equivalent figure in the full model with both TA and STD.]{%
+        The equivalent of \FIG{6} in the full model with both TA and STD.
+        \textbf{A} The sample network, colored by latency rank, i.e.
+the rank order of the time to first spike in response to stimulation at site B in any sequence.
+The response develops asymmetrically due to structural properties, not due to interactions with A.
+Note that non-responding neurons are ranked last and appear dark blue.
+        \textbf{B} Relationship across excitatory neurons of the sample network between $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$, and associated linear regression.
+Neurons are colored by latency rank as in panel A.
+Additionally, excitatory neurons among the earliest 100 responders are shown in the left plot, while later responders are shown in the right plot.
+Inset histograms show the distributions of the plotted values, with the confidence intervals of their medians, calculated by bootstrap with 10000 iterations, in darker color.
+        \textbf{C} Median non-target response contrast $\Delta \R{N}{}$ among the first 100 neurons to respond to stimulation in B, labeled ``early'', and among the later neurons, labeled ``late'', across networks and stimuli.
+        \textbf{D} Median contrast in $\T{B}{}$ among the early and late neurons, respectively, across networks and stimuli.
+        \textbf{E} Relationship across networks and stimuli between $\Delta \T{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression.
+    }
+    {\includegraphics[width=\linewidth]{fig6-withSTD}}\label{figsupp:withSTD}
+\end{figure}
 
@@
 In a given network and stimulus assignment, we ranked each neuron by the minimum response time (i.e., time to first spike) across B trials in all sequences.
-The ranking in the sample network is shown in \FIG{6}a.
+The ranking in the sample network is shown in \FIG{6}A.
 We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site.
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
-In \FIG{6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
-Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
+In \FIG{6}B, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{5}C, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
 As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red, left plot) increasing their response in the \dev{} sequence, while the response of the later portion of the network (blue, right plot) decreased.
@@
 We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992: T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
-In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{6}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{6}d) within these two portions.
-We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{5}d).
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{6}C) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{6}D) within these two portions.
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{5}D).
 $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
 
-Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \FIG{6}e.
+Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \FIG{6}E.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
@@
 To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD.
-On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\FIG{7}a).
+On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\FIG{7}A).
 However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (more than \msc{}: T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (less than \msc{}: T = 2.66e+03, p = 0.00554).
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig7.pdf}
-    \caption{}
-    \label{fig:7}
-\end{figure*}
-\textbf{Fig. 7} Adding STD to the model slightly enhances the effects of TA.
-\textbf{a} Reductions in mean response size $\mean{\R{A}{}}$ in each sequence as a consequence of adding STD to the model, across networks and stimuli.
-\textbf{b} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA. Compare to \FIG{6}c-d.
-\textbf{c} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig7}
+    \caption{%
+        Adding STD to the model slightly enhances the effects of TA.
+        \textbf{A} Reductions in mean response size $\mean{\R{A}{}}$ in each sequence as a consequence of adding STD to the model, across networks and stimuli.
+        \textbf{B} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+Compare to \FIG{6}C-d.
+        \textbf{C} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
 The red line is the identity function; points above this line represent networks with greater response reduction (\dev{} - \msc{}) in the full model than in the ablated model.
 The subtraction of the two quantities in the box plot below shows that the non-target response tends to be reduced more in the full model.
-\textbf{d} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
-Compare to \FIG{5}d.
-\textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
+        \textbf{D} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
+Compare to \FIG{5}D.
+        \textbf{E} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel C.
 The subtraction of the two quantities in the box plot below shows that $\T{A}{}$ tends to be reduced more in the full model.
-\textbf{f} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
-Compare to \FIG{4}e.
-\textbf{g} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
-The subtraction of the two quantities in the box plot below shows that the target response tends to be increased more in the full model
+        \textbf{F} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
+Compare to \FIG{4}E.
+        \textbf{G} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel C.
+The subtraction of the two quantities in the box plot below shows that the target response tends to be increased more in the full model.
+    }
+    \label{fig:7}
+\end{figure}
 
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
-Starting at the cause of the reduced response in non-target trials, we show in \FIG{7}b the full model equivalent of \FIG{6}c-d.
+Starting at the cause of the reduced response in non-target trials, we show in \FIG{7}B the full model equivalent of \FIG{6}C-d.
 The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
@@
 
-Comparing the non-target response reductions in \FIG{7}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
+Comparing the non-target response reductions in \FIG{7}C, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
 Based on the ablated model analysis in \FIG{5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{7}d.
-Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \FIG{7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{7}D.
+Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \FIG{7}E, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
 Finally, based on the ablated model analysis in \FIG{4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
-As shown in \FIG{7}f, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{7}g, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
+As shown in \FIG{7}F, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{7}G, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
 This greater contrast, of course, directly reflects the greater DDI reported in \FIG{3}.
@@
 Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
-The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\FIG{7}c), which was where we noted the first clear shift towards increased deviance detection.
+The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\FIG{7}C), which was where we noted the first clear shift towards increased deviance detection.
 
@@
 We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
-Similar behavior was also seen in the \msc{} sequence, as shown in \ref{sfig-STD-correlation-decay}.
+Similar behavior was also seen in the \msc{} sequence, as shown in \FIGSUPP[8]{correlation-decay}.
 
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig8.pdf}
-    \caption{}
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig8}
+    \caption{%
+        In the full model, STD reduces \dev{} non-target and increases \dev{} target responses relative to \msc{}.
+        \textbf{A} Average depression $\D{}{i}$ in target (A) trials (top row) and B trials (bottom row) in the \dev{} (left) and \msc{} sequence (center), and their contrast (right) in the sample network.
+The color scale is shared across columns (and between \dev{} and \msc{}) for direct comparability.
+        \textbf{B} Relationship across neurons of the sample network between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$, and associated linear regression.
+        \textbf{C} $\Delta \D{A}{}$ across networks and stimuli, plotted against the corresponding Pearson coefficient between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$.
+        \textbf{D} $\Delta \D{B}{}$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
+        \textbf{E} Relationship across networks and stimuli between $\Delta \D{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression.
+    }
     \label{fig:8}
-\end{figure*}
-\textbf{Fig. 8} In the full model, STD reduces \dev{} non-target and increases \dev{} target responses relative to \msc{}.
-\textbf{a} Average depression $\D{}{i}$ in target (A) trials (top row) and B trials (bottom row) in the \dev{} (left) and \msc{} sequence (center), and their contrast (right) in the sample network.
+    % 
+    \figsupp[The equivalent figure in the ablated model with only STD.]{%
+        The equivalent of \FIG{8} in the ablated model with only STD.
+        \textbf{A} Average depression $\D{}{i}$ in target (A) trials (top row) and B trials (bottom row) in the \dev{} (left) and \msc{} sequence (center), and their contrast (right) in the sample network.
 The color scale is shared across columns (and between \dev{} and \msc{}) for direct comparability.
-\textbf{b} Relationship across neurons of the sample network between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$, and associated linear regression.
-\textbf{c} $\Delta \D{A}{}$ across networks and stimuli, plotted against the corresponding Pearson coefficient between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$.
-\textbf{d} $\Delta \D{B}{}$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
-\textbf{e} Relationship across networks and stimuli between $\Delta \D{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression
+        \textbf{B} Relationship across neurons of the sample network between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$, and associated linear regression.
+        \textbf{C} $\Delta \D{A}{}$ across networks and stimuli, plotted against the corresponding Pearson coefficient between $\Delta \D{A}{i}$ and $\Delta \R{A}{i}$.
+        \textbf{D} $\Delta \D{B}{}$ in the first 100 neurons to respond to B, and in the remaining, late portion of the network, across networks and stimuli.
+        \textbf{E} Relationship across networks and stimuli between $\Delta \D{B}{early}$ and $\Delta \R{B}{}$, and associated linear regression.
+    }
+    {\includegraphics[width=\linewidth]{fig8-noTA}}\label{figsupp:noTA}
+    %
+    \figsupp[The STD variable $x_i$ only correlates with the activity of the immediately preceding trial.]{%
+        The STD variable $x_i$ only correlates with the activity of the immediately preceding trial.
+        Each marker represents the coefficient $\rho$ of the Pearson correlation between the spike counts in a target trial and the corresponding values $x_i$ at the beginning of the immediately following trial ($\Delta$ trials = 1) or after one intervening trial ($\Delta$ trials = 2).
+Recall that $x$ is reduced upon spiking, which implies that an effective correlation is negative.
+Color represents the p-value of the alternative hypothesis that $\rho < 0$.
+Horizontal lines indicate the median $\rho$ and p.
+Horizontal jitter is introduced to aid visual separability of the markers.
+    }
+    {\includegraphics[width=8cm]{sfig-STD-correlation-decay}}\label{figsupp:correlation-decay}
+\end{figure}
 
@@
 To show this, we calculated $\D{}{}$ for A and B trials and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
-In \FIG{8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
+In \FIG{8}A, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
 The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast $\Delta \D{A}{i}$, which was negative throughout.
@@
 Notice that this mirrors the correlation shown in \FIG{4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
-As shown in \FIG{8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
-We confirmed this relationship across networks and stimuli, see \FIG{8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
-We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \ref{sfig8}).
+As shown in \FIG{8}B, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
+We confirmed this relationship across networks and stimuli, see \FIG{8}C, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
+We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \FIGSUPP[8]{noTA}).
 
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in \FIG{8}d, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
+First, in \FIG{8}D, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
-We can relate this directly to the non-target response contrasts in the early and late portions shown in \FIG{7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
+We can relate this directly to the non-target response contrasts in the early and late portions shown in \FIG{7}A, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
 % Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
 
-To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{6}e.
-As we show in \FIG{8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
+To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{6}E.
+As we show in \FIG{8}E, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
 This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
@@
 If so, we should expect to see less $\D{A}{}$ in \dev{} than \msc{}.
-To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \FIG{8}c.
+To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \FIG{8}C.
 We conclude that there is clear evidence for greater STD-mediated global suppression in \msc{} than \dev{} in the full model.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{8}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}E, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{8}E), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig10.pdf}
-    \caption{}
-    \label{fig:10}
-\end{figure*}
-\textbf{Fig. 9} Suppression acts in spatially distinguishable subpopulations.
-\textbf{a} Response size $\R{A}{}$ as a function of distance from the center of the stimulation site, evaluated across all networks and stimuli in 200 bins between 0 mm and 6.5 mm, and normalized by the response size in recovered networks (initial trial).
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig10}
+    \caption{%
+        Suppression acts in spatially distinguishable subpopulations.
+        \textbf{A} Response size $\R{A}{}$ as a function of distance from the center of the stimulation site, evaluated across all networks and stimuli in 200 bins between 0 mm and 6.5 mm, and normalized by the response size in recovered networks (initial trial).
 Each of the 200 data points represents the mean response across all relevant trials of all neurons in the corresponding distance bin.
@@
 The vertical line indicates the chosen local/global divider.
-\textbf{b} As in panel a, but showing $\T{A}{}$ without normalization.
-\textbf{c} As in panel a, but showing $\D{A}{}$ without normalization.
-\textbf{d} Mean $\T{A}{}$ in neurons <2 mm from the stimulation site (local) and more distant neurons (global).
+        \textbf{B} As in panel A, but showing $\T{A}{}$ without normalization.
+        \textbf{C} As in panel A, but showing $\D{A}{}$ without normalization.
+        \textbf{D} Mean $\T{A}{}$ in neurons <2 mm from the stimulation site (local) and more distant neurons (global).
 Data are averaged across relevant neurons and trials in each network and stimulus, leaving 120 data points to be summarized by box plot.
 Comparisons of interest are between \std{} and \dev{} in the local subset, and between \msc{} and \dev{} in the global subset.
-\textbf{e} Corresponding mean $\D{A}{}$ in local and global portions of the network.
+        \textbf{E} Corresponding mean $\D{A}{}$ in local and global portions of the network.
+    }
+    \label{fig:10}
+\end{figure}
 
@@
 To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.
-\FIG{10}a-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
-Responses, normalized by the response to stimulation in the recovered networks (\FIG{10}a), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
-For thresholds, $\T[std]{A}{}$ (\FIG{10}b) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
-A similar pattern emerged with $\D{A}{}$ (\FIG{10}c), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+\FIG{10}A-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (\FIG{10}A), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, $\T[std]{A}{}$ (\FIG{10}B) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with $\D{A}{}$ (\FIG{10}C), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
 
 To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
-We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{10}d.
+We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{10}D.
 In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.
@@
 
-\begin{figure*}%
-    \centering
-    \includegraphics{fig9.pdf}
-    \caption{}
-    \label{fig:9}
-\end{figure*}
-\textbf{Fig. 10} Illustration of the origin of DD in our model.
-\textbf{a} Left: In an oddball context, the regular inputs are encoded in the network by high local suppression, which reduces their overall response, leading to low global suppression.
+\begin{figure}
+    \includegraphics[width=\linewidth]{fig9}
+    \caption{%
+        Illustration of the origin of DD in our model.
+        \textbf{A} Left: In an oddball context, the regular inputs are encoded in the network by high local suppression, which reduces their overall response, leading to low global suppression.
 Deviant inputs encounter low local and global suppression and therefore evoke a large response.
@@
 Deviant inputs encounter this high global suppression and therefore evoke smaller responses.
-\textbf{b} Adding additional suppression (e.g., STD) affects these situations differently.
+        \textbf{B} Adding additional suppression (e.g., STD) affects these situations differently.
 Left: In an oddball context, the additional suppression further reduces the regular response, permitting even greater recovery of the wider network.
 Right: In an irregular context, global suppression is instead increased, and all responses reduced.
+    }
+    \label{fig:9}
+\end{figure}
 
@@
 
-\backmatter
-
-\bmhead{Supplementary information}
-TODO
-
-\bmhead{Author contributions}
-All authors contributed to the study conception and design.
-Material preparation, data collection and analysis were performed by Felix Benjamin Kern.
-The first draft of the manuscript was written by Felix Benjamin Kern and all authors commented on previous versions of the manuscript.
-All authors read and approved the final manuscript.
-
-\bmhead{Code availability}
-TODO
-
-\section*{Declarations}
-
-\bmhead{Funding}
-This study was funded by a World Premier International Research Center Initiative research startup grant.
-
-\bmhead{Competing interests}
-The authors have no relevant financial or non-financial interests to disclose.
-
-\bmhead{Ethics approval}
-Not applicable.
-
-\bmhead{Consent to participate}
-Not applicable.
-
-\bmhead{Consent for publication}
-Not applicable.
-
-\begin{appendices}
-
-\section{Supplementary information}\label{secA1}
-
-\begin{figure*}%
-    \centering
-    \includegraphics{fig4-withSTD.pdf}
-    \caption{}
-    \label{sfig4}
-\end{figure*}
-
-\begin{figure*}%
-    \centering
-    \includegraphics{fig5-withSTD.pdf}
-    \caption{}
-    \label{sfig5}
-\end{figure*}
-
-\begin{figure*}%
-    \centering
-    \includegraphics{fig6-withSTD.pdf}
-    \caption{}
-    \label{sfig6}
-\end{figure*}
-
-\begin{figure*}%
-    \centering
-    \includegraphics{sfig-STD-correlation-decay.pdf}
-    \caption{}
-    \label{sfig-STD-correlation-decay}
-\end{figure*}
-\textbf{Fig.
-S?} The STD variable $x_i$ only correlates with the activity of the immediately preceding trial.
-Each marker represents the coefficient $\rho$ of the Pearson correlation between the spike counts in a target trial and the corresponding values $x_i$ at the beginning of the immediately following trial ($\Delta$ trials = 1) or after one intervening trial ($\Delta$ trials = 2).
-Recall that $x$ is reduced upon spiking, which implies that an effective correlation is negative.
-Color represents the p-value of the alternative hypothesis that $\rho < 0$.
-Horizontal lines indicate the median $\rho$ and p.
-Horizontal jitter is introduced to aid visual separability of the markers.
-
-\begin{figure*}%
-    \centering
-    \includegraphics{fig8-noTA.pdf}
-    \caption{}
-    \label{sfig8}
-\end{figure*}
-
-% %%=============================================%%
-% %% For submissions to Nature Portfolio Journals %%
-% %% please use the heading ``Extended Data''.   %%
-% %%=============================================%%
-
-% %%=============================================================%%
-% %% Sample for another appendix section			       %%
-% %%=============================================================%%
-
-% %% \section{Example of another appendix section}\label{secA2}%
-% %% Appendices may be used for helpful, supporting or essential material that would otherwise 
-% %% clutter, break up or be distracting to the text. Appendices can consist of sections, figures, 
-% %% tables and equations etc.
-
-\end{appendices}
-
-%%===========================================================================================%%
-%% If you are submitting to one of the Nature Portfolio journals, using the eJP submission   %%
-%% system, please include the references within the manuscript file itself. You may do this  %%
-%% by copying the reference list from your .bbl file, paste it into the main manuscript .tex %%
-%% file, and delete the associated \verb+\bibliography+ commands.                            %%
-%%===========================================================================================%%
-
-\bibliography{refs}% common bib file
-%% if required, the content of .bbl file can be included here once bbl is generated
-%%\input sn-article.bbl
-
-%% Default %%
-%%\input sn-sample-bib.tex%
+\bibliography{refs}
 

### 2023-03-20 paper-1/elife.tex
Set to 1.5 line spacing
Added WPI funding ack
Fixed hat t_j appearance
Aligned fig file naming & fix references in tex

@@
 %%% PREAMBLE 
-\documentclass[9pt,lineno]{elife}
+\documentclass[9pt,lineno,onehalfspacing]{elife}
 % Use the onehalfspacing option for 1.5 line spacing
@@
 
-\usepackage{lipsum} % Required to insert dummy text
 \usepackage[version=4]{mhchem}
@@
 \begin{align} 
-    \tau_e \frac{dg_e}{dt} &= -g_e + \sum_{j \in \boldsymbol E} U x_j w \delta(t - \hat{t_j}) \nonumber \\
-    \tau_i \frac{dg_i}{dt} &= -g_i + \sum_{j \in \boldsymbol I} w \delta(t - \hat{t_j}) \label{eq:gsyn}
+    \tau_e \frac{dg_e}{dt} &= -g_e + \sum_{j \in \boldsymbol E} U x_j w \delta(t - \hat{t}_j) \nonumber \\
+    \tau_i \frac{dg_i}{dt} &= -g_i + \sum_{j \in \boldsymbol I} w \delta(t - \hat{t}_j) \label{eq:gsyn}
 \end{align}
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig10}
+    \includegraphics[width=\linewidth]{fig9}
     \caption{%
@@
     }
-    \label{fig:10}
+    \label{fig:9}
 \end{figure}
@@
 To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.
-\FIG{10}A-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
-Responses, normalized by the response to stimulation in the recovered networks (\FIG{10}A), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
-For thresholds, $\T[std]{A}{}$ (\FIG{10}B) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
-A similar pattern emerged with $\D{A}{}$ (\FIG{10}C), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+\FIG{9}A-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (\FIG{9}A), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, $\T[std]{A}{}$ (\FIG{9}B) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with $\D{A}{}$ (\FIG{9}C), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
 
 To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
-We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{10}D.
+We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{9}D.
 In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig9}
+    \includegraphics[width=\linewidth]{fig-conclusion}
     \caption{%
@@
     }
-    \label{fig:9}
+    \label{fig:conclusion}
 \end{figure}
@@
 In this study, we showed that non-synaptic neuronal adaptation can induce true deviance detection in a randomly wired, but spatially organized network of simple leaky integrate-and-fire neurons.
-Neuronal adaptation achieves this, as illustrated in Figure 10a, by suppressing the response of neurons local to the frequently stimulated site.
+Neuronal adaptation achieves this, as illustrated in \FIG{conclusion}A, by suppressing the response of neurons local to the frequently stimulated site.
 This drastically reduces the response to frequent stimulations at an early stage of signal propagation through the network, allowing the more distant parts of the network to recover from adaptation and respond vigorously to infrequent deviant stimuli.
@@
 Our second main finding is that synaptic short-term depression, which acts in similar fashion to neuronal adaptation both in terms of how it is induced and in how it affects activity flow in the network, enhances the DD effect of neuronal adaptation, and vice versa.
-We demonstrated that this synergy is a result of increased local suppression of the frequently presented stimulus by STD, which decreases global suppression by both mechanisms in oddball sequences and enhances the deviant response, as illustrated in Figure 10b.
+We demonstrated that this synergy is a result of increased local suppression of the frequently presented stimulus by STD, which decreases global suppression by both mechanisms in oddball sequences and enhances the deviant response, as illustrated in \FIG{conclusion}B.
 Conversely, without strong local suppression in control sequences, global suppression is increased, leading to smaller individual responses.
@@
 
+\section{Acknowledgments}
+
+This work was supported by World Premier International Research Center Initiative (WPI), MEXT, Japan.
+
 \bibliography{refs}

### 2023-03-20 paper-1/manuscript.tex
Change R/T/D from \mathbf to oblique \boldsymbol
Added WPI funding ack
Fixed hat t_j appearance
Aligned fig file naming & fix references in tex

@@
 % Common notation
-\renewcommand{\R}[3][]{{}^{#1}_{}\mathbf{R}^{#2}_{#3}}
-\renewcommand{\T}[3][]{{}^{#1}_{}\mathbf{T}^{#2}_{#3}}
-\newcommand{\D}[3][]{{}^{#1}_{}\mathbf{D}^{#2}_{#3}}
+\renewcommand{\R}[3][]{{}^{#1}_{}\boldsymbol R^{#2}_{#3}}
+\renewcommand{\T}[3][]{{}^{#1}_{}\boldsymbol T^{#2}_{#3}}
+\newcommand{\D}[3][]{{}^{#1}_{}\boldsymbol D^{#2}_{#3}}
 \newcommand{\mean}[1]{\overline{#1}}
-\newcommand{\Rhat}[3][]{{}^{#1}_{}\mathbf{\widehat{R}}^{#2}_{#3}}
+\newcommand{\Rhat}[3][]{{}^{#1}_{}\widehat{\boldsymbol R}^{#2}_{#3}}
 
@@
     \centering
-    \includegraphics{fig10.pdf}
+    \includegraphics{fig9.pdf}
     \caption{}
-    \label{fig:10}
+    \label{fig:9}
 \end{figure*}
@@
 To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.
-\FIG{10}a-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
-Responses, normalized by the response to stimulation in the recovered networks (\FIG{10}a), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
-For thresholds, $\T[std]{A}{}$ (\FIG{10}b) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
-A similar pattern emerged with $\D{A}{}$ (\FIG{10}c), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+\FIG{9}a-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (\FIG{9}a), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, $\T[std]{A}{}$ (\FIG{9}b) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with $\D{A}{}$ (\FIG{9}c), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
 
 To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
-We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{10}d.
+We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{9}d.
 In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.
@@
     \centering
-    \includegraphics{fig9.pdf}
+    \includegraphics{fig-conclusion.pdf}
     \caption{}
-    \label{fig:9}
+    \label{fig:conclusion}
 \end{figure*}
@@
 In this study, we showed that non-synaptic neuronal adaptation can induce true deviance detection in a randomly wired, but spatially organized network of simple leaky integrate-and-fire neurons.
-Neuronal adaptation achieves this, as illustrated in Figure 10a, by suppressing the response of neurons local to the frequently stimulated site.
+Neuronal adaptation achieves this, as illustrated in \FIG{conclusion}a, by suppressing the response of neurons local to the frequently stimulated site.
 This drastically reduces the response to frequent stimulations at an early stage of signal propagation through the network, allowing the more distant parts of the network to recover from adaptation and respond vigorously to infrequent deviant stimuli.
@@
 Our second main finding is that synaptic short-term depression, which acts in similar fashion to neuronal adaptation both in terms of how it is induced and in how it affects activity flow in the network, enhances the DD effect of neuronal adaptation, and vice versa.
-We demonstrated that this synergy is a result of increased local suppression of the frequently presented stimulus by STD, which decreases global suppression by both mechanisms in oddball sequences and enhances the deviant response, as illustrated in Figure 10b.
+We demonstrated that this synergy is a result of increased local suppression of the frequently presented stimulus by STD, which decreases global suppression by both mechanisms in oddball sequences and enhances the deviant response, as illustrated in \FIG{conclusion}b.
 Conversely, without strong local suppression in control sequences, global suppression is increased, leading to smaller individual responses.
@@
 \bmhead{Funding}
-This study was funded by a World Premier International Research Center Initiative research startup grant.
+This work was supported by World Premier International Research Center Initiative (WPI), MEXT, Japan.
 

### 2023-03-20 paper-1/elife.tex
Better fig name: 9 -> space

@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig9}
+    \includegraphics[width=\linewidth]{fig-space}
     \caption{%
@@
     }
-    \label{fig:9}
+    \label{fig:space}
 \end{figure}
@@
 To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.
-\FIG{9}A-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
-Responses, normalized by the response to stimulation in the recovered networks (\FIG{9}A), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
-For thresholds, $\T[std]{A}{}$ (\FIG{9}B) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
-A similar pattern emerged with $\D{A}{}$ (\FIG{9}C), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+\FIG{space}A-C shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (\FIG{space}A), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, $\T[std]{A}{}$ (\FIG{space}B) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with $\D{A}{}$ (\FIG{space}C), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
 
 To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
-We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{9}D.
+We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{space}D.
 In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 9 -> space

@@
     \centering
-    \includegraphics{fig9.pdf}
+    \includegraphics{fig-space.pdf}
     \caption{}
-    \label{fig:9}
+    \label{fig:space}
 \end{figure*}
@@
 To do so, we assessed the distance between the center of the target stimulation site and each neuron in all networks and stimulus combinations, and calculated $\R{A}{}$, $\T{A}{}$ and $\D{A}{}$ as a function of this distance.
-\FIG{9}a-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
-Responses, normalized by the response to stimulation in the recovered networks (\FIG{9}a), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
-For thresholds, $\T[std]{A}{}$ (\FIG{9}b) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
-A similar pattern emerged with $\D{A}{}$ (\FIG{9}c), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
+\FIG{space}a-c shows that there were indeed clear spatially defined differences between \dev{}, \msc{} and \std{}.
+Responses, normalized by the response to stimulation in the recovered networks (\FIG{space}a), clearly spread out farthest in \dev{}, and least in \std{}, despite a nearly indistinguishable response at the immediate stimulation site.
+For thresholds, $\T[std]{A}{}$ (\FIG{space}b) was much stronger near the stimulation site, and somewhat weaker distally, than either of the less frequent paradigms.
+A similar pattern emerged with $\D{A}{}$ (\FIG{space}c), though here, the local patch covered a larger area, as we might expect from a synaptic effect.
 
 To quantify this statistically, we split the neurons into local and distal (global) subsets, cutting at a radius of 2 mm from the stimulus site, yielding an average local subset size of 226.8 $\pm$ 13.2 neurons, out of 1000.
-We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{9}d.
+We then assessed the difference of the mean values of $\T{A}{}$ and $\D{A}{}$ across the two subsets in each sequence, as shown in \FIG{space}d.
 In keeping with the previously stated analysis, we found that local suppression was higher in \std{} than in \dev{} both for $\T{A}{}$ (T = 482, p = 8.31e-17; one-sided Wilcoxon, n = 120) and for $\D{A}{}$ (T = 735, p = 1.71e-14), indicating that the network response to the non-target stimulus in the \dev{} sequence was indeed kept in check by high local suppression.

### 2023-03-20 paper-1/elife.tex
Better fig name: 8 -> synergy

@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig8}
+    \includegraphics[width=\linewidth]{fig-synergy}
     \caption{%
@@
     \figsupp[The equivalent figure in the ablated model with only STD.]{%
-        The equivalent of \FIG{8} in the ablated model with only STD.
+        The equivalent of \FIG{synergy} in the ablated model with only STD.
         \textbf{A} Average depression $\D{}{i}$ in target (A) trials (top row) and B trials (bottom row) in the \dev{} (left) and \msc{} sequence (center), and their contrast (right) in the sample network.
@@
     }
-    {\includegraphics[width=\linewidth]{fig8-noTA}}\label{figsupp:noTA}
+    {\includegraphics[width=\linewidth]{fig-synergy-noTA}}\label{figsupp:noTA}
     %
@@
 To show this, we calculated $\D{}{}$ for A and B trials and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
-In \FIG{8}A, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
+In \FIG{synergy}A, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
 The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast $\Delta \D{A}{i}$, which was negative throughout.
@@
 Notice that this mirrors the correlation shown in \FIG{4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
-As shown in \FIG{8}B, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
-We confirmed this relationship across networks and stimuli, see \FIG{8}C, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
+As shown in \FIG{synergy}B, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
+We confirmed this relationship across networks and stimuli, see \FIG{synergy}C, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
 We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \FIGSUPP[8]{noTA}).
@@
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in \FIG{8}D, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
+First, in \FIG{synergy}D, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
@@
 To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{6}E.
-As we show in \FIG{8}E, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
+As we show in \FIG{synergy}E, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
 This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
@@
 If so, we should expect to see less $\D{A}{}$ in \dev{} than \msc{}.
-To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \FIG{8}C.
+To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \FIG{synergy}C.
 We conclude that there is clear evidence for greater STD-mediated global suppression in \msc{} than \dev{} in the full model.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}E, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{8}E), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}E, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{synergy}E), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 8 -> synergy

@@
     \centering
-    \includegraphics{fig8.pdf}
+    \includegraphics{fig-synergy.pdf}
     \caption{}
-    \label{fig:8}
+    \label{fig:synergy}
 \end{figure*}
@@
 To show this, we calculated $\D{}{}$ for A and B trials and contrasted it across \dev{} and \msc{} sequences as with $\T{}{}$ before.
-In \FIG{8}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
+In \FIG{synergy}a, we see that target trials (top) experienced depression across large parts of the sample network in both sequences.
 The \msc{} sequence, however, clearly evoked greater depression, as borne out in the contrast $\Delta \D{A}{i}$, which was negative throughout.
@@
 Notice that this mirrors the correlation shown in \FIG{4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
-As shown in \FIG{8}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
-We confirmed this relationship across networks and stimuli, see \FIG{8}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
-We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \ref{sfig8}).
+As shown in \FIG{synergy}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
+We confirmed this relationship across networks and stimuli, see \FIG{synergy}c, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
+We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \ref{sfig-synergy}).
 
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in \FIG{8}d, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
+First, in \FIG{synergy}d, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
@@
 To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{6}e.
-As we show in \FIG{8}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
+As we show in \FIG{synergy}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
 This confirms that STD-mediated local suppression played a role in reducing the non-target response in the \dev{} sequence.
@@
 If so, we should expect to see less $\D{A}{}$ in \dev{} than \msc{}.
-To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \FIG{8}c.
+To confirm this, we analyzed $\Delta \D{A}{}$ across networks and stimuli and found it to be robustly negative (T = 757, p = 2.66e-14; one-sided Wilcoxon, n = 120), as can be seen in the inset histogram in \FIG{synergy}c.
 We conclude that there is clear evidence for greater STD-mediated global suppression in \msc{} than \dev{} in the full model.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{8}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{synergy}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.
@@
     \centering
-    \includegraphics{fig8-noTA.pdf}
+    \includegraphics{fig-synergy-noTA.pdf}
     \caption{}
-    \label{sfig8}
+    \label{sfig-synergy}
 \end{figure*}

### 2023-03-20 paper-1/elife.tex
Better fig name: 7 -> addingSTD

@@
 To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD.
-On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\FIG{7}A).
+On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\FIG{addingSTD}A).
 However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (more than \msc{}: T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (less than \msc{}: T = 2.66e+03, p = 0.00554).
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig7}
+    \includegraphics[width=\linewidth]{fig-addingSTD}
     \caption{%
@@
         \textbf{B} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
-Compare to \FIG{6}C-d.
+Compare to \FIG{6}C-D.
         \textbf{C} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
@@
     }
-    \label{fig:7}
+    \label{fig:addingSTD}
 \end{figure}
@@
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
-Starting at the cause of the reduced response in non-target trials, we show in \FIG{7}B the full model equivalent of \FIG{6}C-d.
+Starting at the cause of the reduced response in non-target trials, we show in \FIG{addingSTD}B the full model equivalent of \FIG{6}C-D.
 The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
@@
 
-Comparing the non-target response reductions in \FIG{7}C, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
+Comparing the non-target response reductions in \FIG{addingSTD}C, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
 Based on the ablated model analysis in \FIG{5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{7}D.
-Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \FIG{7}E, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{addingSTD}D.
+Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \FIG{addingSTD}E, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
 Finally, based on the ablated model analysis in \FIG{4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
-As shown in \FIG{7}F, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{7}G, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
+As shown in \FIG{addingSTD}F, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{addingSTD}G, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
 This greater contrast, of course, directly reflects the greater DDI reported in \FIG{3}.
@@
 Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
-The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\FIG{7}C), which was where we noted the first clear shift towards increased deviance detection.
+The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\FIG{addingSTD}C), which was where we noted the first clear shift towards increased deviance detection.
 
@@
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
-We can relate this directly to the non-target response contrasts in the early and late portions shown in \FIG{7}A, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
+We can relate this directly to the non-target response contrasts in the early and late portions shown in \FIG{addingSTD}A, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
 % Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
@@
 To sum up, in the full model, STD acted in much the same way as TA, causing both a reduction in \dev{} of the response to the frequently presented non-target stimuli (lower $\R[dev]{B}{}$) by local suppression, and an increase in the response to target stimuli (higher $\R[dev]{A}{}$) by relief from global suppression, relative to \msc{}.
-Yet, as we showed in \FIG{7} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower thresholds $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
+Yet, as we showed in \FIG{addingSTD} and in the previous section, a decrease in $\R[dev]{B}{}$ also entailed lower thresholds $\T[dev]{A}{}$, which independently led to higher target responses $\R[dev]{A}{}$.
 Thus, adding STD to the model increased the DDI both directly, through its intrinsic effects in local and global suppression, and indirectly, by altering the pattern of TA by way of local suppression of B.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}E, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{synergy}E), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{addingSTD}E, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{synergy}E), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 7 -> addingSTD

@@
 To understand the basis of this apparent synergy, we contrasted the TA only condition examined in the previous section against the same networks evaluated in the full model with both TA and STD.
-On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\FIG{7}a).
+On average, adding STD caused a reduction in response magnitude across the board, as we would expect from a depression mechanism (\FIG{addingSTD}a).
 However, this reduction was not uniform: Target responses $\mean{\R{A}{}}$ were reduced most in the \std{} sequence (more than \msc{}: T = 2.99e+03, p = 0.0474; one-sided Wilcoxon, n = 120), and least in the \dev{} sequence (less than \msc{}: T = 2.66e+03, p = 0.00554).
@@
     \centering
-    \includegraphics{fig7.pdf}
+    \includegraphics{fig-addingSTD.pdf}
     \caption{}
-    \label{fig:7}
+    \label{fig:addingSTD}
 \end{figure*}
@@
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
-Starting at the cause of the reduced response in non-target trials, we show in \FIG{7}b the full model equivalent of \FIG{6}c-d.
+Starting at the cause of the reduced response in non-target trials, we show in \FIG{addingSTD}b the full model equivalent of \FIG{6}c-d.
 The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
@@
 
-Comparing the non-target response reductions in \FIG{7}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
+Comparing the non-target response reductions in \FIG{addingSTD}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
 Based on the ablated model analysis in \FIG{5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
-Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{7}d.
-Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \FIG{7}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
+Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{addingSTD}d.
+Correspondingly, $\T[dev]{A}{exc}$ was reduced over $\T[msc]{A}{exc}$, as shown by way of their contrast in \FIG{addingSTD}e, and this too by a slightly greater amount in the full model than in the ablated model (T = 4.37e+03, p = 0.0257).
 
 Finally, based on the ablated model analysis in \FIG{4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
-As shown in \FIG{7}f, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
-Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{7}g, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
+As shown in \FIG{addingSTD}f, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
+Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{addingSTD}g, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
 This greater contrast, of course, directly reflects the greater DDI reported in \FIG{3}.
@@
 Taken together, the data indicate that, with STD added to the model, the overall chain of causality -- increased local suppression to B resulting in reduced B responses and reduced global suppression, which in turn allows greater target responses -- remained intact with only minor quantitative changes.
-The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\FIG{7}c), which was where we noted the first clear shift towards increased deviance detection.
+The primary cause for the greater target response, therefore, appeared to be the smaller non-target response (\FIG{addingSTD}c), which was where we noted the first clear shift towards increased deviance detection.
 
@@
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
-We can relate this directly to the non-target response contrasts in the early and late portions shown in \FIG{7}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
+We can relate this directly to the non-target response contrasts in the early and late portions shown in \FIG{addingSTD}a, which followed a similar pattern, and take this to indicate that the increased presentation of B in \dev{} causes local synaptic suppression.
 % Possible concern: The relationship dXnotA ~ dD^B is not clearly established.
@@
 We note that the direct and indirect effects of STD on global suppression appear to be on the same order of magnitude.
-To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{7}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{synergy}e), both of which are on the order of $\approx$ 0.1 mV.
+To see this, compare the change in $\Delta \T{A}{}$ as a result of adding STD to the model (\FIG{addingSTD}e, bottom) with the corresponding value of $\Delta \D{A}{}$ (\FIG{synergy}e), both of which are on the order of $\approx$ 0.1 mV.
 Conversely, notice that, since TA and STD work in similar fashion to increase the \dev{} target response, it is very plausible that TA likewise increases the effect of STD, by enhancing the STD-mediated local suppression of B and thereby increasing the contrast in STD-mediated global suppression.

### 2023-03-20 paper-1/elife.tex
Better fig name: 6 -> early adaptation

@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig6}
+    \includegraphics[width=\linewidth]{fig-early-adaptation}
     \caption{%
@@
     }
-    \label{fig:6}
+    \label{fig:early-adaptation}
     % 
     \figsupp[The equivalent figure in the full model with both TA and STD.]{%
-        The equivalent of \FIG{6} in the full model with both TA and STD.
+        The equivalent of \FIG{early-adaptation} in the full model with both TA and STD.
         \textbf{A} The sample network, colored by latency rank, i.e.
@@
     }
-    {\includegraphics[width=\linewidth]{fig6-withSTD}}\label{figsupp:withSTD}
+    {\includegraphics[width=\linewidth]{fig-early-adaptation-withSTD}}\label{figsupp:withSTD}
 \end{figure}
@@
 In a given network and stimulus assignment, we ranked each neuron by the minimum response time (i.e., time to first spike) across B trials in all sequences.
-The ranking in the sample network is shown in \FIG{6}A.
+The ranking in the sample network is shown in \FIG{early-adaptation}A.
 We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site.
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
-In \FIG{6}B, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
+In \FIG{early-adaptation}B, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
 Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{5}C, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
@@
 We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992: T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
-In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{6}C) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{6}D) within these two portions.
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{early-adaptation}C) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{early-adaptation}D) within these two portions.
 We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{5}D).
@@
 
-Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \FIG{6}E.
+Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \FIG{early-adaptation}E.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
@@
         \textbf{B} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
-Compare to \FIG{6}C-D.
+Compare to \FIG{early-adaptation}C-D.
         \textbf{C} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
@@
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
-Starting at the cause of the reduced response in non-target trials, we show in \FIG{addingSTD}B the full model equivalent of \FIG{6}C-D.
+Starting at the cause of the reduced response in non-target trials, we show in \FIG{addingSTD}B the full model equivalent of \FIG{early-adaptation}C-D.
 The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
@@
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in \FIG{synergy}D, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
+First, in \FIG{synergy}D, we recapitulate the logic established in \FIG{early-adaptation}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
@@
 
-To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{6}E.
+To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{early-adaptation}E.
 As we show in \FIG{synergy}E, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 6 -> early adaptation

@@
     \centering
-    \includegraphics{fig6.pdf}
+    \includegraphics{fig-early-adaptation.pdf}
     \caption{}
-    \label{fig:6}
+    \label{fig:early-adaptation}
 \end{figure*}
@@
 In a given network and stimulus assignment, we ranked each neuron by the minimum response time (i.e., time to first spike) across B trials in all sequences.
-The ranking in the sample network is shown in \FIG{6}a.
+The ranking in the sample network is shown in \FIG{early-adaptation}a.
 We found that the response developed in an approximately radial pattern, spreading outward in all directions from the stimulus site.
@@
 We then notionally split the network into an early portion, containing the first 100 neurons to respond to B, and a late portion, containing the remaining 900 neurons.
-In \FIG{6}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
+In \FIG{early-adaptation}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
 Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
@@
 We first confirmed that $\Delta \R{N}{i}$ and $\Delta \T{B}{i}$ were robustly positively correlated (median Pearson's $\rho$ 0.992: T = 7.26e+03, p = 9.86e-22; one-sided Wilcoxon, n = 120).
-In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{6}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{6}d) within these two portions.
+In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{early-adaptation}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{early-adaptation}d) within these two portions.
 We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{5}d).
@@
 
-Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \FIG{6}e.
+Finally, to show that high $\T[dev]{B}{early}$ was causal in reducing the response to B, we correlated the median contrast in $\T{B}{}$ among the early portion, $\Delta \T{B}{early}$, with the network-wide median contrast of the response to B, $\Delta \R{B}{}$, as shown in \FIG{early-adaptation}e.
 Across networks and stimuli, we found a strong negative correlation ($\rho$ = -0.587, p = 9.56e-13, Wald test, n = 120), indicating that greater TA in the neurons responding early was indeed predictive of a lowered response to B in the overall network.
@@
 \textbf{a} Reductions in mean response size $\mean{\R{A}{}}$ in each sequence as a consequence of adding STD to the model, across networks and stimuli.
-\textbf{b} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA. Compare to \FIG{6}c-d.
+\textbf{b} $\Delta \R{N}{}$ (top) and $\Delta \T{B}{}$ (bottom) among early and late neurons, respectively, across networks and stimuli, in the full model with both STD and TA.
+Compare to \FIG{early-adaptation}c-d.
 \textbf{c} $\Delta \R{N}{exc}$ across networks and stimuli in the full model (horizontal axis) and the ablated model with TA only (vertical axis).
@@
 To understand how STD affects deviance detection and causes this difference in response reduction, we first examined the effects of TA in the full model, following the analysis on the ablated model laid out in the previous section.
-Starting at the cause of the reduced response in non-target trials, we show in \FIG{addingSTD}b the full model equivalent of \FIG{6}c-d.
+Starting at the cause of the reduced response in non-target trials, we show in \FIG{addingSTD}b the full model equivalent of \FIG{early-adaptation}c-d.
 The contrasts (\dev{} - \msc{}) in the non-target response magnitude, $\Delta \R{N}{}$, and in the resulting $\T{B}{}$ levels both retained a clear separation between the portion of neurons responding early, which showed increased responses (T = 5.44e+03, p = 1.02e-06; one-sided Wilcoxon, n = 120) and increased TA levels (T = 3.41e+03, p = 9.51e-08), and the neurons responding later, whose response (T = 282, p = 9.19e-19) and TA levels (T = 72, p = 2.49e-17) decreased in the \dev{} sequence.
@@
 Having gained a rough intuition of how individual neurons are affected by STD, we now turn to its effect on the network, working with statistics across networks and stimuli.
-First, in \FIG{synergy}d, we recapitulate the logic established in \FIG{6}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
+First, in \FIG{synergy}d, we recapitulate the logic established in \FIG{early-adaptation}: Given the spatial clustering of $\D{B}{i}$, we surmised a similar local suppression effect as with TA.
 We found that, indeed, neurons responding early to stimulation in B showed greater depression in \dev{} than \msc{} ($\Delta \D{B}{early} > 0$: T = 7.26e+03, p = 1.01e-21; one-sided Wilcoxon, n = 120), while the late portion had slightly reduced depression ($\Delta \D{B}{late} < 0$: T = 2.24e+03, p = 0.000133).
@@
 
-To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{6}e.
+To show that increased local suppression by STD reduces the non-target response, we correlated $\Delta \D{B}{early}$ with $\Delta \R{B}{}$, echoing \FIG{early-adaptation}e.
 As we show in \FIG{synergy}e, higher $\Delta \D{B}{early}$ was clearly predictive of a lower network response ($\rho$ = -0.581, p = 1.81e-12; Wald test, n = 120).
@@
     \centering
-    \includegraphics{fig6-withSTD.pdf}
+    \includegraphics{fig-early-adaptation-withSTD.pdf}
     \caption{}

### 2023-03-20 paper-1/elife.tex
Better fig name: 5 -> thresholds follow std

@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig5}
+    \includegraphics[width=\linewidth]{fig-thresholds-follow-std}
     \caption{%
@@
     }
-    \label{fig:5}
+    \label{fig:thresholds-follow-std}
     % 
     \figsupp[The equivalent figure in the full model with both TA and STD.]{%
-        The equivalent of \FIG{5} in the full model with both TA and STD.
+        The equivalent of \FIG{thresholds-follow-std} in the full model with both TA and STD.
         \textbf{A} $\T{A}{i}$ in the sample network, with each neuron shown in its spatial location.
@@
     }
-    {\includegraphics[width=\linewidth]{fig5-withSTD}}\label{figsupp:withSTD}
+    {\includegraphics[width=\linewidth]{fig-thresholds-follow-std-withSTD}}\label{figsupp:withSTD}
 \end{figure}
 
-First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \FIG{5}A.
+First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \FIG{thresholds-follow-std}A.
 In the \msc{} sequence, $\T[msc]{A}{i}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
@@
 
-Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\FIG{5}B), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
+Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\FIG{thresholds-follow-std}B), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
 Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.
 
-Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\FIG{5}C), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in \FIG{5}D, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
+Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\FIG{thresholds-follow-std}C), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in \FIG{thresholds-follow-std}D, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
 
@@
 In \FIG{early-adaptation}B, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
-Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{5}C, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{thresholds-follow-std}C, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
 As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red, left plot) increasing their response in the \dev{} sequence, while the response of the later portion of the network (blue, right plot) decreased.
@@
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{early-adaptation}C) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{early-adaptation}D) within these two portions.
-We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{5}D).
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{thresholds-follow-std}D).
 $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
@@
         \textbf{D} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
-Compare to \FIG{5}D.
+Compare to \FIG{thresholds-follow-std}D.
         \textbf{E} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel C.
@@
 Comparing the non-target response reductions in \FIG{addingSTD}C, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
-Based on the ablated model analysis in \FIG{5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
+Based on the ablated model analysis in \FIG{thresholds-follow-std}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
 Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{addingSTD}D.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 5 -> thresholds follow std

@@
     \centering
-    \includegraphics{fig5.pdf}
+    \includegraphics{fig-thresholds-follow-std.pdf}
     \caption{}
-    \label{fig:5}
+    \label{fig:thresholds-follow-std}
 \end{figure*}
@@
 
-First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \FIG{5}a.
+First, we mapped the average value of $\theta$ at the start of target trials, $\T{A}{i}$, to the spatial location of the neurons, as shown in \FIG{thresholds-follow-std}a.
 In the \msc{} sequence, $\T[msc]{A}{i}$ was roughly evenly distributed across the network, consistent with the random, distributed nature of stimulation.
@@
 
-Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\FIG{5}b), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
+Visualizing the likely cause of this difference, the non-target responses, in the same fashion (\FIG{thresholds-follow-std}b), we see that the pattern of activity in non-target trials, constituting 80\% of all trials, closely matched that of the thresholds, as expected.
 Notably, although the comparison is between 400 B trials in \dev{}, and only 100 B trials (alongside the same number of C, D and E trials) in \msc{}, only a small handful of neurons responded more to non-target stimulation in \dev{}, most of these very close to the stimulation site of B.
 
-Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\FIG{5}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
-This finding held across networks and stimuli, as shown in \FIG{5}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
+Correlating $\Delta \R{N}{i}$ with $\Delta \T{A}{i}$ across excitatory neurons (\FIG{thresholds-follow-std}c), we found that the reduced non-target response magnitude in \dev{} almost perfectly predicted the resulting reduction in TA levels before target stimuli (Pearson's $\rho$ = 0.99).
+This finding held across networks and stimuli, as shown in \FIG{thresholds-follow-std}d, with a median correlation coefficient of 0.99 and a consistent reduction in the median non-target response $\R{N}{}$ between \msc{} and \dev{} sequences (T = 525, p = 2.12e-16, median = -0.094; one-sided Wilcoxon, n = 120), clearly validating our notion that the reduced non-target trial response was the primary driver of lower TA, which we showed above to be responsible for the higher target response.
 
@@
 In \FIG{early-adaptation}b, we related the non-target response contrast, $\Delta \R{N}{i}$, to the contrast in $\theta$ at the start of B trials, $\Delta \T{B}{i}$.
-Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{5}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
+Across excitatory neurons, a strong linear relationship (Pearson's $\rho$ = 0.993, n = 800) echoed the related finding in \FIG{thresholds-follow-std}c, where we correlated $\Delta \R{N}{i}$ to $\Delta \T{A}{i}$.
 As the color coding and inset histograms show, however, there was a clear separation between the early and late portions of the network, with many early neurons (red, left plot) increasing their response in the \dev{} sequence, while the response of the later portion of the network (blue, right plot) decreased.
@@
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{early-adaptation}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{early-adaptation}d) within these two portions.
-We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{5}d).
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{thresholds-follow-std}d).
 $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
@@
 \textbf{d} $\Delta \R{N}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \R{N}{i}$ and $\Delta \T{A}{i}$ across excitatory neurons.
-Compare to \FIG{5}d.
+Compare to \FIG{thresholds-follow-std}d.
 \textbf{e} $\Delta \T{A}{exc}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
@@
 Comparing the non-target response reductions in \FIG{addingSTD}c, we found that they were on average larger in the full model than in the ablated model (T = 4.16e+03, p = 0.0175; one-sided Wilcoxon, n = 120), with the grand medians of $\Delta \R{N}{exc}$ differing by 36\%.
-Based on the ablated model analysis in \FIG{5}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
+Based on the ablated model analysis in \FIG{thresholds-follow-std}, we expected reduced non-target responses to similarly reduce the TA levels at the start of A trials.
 Indeed, $\Delta \R{N}{i}$ remained tightly correlated with $\Delta \T{A}{i}$ (median Pearson's $\rho$ = 0.989), as shown in \FIG{addingSTD}d.
@@
     \centering
-    \includegraphics{fig5-withSTD.pdf}
+    \includegraphics{fig-thresholds-follow-std-withSTD.pdf}
     \caption{}

### 2023-03-20 paper-1/elife.tex
Better fig name: 4 -> threshold initial

@@
 To understand why the same stimulus A evoked a higher response in \dev{} than in \msc{} with TA only, we first compared the average responses in the two sequences, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
-As shown in \FIG{4}A (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
+As shown in \FIG{threshold-initial}A (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
 The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, which indicate higher \dev{} activity.
@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig4}
+    \includegraphics[width=\linewidth]{fig-threshold-initial}
     \caption{%
@@
     }
-    \label{fig:4}
+    \label{fig:threshold-initial}
     % 
     \figsupp[The equivalent figure in the full model with both TA and STD.]{%
-        The equivalent of \FIG{4} in the full model with both TA and STD.
+        The equivalent of \FIG{threshold-initial} in the full model with both TA and STD.
         \textbf{A} Post-stimulus spike histogram across target trials in the sample network, showing trial time along the horizontal axis, and neurons along the vertical, sorted by the time of the first recorded spike across all trials and sequences.
@@
     }
-    {\includegraphics[width=\linewidth]{fig4-withSTD}}\label{figsupp:withSTD}
+    {\includegraphics[width=\linewidth]{fig-threshold-initial-withSTD}}\label{figsupp:withSTD}
 \end{figure}
@@
 We quantified the former with the strength of adaptation $\theta$ (\EQ{TA}), i.e., the amount by which the firing threshold was raised due to TA.
-As shown in \FIG{4}B, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+As shown in \FIG{threshold-initial}B, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
@@
 To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we refer to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
-As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\FIG{4}C) largely reflected that seen in spiking activity.
+As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\FIG{threshold-initial}C) largely reflected that seen in spiking activity.
 The contrast (right panel) shows not only a notable relative increase in membrane voltages in \dev{} around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
@@
 We first focused on the contribution of $\theta$ alone to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \FIG{4}B, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
-In the sample network (\FIG{4}D), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in \FIG{4}E, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \FIG{threshold-initial}B, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
+In the sample network (\FIG{threshold-initial}D), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+As shown in \FIG{threshold-initial}E, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
 In addition, the correlation across excitatory neurons between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
@@
 
-Yet, as we saw in \FIG{4}C, the membrane voltages also differed between the two sequences.
+Yet, as we saw in \FIG{threshold-initial}C, the membrane voltages also differed between the two sequences.
 In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
 We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference and summed across neurons.
-The resulting time courses, shown in \FIG{4}F, show how much these two measures contributed to the increased \dev{} response.
+The resulting time courses, shown in \FIG{threshold-initial}F, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (\FIG{4}G), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
-A similar pattern emerged across networks and stimuli, as shown in \FIG{4}H.
+In relative terms (\FIG{threshold-initial}G), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+A similar pattern emerged across networks and stimuli, as shown in \FIG{threshold-initial}H.
 Here, only the initial response was driven mainly by $\theta$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
@@
 Secondly, the resulting  frequent presentation of B changes TA values, likely causing adaptation to B and thereby decreasing the average response.
-We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B (and therefore more to non-target stimuli in \dev{} than \msc{}), and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in \FIG{4}.
+We hypothesized that this adaptation is mediated by a small subset of neurons that (1) responds very quickly and rather strongly to stimulation in B (and therefore more to non-target stimuli in \dev{} than \msc{}), and (2) has increased TA levels as a result, reducing the response of the remaining downstream neurons in the same way as demonstrated in reverse in \FIG{threshold-initial}.
 
@@
 
-To sum up, we showed in \FIG{4} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
+To sum up, we showed in \FIG{threshold-initial} that lower $\T[dev]{A}{}$ in the portion of the network responding earliest to A played an important role in increasing the overall response to A.
 Here, we showed that the neurons responding earliest to B exhibited higher $\T[dev]{B}{}$, which analogously reduced the response of the network as a whole to B.
@@
         \textbf{F} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
-Compare to \FIG{4}E.
+Compare to \FIG{threshold-initial}E.
         \textbf{G} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel C.
@@
 
-Finally, based on the ablated model analysis in \FIG{4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
+Finally, based on the ablated model analysis in \FIG{threshold-initial}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
 As shown in \FIG{addingSTD}F, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
@@
 To show that $\D{}{}$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated $\Delta \D{A}{i}$ with $\Delta \R{A}{i}$ across neurons of the sample network.
-Notice that this mirrors the correlation shown in \FIG{4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
+Notice that this mirrors the correlation shown in \FIG{threshold-initial} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
 As shown in \FIG{synergy}B, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 4 -> threshold initial

@@
 To understand why the same stimulus A evoked a higher response in \dev{} than in \msc{} with TA only, we first compared the average responses in the two sequences, sorting neurons by their response onset to permit a high-resolution view of the stimulus response as it propagates through the network.
-As shown in \FIG{4}a (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
+As shown in \FIG{threshold-initial}a (left and middle), we found that the response developed in a qualitatively similar fashion in both sequences.
 The contrast (right) revealed a clear increase in \dev{} spikes, particularly in the later parts of the response, as shown by the predominance of orange areas, which indicate higher \dev{} activity.
@@
     \centering
-    \includegraphics{fig4.pdf}
+    \includegraphics{fig-threshold-initial.pdf}
     \caption{}
-    \label{fig:4}
+    \label{fig:threshold-initial}
 \end{figure*}
@@
 We quantified the former with the strength of adaptation $\theta$ (\EQ{TA}), i.e., the amount by which the firing threshold was raised due to TA.
-As shown in \FIG{4}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
+As shown in \FIG{threshold-initial}b, there was a clear and widespread lowering of thresholds at the start of target \dev{} trials relative to \msc{}, visible both in the raw data (left and middle panels) and in the contrast (right panel, early blue portions).
 On the other hand, once the bulk of activity had occurred, many neurons showed higher thresholds in \dev{} (right panel, later orange portions), since they spiked more often.
@@
 To assess the difference in synaptic currents in a way that allowed direct comparison to $\theta$, we refer to the membrane voltage $V_m$, which is largely independent from and complementary to the firing threshold, and driven mostly by presynaptic inputs, resets after spikes notwithstanding.
-As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\FIG{4}c) largely reflected that seen in spiking activity.
+As the immediate cause of neuronal spiking, the overall pattern of $V_m$ (\FIG{threshold-initial}c) largely reflected that seen in spiking activity.
 The contrast (right panel) shows not only a notable relative increase in membrane voltages in \dev{} around the typical response time (note the orange band running along the response front), but also an increased hyperpolarization starting immediately after due to post-spike resets.
@@
 We first focused on the contribution of $\theta$ alone to confirm that thresholds are lower in \dev{}, and that this reduction in thresholds causes increased firing.
-We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \FIG{4}b, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
-In the sample network (\FIG{4}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
-As shown in \FIG{4}e, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
+We quantified the contrast in $\theta$ at the start of target trials as $\Delta \T{A}{i} = \T[dev]{A}{i} - \T[msc]{A}{i}$, where $\T{}{i} = \theta_i(t=0)$, corresponding to the leftmost datapoints of the contrast histogram in \FIG{threshold-initial}b, and correlated this with the target trial response size contrast $\Delta \R{A}{i}$ across excitatory neurons.
+In the sample network (\FIG{threshold-initial}d), we found that thresholds were clearly lower in \dev{} than \msc{} (T = 2.69e+04, p = 1.58e-92, median = -0.166 mV; one-sided Wilcoxon, n = 800 excitatory neurons), and that this decrease was accompanied by an increase in the target trial response at the level of individual neurons (Pearson's $\rho$ = -0.424, p = 1.6e-36; one-sided Wald test, n = 800).
+As shown in \FIG{threshold-initial}e, this relationship held across networks and stimuli: The median of $\T{A}{i}$ among excitatory neurons was lower in \dev{} than \msc{} ($\Delta \T{A}{exc} < 0$: T = 514, p = 1.67e-16, grand median = -0.094 mV; one-sided Wilcoxon, n = 120), indicating that networks were less adapted during oddball sequences.
 In addition, the correlation across excitatory neurons between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ was almost universally negative (T = 40, p = 2.68e-21, median = -0.357).
@@
 
-Yet, as we saw in \FIG{4}c, the membrane voltages also differed between the two sequences.
+Yet, as we saw in \FIG{threshold-initial}c, the membrane voltages also differed between the two sequences.
 In order to assess the relative contribution of $\theta$ to the increased \dev{} response, we chose to focus exclusively on time points with higher firing probability in the \dev{} sequence, discarding data points with equal or lower firing probability relative to the \msc{} sequence.
 We then weighted the remaining contrasts in $\theta$ and $V_m$ by the magnitude of the response difference and summed across neurons.
-The resulting time courses, shown in \FIG{4}f, show how much these two measures contributed to the increased \dev{} response.
+The resulting time courses, shown in \FIG{threshold-initial}f, show how much these two measures contributed to the increased \dev{} response.
 While thresholds were slightly lowered throughout the response, the membrane potential was dramatically increased.
-In relative terms (\FIG{4}g), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
-A similar pattern emerged across networks and stimuli, as shown in \FIG{4}h.
+In relative terms (\FIG{threshold-initial}g), $\theta$ constituted a major driver of increased firing right at the onset of the response, as well as briefly towards the end of the response.
+A similar pattern emerged across networks and stimuli, as shown in \FIG{threshold-initial}h.
 Here, only the initial response was driven mainly by $\theta$, while later differences were largely explained by differences in $V_m$, leading to a relative TA contribution of less than 20\% on average.
@@
 \textbf{f} $\Delta \T{A}{exc}$ across networks and stimuli in the full model, plotted against the corresponding Pearson coefficient of the correlation between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$ across excitatory neurons.
-Compare to \FIG{4}e.
+Compare to \FIG{threshold-initial}e.
 \textbf{g} $\Delta \R{A}{}$ across networks and stimuli in the full and ablated models, arranged as in panel c.
@@
 
-Finally, based on the ablated model analysis in \FIG{4}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
+Finally, based on the ablated model analysis in \FIG{threshold-initial}, we sought to confirm that lowered $\T{A}{}$ was responsible for the increased target response.
 As shown in \FIG{addingSTD}f, $\Delta \T{A}{exc}$ and $\Delta \R{A}{exc}$ were negatively correlated (T = 30, p = 2.09e-21, median = -0.268; one-sided Wilcoxon, n = 120), though the correlation was weaker in the full model (T = 1.67e+03, p = 1.49e-07).
@@
 To show that $\D{}{}$ captures not only changes in $V_m$, by definition, but also in the resulting response, we correlated $\Delta \D{A}{i}$ with $\Delta \R{A}{i}$ across neurons of the sample network.
-Notice that this mirrors the correlation shown in \FIG{4} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
+Notice that this mirrors the correlation shown in \FIG{threshold-initial} between $\Delta \T{A}{i}$ and $\Delta \R{A}{i}$.
 As shown in \FIG{synergy}b, the relationship was clearly negative (Pearson's $\rho$ = -0.541, p = 2.28e-77; Wald test, n = 1000), indicating that higher $\D{}{}$ causes reduced responses as expected.
@@
     \centering
-    \includegraphics{fig4-withSTD.pdf}
+    \includegraphics{fig-threshold-initial-withSTD.pdf}
     \caption{}

### 2023-03-20 paper-1/elife.tex
Better fig name: 3 -> ablation

@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig3}
+    \includegraphics[width=\linewidth]{fig-ablation}
     \caption{%
@@
     }
-    \label{fig:3}
+    \label{fig:ablation}
 \end{figure}
@@
 Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-\FIG{3}A shows the responses to target stimulation of the sample network under ablation.
+\FIG{ablation}A shows the responses to target stimulation of the sample network under ablation.
 With no plasticity, the responses are indistinguishable between sequences, yielding a DDI of 0, and indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
@@
 
-Finally, across networks and stimuli, the DDI (\FIG{3}C) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Finally, across networks and stimuli, the DDI (\FIG{ablation}C) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
 To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: T = 6.1e+03, p = 4.95e-11; TA: T = 4.87e+03, p = 0.000604), and greater too than the sum of the DDI values across both STD only and TA only models (T = 5.12e+03, p = 4.52e-05).
@@
 Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{addingSTD}G, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
-This greater contrast, of course, directly reflects the greater DDI reported in \FIG{3}.
+This greater contrast, of course, directly reflects the greater DDI reported in \FIG{ablation}.
 

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 3 -> ablation

@@
     \centering
-    \includegraphics{fig3.pdf}
+    \includegraphics{fig-ablation.pdf}
     \caption{}
-    \label{fig:3}
+    \label{fig:ablation}
 \end{figure*}
@@
 Having confirmed the presence of deviance detection in our model, we turned to an ablation approach to try to identify how the two short-term plasticity mechanisms in the model, STD and TA, enabled the networks to learn relevant regularities.
-\FIG{3}a shows the responses to target stimulation of the sample network under ablation.
+\FIG{ablation}a shows the responses to target stimulation of the sample network under ablation.
 With no plasticity, the responses are indistinguishable between sequences, yielding a DDI of 0, and indicating that stimulation did not have any long-lasting effects in membrane voltage or synaptic conductance.
@@
 
-Finally, across networks and stimuli, the DDI (\FIG{3}c) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Finally, across networks and stimuli, the DDI (\FIG{ablation}c) was exactly 0 as expected in the ablation with no plasticity, not significantly different from 0 with STD only (T = 3.2e+03, p = 0.262, median -2.1e-04; two-sided Wilcoxon, n = 120), but clearly positive for TA only (T = 5.85e+03, p = 3.1e-09, median 0.077; one-sided Wilcoxon, n = 120) and the full model (T = 5.97e+03, p = 4.75e-10, median 0.114).
 To our surprise, the DDI of the full model was significantly greater than that of either plasticity mechanism alone (STD: T = 6.1e+03, p = 4.95e-11; TA: T = 4.87e+03, p = 0.000604), and greater too than the sum of the DDI values across both STD only and TA only models (T = 5.12e+03, p = 4.52e-05).
@@
 Despite the weaker correlation, however, the full model target response remained clearly increased in \dev{} over \msc{}, as shown in \FIG{addingSTD}g, and this increase was greater in the full model than in the ablated model (T = 1.94e+03, p = 0.000723), with the grand medians of $\Delta \R{A}{}$ differing by 94\%.
-This greater contrast, of course, directly reflects the greater DDI reported in \FIG{3}.
+This greater contrast, of course, directly reflects the greater DDI reported in \FIG{ablation}.
 

### 2023-03-20 paper-1/elife.tex
Better fig name: 2 -> DD

@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig2}
+    \includegraphics[width=\linewidth]{fig-dd}
     \caption{%
@@
     }
-    \label{fig:2}
+    \label{fig:DD}
 \end{figure}
 
-To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{2}A.
+To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{DD}A.
 Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color.
@@
 The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster.
-We sum these patterns up in \FIG{2}B, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
+We sum these patterns up in \FIG{DD}B, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
 Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.
@@
 To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences.
-\FIG{2}C shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
+\FIG{DD}C shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
 The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15).
-Accordingly, the DDI (\FIG{2}D) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Accordingly, the DDI (\FIG{DD}D) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
 This shows that our model was well behaved and suitable for an investigation of deviance detection.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 2 -> DD

@@
     \centering
-    \includegraphics{fig2.pdf}
+    \includegraphics{fig-dd.pdf}
     \caption{}
-    \label{fig:2}
+    \label{fig:DD}
 \end{figure*}
@@
 
-To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{2}a.
+To show that our model is capable of deviance detection, we first show the responses of the sample network in all sequences (\dev{}, \msc{}, and \std{}) in \FIG{DD}a.
 Here, each row corresponds to one trial, aligned to stimulus input at t = 0, with the number of spikes in each millisecond bin indicated by color.
@@
 The \std{} sequence responses show a later, more diffuse activity peak around 20 ms, and the \msc{} sequence responses lack any obvious activity cluster.
-We sum these patterns up in \FIG{2}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
+We sum these patterns up in \FIG{DD}b, where we show the mean network activity, in terms of spikes per millisecond bin, across trials of each type.
 Clearly, the target response in the \dev{} sequence is greater than either the \msc{} or the \std{} response.
@@
 To show that deviance detection is robust to network and stimulus identity, we calculated the average spike count per neuron per target trial for each network and stimulus assignment, and compared these across sequences.
-\FIG{2}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
+\FIG{DD}c shows a summary of the average \dev{} and \std{} response sizes, normalized by the response in the \msc{} sequence.
 The data clearly confirmed that \dev{} responses were larger than \msc{} (T = 6.38e+03, p = 2.75e-13; one-sided Wilcoxon signed-rank test, n = 120), and \std{} responses smaller (T = 687, p = 6.42e-15).
-Accordingly, the DDI (\FIG{2}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
+Accordingly, the DDI (\FIG{DD}d) was positive in most networks (T = 5.97e+03, p = 4.75e-10, median 0.114).
 This shows that our model was well behaved and suitable for an investigation of deviance detection.

### 2023-03-20 paper-1/elife.tex
Better fig name: 1 -> setup

@@
 \begin{figure}
-    \includegraphics[width=\linewidth]{fig1}
+    \includegraphics[width=\linewidth]{fig-setup}
     \caption{%
@@
     }
-    \label{fig:1}
+    \label{fig:setup}
 \end{figure}
 
-\FIG{1}A shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
+\FIG{setup}A shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
 The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \citep{Stimberg2020-go}, and simulated with an integration time step of 1~ms.
@@
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-\FIG{1}B shows a representative example of the resulting network structure.
-In \FIG{1}C, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
+\FIG{setup}B shows a representative example of the resulting network structure.
+In \FIG{setup}C, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
 Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response: Networks where fewer than 500 neurons responded to stimulation at any of the five sites after full recovery were discarded.
@@
 
-To investigate deviance detection, we used a classical oddball paradigm as illustrated in \FIG{1}D, presenting a total of 500 stimuli at regular intervals of 500~ms.
+To investigate deviance detection, we used a classical oddball paradigm as illustrated in \FIG{setup}D, presenting a total of 500 stimuli at regular intervals of 500~ms.
 Target stimuli (labeled ``A'' or ``target'' hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The ``standard'' sequence (\std{}), consisting of 400 presentations of A and 100 presentations of B; the ``deviant'' sequence (\dev{}) with 100 A and 400 B, and the many-standards control sequence (\msc{}) with 100 presentations of each of the stimuli A through E.
@@
 Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
-A and B stimulus locations are highlighted in the left plot; see \FIG{1} for the locations of the remaining stimuli used in the \msc{} sequence.
+A and B stimulus locations are highlighted in the left plot; see \FIG{setup} for the locations of the remaining stimuli used in the \msc{} sequence.
         \textbf{B} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel A.

### 2023-03-20 paper-1/manuscript.tex
Better fig name: 1 -> setup

@@
     \centering
-    \includegraphics{fig1.pdf}
+    \includegraphics{fig-setup.pdf}
     \caption{}
-    \label{fig:1}
+    \label{fig:setup}
 \end{figure*}
@@
 
-\FIG{1}a shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
+\FIG{setup}a shows the internal and presynaptic dynamics of an excitatory neuron in a contrived example for the purpose of illustration.
 The model was implemented in Brian2 \citep{Stimberg2019-tc}, accelerated with Brian2GeNN \citep{Stimberg2020-go}, and simulated with an integration time step of 1~ms.
@@
 Stimulation sites were evenly distributed 2.5~mm from the center of the dish, and stimulation delivered a one-time increase in $g_e$ to the 10 closest neurons, sufficient to trigger 2-3 spikes each.
-\FIG{1}b shows a representative example of the resulting network structure.
-In \FIG{1}c, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
+\FIG{setup}b shows a representative example of the resulting network structure.
+In \FIG{setup}c, we show the distance from the neurons stimulated at site 1 to each neuron in this network, in terms of the minimum number of excitatory synapses that need to be traversed to reach the target, in order to give an intuition for how activity may spread.
 Networks were pseudo-randomly generated according to the above scheme and screened for minimal stimulus response: Networks where fewer than 500 neurons responded to stimulation at any of the five sites after full recovery were discarded.
@@
 
-To investigate deviance detection, we used a classical oddball paradigm as illustrated in \FIG{1}d, presenting a total of 500 stimuli at regular intervals of 500~ms.
+To investigate deviance detection, we used a classical oddball paradigm as illustrated in \FIG{setup}d, presenting a total of 500 stimuli at regular intervals of 500~ms.
 Target stimuli (labeled ``A'' or ``target'' hereafter) and distractor stimuli (labeled B to E) were presented in three randomized sequences: The ``standard'' sequence (\std{}), consisting of 400 presentations of A and 100 presentations of B; the ``deviant'' sequence (\dev{}) with 100 A and 400 B, and the many-standards control sequence (\msc{}) with 100 presentations of each of the stimuli A through E.
@@
 Left and middle column, raw averages in the \dev{} sequence ($\T[dev]{A}{i}$) and \msc{} sequence ($\T[msc]{A}{i}$), respectively; right column, contrast between these two ($\Delta \T{A}{i}$).
-A and B stimulus locations are highlighted in the left plot; see \FIG{1} for the locations of the remaining stimuli used in the \msc{} sequence.
+A and B stimulus locations are highlighted in the left plot; see \FIG{setup} for the locations of the remaining stimuli used in the \msc{} sequence.
 \textbf{b} Average response (in spikes per trial) to non-target trials ($\R{N}{i}$, with $N$ meaning B in \dev{}, and B, C, D, and E in \msc{}). Columns are arranged as in panel a.

### 2023-03-20 paper-1/elife.tex
Missed fig8 label
Missed fig8 supplement references

@@
 We found that, while the correlation was strong and very clearly positive in the trial immediately following target (Pearson's $\rho$ = -0.984, p = 0; Wald test, n = 800), it was almost completely lost by the next trial ($\rho$ = -0.0525, p = 0.138) under the combined influence of decay and unrelated activity in the intervening trial.
-Similar behavior was also seen in the \msc{} sequence, as shown in \FIGSUPP[8]{correlation-decay}.
+Similar behavior was also seen in the \msc{} sequence, as shown in \FIGSUPP[synergy]{correlation-decay}.
 
@@
     }
-    \label{fig:8}
+    \label{fig:synergy}
     % 
@@
 We confirmed this relationship across networks and stimuli, see \FIG{synergy}C, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
-We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \FIGSUPP[8]{noTA}).
+We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \FIGSUPP[synergy]{noTA}).
 

### 2023-03-20 paper-1/elife.tex
Fixed figsupp labels for uniqueness

@@
     }
-    {\includegraphics[width=\linewidth]{fig-threshold-initial-withSTD}}\label{figsupp:withSTD}
+    {\includegraphics[width=\linewidth]{fig-threshold-initial-withSTD}}\label{figsupp:threshold-initial-withSTD}
 \end{figure}
@@
     }
-    {\includegraphics[width=\linewidth]{fig-thresholds-follow-std-withSTD}}\label{figsupp:withSTD}
+    {\includegraphics[width=\linewidth]{fig-thresholds-follow-std-withSTD}}\label{figsupp:thresholds-follow-std-withSTD}
 \end{figure}
@@
     }
-    {\includegraphics[width=\linewidth]{fig-early-adaptation-withSTD}}\label{figsupp:withSTD}
+    {\includegraphics[width=\linewidth]{fig-early-adaptation-withSTD}}\label{figsupp:adaptation-withSTD}
 \end{figure}
@@
     }
-    {\includegraphics[width=\linewidth]{fig-synergy-noTA}}\label{figsupp:noTA}
+    {\includegraphics[width=\linewidth]{fig-synergy-noTA}}\label{figsupp:synergy-noTA}
     %
@@
 We confirmed this relationship across networks and stimuli, see \FIG{synergy}C, and found that greater depression in \dev{} predicted a lower response in most networks (T = 724, p = 1.37e-14; one-sided Wilcoxon, n = 120).
-We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \FIGSUPP[synergy]{noTA}).
+We note that calculating this same correlation in the model with STD only yields no significantly negative dependence (T = 3.99e+03, p = 0.828; see \FIGSUPP[synergy]{synergy-noTA}).
 

### 2023-03-20 conf/isi3.py
Cleanup

@@ -1,17 +1,0 @@
-from .params import params
-
-N_networks = 30
-N_templates = 5
-N_templates_with_dynamics = 1
-STDs = (0,1)
-TAs = (0,1)
-ISIs = (100, 250, 500, 1000, 2000)
-fbase = '/data/felix/culture/isi3_'
-fname = fbase + 'net{net}_isi{isi}_STD{STD}_TA{TA}_templ{templ}.h5'
-netfile = fbase + 'net{net}.h5'
-digestfile = fbase + '{kind}.h5'
-
-stimuli = {key: j for j, key in enumerate('ABCDE')}
-pairings=(('A','B'), ('C','E'))
-
-contrib_cutoff = .8

### 2023-03-20 conf/isi3_gpu0.py

@@ -1,5 +1,0 @@
-from .isi3 import *
-
-gpuid = 0
-N_templates = 4  # stop_at
-start_at = dict(net=26, isi=100, STD=0, TA=0, templ=2)

### 2023-03-20 conf/isi3_gpu1.py

@@ -1,4 +1,0 @@
-from .isi3 import *
-
-gpuid = 1
-start_at = dict(net=0, isi=100, STD=0, TA=0, templ=4)

### 2023-03-20 conf/isi3_mouse.py

@@ -1,7 +1,0 @@
-from .isi3 import *
-
-fbase_mouse = 'data/isi3_'
-fname = fname.replace(fbase, fbase_mouse)
-netfile = netfile.replace(fbase, fbase_mouse)
-digestfile = digestfile.replace(fbase, fbase_mouse)
-fbase = fbase_mouse

### 2023-03-20 conf/isi4.py

@@ -1,19 +1,0 @@
-from .params import params
-
-N_networks = 1
-N_templates = 5
-N_templates_with_dynamics = 1
-STDs = (0,1)
-TAs = (0,1)
-ISIs = (100, 500, 1000)
-fbase = 'data/isi4_'
-fname = fbase + 'net{net}_isi{isi}_STD{STD}_TA{TA}_templ{templ}.h5'
-netfile = fbase + 'net{net}.h5'
-digestfile = fbase + '{kind}.h5'
-
-nets = {0: 'presynaptic_events_singular.h5'}
-
-stimuli = {key: j for j, key in enumerate('BCDEA')}
-pairings=(('A','B'), ('C','E'))
-
-contrib_cutoff = .8

### 2023-03-20 conf/isi4_brutus.py

@@ -1,9 +1,0 @@
-from .isi4 import *
-
-gpuid=0
-
-fbase_brutus = '/data/felix/culture/isi4_'
-fname = fname.replace(fbase, fbase_brutus)
-netfile = netfile.replace(fbase, fbase_brutus)
-digestfile = digestfile.replace(fbase, fbase_brutus)
-fbase = fbase_brutus

### 2023-03-20 dish.py

@@ -1,87 +1,0 @@
-import numpy as np
-from brian2.only import ms
-dish = {
-    'dt': 10*ms,
-    'dev': np.asarray([
-        3.587385019710907,
-        5.153745072273325,
-        8.885676741130093,
-        9.579500657030225,
-        9.442838370565047,
-        9.127463863337715,
-        8.643889618922472,
-        7.971090670170829,
-        6.909329829172142,
-        5.816031537450724,
-        4.743758212877793,
-        3.608409986859395,
-        2.5676741130091987,
-        1.6215505913271997,
-        0.9382391590013146,
-        0.5072273324572922,
-        0.27595269382391585,
-        0.16031537450722766,
-        0.07621550591327164,
-        0.03416557161629363,
-        0.023653088042049575,
-        0.0026281208935596823,
-        0,0,0,0,0,
-    ]),
-
-    'msc': np.asarray([
-        3.0722733245729312,
-        1.8948751642575559,
-        1.8738501971090678,
-        2.052562417871222,
-        2.4730617608409986,
-        2.83048620236531,
-        2.977660972404731,
-        2.9671484888304853,
-        2.756898817345598,
-        2.4730617608409986,
-        2.157687253613666,
-        1.7792378449408677,
-        1.3902759526938233,
-        0.9592641261498027,
-        0.6859395532194483,
-        0.423127463863338,
-        0.27595269382391585,
-        0.18134034165571578,
-        0.13929040735873777,
-        0.0867279894875157,
-        0.055190538764783525,
-        0.04467805519053769,
-        0.013140604467805517,
-        0,0,0
-    ]),
-
-    'std': np.asarray([
-        2.935611038107753,
-        1.4848883048620234,
-        1.2956636005256232,
-        1.264126149802891,
-        1.3377135348226012,
-        1.3902759526938233,
-        1.4323258869908013,
-        1.4638633377135353,
-        1.4743758212877793,
-        1.4428383705650472,
-        1.3797634691195793,
-        1.2431011826544012,
-        1.0854139290407367,
-        0.9382391590013146,
-        0.7805519053876466,
-        0.6438896189224703,
-        0.5282522996057821,
-        0.41261498028909216,
-        0.32851511169513614,
-        0.2654402102496718,
-        0.1918528252299616,
-        0.1498028909329836,
-        0.10775295663600382,
-        0.07621550591327164,
-        0.04993429697766061,
-        0.03416557161629363,
-        0.023653088042049575,
-    ])
-}

### 2023-03-20 process_nspikes.py
Scripted nspikes processing

@@ -1,0 +1,36 @@
+import numpy as np
+import deepdish as dd
+
+from readout import load_results
+from digest import conds
+
+out_fname = 'nspikes.h5'
+
+
+def get_nspikes(cfg, isi, templ):
+    '''
+    Retrieves the average number of spikes per trial, summed across the entire network.
+    Structure: {STD: {TA: {cond: ndarray(nets*pairs*2)}}}
+    '''
+    nspikes = {STD: {TA: {cond: [] for cond in conds} for TA in cfg.TAs} for STD in cfg.STDs}
+    for STD in cfg.STDs:
+        for TA in cfg.TAs:
+            for net in range(cfg.N_networks):
+                res = load_results(cfg.fname.format(net=net, isi=isi, STD=STD, TA=TA, templ=templ), compress=True, process_dynamics=False)
+                for ipair, pair in enumerate(cfg.pairings):
+                    for istim, stim in enumerate(pair):
+                        for cond in conds:
+                            data = res['spikes'][ipair][stim][cond]
+                            nspikes[STD][TA][cond].append(data['nspikes'].sum(1).mean())
+                            
+            nspikes[STD][TA] = {cond: np.asarray(x) for cond, x in nspikes[STD][TA].items()}
+    return nspikes
+
+
+if __name__ == '__main__':
+    import conf.isi5_500 as cfg
+    isi = cfg.ISIs[0]
+    templ = 0
+
+    nspikes = get_nspikes(cfg, isi, templ)
+    dd.io.save(out_fname, nspikes)

### 2023-03-20 digest.py
rm Reset as a voltage measure

@@ -22,13 +22,11 @@ def get_voltages(params, dynamics, overflow=None):
     depression = ensure_unit(dynamics['u'] - dynamics['v'], volt)
     threshold = ensure_unit(dynamics['th_adapt'], volt)
     synapses = ensure_unit(dynamics['vsyn'], volt) + depression  # vsyn + depr = vsyn + u-v = usyn
-    reset = ensure_unit(dynamics['v'] - dynamics['vsyn'], volt) - params['v_rest']
     vm = ensure_unit(dynamics['v'], volt)
     return {
         'Depression': depression,
         'Threshold': threshold,
         'Synapses': synapses,
-        'Reset': reset,
         'Vm': vm}
 
 

### 2023-03-20 plot_utils.py

@@ -16,7 +16,6 @@ colors = {
     'Depression': '#9b19f5',
     'Threshold': 'C2',
     'Synapses': 'orange',
-    'Reset': 'slategrey',
     'Vm': 'orange'
 }
 
@@ -42,7 +41,6 @@ labels = {
     'Depression': '$V_{STD}$ (mV)',
     'Threshold': f'${T()}$ (mV)',
     'Synapses': '$V_{syn}$ (mV)',
-    'Reset': '$V_{reset}$ (mV)',
     'Vm': '$V_m$ (mV)'
 }
 

### 2023-03-20 digest.py
rm Synapses as a voltage measure

@@ -15,18 +15,16 @@ from readout import load_results
 
 
 conds = ('std', 'msc', 'dev')
-voltage_measures = ('Activity', 'Depression', 'Threshold')#, 'Synapses')
+voltage_measures = ('Activity', 'Depression', 'Threshold')
 
 
 def get_voltages(params, dynamics, overflow=None):
     depression = ensure_unit(dynamics['u'] - dynamics['v'], volt)
     threshold = ensure_unit(dynamics['th_adapt'], volt)
-    synapses = ensure_unit(dynamics['vsyn'], volt) + depression  # vsyn + depr = vsyn + u-v = usyn
     vm = ensure_unit(dynamics['v'], volt)
     return {
         'Depression': depression,
         'Threshold': threshold,
-        'Synapses': synapses,
         'Vm': vm}
 
 

### 2023-03-20 plot_utils.py

@@ -15,7 +15,6 @@ colors = {
     'msc': 'b',
     'Depression': '#9b19f5',
     'Threshold': 'C2',
-    'Synapses': 'orange',
     'Vm': 'orange'
 }
 
@@ -40,7 +39,6 @@ labels = {
     'pspike': 'Spike probability',
     'Depression': '$V_{STD}$ (mV)',
     'Threshold': f'${T()}$ (mV)',
-    'Synapses': '$V_{syn}$ (mV)',
     'Vm': '$V_m$ (mV)'
 }
 

### 2023-03-20 process_contrasts.py
Scripted contrasts processing

@@ -1,0 +1,70 @@
+import numpy as np
+from brian2.only import *
+import deepdish as dd
+
+from digest import conds, get_voltages
+from readout import load_results
+
+out_fname = 'contrasts.h5'
+
+
+def get_contrasts(cfg, isi, templ):
+    '''
+    Calculate voltage measure contrasts (dev - msc), weighted by pspike contrast (dev - msc),
+    and masked by pspike contrast (dev - msc > 0.05)
+    Structure: {STD: {TA: {measure: ndarray((nets*pairs*2, time))}}}
+    '''
+    measures = ('Vm', 'Depression', 'Threshold')
+    contrasts = {STD: {TA: {measure: [] for measure in measures} for TA in cfg.TAs} for STD in cfg.STDs}
+    tmax_v = {STD: {TA: 0 for TA in cfg.TAs} for STD in cfg.STDs}
+    print('get_contrasts')
+    for STD in cfg.STDs:
+        for TA in cfg.TAs:
+            print(f'STD {STD}, TA {TA} ...')
+            for net in range(cfg.N_networks):
+                res = load_results(
+                    cfg.fname.format(net=net, isi=isi, STD=STD, TA=TA, templ=templ),
+                    compress=True, dynamics_supplements={'u': 'v', 'th_adapt': 0})
+                for ipair, pair in enumerate(cfg.pairings):
+                    for istim, stim in enumerate(pair):
+                        hist = {cond: {} for cond in conds}
+                        for cond in conds:
+                            data = res['dynamics'][ipair][stim][cond]
+                            # Average over trials, leaving measures shaped (neuron, time):
+                            V = {measure: v.mean(1) for measure, v in get_voltages(cfg.params, data).items()}
+                            for measure, v in V.items():
+                                hist[cond][measure] = v
+                                tmax_v[STD][TA] = max(tmax_v[STD][TA], v.shape[1])
+                            hist[cond]['pspike'] = res['spikes'][ipair][stim][cond]['spike_hist']
+                        
+                        # Spike probability contrast
+                        pdiff = hist['dev']['pspike'] - hist['msc']['pspike']
+                        # Mask out entries where the pdiff contrast <= 0.05
+                        pdiff = np.where(pdiff > .05, pdiff, np.nan)
+
+                        # Weight & mask measures by pspike
+                        for measure in measures:
+                            # Calculate raw contrast (dev - msc)
[…]

### 2023-03-20 process_suppression.py
Scripted suppression processing

@@ -1,0 +1,82 @@
+import numpy as np
+from brian2.only import *
+import deepdish as dd
+
+from readout import load_results
+
+out_fname = 'suppression.TA{TA}.h5'
+
+
+def get_bspikes(res, episode):
+    b = np.zeros(np.asarray(res['raw_dynamics']['v'].shape)[[0,2,3]], bool)
+    for itrial, (i, spike_t) in enumerate(zip(*[res['raw_spikes'][episode][f'pulsed_{k}'] for k in 'it'])):
+        t = (spike_t / res['params']['dt'] + .5).astype(int)
+        b[i, itrial, t] = True
+    return b
+
+
+def get_suppression_data(res, Wb, ipair, istim, with_threshold=True, with_depression=True):
+    Response = {AB: {} for AB in 'AB'}
+    Threshold = {AB: {} for AB in 'AB'}
+    Depression = {AB: {} for AB in 'AB'}
+    
+    target = res['pairs'][ipair][f'S{istim + 1}']
+    nontarget = res['pairs'][ipair][f'S{(istim+1)%2 + 1}']
+
+    # Response
+    for cond in ('dev', 'msc', 'std'):
+        episode = res['pairs'][ipair][cond][target]
+        spikes = get_bspikes(res, episode)
+        for isB, AB in enumerate('AB'):
+            trials = res['sequences'][episode] == res['stimuli'][nontarget if isB else target]
+            Response[AB][cond] = spikes[:, trials].sum(2).mean(1)
+    
+    # Suppression
+    for cond in ('dev', 'msc'):
+        episode = res['pairs'][ipair][cond][target]
+        for isB, AB in enumerate('AB'):
+            trials = res['sequences'][episode] == res['stimuli'][nontarget if isB else target]
+            
+            if with_threshold:
+                Threshold[AB][cond] = res['raw_dynamics']['th_adapt'][:, episode, trials, 0].mean(1)*volt/mV
+            
+            if with_depression:
+                xr = res['raw_dynamics']['neuron_xr'][:, episode, trials, 0].mean(1)
+                Depression[AB][cond] = np.einsum('eo,e->o', Wb, np.mean([R for R in Response[AB].values()], 0) * (1-xr))
+    
+    return Response, Threshold, Depression
+
+
+def get_suppression(cfg, isi, templ, TA):
+    isi, TA, templ = 500, 1, 0
+    Rdata, TAdata, Ddata = [{STD: [] for STD in cfg.STDs} for _ in range(3)]
[…]

### 2023-03-20 process_stats.py
Scripted stats processing

@@ -1,0 +1,123 @@
+import numpy as np
+from brian2.only import *
+import deepdish as dd
+
+from digest import conds
+import isi_indexing as ii
+from readout import load_results
+
+out_fname = 'stats.h5'
+
+
+def get_bspikes(res, episode):
+    b = np.zeros(np.asarray(res['raw_dynamics']['v'].shape)[[0,2,3]], bool)
+    for itrial, (i, spike_t) in enumerate(zip(*[res['raw_spikes'][episode][f'pulsed_{k}'] for k in 'it'])):
+        t = (spike_t / res['params']['dt'] + .5).astype(int)
+        b[i, itrial, t] = True
+    return b
+
+
+def get_stats(cfg, isi, templ, early=100):
+    '''
+    Computes a bunch of stats across networks and stimuli.
+    Requires data coming from `process_suppression.py` in the default file locations.
+    Args:
+        early: int, number of neurons to consider in the "early" subpopulation
+    '''
+    Astim = [stim for net in range(cfg.N_networks) for pair in cfg.pairings for stim in pair]
+    Bstim = [stim for net in range(cfg.N_networks) for pair in cfg.pairings for stim in pair[::-1]]
+    Aidx = np.asarray([cfg.stimuli[stim] for stim in Astim])
+    Bidx = np.asarray([cfg.stimuli[stim] for stim in Bstim])
+
+    TAdata = dd.io.load('suppression.TA1.h5', '/T')
+    Ddata = {0: dd.io.load('suppression.TA0.h5', '/D')[1], 1: dd.io.load('suppression.TA1.h5', '/D')[1]}
+    
+    shape = cfg.N_networks*len(cfg.pairings)*2, cfg.params['N'], cfg.params['sequence_count']*cfg.params['sequence_length']
+    dXnotA, dXA, stimchange, adaptation = [{STD: {TA: np.empty(shape[:2]) for TA in cfg.TAs} for STD in cfg.STDs} for _ in range(4)]
+    dXnotA_early, dXnotA_late = [{STD: {TA: np.empty(shape[0]) for TA in cfg.TAs} for STD in cfg.STDs} for _ in range(2)]
+    dTA_B = {STD: np.empty(shape[:2]) for STD in cfg.STDs}
+    dTA_B_early, dTA_B_late = [{STD: np.empty(shape[0]) for STD in cfg.STDs} for _ in range(2)]
[…]

### 2023-03-20 process_contrasts.py
Better interface for process_* scripts

@@ -61,10 +61,21 @@ def get_contrasts(cfg, isi, templ):
     return contrasts
 
 
-if __name__ == '__main__':
-    import conf.isi5_500 as cfg
-    isi = cfg.ISIs[0]
-    templ = 0
-
+def process_to_disk(cfg, isi = None, templ = 0):
+    if isi is None:
+        isi = cfg.ISIs[0]
     contrasts = get_contrasts(cfg, isi, templ)
     dd.io.save(out_fname, contrasts)
+
+
+if __name__ == '__main__':
+    import sys
+    import importlib
+
+    if len(sys.argv) > 1:
+        conf = '.'.join(sys.argv[1].split('.')[0].split('/'))
+    else:
+        conf = 'conf.isi5_500'
+
+    cfg = importlib.import_module(conf)
+    process_to_disk(cfg)

### 2023-03-20 process_nspikes.py

@@ -27,10 +27,21 @@ def get_nspikes(cfg, isi, templ):
     return nspikes
 
 
-if __name__ == '__main__':
-    import conf.isi5_500 as cfg
-    isi = cfg.ISIs[0]
-    templ = 0
-
+def process_to_disk(cfg, isi = None, templ = 0):
+    if isi is None:
+        isi = cfg.ISIs[0]
     nspikes = get_nspikes(cfg, isi, templ)
     dd.io.save(out_fname, nspikes)
+
+
+if __name__ == '__main__':
+    import sys
+    import importlib
+
+    if len(sys.argv) > 1:
+        conf = '.'.join(sys.argv[1].split('.')[0].split('/'))
+    else:
+        conf = 'conf.isi5_500'
+
+    cfg = importlib.import_module(conf)
+    process_to_disk(cfg)

### 2023-03-20 process_stats.py

@@ -20,7 +20,7 @@ def get_bspikes(res, episode):
 def get_stats(cfg, isi, templ, early=100):
     '''
     Computes a bunch of stats across networks and stimuli.
-    Requires data coming from `process_suppression.py` in the default file locations.
+    Requires that process_suppression.process_to_disk has been run with the same args.
     Args:
         early: int, number of neurons to consider in the "early" subpopulation
     '''
@@ -114,10 +114,21 @@ def get_stats(cfg, isi, templ, early=100):
         'dD_B_late': dD_B_late}
 
 
-if __name__ == '__main__':
-    import conf.isi5_500 as cfg
-    isi = cfg.ISIs[0]
-    templ = 0
-    
+def process_to_disk(cfg, isi = None, templ = 0):
+    if isi is None:
+        isi = cfg.ISIs[0]
     stats = get_stats(cfg, isi, templ)
     dd.io.save(out_fname, stats)
+
+
+if __name__ == '__main__':
+    import sys
+    import importlib
+
+    if len(sys.argv) > 1:
+        conf = '.'.join(sys.argv[1].split('.')[0].split('/'))
+    else:
+        conf = 'conf.isi5_500'
+
+    cfg = importlib.import_module(conf)
+    process_to_disk(cfg)

### 2023-03-20 process_suppression.py

@@ -70,13 +70,25 @@ def get_suppression(cfg, isi, templ, TA):
     return Rdata, TAdata, Ddata
 
 
-if __name__ == '__main__':
-    import conf.isi5_500 as cfg
-    isi = cfg.ISIs[0]
-    templ = 0
+def process_to_disk(cfg, isi = None, templ = 0):
+    if isi is None:
+        isi = cfg.ISIs[0]
 
     R, T, D = get_suppression(cfg, isi, templ, 0)
     dd.io.save(out_fname.format(TA=0), {'R': R, 'D': D})
 
     R, T, D = get_suppression(cfg, isi, templ, 1)
     dd.io.save(out.fname.format(TA=1), {'R': R, 'T': T, 'D': D})
+
+
+if __name__ == '__main__':
+    import sys
+    import importlib
+
+    if len(sys.argv) > 1:
+        conf = '.'.join(sys.argv[1].split('.')[0].split('/'))
+    else:
+        conf = 'conf.isi5_500'
+
+    cfg = importlib.import_module(conf)
+    process_to_disk(cfg)

### 2023-03-20 process_all.py
Added process_all for good measure.

@@ -1,0 +1,32 @@
+import process_nspikes as nspikes
+import process_contrasts as contrasts
+import process_suppression as suppr
+import process_stats as stats
+
+
+def process_to_disk(cfg, isi, templ):
+    for module in (nspikes, contrasts, suppr, stats):
+        module.process_to_disk(cfg, isi, templ)
+
+
+if __name__ == '__main__':
+    import sys
+    import importlib
+
+    if len(sys.argv) > 1:
+        conf = '.'.join(sys.argv[1].split('.')[0].split('/'))
+    else:
+        conf = 'conf.isi5_500'
+    
+    if len(sys.argv) > 2:
+        isi = int(sys.argv[2])
+    else:
+        isi = None
+    
+    if len(sys.argv) > 3:
+        templ = int(sys.argv[3])
+    else:
+        templ = 0
+
+    cfg = importlib.import_module(conf)
+    process_to_disk(cfg, isi, templ)

### 2023-03-20 process_all.py
Added dist/spatial stuff to processing scripts

@@ -2,10 +2,11 @@ import process_nspikes as nspikes
 import process_contrasts as contrasts
 import process_suppression as suppr
 import process_stats as stats
+import process_distance as distance
 
 
 def process_to_disk(cfg, isi, templ):
-    for module in (nspikes, contrasts, suppr, stats):
+    for module in (nspikes, contrasts, suppr, distance, stats):
         module.process_to_disk(cfg, isi, templ)
 
 

### 2023-03-20 process_distance.py

@@ -1,0 +1,248 @@
+import numpy as np
+from brian2.only import *
+import deepdish as dd
+
+from readout import load_results
+from digest import conds
+import spatial
+
+out_fname = 'distance.h5'
+
+
+def get_bspikes(res, episode):
+    b = np.zeros(np.asarray(res['raw_dynamics']['v'].shape)[[0,2,3]], bool)
+    for itrial, (i, spike_t) in enumerate(zip(*[res['raw_spikes'][episode][f'pulsed_{k}'] for k in 'it'])):
+        t = (spike_t / res['params']['dt'] + .5).astype(int)
+        b[i, itrial, t] = True
+    return b
+
+
+def get_neuron_distances(cfg, netf, stim):
+    x,y = [xy[cfg.stimuli[stim]] for xy in spatial.create_stimulus_locations(cfg.params)]
+    return np.sqrt((netf['X']*meter-x)**2 + (netf['Y']*meter-y)**2)
+
+
+def get_trial_average(res, episode, stim_distances, stimidx=None):
+    spike_distances, spike_counts, target = np.empty((3, len(res['sequences'][episode])))
+    target = target.astype(bool)
+    spikes = get_bspikes(res, episode)
+    for trial, idx_stimulated in enumerate(res['sequences'][episode]):
+        dist_from_stimulated = stim_distances[idx_stimulated]
+        trial_spikes = spikes[:, trial]
+        trial_distances = np.broadcast_to(dist_from_stimulated[:, None], trial_spikes.shape)[trial_spikes]
+        spike_distances[trial] = np.median(trial_distances)
+        target[trial] = idx_stimulated == stimidx
+        spike_counts[trial] = trial_spikes.sum()
+    return spike_distances, spike_counts, target
+
+
+def get_distconf(cfg, K = 1.4, distbins = 200, rough_cutoff = 2*mm):
+    maxdist = cfg.params['r_dish'] + cfg.params['stim_distribution_radius']
+    
+    histargs = {'bins': distbins, 'range': (0, maxdist/mm)}
+    _, histedges = np.histogram([0], **histargs)
+
+    histargs_rough = {'bins': [0, rough_cutoff/mm, maxdist/mm]}
+    _, histedges_rough = np.histogram([0], **histargs_rough)
+    
+    return {'full': {'args': histargs, 'edges': histedges},
+            'rough': {'args': histargs_rough, 'edges': histedges_rough},
+            'K': K}
+
+
[…]

### 2023-03-22 process_distance.py
Take means over trials/nets in process_distance
This shrinks distance.h5 to a githubbable size.

@@ -138,8 +138,6 @@ def get_distance(cfg, isi, templ, **kwargs):
     TA, STD = 1, 1
     distconf = get_distconf(cfg, **kwargs)
 
-    distances, counts = [{cond: np.full((cfg.N_networks*len(cfg.pairings)*2, cfg.params['sequence_length']*cfg.params['sequence_count']), np.nan)
-                                for cond in conds} for _ in range(2)]
     disthist = {cond: np.full((cfg.N_networks*len(cfg.pairings)*2, cfg.params['sequence_length']*cfg.params['sequence_count'], distconf['full']['args']['bins']), np.nan)
                 for cond in conds}
     distrough = {cond: np.full((cfg.N_networks*len(cfg.pairings)*2, cfg.params['sequence_length']*cfg.params['sequence_count'], 2), np.nan)
@@ -180,8 +178,6 @@ def get_distance(cfg, isi, templ, **kwargs):
                 ep_fine_STD, ep_rough_STD = get_STD_distrib(distconf, res, episode, stim_distances, R, Wb)
                 i = 4*net + 2*ipair + istim
 
-                distances['dev'][i, target] = dist[target]
-                counts['dev'][i, target] = count[target]
                 disthist['dev'][i, target] = hist[target]
                 distrough['dev'][i, target] = rough[target]
                 fine_TA['dev'][i, target] = ep_fine_TA[target]
@@ -189,8 +185,6 @@ def get_distance(cfg, isi, templ, **kwargs):
                 fine_STD['dev'][i, target] = ep_fine_STD[target]
                 rough_STD['dev'][i, target] = ep_rough_STD[target]
 
-                distances['std'][i, ~target] = dist[~target]
-                counts['std'][i, ~target] = count[~target]
                 disthist['std'][i, ~target] = hist[~target]
                 distrough['std'][i, ~target] = rough[~target]
                 fine_TA['std'][i, ~target] = ep_fine_TA[~target]
@@ -199,8 +193,6 @@ def get_distance(cfg, isi, templ, **kwargs):
                 rough_STD['std'][i, ~target] = ep_rough_STD[~target]
 
                 msc_target = res['sequences'][0] == cfg.stimuli[stim]
-                distances['msc'][i, msc_target] = msc_dist[msc_target]
-                counts['msc'][i, msc_target] = msc_count[msc_target]
[…]

### 2023-03-22 process_contrasts.py
Ensure process scripts flush stdout on net report

@@ -48,7 +48,7 @@ def get_contrasts(cfg, isi, templ):
                             c = (hist['dev'][measure] - hist['msc'][measure])/mV
                             # Weight with pdiff contrast, and sum valid entries over neurons
                             contrasts[STD][TA][measure].append(np.nansum(c*pdiff, 0))
-                print(net, end=' ')
+                print(net, end=' ', flush=True)
             
             # Stack everything in {STD: {TA: {measure: ...}}} into a (nets*pairs*2, tmax_v) ndarray
             contrasts[STD][TA] = {

### 2023-03-22 process_distance.py

@@ -201,7 +201,7 @@ def get_distance(cfg, isi, templ, **kwargs):
                 rough_STD['msc'][i, msc_target] = msc_rough_STD[msc_target]
 
                 histbase[i], roughbase[i] = get_hist_base(distconf, stim_distances, cfg.stimuli[stim])
-        print(net, end=' ')
+        print(net, end=' ', flush=True)
     
     return {
         'disthist_initial': np.nanmean(np.concatenate([d[:, 0] for d in disthist.values()], 0), 0),

### 2023-03-22 process_stats.py

@@ -92,7 +92,7 @@ def get_stats(cfg, isi, templ, early=100):
                             dD_B[TA][k] = dD
                             dD_B_early[TA][k] = np.median(dD[early_B])
                             dD_B_late[TA][k] = np.median(dD[late_B])
-                print(net, end=' ')
+                print(net, end=' ', flush=True)
     dTA_A = {STD: [] for STD in cfg.STDs}
     for STD in cfg.STDs:
         for k, ta in enumerate(TAdata[STD]):

### 2023-03-22 process_suppression.py

@@ -66,7 +66,7 @@ def get_suppression(cfg, isi, templ, TA):
                     Rdata[STD].append(data[0])
                     TAdata[STD].append(data[1])
                     Ddata[STD].append(data[2])
-            print(net, end=' ')
+            print(net, end=' ', flush=True)
     return Rdata, TAdata, Ddata
 
 

### 2023-03-22 process_nspikes.py
Added progress output to process_nspikes

@@ -13,8 +13,10 @@ def get_nspikes(cfg, isi, templ):
     Structure: {STD: {TA: {cond: ndarray(nets*pairs*2)}}}
     '''
     nspikes = {STD: {TA: {cond: [] for cond in conds} for TA in cfg.TAs} for STD in cfg.STDs}
+    print('get_nspikes ...')
     for STD in cfg.STDs:
         for TA in cfg.TAs:
+            print(f'STD {STD}, TA {TA} ...')
             for net in range(cfg.N_networks):
                 res = load_results(cfg.fname.format(net=net, isi=isi, STD=STD, TA=TA, templ=templ), compress=True, process_dynamics=False)
                 for ipair, pair in enumerate(cfg.pairings):
@@ -22,7 +24,8 @@ def get_nspikes(cfg, isi, templ):
                         for cond in conds:
                             data = res['spikes'][ipair][stim][cond]
                             nspikes[STD][TA][cond].append(data['nspikes'].sum(1).mean())
-                            
+                print(net, end=' ', flush=True)
+
             nspikes[STD][TA] = {cond: np.asarray(x) for cond, x in nspikes[STD][TA].items()}
     return nspikes
 

### 2023-03-22 process_contrasts.py
Added progress report newlines where appropriate

@@ -57,6 +57,7 @@ def get_contrasts(cfg, isi, templ):
                     for c in contrast
                 ])
                 for measure, contrast in contrasts[STD][TA].items()}
+            print()
     
     return contrasts
 

### 2023-03-22 process_distance.py

@@ -202,6 +202,7 @@ def get_distance(cfg, isi, templ, **kwargs):
 
                 histbase[i], roughbase[i] = get_hist_base(distconf, stim_distances, cfg.stimuli[stim])
         print(net, end=' ', flush=True)
+    print()
     
     return {
         'disthist_initial': np.nanmean(np.concatenate([d[:, 0] for d in disthist.values()], 0), 0),

### 2023-03-22 process_nspikes.py

@@ -13,7 +13,7 @@ def get_nspikes(cfg, isi, templ):
     Structure: {STD: {TA: {cond: ndarray(nets*pairs*2)}}}
     '''
     nspikes = {STD: {TA: {cond: [] for cond in conds} for TA in cfg.TAs} for STD in cfg.STDs}
-    print('get_nspikes ...')
+    print('get_nspikes')
     for STD in cfg.STDs:
         for TA in cfg.TAs:
             print(f'STD {STD}, TA {TA} ...')
@@ -27,6 +27,7 @@ def get_nspikes(cfg, isi, templ):
                 print(net, end=' ', flush=True)
 
             nspikes[STD][TA] = {cond: np.asarray(x) for cond, x in nspikes[STD][TA].items()}
+            print()
     return nspikes
 
 

### 2023-03-22 process_stats.py

@@ -93,6 +93,7 @@ def get_stats(cfg, isi, templ, early=100):
                             dD_B_early[TA][k] = np.median(dD[early_B])
                             dD_B_late[TA][k] = np.median(dD[late_B])
                 print(net, end=' ', flush=True)
+            print()
     dTA_A = {STD: [] for STD in cfg.STDs}
     for STD in cfg.STDs:
         for k, ta in enumerate(TAdata[STD]):

### 2023-03-22 process_suppression.py

@@ -67,6 +67,7 @@ def get_suppression(cfg, isi, templ, TA):
                     TAdata[STD].append(data[1])
                     Ddata[STD].append(data[2])
             print(net, end=' ', flush=True)
+        print()
     return Rdata, TAdata, Ddata
 
 

### 2023-03-22 readout.py
Ensure path to conf'ed fname exists on save

@@ -1,3 +1,4 @@
+import os
 from distutils.log import warn
 from importlib import import_module
 from itertools import count
@@ -286,6 +287,7 @@ def save_results(fname, rundata):
     elide = ['dynamics', 'spikes']
     if rundata.get('raw_fbase', None) is not None:
         elide += ['raw_dynamics']
+    os.makedirs(os.path.dirname(os.path.abspath(fname)), exist_ok=True)
     dd.io.save(fname, {k:v for k,v in rundata.items() if k not in elide})
 
 

### 2023-03-22 conf/isi5.py
Leave only isi5_500 conf

@@ -1,20 +1,0 @@
-from .params import params
-
-gpuid=0
-
-N_networks = 30
-N_templates = 5
-N_templates_with_dynamics = 1
-STDs = (0,1)
-TAs = (0,1)
-ISIs = (100, 250, 500, 1000, 2000)
-fbase = '/data/felix/culture/isi5_'
-fname = fbase + 'net{net}_isi{isi}_STD{STD}_TA{TA}_templ{templ}.h5'
-netfile = fbase + 'net{net}.h5'
-digestfile = fbase + '{kind}.h5'
-
-stimuli = {key: j for j, key in enumerate('ABCDE')}
-pairings=(('A','B'), ('C','E'))
-
-contrib_cutoff = .8
-minimum_active_fraction = .5

### 2023-03-22 conf/isi5_500.py

@@ -1,16 +1,19 @@
-from .isi5 import *
+from .params import params
 
-del gpuid
-
-fbase_500 = '/data/felix/culture/isi5-500/isi5_'
-fname = fname.replace(fbase, fbase_500)
-netfile = netfile.replace(fbase, fbase_500)
-digestfile = digestfile.replace(fbase, fbase_500)
-fbase = fbase_500
-
-raw_fbase = fname[:-3]
-
-ISIs = (500,)
+N_networks = 30
 N_templates = 1
+N_templates_with_dynamics = 1
+STDs = (0,1)
+TAs = (0,1)
+ISIs = (500,)
+fbase = 'data/isi5-500/isi5_'
+raw_fbase = fbase + 'net{net}_isi{isi}_STD{STD}_TA{TA}_templ{templ}'
+fname = raw_fbase + '.h5'
+netfile = fbase + 'net{net}.h5'
+digestfile = fbase + '{kind}.h5'
 
-start_at = {'newnet': False, 'newtempl': False}
+stimuli = {key: j for j, key in enumerate('ABCDE')}
+pairings=(('A','B'), ('C','E'))
+
+contrib_cutoff = .8
+minimum_active_fraction = .5

### 2023-03-22 conf/isi5_mouse.py

@@ -1,9 +1,0 @@
-from .isi5 import *
-
-fbase_mouse = 'data/isi5_'
-fname = fname.replace(fbase, fbase_mouse)
-netfile = netfile.replace(fbase, fbase_mouse)
-digestfile = digestfile.replace(fbase, fbase_mouse)
-fbase = fbase_mouse
-
-raw_fbase = fname[:-3]

### 2023-03-22 process_suppression.py
Bugfix

@@ -79,7 +79,7 @@ def process_to_disk(cfg, isi = None, templ = 0):
     dd.io.save(out_fname.format(TA=0), {'R': R, 'D': D})
 
     R, T, D = get_suppression(cfg, isi, templ, 1)
-    dd.io.save(out.fname.format(TA=1), {'R': R, 'T': T, 'D': D})
+    dd.io.save(out_fname.format(TA=1), {'R': R, 'T': T, 'D': D})
 
 
 if __name__ == '__main__':

### 2023-03-22 conf/isi5_500.py
Documented conf module

@@ -1,19 +1,35 @@
 from .params import params
 
+# Number of networks to include:
 N_networks = 30
+
+# Number of sequence templates to run:
 N_templates = 1
+
+# Number of sequence templates that save dynamics data:
 N_templates_with_dynamics = 1
+
+# Toggle STD - always an iterable of 0 and/or 1
 STDs = (0,1)
+
+# Toggle TA - always an iterable of 0 and/or 1
 TAs = (0,1)
+
+# Sequence inter-trial intervals to run (ms):
 ISIs = (500,)
+
+# Output file locations
 fbase = 'data/isi5-500/isi5_'
 raw_fbase = fbase + 'net{net}_isi{isi}_STD{STD}_TA{TA}_templ{templ}'
 fname = raw_fbase + '.h5'
 netfile = fbase + 'net{net}.h5'
 digestfile = fbase + '{kind}.h5'
 
+# Iterable of stimulus names
 stimuli = {key: j for j, key in enumerate('ABCDE')}
+
+# Iterable of stimulus pairings (each pair is run as both std/dev and dev/std)
 pairings=(('A','B'), ('C','E'))
 
-contrib_cutoff = .8
+# Network validation: minimum number of neurons spiking in response to each stimulus
 minimum_active_fraction = .5

### 2023-03-22 process_suppression.py
Bugfix

@@ -48,7 +48,6 @@ def get_suppression_data(res, Wb, ipair, istim, with_threshold=True, with_depres
 
 
 def get_suppression(cfg, isi, templ, TA):
-    isi, TA, templ = 500, 1, 0
     Rdata, TAdata, Ddata = [{STD: [] for STD in cfg.STDs} for _ in range(3)]
     print('get_suppression')
     for STD in cfg.STDs:

### 2023-03-23 demonstrator.py
Added templ, ISI to demonstrator

@@ -1,5 +1,7 @@
 demonstrator = {
     'net': 7,
     'ipair': 0,
-    'istim': 1
+    'istim': 1,
+    'isi': 500,
+    'templ': 0
 }

### 2023-03-23 paper-1/elife.tex
Manuscript bugfix

@@
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{early-adaptation}C) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{early-adaptation}D) within these two portions.
-We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{thresholds-follow-std}D).
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{N}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{N}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{thresholds-follow-std}D).
 $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).

### 2023-03-23 paper-1/manuscript.tex
Manuscript bugfix

@@
 In each network, we then identified the early and late portions with respect to stimulation in B and quantified the median non-target response contrast ($\Delta \R{N}{}$, \FIG{early-adaptation}c) and the median contrast in $\T{B}{}$ ($\Delta \T{B}{}$, \FIG{early-adaptation}d) within these two portions.
-We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{B}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{B}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{thresholds-follow-std}d).
+We found that the response in the early portion was consistently greater in \dev{} than \msc{} ($\Delta \R{N}{early} > 0$: T = 6.98e+03, p = 8.48e-19), unlike the late portion, which decreased ($\Delta \R{N}{late} < 0$: T = 295, p = 1.23e-18; see also \FIG{thresholds-follow-std}d).
 $\Delta \T{B}{}$ followed the same pattern, with the early portion exhibiting stronger TA in \dev{} (T = 4.63e+03, p = 2.21e-17), while the late portion was instead more recovered (T = 26, p = 2.97e-17).
